---
title: "Diagnostics"
author: "Mohamed Albirair"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Data

```{r}
load("R/inputs/RData/perform_testing.RData")
```

# Pre-Processing

```{r}
gbd_epi_1yr_predict %>%
      filter(sex == "Female",
             cause %in% c(cause2[-2], cause3[2]),
             location == "Uganda",
             year == 2019) %>%

      ggplot(aes(x = age,
                 # y = val,
                 y = log(pred_rates)
                 # color = sex
                 )) +
                 # shape = location,
                 # group = interaction(sex, location))) +
      # geom_point(alpha = 0.6) +
      geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
      geom_line(aes(x = age, y = log(pred_rates)), linetype = 1, color = "green") +
      # geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
      facet_grid(cause ~ measure, scales = "free_y") +  # Free y-axis scales
      labs(title = "Age-Value Regression Lines by Group",
           x = "Age", y = "Value") +
      theme_bw()
```

```{r}
gbd_epi_1yr_predict %>%
      filter(sex == "Female",
             cause %in% c(cause2[-2], cause3[2]),
             location == "Uganda",
             # measure == "Deaths",
             year == 2019) %>%

      ggplot(aes(x = age,
                 # y = val,
                 y = pred_cases
                 # color = sex
                 )) +
                 # shape = location,
                 # group = interaction(sex, location))) +
      # geom_point(alpha = 0.6) +
      geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
      geom_line(aes(x = age, y = pred_cases), linetype = 1, color = "green") +
      # geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
      facet_grid(cause ~ measure, scales = "free_y") +  # Free y-axis scales
      labs(title = "Age-Value Regression Lines by Group",
           x = "Age", y = "Value") +
      theme_bw()
```

```{r}
gbd_calc_compr <- left_join(

      # Linear model predicted counts (incid and deaths), all ages
      x  = gbd_epi_1yr_predict %>% 
            filter(year == 2019) %>% 
            mutate(calc_pred_cases = pred_rates * pop / 1e5) %>% 
            select(!pop),

      # GBD calculated counts (incid and deaths), all ages
      y = full_join(

            # GBD epi data, 5YAGs
            x  = gbd_epi_5yr %>% filter(year == 2019),

            # GBD pop data, 5YAGs
            y  = gbd_pop_5yr %>% filter(year == 2019) %>% 
                  mutate(age_gp = as.character(age_gp)),

            by = join_by(location, sex, age, age_gp, year)
      ) %>% 
            filter(metric == "Number") %>% 
            select(!c(pop, upper, lower, age_cat, age_mid, pop)),

      by = join_by(measure, location, year, cause, sex, age, age_gp)
) %>% 
      mutate(counts_diff = pred_cases - val,
             counts_ratio = pred_cases / val)
```

```{r}
gbd_calc_compr %>% 
      filter(sex == "Female",
             cause %in% c(cause2[-2], cause3[2]),
             location == "Uganda") %>% 
      filter(age <= 75) %>% 
      filter(measure != "Prevalence") %>% 

      ggplot() +
      geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
      geom_point(aes(x = age, y = val), color = "blue", alpha = 0.6) +
      geom_line(aes(x = age, y = calc_pred_cases), linetype = 1, color = "green") +
      # geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
      facet_grid(cause ~ measure, scales = "free_y") +  # Free y-axis scales
      labs(title = "Age-Value Regression Lines by Group",
           x = "Age", y = "Value") +
      theme_bw()
```

```{r}
gbd_calc_compr %>% 
      filter(sex == "Female",
             cause %in% c(cause2[-2], cause3[2]),
             location == "Uganda") %>% 
      filter(age <= 75) %>% 
      filter(measure != "Prevalence") %>% 

      ggplot() +
      geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
      geom_point(aes(x = age, y = log(val)), color = "blue", alpha = 0.6) +
      geom_line(aes(x = age, y = log(calc_pred_cases)), linetype = 1, color = "green") +
      # geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
      facet_grid(cause ~ measure, scales = "free_y") +  # Free y-axis scales
      labs(title = "Age-Value Regression Lines by Group",
           x = "Age", y = "Value") +
      theme_bw()
```

```{r}
gbd_calc_sum_compr <- left_join(

      # Linear model predicted counts (incid and deaths), all ages
      x  = gbd_epi_1yr_predict %>% 
            filter(year == 2019) %>% 
            group_by(measure, location, cause, sex) %>%
            reframe(tot_cases_calc = sum(pred_cases)),

      # GBD calculated counts (incid and deaths), all ages
      y = full_join(

            # GBD epi data, 5YAGs
            x  = gbd_epi_5yr %>% filter(year == 2019) %>% 
                  rename(age_gp = age_cat),

            # GBD pop data, 5YAGs
            y  = gbd_pop_5y %>% filter(year == 2019) %>% 
                  mutate(age_gp = as.character(age_gp)),

            by = join_by(location, sex, age_gp, year)
      ) %>%
            mutate(gbd_cases = exp(log_val) * pop) %>%
            group_by(measure, location, cause, sex) %>%
            reframe(tot_cases_gbd = sum(gbd_cases)),

      by = join_by(measure, location, cause, sex)
) %>%
      mutate(counts_diff = tot_cases_calc - tot_cases_gbd,
             counts_ratio = tot_cases_calc / tot_cases_gbd)
```


```{r}
gbd_calc_compr %>%
      ggplot() +
      geom_histogram(aes(x = counts_diff), alpha = 0.5, position = "identity", fill = "red") +
      geom_histogram(aes(x = counts_ratio), alpha = 0.5, position = "identity", fill = "blue") +
      facet_wrap(. ~ cause) +
      theme_minimal()
```


```{r}
gbd_epi %>%
      filter(year == reference_year,
             age < 95,
             measure != "Prevalence",
             metric == "Rate",
             location == "Uganda",
             cause %in% cause2) %>%
      group_by(location, sex, cause, measure) %>%
      filter(n() > 1) %>%
      ungroup() %>%
      # mutate(group_label = paste(location, sex, cause, measure, sep = "\n")) %>%
      ggplot(aes(x = age, y = log(val), color = sex)) +
      geom_point(alpha = 0.6) +
      geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
      facet_grid(cause ~ measure, scales = "free_y") +  # Free y-axis scales
      labs(title = "Age-Value Regression Lines by Group",
           x = "Age", y = "Value") +
      theme_bw()
```

# Count Calculations

## Comparison of Rates between GBD and Outputs using the `ungroup::pclm()`

```{r}
full_join(
      x  = gbd$Uganda$`Bladder cancer`$Female$`2019` %>% 
            filter(measure == "Deaths",
                   metric == "Rate",
                   age >= 15) %>% 
            complete(age = 15:95, nesting(sex, measure, year, cause, location, metric)) %>% 
            mutate(val = zoo::na.approx(val)) %>% 
            rename(cmx = val) %>% 
            select(!upper, !lower) %>% 
            mutate(source = "GBD"),
      # y  = n_cmx_deaths$`Bladder cancer`$Female %>% 
      #       filter(age >= 15) %>% 
      #       mutate(cmx = cmx * 1e5) %>% 
      #       mutate(source = "PCLM"),
      y  = full_join(
            x  = n_cases$`Bladder cancer`$Female %>% 
                  filter(age >= 15),
            y  = n_mx_N_acm$`All causes`$Female %>% 
                  filter(age >= 15),
            by = join_by(age, year, sex, location)
      ) %>% 
            mutate(cmx = counts / Nx * 1e5,
                   source = "PCLM"),
      by = join_by(age, sex, year, location, cmx, source)
) %>% 
      
      ggplot(aes(x = age, y = cmx, color = source)) +
      geom_line() +
      theme_minimal()
```

## Markov states over time (40 YOF, breast cancer)

```{r}
bsln_markov_trace %>% 
      filter(cause == cause2[1],
             # between(age, 40, 70),
             age %in% seq(40, 70, 5),
             sex == "Female") %>% 
      pivot_longer(cols = c(plcl, pcdx, pbgdx),
                   names_to = "trans",
                   values_to = "prob") %>% 
      mutate(age = as.character(age)) %>% 
      ggplot(aes(x = year, y = prob, group = age, color = age)) +
      geom_line() +
      scale_y_continuous(name = "Probability",
                         labels = scales::label_percent(accuracy = 0.01, scale = 1e2)) +
      facet_wrap(. ~ trans, scales = "free_y") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

## cause-specific Markov states, incidence and mortality rates by time (40 YOF)

```{r}
bsln_markov_trace %>% 
      filter(cause %in% target_cancers[-5],
             # age %in% seq(40, 70, 5),
             age == 40,
             sex == "Female") %>% 
      pivot_longer(cols = c(plcl, pcdx, inc_rate, cdeath_rate),
                   names_to = "measure",
                   values_to = "rate") %>% 
      # mutate(# age = as.character(age),
      #        measure = ifelse(measure == "inc_rate", "Incidence",
      #                         "Cause-specific mortality")) %>% 
      ggplot(aes(x = year, y = rate, group = age, color = age)) +
      geom_line() +
      scale_y_continuous(name = "Rate (per 100,000)",
                         labels = scales::label_comma(accuracy = 0.1, scale = 1e5)) +
      facet_grid(cause ~ measure, scales = "free_y") +
      theme_minimal()
```

## Cause-specific changes in intervention TPs by time

```{r}
intv_modif_rr %>% 
      filter(cause %in% target_cancers) %>% 
      pivot_longer(cols = starts_with("effect_on_"), names_to = "cat", values_to = "tp") %>% 
      ggplot(aes(x = year, y = tp, color = cat)) +
      geom_line() +
      facet_wrap(. ~ cause) +
      theme_minimal()
```

```{r}
country_epi_2000_2010 <- calc_metric(country_dt   = country_epi,
            measure_inpt = "Prevalence",
            counts_name  = "cPx",
            rate_name    = "cpx",
            add_Nx       = FALSE,
            year_inpt    = 2000:2010)

country_epi_2000_2010 %>% 
      filter(age == 40,
             sex == "Female",
             cause %in% target_cancers[-5]) %>% 
      ggplot(aes(x = year, y = cmx * 1e5)) +
      geom_line() +
      facet_wrap(. ~ cause, scales = "free_y") +
      theme_minimal()
```

## Cause-specific baseline TP values

```{r}
bsln_markov_trace %>% 
      filter(age == 40,
             sex == "Female",
             cause %in% target_cancers[-5]) %>% 
      ggplot(aes(x = year, y = pdst)) +
      geom_point() +
      facet_wrap(. ~ cause, scales = "free_y") +
      theme_minimal()
```

## Baseline vs. intervention Markov health state distribution

```{r}
# scen 1: 1, 1, 0.9, 0.9, 0.9
# intv_modif_rr_1 <- intv_modif_rr
# intv_tps_1 <- intv_tps
# intv_markov_trace_1 <- intv_markov_trace

# scen def: 1, 1, 0.05, 0.5, 0.9
# intv_modif_rr_cfr1 <- intv_modif_rr
# intv_tps_cfr1 <- intv_tps
# intv_markov_trace_cfr1 <- intv_markov_trace

# scen def: 0.05, 1, 0.05, 0.5, 0.9
# intv_modif_def <- intv_modif_rr
# intv_tps_def <- intv_tps
# intv_markov_trace_def <- intv_markov_trace

full_join(
      x  = full_join(
            x  = bsln_markov_trace %>% 
                  select(age, year, sex, cause, location, plcl, pcdx, inc_rate, cdeath_rate) %>% 
                  mutate(cohort = year - age) %>% 
                  mutate(lplcl = plcl - lag(plcl, default = first(plcl)),
                         lpcdx = pcdx - lag(pcdx, default = first(pcdx)),
                         .by = c(cohort, sex, cause, location)) %>% 
                  mutate(source = "blsn"),
            y  = intv_markov_trace %>% 
                  select(age, year, sex, cause, location, plcl, pcdx, inc_rate, cdeath_rate) %>% 
                  mutate(cohort = year - age) %>% 
                  mutate(lplcl = plcl - lag(plcl, default = first(plcl)),
                         lpcdx = pcdx - lag(pcdx, default = first(pcdx)),
                         .by = c(cohort, sex, cause, location)) %>% 
                  mutate(source = "intv"),
            by = join_by(age, year, sex, cause, location, plcl, pcdx,
                         lplcl, lpcdx, inc_rate, cdeath_rate, source)
      ),
      
      y = full_join(
            x  = intv_markov_trace_1 %>% 
                  select(age, year, sex, cause, location, plcl, pcdx, inc_rate, cdeath_rate) %>% 
                  mutate(cohort = year - age) %>% 
                  mutate(lplcl = plcl - lag(plcl, default = first(plcl)),
                         lpcdx = pcdx - lag(pcdx, default = first(pcdx)),
                         .by = c(cohort, sex, cause, location)) %>% 
                  mutate(source = "secn 1"),
            y  = intv_markov_trace_cfr0.05 %>% 
                  select(age, year, sex, cause, location, plcl, pcdx, inc_rate, cdeath_rate) %>% 
                  mutate(cohort = year - age) %>% 
                  mutate(lplcl = plcl - lag(plcl, default = first(plcl)),
                         lpcdx = pcdx - lag(pcdx, default = first(pcdx)),
                         .by = c(cohort, sex, cause, location)) %>% 
                  mutate(source = "cfr 0.05"),
            by = join_by(age, year, sex, cause, location, plcl, pcdx,
                         lplcl, lpcdx, inc_rate, cdeath_rate, source)
      ),
      
      by = join_by(age, year, sex, cause, location, plcl, pcdx,
                   lplcl, lpcdx, inc_rate, cdeath_rate, source)
) %>% 
      pivot_longer(cols = c(plcl, pcdx, lplcl, lpcdx, inc_rate, cdeath_rate),
                   names_to = "metric", values_to = "val") %>% 
      filter(age == 50,
             sex == "Female",
             cause %in% target_cancers[2],
             # metric %in% c("cdeath_rate")) %>%
             # metric %in% c("pcdx")) %>% 
             metric %in% c("cdeath_rate", "lpcdx")) %>%
      # pivot_wider(id_cols = c(age, year, sex, cause, location, source),
      #             names_from = metric, values_from = val) %>% 
      # view()
      ggplot(aes(x = year, y = val, col = source)) +
      geom_line() +
      scale_y_continuous(labels = scales::label_comma(scale = 1 * 1e5)) +
      facet_grid(cause ~ metric, scales = "free_y") +
      theme_minimal() +
      theme(strip.text.y = element_text(angle = 0),
            legend.position = "bottom")
```

```{r}
full_join(
      x  = bsln_markov_trace %>% 
            select(age, year, sex, cause, location, plcl, pcdx, inc_rate, cdeath_rate) %>% 
            mutate(cohort = year - age) %>% 
            mutate(lplcl = plcl - lag(plcl, default = first(plcl)),
                   lpcdx = pcdx - lag(pcdx, default = first(pcdx)),
                   .by = c(cohort, sex, cause, location)) %>% 
            mutate(source = "blsn"),
      y  = intv_markov_trace %>% 
            select(age, year, sex, cause, location, plcl, pcdx, inc_rate, cdeath_rate) %>% 
            mutate(cohort = year - age) %>% 
            mutate(lplcl = plcl - lag(plcl, default = first(plcl)),
                   lpcdx = pcdx - lag(pcdx, default = first(pcdx)),
                   .by = c(cohort, sex, cause, location)) %>% 
            mutate(source = "intv"),
      by = join_by(age, year, sex, cause, location, plcl, pcdx,
                   lplcl, lpcdx, inc_rate, cdeath_rate, source)
) %>% 
      pivot_longer(cols = c(plcl, pcdx, lplcl, lpcdx, inc_rate, cdeath_rate),
                   names_to = "metric", values_to = "val") %>% 
      filter(age == 50,
             sex == "Female",
             cause %in% target_cancers[2],
             # metric %in% c("cdeath_rate")) %>%
             # metric %in% c("pcdx")) %>% 
             metric %in% c("cdeath_rate", "lpcdx")) %>%
      # pivot_wider(id_cols = c(age, year, sex, cause, location, source),
      #             names_from = metric, values_from = val) %>% 
      # view()
      ggplot(aes(x = year, y = val, col = source)) +
      geom_line() +
      scale_y_continuous(labels = scales::label_comma(scale = 1 * 1e5)) +
      facet_grid(cause ~ metric, scales = "free_y") +
      theme_minimal() +
      theme(strip.text.y = element_text(angle = 0),
            legend.position = "bottom")
```

## Testing how BGMx changes in the intv scen

```{r}
tps_sbst <- intv_tps %>% 
      filter(year == 2030,
             sex == "Female",
             cause == cause3[1]) %>% 
      rowwise() %>% 
      mutate(w_sum = sum(c_across(starts_with("dst_")), na.rm = TRUE),
             adds_to_one = abs(w_sum - 1) < 1e-8) %>% 
      ungroup() %>% 
      # select()
      view()


rowSums(tps_sbst[, grepl("^w_", names(tps_sbst))])

w_cols <- tps_sbst[ , grepl("^w_", names(tps_sbst))]

tps_sbst[, TRUE]
```

```{r}
full_join(
      x  = bsln_tps %>% 
            filter(age %in% seq(50, 90, 5),
                   sex == "Female",
                   cause == cause3[1]) %>% 
            rowwise() %>% 
            mutate(dst_sum = sum(c_across(starts_with("dst_")), na.rm = TRUE),
                   adds_to_one = abs(dst_sum - 1) < 1e-8,
                   scen = "bsln") %>% 
            ungroup(),
      y  = intv_tps %>% 
            filter(age %in% seq(50, 90, 5),
                   sex == "Female",
                   cause == cause3[1]) %>% 
            rowwise() %>% 
            mutate(dst_sum = sum(c_across(starts_with("dst_")), na.rm = TRUE),
                   adds_to_one = abs(dst_sum - 1) < 1e-8,
                   scen = "intv") %>% 
            ungroup()
) %>% 
      # filter(scen == "intv",
      #        year >= 2019) %>% 
      # select(year, starts_with("dst_"))

      ggplot(aes(x = year, y = dst_sum, color = scen)) +
      geom_line() +
      facet_wrap(. ~ age) +
      theme_minimal()
```

```{r}
country_epi %>% 
      filter(sex == "Female",
             age %in% seq(40, 70, 5),
             cause %in% target_cancers,
             measure == "Incidence",
             metric == "Rate") %>% 
      mutate(age = as.character(age)) %>% 
      ggplot(aes(x = year, y = val, group = age, col = age)) +
      geom_line() +
      facet_wrap(. ~ cause) +
      theme_minimal()
```

## Fertility Charts

```{r}
full_join(
      x = gbd_fert %>% 
            filter(location %in% c("United States of America", "Uganda", "India", "China"),
                   sex == "Female",
                   year == 2000, sex == "Female"),
      y = gbd_fert_5y %>% 
            filter(location %in% c("United States of America", "Uganda", "India", "China"),
                   sex == "Female",
                   year == 2000, sex == "Female")
      ) %>% 
      # pivot_longer(cols = c(asfr, asfr_5)) %>% 
      ggplot(aes(x = age)) +
      geom_line(aes(y = asfr)) +
      geom_point(aes(y = asfr_5)) +
      facet_wrap(.~ location) +
      theme_minimal()
```

## Cancer-Specific Markov Trace Plost

```{r}
bsln_tps %>% 
      filter(cause %in% c(cause1, cause2, cause3),
             # sex == "Female",
             case_when(cause == cause2[2] ~ sex == "Male",
                       TRUE ~ sex == "Female"),
             age == 60) %>% 
      pivot_longer(cols = c(w_prc, w_lcl, prc_lcl, lcl_rgn, rgn_dst,
                            dst_d, lcl_d),
                   names_to = "tp",
                   values_to = "val") %>% 
      ggplot(aes(x = year, y = val, col = tp)) +
      geom_line() +
      facet_wrap(. ~ cause, scales = "free_y") +
      theme_minimal()
```

```{r}
bsln_markov_trace %>% 
      filter(cause %in% c(cause1, cause2, cause3),
             # sex == "Female",
             case_when(cause == cause2[2] ~ sex == "Male",
                       TRUE ~ sex == "Female"),
             age == 60) %>% 
      pivot_longer(cols = c(pprc, plcl, prgn, pdst, pcdx),
                   names_to = "tp",
                   values_to = "val") %>% 
      ggplot(aes(x = year, y = val, col = tp)) +
      geom_line() +
      facet_wrap(. ~ cause, scales = "free_y") +
      theme_minimal()
```

## Jan 2026

```{r}
`%between%` <- data.table::`%between%`

calc_markov_state_deltas_dt <- function(
            markov_trace,
            cols = c("pwell","pprc","plcl","prgn","pdst","pcdx","pbgdx"),
            year_range = c(2000L, 2050L),
            age_range  = c(1L, 75L),
            year_col = "year",
            age_col  = "age",
            delta_prefix = "d_",
            check_unique_year_age = FALSE
) {
      dt <- data.table::as.data.table(markov_trace)
      
      miss_keys <- setdiff(c(year_col, age_col), names(dt))
      if (length(miss_keys)) stop("Missing key columns: ", paste(miss_keys, collapse = ", "))
      
      miss_cols <- setdiff(cols, names(dt))
      if (length(miss_cols)) stop("Missing state columns: ", paste(miss_cols, collapse = ", "))
      
      dt[, (year_col) := as.integer(get(year_col))]
      dt[, (age_col)  := as.integer(get(age_col))]
      
      y0 <- as.integer(year_range[1]); y1 <- as.integer(year_range[2])
      a0 <- as.integer(age_range[1]);  a1 <- as.integer(age_range[2])
      
      
      dt_sub <- dt[
            get(year_col) %between% c(y0, y1 + 1L) &
                  get(age_col)  %between% c(a0, a1 + 1L)
      ]
      
      if (check_unique_year_age) {
            dupN <- dt_sub[, .N, by = c(year_col, age_col)][N > 1L, .N]
            if (nrow(dupN) > 0) {
                  stop("Found duplicate rows for the same (year, age). ",
                       "Delta merge would expand rows. Fix upstream or aggregate first.")
            }
      }
      
      # shifted "next" table: align (year+1, age+1) back onto (year, age)
      next_dt <- dt_sub[, c(year_col, age_col, cols), with = FALSE]
      next_dt[, (year_col) := get(year_col) - 1L]
      next_dt[, (age_col)  := get(age_col)  - 1L]
      data.table::setnames(next_dt, cols, paste0("next_", cols))
      
      # merge by (year, age)
      data.table::setkeyv(dt_sub, c(year_col, age_col))
      data.table::setkeyv(next_dt, c(year_col, age_col))
      out <- next_dt[dt_sub]  
      
      # compute deltas
      for (j in cols) {
            out[, (paste0(delta_prefix, j)) := get(paste0("next_", j)) - get(j)]
      }
      
      # drop helper next_*
      out[, (paste0("next_", cols)) := NULL]
      
      # restrict to requested output window
      out <- out[
            get(year_col) %between% c(y0, y1) &
                  get(age_col)  %between% c(a0, a1)
      ]
      
      out
}

```

```{r}
calc_markov_state_deltas_dt(markov_trace = bsln_markov_trace %>% 
                                  filter(sex == "Female", cause == cause2[1])) %>% 
      filter(age %in% seq(45, 75, 5)) %>%
      # mutate(cohort = year - age) %>% 
      # filter(cohort %in% seq(1945, 1975, 5)) %>% 
      # filter(age %in% seq(0, 40, 5)) %>% 
      mutate(age = as.character(age)) %>% 
      # mutate(cohort = as.character(cohort)) %>% 
      pivot_longer(cols = ends_with("_rate"),
                   names_to = "metric",
                   values_to = "rate") %>% 
      ggplot(aes(x = year, y = rate, col = age)) +
      geom_line() +
      facet_wrap(. ~ metric, scales = "free_y") +
      theme_caviz
```

```{r}
calc_markov_state_deltas_dt(markov_trace = bsln_markov_trace %>% 
                                  filter(sex == "Female", cause == cause2[1])) %>% 
      # filter(age %in% seq(45, 75, 5)) %>% 
      mutate(cohort = year - age) %>% 
      filter(cohort %in% seq(1945, 1975, 5)) %>% 
      # filter(age %in% seq(0, 40, 5)) %>% 
      # mutate(age = as.character(age)) %>% 
      mutate(cohort = as.character(cohort)) %>% 
      pivot_longer(cols = ends_with("_rate"),
                   names_to = "metric",
                   values_to = "rate") %>% 
      ggplot(aes(x = year, y = rate, col = cohort)) +
      geom_line() +
      facet_wrap(. ~ metric, scales = "free_y") +
      theme_caviz
```



### Session with Renu 2/4/2026

```{r}
test_gbd <- explore_rdata_obj(file_path = "../inputs/gbd_dt.RData",
                              obj_nm    = "gbd_epi_1yr_predict")
```

```{r}
mod_proj <- explore_rdata_obj(file_path = "../outputs/RData/Uganda/Uganda_top_5.RData")
                              # obj_nm    = "gbd_epi_1yr_predict")
```

```{r}
full_join(
      x  = test_gbd %>% 
            filter(measure == "Deaths",
                   location == "Uganda",
                   sex == "Female",
                   cause == cause2[1]) %>% head(),
      
      y  = bsln_pop_projection %>% names()
            filter(location == "Uganda",
                   sex == "Female",
                   cause == cause2[1]) %>% names(),
      
      by = join_by()
)
```


```{r}
plot_age_specific_rates <- function(data,
                                    measure_name,
                                    location_name,
                                    cause_name) {
      # Filter the data
      plot_data <- data %>%
            filter(
                  measure == measure_name,
                  location == location_name,
                  sex == "Female",
                  age %in% c("40-44 years", "50-54 years", "60-64 years"),
                  cause == cause_name,
                  metric == "Rate",
                  year >= 2000 & year <= 2019
            )
      # Create the plot
      ggplot(plot_data, aes(x = year, y = val, color = age)) +
            geom_line(linewidth = 1) +
            geom_point(size = 2) +
            labs(
                  title = paste(measure_name, "Rate by Age Group -", cause_name),
                  subtitle = paste(location_name, "| Females | 2000-2019"),
                  x = "Year",
                  y = paste(measure_name, "Rate (per 100,000)"),
                  color = "Age Group"
            ) +
            scale_x_continuous(breaks = seq(2000, 2019, by = 2)) +
            theme_minimal() +
            theme(
                  legend.position = "bottom",
                  plot.title = element_text(face = "bold")
            )
}
```


```{r}
gbd_epi_raw <- explore_rdata_obj(file_path = "../inputs/gbd_dt.RData",
                                 obj_nm    = "gbd_epi_raw")

plot_data <- gbd_epi_raw %>%
      filter(
            measure == "Incidence",
            location == "Uganda",
            sex == "Female",
            age %in% c("40-44 years", "50-54 years", "60-64 years"),
            cause == "Breast cancer",
            metric == "Rate",
            year >= 2000 & year <= 2019
      )
 

plot_age_specific_rates(
      data = gbd_epi_raw,
      measure_name = "Incidence",
      location_name = "Uganda",
      cause_name = "Breast cancer"
)
```

## GBD Cohort Explorations

```{r}
country_epi %>% 
      filter(location == "Uganda",
             sex      == "Female",
             # cause %in% c("Breast cancer", "Cervical cancer", "Pancreatic cancer"),
             measure %in% c("Incidence", "Deaths")) %>% 
      mutate(cohort = year - age) %>% 
      # filter(cohort %in% seq(1960, 2000, 5)) %>%
      # filter(between(age, 40, 79)) %>%
      mutate(measure = factor(measure, levels = c("Incidence", "Deaths"))) %>%
      mutate(cohort = as.character(cohort)) %>% 
      
      ggplot(aes(x = age, y = pred_rates,
                 # color = cohort,
                 group = cohort)) +
      geom_line(alpha = 0.05) +
      facet_grid(cause ~ measure, scales = "free_y") +
      scale_y_continuous(name     = "Cancer-specific rates (per 100,000)",
                         labels   = scales::label_comma(scale = 1e5),
                         n.breaks = 3) +
      # theme_minimal() +
      theme_caviz +
      theme(aspect.ratio    = 1 / 3,
            legend.position = "bottom",
            strip.text.y    = element_text(color = "black", angle = 0))

ggsave(filename = "../outputs/diagnostics/rates_1year_all_cohorts.png",
       width = 6, height = 10)
```

```{r}
country_epi %>% 
      filter(location == "Uganda",
             sex      == "Female",
             # cause %in% c("Breast cancer", "Cervical cancer", "Pancreatic cancer"),
             measure %in% c("Incidence", "Deaths")) %>% 
      mutate(cohort = year - age) %>% 
      filter(cohort %in% seq(1960, 2000, 5)) %>%
      # filter(between(age, 40, 79)) %>%
      mutate(measure = factor(measure, levels = c("Incidence", "Deaths"))) %>%
      mutate(cohort = as.character(cohort)) %>% 
      
      ggplot(aes(x = age,
                 y = pred_rates,
                 color = cohort,
                 group = cohort)) +
      geom_line(alpha = 0.3) +
      facet_grid(cause ~ measure, scales = "free_y") +
      scale_y_continuous(name     = "Cancer-specific rates (per 100,000)",
                         labels   = scales::label_comma(scale = 1e5),
                         # trans    = 'log10',
                         n.breaks = 3) +
      # theme_minimal() +
      theme_caviz +
      theme(aspect.ratio    = 1 / 3,
            legend.position = "bottom",
            strip.text.y    = element_text(color = "black", angle = 0))

ggsave(filename = "../outputs/diagnostics/rates_1year_all_cohorts_col_log.png",
       width = 6, height = 12)
```

```{r}
full_join(
      x  = country_epi %>% 
            # filter(cause == cause2[3]) %>% 
            filter(location == "Uganda",
                   sex      == "Female",
                   measure %in% c("Incidence", "Deaths")) %>% 
            rename(rate = pred_rates) %>% 
            # pivot_wider(id_cols = c(location, sex, year, age, pop, cause),
            #             names_from = measure,
            #             values_from = pred_rates) %>% 
            # rename(inc_rate = Incidence, cdeath_rate = Deaths) %>% 
            mutate(source = "GBD"),
      
      y  = bsln_markov_trace %>% 
            filter(sex == "Female") %>% 
            split(.$cause) %>%
            map(.x = .,
                .f = ~ calc_markov_state_deltas_dt(markov_trace = .x)) %>%
            list_rbind() %>%
            pivot_longer(cols = c(inc_rate, cdeath_rate),
                         names_to = "measure",
                         values_to = "rate") %>% 
            mutate(measure = recode_values(measure,
                                           "inc_rate"    ~ "Incidence",
                                           "cdeath_rate" ~ "Deaths",
                                           default = NA_character_)) %>% 
            # filter(cause == cause2[3]) %>% 
            # mutate(rate = rate / 1e5) %>%
            mutate(source = "Markov"),
      
      by = join_by(location, sex, year, age, cause, measure, rate, source)
) %>% 
      select(location, sex, year, age, cause, measure, rate, source) %>% 
      mutate(measure = factor(measure, levels = c("Incidence", "Deaths"))) %>%
      mutate(cohort = as.character(year - age)) %>% 
      
      # filter(age == 0) %>% 
      filter(cause %in% c(cause1[2],
                          cause1[4],
                          cause2[1],
                          cause2[3],
                          cause3[1]),
             sex == "Female",
             cohort %in% c(1960, 1980, 2000),
             year <= reference_year) %>% 
      
      mutate(cause = recode_values(cause,
                                   cause1[2] ~ "Esoph.",
                                   cause1[4] ~ "Liver",
                                   cause2[1] ~ "Breast",
                                   cause2[3] ~ "Lung",
                                   cause3[1] ~ "Cervix",
                                   cause3[2] ~ "CRC",
                                   default = NA_character_)) %>% 
      
      # filter(source == "Markov") %>% 
      # filter(measure == "Deaths") %>% 
      
      # filter(age == 1,
      #        sex == "Female",
      #        year == 2001,
      #        cause == "Breast cancer",
      #        source == "GBD")
      
      # count(age, sex, year, cause, measure, source) %>% 
      
      ggplot(aes(x = year - 2000,
                 y = rate,
                 color = source,
                 group = interaction(source, cohort))) +
      geom_line() +
      scale_y_continuous(name     = "Cancer-specific rates\n(per 100,000)",
                         labels   = scales::label_comma(scale = 1e5),
                         trans    = 'log10',
                         n.breaks = 3) +
      facet_grid(measure + cohort ~ cause, scale = "free_y") +
      # facet_wrap(. ~ measure, scale = "free_y") +
      labs(x = "Year (post 2000)") +
      theme_caviz +
      theme(aspect.ratio    = 1 / 3,
            legend.position = "bottom",
            strip.text.y    = element_text(color = "black", angle = 0))
```

#### Without Markov Correction

```{r}
full_join(
      x  = country_epi %>% 
            filter(location == "Uganda",
                   sex      == "Female",
                   measure %in% c("Incidence", "Deaths")) %>% 
            rename(rate = Rate) %>% #pred_rates) %>% 
            mutate(cohort = as.character(year - age)) %>% 
            mutate(source = "GBD"),
      
      y  = bsln_markov_trace %>% 
            filter(sex == "Female") %>% 
            calc_epi_rates_markov(markov_inpt = .) %>% 
            pivot_longer(cols = c(inc_rate, cfr),
                         names_to = "measure",
                         values_to = "rate") %>% 
            mutate(measure = recode_values(measure,
                                           "inc_rate" ~ "Incidence",
                                           "cfr"      ~ "Deaths",
                                           default = NA_character_)) %>% 
            mutate(source = "Markov"),
      
      by = join_by(location, sex, year, age, cohort, cause, measure, rate, source)
) %>% 
      select(location, sex, year, age, cohort, cause, measure, rate, source) %>% 
      mutate(measure = factor(measure, levels = c("Incidence", "Deaths"))) %>%
      
      filter(cause %in% c(cause1[1],
                          cause1[2],
                          cause1[4],
                          cause1[7],
                          cause1[8],
                          cause1[9],
                          cause1[10] #,
                          # cause2[1],
                          # cause2[3],
                          # cause3[1],
                          ),
             sex == "Female",
             cohort %in% seq(1960, 2000, 10),
             measure == "Incidence", # "Deaths", # 
             year <= reference_year) %>% 
      
      mutate(cause = recode_values(cause,
                                   cause1[1] ~ "Bladder",
                                   cause1[2] ~ "Esoph.",
                                   cause1[4] ~ "Liver",
                                   cause1[7] ~ "Ovary",
                                   cause1[8] ~ "Pancreas",
                                   cause1[9] ~ "Stomach",
                                   cause1[10] ~ "Thyroid",
                                   cause2[1] ~ "Breast",
                                   cause2[3] ~ "Lung",
                                   cause3[1] ~ "Cervix",
                                   cause3[2] ~ "CRC",
                                   default = NA_character_)) %>% 
      
      ggplot(aes(x = year - 2000,
                 y = rate,
                 color = source,
                 group = interaction(source, cohort))) +
      geom_line() +
      scale_y_continuous(name     = "Cancer-specific rates\n(per 100,000)",
                         labels   = scales::label_comma(scale = 1e5),
                         trans    = 'log10',
                         n.breaks = 3) +
      facet_grid(cause ~ measure + cohort, scale = "free_y") +
      # facet_wrap(. ~ measure, scale = "free_y") +
      labs(x = "Year (post 2000)") +
      theme_caviz +
      theme(aspect.ratio    = 1 / 4,
            legend.position = "bottom",
            strip.text.y    = element_text(color = "black", angle = 0))
```

```{r}
## Deep dive
bsln_markov_trace %>% 
      mutate(cohort = year - age) %>% 
      filter(sex == "Female" & cause == cause1[4] & cohort == 1980) %>% 
      # mutate(plcl = plcl - lag(plcl, default = 0)) %>% 
      filter(year <= 2020) %>% 
      
      calc_epi_rates_markov(markov_inpt = .) %>% 
      
      # ggplot(aes(x = year, y = plcl / pwell)) +
      ggplot(aes(x = year, y = inc_rate)) +
      geom_line() +
      scale_y_continuous(trans = "log10") +
      theme_caviz
```

```{r}
read.csv(paste0("../../Data/GBD 2021/16 cancer types/split/gbd_epi_",
                "China",
                ".csv")) %>% 
      
# country_epi %>% 
      mutate(cohort = year - age) %>% 
      filter(sex      == "Female",
             measure == "Incidence") %>% 
      filter(cause %in% cause1[c(1, 2, 4, 7, 8, 9, 10)]) %>% 
      # filter(cohort %in% seq(1960, 2000, 5)) %>%
      filter(age %in% seq(40, 75, 5)) %>%
      filter(year <= reference_year) %>% 
      # mutate(cohort = as.character(cohort)) %>%
      mutate(age = as.character(age)) %>%
      rename(rate = pred_rates) %>% # Rate) %>% #
      
      mutate(cause = recode_values(cause,
                                   cause1[1] ~ "Bladder",
                                   cause1[2] ~ "Esoph.",
                                   cause1[4] ~ "Liver",
                                   cause1[7] ~ "Ovary",
                                   cause1[8] ~ "Pancreas",
                                   cause1[9] ~ "Stomach",
                                   cause1[10] ~ "Thyroid",
                                   cause2[1] ~ "Breast",
                                   cause2[3] ~ "Lung",
                                   cause3[1] ~ "Cervix",
                                   cause3[2] ~ "CRC",
                                   default = NA_character_)) %>% 
      
      # ggplot(aes(x = year, y = plcl / pwell)) +
      ggplot(aes(x = year, y = rate, colour = cause, group = cause)) +
      geom_line() +
      scale_x_continuous(breaks = seq(2000, 2020, 10)) +
      scale_y_continuous(n.breaks = 3,
                         trans = "log10") +
      # facet_grid(cause ~ cohort, scale = "free_y") +
      facet_grid(cause ~ age, scale = "free_y") +
      guides(color = "none") +
      theme_caviz +
      theme(aspect.ratio = 1 / 3,
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10),
            legend.position = "bottom",
            strip.text.y    = element_text(color = "black", angle = 0))
```

```{r}
reduce(.x = list(read.csv(paste0("../../Data/GBD 2021/16 cancer types/split/gbd_epi_",
                                 "China",
                                 ".csv")),
                 read.csv(paste0("../../Data/GBD 2021/16 cancer types/split/gbd_epi_",
                                 "United States of America",
                                 ".csv")),
                read.csv(paste0("../../Data/GBD 2021/16 cancer types/split/gbd_epi_",
                                "Uganda",
                                ".csv")),
                read.csv(paste0("../../Data/GBD 2021/16 cancer types/split/gbd_epi_",
                                "India",
                                ".csv"))),
       .f = full_join) %>% 
      
      mutate(cohort = year - age) %>% 
      filter(sex     == "Female",
             measure == "Incidence") %>% 
      filter(cause %in% cause1[c(1, 2, 4, 7, 8, 9, 10)]) %>% 
      # filter(cohort %in% seq(1960, 2000, 5)) %>%
      filter(age %in% seq(40, 75, 5)) %>%
      filter(year == reference_year) %>% 
      # mutate(cohort = as.character(cohort)) %>%
      mutate(age = as.character(age)) %>%
      rename(rate = pred_rates) %>% # Rate) %>% #
      
      mutate(cause = recode_values(cause,
                                   cause1[1] ~ "Bladder",
                                   cause1[2] ~ "Esoph.",
                                   cause1[4] ~ "Liver",
                                   cause1[7] ~ "Ovary",
                                   cause1[8] ~ "Pancreas",
                                   cause1[9] ~ "Stomach",
                                   cause1[10] ~ "Thyroid",
                                   cause2[1] ~ "Breast",
                                   cause2[3] ~ "Lung",
                                   cause3[1] ~ "Cervix",
                                   cause3[2] ~ "CRC",
                                   default = NA_character_)) %>% 
      
      ggplot(aes(x = rate, colour = location, group = location, fill = location)) +
      geom_histogram(alpha = 0.3) +
      scale_x_continuous(name = "Incidence rate (per 100,000)",
                         labels = scales::label_number(accuracy = 1, scale = 1e5),
                         n.breaks = 3) +
      scale_y_continuous(n.breaks = 2) +
      # facet_grid(cause ~ cohort, scale = "free_y") +
      facet_grid(cause ~ age, scale = "free_y") +
      # guides(color = "none") +
      theme_caviz +
      theme(aspect.ratio = 1 / 4,
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10),
            legend.position = "bottom",
            strip.text.y    = element_text(color = "black", angle = 0))
```

```{r}
reduce(.x = list(read.csv(paste0("../../Data/GBD 2021/16 cancer types/split/gbd_epi_",
                                 "China",
                                 ".csv")),
                 read.csv(paste0("../../Data/GBD 2021/16 cancer types/split/gbd_epi_",
                                 "United States of America",
                                 ".csv")),
                read.csv(paste0("../../Data/GBD 2021/16 cancer types/split/gbd_epi_",
                                "Uganda",
                                ".csv")),
                read.csv(paste0("../../Data/GBD 2021/16 cancer types/split/gbd_epi_",
                                "India",
                                ".csv"))),
       .f = full_join) %>% 
      
      mutate(cohort = year - age) %>% 
      filter(sex     == "Female",
             measure == "Incidence") %>% 
      filter(cause %in% cause1[c(1, 2, 4, 7, 8, 9, 10)]) %>%
      # filter(cohort %in% seq(1960, 2000, 5)) %>%
      filter(age %in% seq(40, 75, 5)) %>%
      filter(year <= reference_year) %>% 
      # mutate(cohort = as.character(cohort)) %>%
      mutate(age = as.character(age)) %>%
      rename(rate = pred_rates) %>% # Rate) %>% #
      
      mutate(cause = recode_values(cause,
                                   cause1[1] ~ "Bladder",
                                   cause1[2] ~ "Esoph.",
                                   cause1[4] ~ "Liver",
                                   cause1[7] ~ "Ovary",
                                   cause1[8] ~ "Pancreas",
                                   cause1[9] ~ "Stomach",
                                   cause1[10] ~ "Thyroid",
                                   cause2[1] ~ "Breast",
                                   cause2[3] ~ "Lung",
                                   cause3[1] ~ "Cervix",
                                   cause3[2] ~ "CRC",
                                   default = NA_character_)) %>% 
      
      group_by(cohort, sex, cause) %>% 
      mutate(diff = rate - rate[location == "United States of America"]) %>% 
      
      ggplot(aes(x = year, y = diff,
                 colour = location, group = location, fill = location)) +
      # geom_histogram(alpha = 0.3) +
      geom_line() +
      scale_x_continuous(name = "Year",
                         n.breaks = 3) +
      scale_y_continuous(name = "Difference in incidence rates (from the US)",
                         n.breaks = 2,
                         labels = scales::label_number(accuracy = 1, scale = 1e5),
                         # trans = "log10"
                         ) +
      # facet_grid(cause ~ cohort, scale = "free_y") +
      facet_grid(cause ~ age, scale = "free_y") +
      # guides(color = "none") +
      theme_caviz +
      theme(aspect.ratio = 1 / 4,
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10),
            legend.position = "bottom",
            strip.text.y    = element_text(color = "black", angle = 0))
```

```{r}
full_join(
      x  = country_epi %>% 
            # filter(cause == cause2[3]) %>% 
            filter(location == "Uganda",
                   sex      == "Female",
                   measure %in% c("Incidence", "Deaths")) %>% 
            rename(rate = pred_rates) %>% 
            # pivot_wider(id_cols = c(location, sex, year, age, pop, cause),
            #             names_from = measure,
            #             values_from = pred_rates) %>% 
            # rename(inc_rate = Incidence, cdeath_rate = Deaths) %>% 
            mutate(source = "GBD"),
      
      y  = bsln_pop_proj_markov %>% # bsln_markov_trace %>% 
            filter(sex == "Female") %>% 
            # split(.$cause) %>% 
            # map(.x = .,
            #     .f = ~ calc_markov_state_deltas_dt(markov_trace = .x)) %>% 
            # list_rbind() %>% 
            pivot_longer(cols = c(i_rate, d_rate),
                         names_to = "measure",
                         values_to = "rate") %>% 
            mutate(measure = recode_values(measure,
                                           "i_rate" ~ "Incidence",
                                           "d_rate" ~ "Deaths",
                                           default = NA_character_)) %>% 
            # filter(cause == cause2[3]) %>% 
            # mutate(rate = rate / 1e5) %>%
            mutate(source = "Markov"),
      
      by = join_by(location, sex, year, age, cause, measure, rate, source)
) %>% 
      select(location, sex, year, age, cause, measure, rate, source) %>% 
      mutate(measure = factor(measure, levels = c("Incidence", "Deaths"))) %>%
      mutate(cohort = as.character(year - age)) %>% 
      
      # filter(age == 0) %>% 
      filter(cause %in% cause2,
             sex == "Female",
             cohort %in% c(1960, 1980, 2000),
             year < reference_year) %>% 
      
      # filter(source == "Markov") %>% 
      
      # filter(age == 1,
      #        sex == "Female",
      #        year == 2001,
      #        cause == "Breast cancer",
      #        source == "GBD")
      
      # count(age, sex, year, cause, measure, source) %>% 
      
      ggplot(aes(x = year, y = rate, color = source, group = interaction(source, cohort))) +
      geom_line() +
      scale_y_continuous(name     = "Cancer-specific rates (per 100,000)",
                         labels   = scales::label_comma(scale = 1e5),
                         n.breaks = 3) +
      facet_grid(cause ~ measure, scale = "free_y") +
      # facet_wrap(. ~ measure, scale = "free_y") +
      theme_caviz +
      theme(#aspect.ratio    = 1 / 3,
            legend.position = "bottom",
            strip.text.y    = element_text(color = "black", angle = 0))
```

```{r}
full_join(
      x  = country_epi %>% 
            # filter(cause == cause2[3]) %>% 
            filter(location == "Uganda",
                   sex      == "Female",
                   measure %in% c("Incidence", "Deaths")) %>% 
            rename(rate = pred_rates) %>% 
            mutate(source = "GBD"),
      
      y  = bsln_pop_proj_markov %>% # bsln_markov_trace %>% 
            filter(sex == "Female") %>% 
            rename(pop = Nx) %>% 
            mutate(source = "Markov"),
      
      by = join_by(location, sex, year, age, cause, pop, source)
) %>% 
      select(location, sex, year, age, pop, source) %>% 
      mutate(cohort = as.character(year - age)) %>% 
      
      unique() %>% 
      
      # filter(age == 0) %>% 
      filter(sex == "Female",
             cohort %in% seq(1960, 2000, 5),
             year <= reference_year) %>% 
      
      # filter(source == "Markov") %>% 
      
      # filter(age == 1,
      #        sex == "Female",
      #        year == 2001,
      #        cause == "Breast cancer",
      #        source == "GBD")
      
      # count(age, sex, year, cause, measure, source) %>% 
      
      ggplot(aes(x = year - 2000,
                 y = pop,
                 color = source,
                 group = source)) +
      geom_line() +
      scale_x_continuous(name   = "Year (post 2000)") +
      scale_y_continuous(name   = "Total pop counts (in thousands)",
                         labels = scales::label_number(scale = 1e-3)) +
      facet_wrap(. ~ cohort, scale = "free_y") +
      # facet_wrap(. ~ measure, scale = "free_y") +
      theme_caviz +
      theme(#aspect.ratio    = 1 / 3,
            legend.position = "bottom",
            strip.text.y    = element_text(color = "black", angle = 0))
```

```{r}
bsln_tps_raw %>% 
      select(cause:age, lcl_d, dst_d) %>% 
      pivot_longer(cols = ends_with("_d"),
                   names_to = "tp") %>% 
      filter(sex == "Female",
             case_when(cause %in% cause1 ~ tp == "lcl_d",
                       TRUE              ~ tp == "dst_d")) %>% 
      # filter(dst_d < 0.01)
      
      ggplot(aes(x = age, y = value)) +
      geom_point(alpha = 0.01) +
      facet_wrap(. ~ cause, scales = "free_y") +
      labs(x = "Age", y = "TP to death\n(lcl_d in category 1, dst_d otherwise)") +
      theme_minimal()
```

```{r}
bsln_tps_raw %>% 
      filter(year == 2019) %>% 
      select(cause:age, lcl_d, dst_d) %>% 
      pivot_longer(cols = ends_with("_d"),
                   names_to = "tp") %>% 
      filter(sex == "Female",
             case_when(cause %in% cause1 ~ tp == "lcl_d",
                       TRUE              ~ tp == "dst_d")) %>% 
      # filter(dst_d < 0.01)
      
      ggplot(aes(x = age, y = value)) +
      geom_point(alpha = 0.2) +
      facet_wrap(. ~ cause, scales = "free_y") +
      labs(x = "Age", y = "TP to death\n(lcl_d in category 1, dst_d otherwise)") +
      theme_minimal()
```

## Call with Renu 2026/2/20

### Checking GBD amcx and prev add up to Nx

```{r}
reduce(.x = list(country_incd %>% 
                       # filter(age == 45,
                       #        # year == 2000,
                       #        sex == "Female") %>% 
                       select(!all_of(c("log_fitted_val", "measure"))),
                 country_prev %>% 
                       # filter(age == 45,
                       #        # year == 2000,
                       #        sex == "Female") %>% 
                       select(!all_of(c("log_fitted_val", "measure"))),
                 country_mort %>% 
                       # filter(age == 45,
                       #        # year == 2000,
                       #        sex == "Female") %>% 
                       select(!all_of(c("log_fitted_val", "measure"))),
                 country_acmx %>% 
                       # filter(age == 45,
                       #        sex == "Female") %>% 
                       select(!all_of(c("cause", "log_fitted_val", "measure")))),
       .f = full_join) %>% 
      mutate(cohort = year - age) %>% 
      filter(sex == "Female",
             cohort == 1980) %>% 
      
      pivot_longer(cols = c(cix, cpx, cmx, acmx),
                   names_to = "rate",
                   values_to = "val") %>% 
      
      mutate(val_p = rate_to_prob(time = 1, rate = val)) %>% 
      
      split(.$cause) %>% 
      map(.x = .,
          .f = ~ .x %>% 
                ggplot(aes(x = year, y = val, col = rate, group = rate)) +
                geom_line() +
                # facet_wrap(. ~ cause, scale = "free_y") +
                scale_y_continuous(name   = "Rate (per 100,000)",
                                   labels = scales::label_number(scale = 1e5),
                                   trans  = "log10"
                                   ) +
                theme_caviz +
                theme(aspect.ratio = 1 / 3,
                      legend.position = "bottom"))
      
      
```

```{r}
reduce(.x = list(country_incd %>% 
                       select(!all_of(c("log_fitted_val", "measure"))),
                 country_prev %>% 
                       select(!all_of(c("log_fitted_val", "measure"))),
                 country_mort %>% 
                       select(!all_of(c("log_fitted_val", "measure"))),
                 country_acmx %>% 
                       select(!all_of(c("cause", "log_fitted_val", "measure")))),
       .f = full_join) %>% 
      mutate(cohort = year - age) %>% 
      filter(case_when(cause == cause2[2] ~ sex == "Male",
                       TRUE               ~ sex == "Female")) %>% 
      
      filter(cohort == 1980) %>% 
      
      # mutate(across(.cols = c(cix, cpx, cmx, acmx),
      #               .fns  = ~ rate_to_prob(time = 1, rate = .))) %>% 
      # mutate(tot_dist = cix + cpx + cmx + acmx)
      
      # filter()
      
      pivot_longer(cols = c(cix, cpx, cmx, acmx),
                   names_to = "rate",
                   values_to = "val") %>%
      mutate(cause = recode(cause, !!!cx_lab)) %>% 
      
      filter(rate != "cix") %>% 
      
      ggplot(aes(x = year, y = val, col = rate, group = rate)) +
      geom_line() +
      facet_wrap(. ~ cause, scale = "free_y") +
      scale_y_continuous(name   = "Rate (per 100,000)",
                         labels = scales::label_number(scale = 1e5),
                         n.breaks = 3,
                         # trans  = "log10"
                         ) +
      scale_color_manual(values = c("#C6DABF", "#88D498", "#114B5F")) +
      theme_caviz +
      theme(aspect.ratio = 1 / 3,
            legend.position = "bottom")
```

```{r}
reduce(.x = list(country_nx %>% 
                       select(all_of(c(uid_vec[-4], "Nx"))),
                 country_prev %>% 
                       select(all_of(c(uid_vec, "cPx"))),
                 country_mort %>% 
                       select(all_of(c(uid_vec, "cDx"))),
                 country_acmx %>% 
                       select(all_of(c(uid_vec[-4], "acDx")))),
       .f = full_join) %>% 
      mutate(cohort = year - age) %>% 
      
      filter(case_when(cause == cause2[2] ~ sex == "Male",
                       TRUE               ~ sex == "Female")) %>% 
      
      filter(cohort == 1980) %>% 
      
      # mutate(across(.cols = c(cix, cpx, cmx, acmx),
      #               .fns  = ~ rate_to_prob(time = 1, rate = .))) %>% 
      
      group_by(across(all_of(uid_vec))) %>% 
      mutate(cpx  = cPx / Nx,
             cmx  = cDx / Nx,
             acmx = acDx / Nx) %>% 
      mutate(well = 1 - acmx,
             bgmx = acmx - cmx) %>% 
      ungroup() %>% 
      
      pivot_longer(cols = c(well, cpx, cmx, bgmx),
                   names_to = "metric",
                   values_to = "prop") %>% 
      mutate(cause = recode(cause, !!!cx_lab)) %>% 
      
      filter(metric != "well") %>% 
      # mutate
      
      ggplot(aes(x = year, y = prop, fill = metric, col = metric, group = metric)) +
      # geom_line() +
      geom_area() +
      facet_wrap(. ~ cause, scale = "free_y") +
      scale_y_continuous(name   = "Proportion (per 1000)",
                         labels = scales::label_number(scale = 1e3),
                         n.breaks = 3,
                         # trans  = "log10"
                         ) +
      scale_fill_manual(values = c("#F3E9D2", "#88D498", "#114B5F")) +
      scale_color_manual(values = c("#F3E9D2", "#88D498", "#114B5F")) +
      theme_caviz +
      theme(aspect.ratio = 1 / 3,
            legend.position = "bottom",
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10))
```

### Checking Numbers instead of Rates

```{r}
reduce(.x = list(country_incd %>% 
                       select(!all_of(c("log_fitted_val", "measure"))),
                 country_prev %>% 
                       select(!all_of(c("log_fitted_val", "measure"))),
                 country_mort %>% 
                       select(!all_of(c("log_fitted_val", "measure"))),
                 country_acmx %>% 
                       select(!all_of(c("cause", "log_fitted_val", "measure")))),
       .f = full_join) %>% 
      mutate(cohort = year - age) %>% 
      filter(sex == "Female",
             cohort == 1980) %>% 
      
      pivot_longer(cols = c(cIx, cPx, cDx, acDx), # Nx,
                   names_to = "metric",
                   values_to = "count") %>% 
      
      mutate(rate = count / Nx) %>% 
      
      split(.$cause) %>% 
      map(.x = .,
          .f = ~ .x %>% 
                ggplot(aes(x = year, y = rate, col = metric, group = metric)) +
                geom_line() +
                # facet_wrap(. ~ cause, scale = "free_y") +
                scale_y_continuous(name   = "Rate (per 100,000)",
                                   labels = scales::label_number(scale = 1e5),
                                   trans  = "log10"
                                   ) +
                theme_caviz +
                theme(aspect.ratio = 1 / 3,
                      legend.position = "bottom"))
```

## Compare HSD with and without Markov Correction

```{r}
full_join(
      x  = start_hstate_dist %>% mutate(markov = TRUE),
      y  = start_hstate_dist_no_markov %>% mutate(markov = FALSE)
) %>% 
      filter(case_when(cause == cause2[2] ~ sex == "Male",
                       TRUE               ~ sex == "Female")) %>% 
      pivot_longer(cols = pwell:pbgdx,
                   names_to = "hstate",
                   values_to = "prop") %>% 
      group_by(age, year, sex, location, cause, hstate) %>% 
      reframe(prop_diff = prop[markov] - prop[!markov]) %>% 
      
      filter(year == sim_start_year,
             cause %in% c(cause1[4], cause2[1], cause3[1])) %>% 
      
      ggplot(aes(x = age, y = prop_diff, color = hstate, group = hstate)) +
      geom_line() +
      scale_y_continuous(name = "Difference in starting HS distribution\nbetween w/out Markov correction") +
      facet_wrap(. ~ cause, nrow = 1) +
      theme_caviz +
      theme(aspect.ratio = 3 / 1)
```

## Compare Markovs

```{r}
full_join(
      x  = bsln_markov_trace %>% mutate(scen = "bsln"),
      y  = intv_markov_trace %>% mutate(scen = "intv")
) %>% 
      filter(cause == cause3[1],
             sex == "Female",
             year == reference_year) %>% 
      
      ggplot(aes(x = age, y = cdeath_rate, col = scen, group = scen)) +
      geom_line() +
      facet_wrap(. ~ scen) +
      theme_caviz
```

```{r}
full_join(
      x  = bsln_pop_projection, #%>% mutate(scen = "bsln"),
      y  = intv_pop_projection #%>% mutate(scen = "intv")
) %>% 
      filter(# cause == cause3[1],
             sex == "Female",
             year == reference_year + 1) %>% 
      
      ggplot(aes(x = age, y = cdeath_rate, col = scen, group = scen)) +
      geom_line() +
      facet_wrap(. ~ scen) +
      theme_caviz
```

## On the plane

```{r}
bsln_markov_trace %>% 
      mutate(cohort = year - age) %>% 
      filter(case_when(cause == cause2[2] ~ sex == "Male",
                       TRUE               ~ sex == "Female"),
             cohort == 1980) %>% 
      pivot_longer(cols = pwell:pbgdx,
                   names_to = "hstate",
                   values_to = "tp") %>% 
      
      filter(hstate %!in% c("pwell", "pbgdx")) %>% 
      
      ggplot(aes(x = year, y = tp, col = hstate, group = hstate)) +
      geom_line() +
      scale_y_continuous(n.breaks = 3) +
      facet_wrap(. ~ cause, scales = "free_y") +
      theme_caviz +
      theme(aspect.ratio = 1 / 4,
            legend.position = "bottom")
```

```{r}
bsln_tps %>% # names()
      mutate(cohort = year - age) %>% 
      filter(sex == "Female",
             # cohort == 1980,
             age == 45,
             cause == cause3[2]) %>% 
      pivot_longer(cols = c(w_lcl, lcl_rgn, rgn_dst, dst_d),
                   names_to = "tp",
                   values_to = "val") %>% 
      
      ggplot(aes(x = year, y = val, col = tp, group = tp)) +
      geom_line() +
      theme_caviz +
      theme(legend.position = "bottom")
```

```{r}
start_hstate_dist %>% 
      mutate(cohort = year - age) %>% 
      filter(cause %in% c(cause1[4], cause2[1], cause3),
             sex == "Female",
             cohort %in% seq(1960, 1990, 10)) %>% 
      pivot_longer(cols = pwell:pbgdx,
                   names_to = "hstate",
                   values_to = "prop") %>% 
      
      ggplot(aes(x = year, y = prop, group = hstate, colour = hstate)) +
      geom_line() +
      scale_y_continuous(n.breaks = 3) +
                         # trans = "log10") +
      facet_grid(cohort ~ cause) +
      theme_caviz +
      theme(legend.position = "bottom",
            aspect.ratio = 1 / 3)
```

```{r}
start_hstate_dist %>% 
      mutate(cohort = year - age) %>% 
      # filter(cause %in% c(cause1[4], cause2[1], cause3),
      #        sex == "Female",
      #        cohort %in% seq(1960, 1990, 10)) %>% 
      # pivot_longer(cols = pwell:pbgdx,
      #              names_to = "hstate",
      #              values_to = "tp") %>% 
      
      filter(case_when(cause == cause2[2] ~ sex == "Male",
                       TRUE               ~ sex == "Female")) %>% 
      # filter(case_when(cause == cause2[2] ~ age <  78,
      #                  TRUE               ~ age <= 95)) %>%
      
      # select(sex) %>% unique()
      
      group_by(age, year, sex, cause, location) %>% 
      mutate(val = pwell + pprc + plcl + prgn + pdst + pcdx + pbgdx,
              plaus = val == 1,
              marker = ifelse(plaus, "o", "x")) %>% 
      
      # filter(cause %in% c(cause1[1], cause3))
      # # filter(year == 2000,
      # #        age == 45) %>%
      # filter(val < 0.5)
      
      ggplot(aes(x = year, y = val, group = marker, colour = marker)) +
      geom_line() +
      scale_y_continuous(n.breaks = 3) +
                         # trans = "log10") +
      facet_wrap(. ~ cause) +
      theme_caviz +
      theme(legend.position = "bottom",
            aspect.ratio = 1 / 3)
```

```{r}
start_hstate_dist %>% 
      filter(case_when(cause == cause2[2] ~ sex == "Male",
                       TRUE               ~ sex == "Female")) %>% 
      # filter(case_when(cause == cause2[2] ~ age <  78,
      #                  TRUE               ~ age <= 95)) %>%
      
      mutate(val = pwell + pprc + plcl + prgn + pdst + pcdx + pbgdx,
             plaus = val == 1,
             marker = ifelse(plaus, "o", "x")) %>% 
      
      filter(year == 2000) %>% 
      # filter(cause == cause2[2]) %>%
      # filter(year == 2000,
      #        age == 45) %>%
      # filter(val < 0.75) %>% 
      filter(pwell < 0) %>% 
      
      count(cause, age)
```

```{r}
start_hstate_dist %>% 
      filter(cause == cause1[1],
             sex == "Female",
             age == 95,
             year == 2000) %>% 
      # mutate(p_cDx = pmin(pmax(cDx / Nx, 1), 0)) %>% 
      mutate(p_cDx = pmin(pmax(cDx / Nx, 0), 1))
```

```{r}
markov_correction %>% 
      filter(cause == cause1[1],
             sex == "Female",
             age == 95,
             year == 2000) #%>% 
      # mutate(p_cDx = pmin(pmax(cDx / Nx, 1), 0)) %>% 
      # mutate(p_cDx = pmin(pmax(cDx / Nx, 0), 1))
```

```{r}
# Cohorts
tp_primers %>% 
      mutate(cohort = year - age) %>% 
      filter(cause %in% c(cause1[4], cause2[1], cause3),
             sex == "Female",
             cohort %in% seq(1960, 2000, 10)) %>% 
      mutate(cause = recode_values(cause,
                                   cause1[4] ~ "Liver",
                                   cause2[1] ~ "Breast",
                                   cause3[1] ~ "Cervix",
                                   cause3[2] ~ "CRC")) %>% 
      pivot_longer(cols = cix:bgmx,
                   names_to = "measure",
                   values_to = "val") %>% 
      
      # filter(measure %!in% c("acmx", "bgmx")) %>% 
      filter(measure %in% c("cix", "cmx", "acmx")) %>% 
      
      ggplot(aes(x = year, y = val, group = measure, colour = measure)) +
      geom_line() +
      scale_y_continuous(name = "Rate (per 100,000)",
                         labels = scales::label_comma(scale = 1e5),
                         n.breaks = 3, #) +
                         trans = "log10") +
      facet_grid(cohort ~ cause, scales = "free_y") +
      theme_caviz +
      theme(legend.position = "bottom",
            aspect.ratio = 1 / 3,
            strip.text.y.right = element_text(colour = "black", size = 10),
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10))
```

```{r}
# Age
bsln_tps %>% 
      mutate(cohort = year - age) %>% 
      filter(cause %in% c(cause1[4], cause2[1], cause3),
             sex == "Female",
             # cohort %in% seq(1960, 2000, 10),
             age %in% seq(40, 70, 10)) %>% 
      mutate(cause = recode_values(cause,
                                   cause1[4] ~ "Liver",
                                   cause2[1] ~ "Breast",
                                   cause3[1] ~ "Cervix",
                                   cause3[2] ~ "CRC")) %>% 
      pivot_longer(cols = cix:bgmx,
                   names_to = "measure",
                   values_to = "val") %>% 
      
      # filter(measure %!in% c("acmx", "bgmx")) %>% 
      filter(measure %in% c("cix", "cmx", "acmx")) %>% 
      
      ggplot(aes(x = year, y = val, group = measure, colour = measure)) +
      geom_line() +
      scale_y_continuous(name = "Rate (per 100,000)",
                         labels = scales::label_comma(scale = 1e5),
                         n.breaks = 3, #) +
                         trans = "log10") +
      facet_grid(age ~ cause, scales = "free_y") +
      theme_caviz +
      theme(legend.position = "bottom",
            aspect.ratio = 1 / 3,
            strip.text.y.right = element_text(colour = "black", size = 10),
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10))
```

```{r}
tp_primers %>% 
      mutate(cohort = year - age) %>% 
      filter(#cause %in% c(cause1[4], cause2[1], cause3),
             sex == "Female",
             cohort %in% seq(1960, 2000, 10)) %>% 
      mutate(cause = recode_values(cause,
                                   cause1[4] ~ "Liver",
                                   cause2[1] ~ "Breast",
                                   cause3[1] ~ "Cervix",
                                   cause3[2] ~ "CRC")) %>% 
      # pivot_longer(cols = cix:bgmx,
      #              names_to = "measure",
      #              values_to = "val") %>% 
      
      mutate(diff = acmx - cmx) %>% 
      select(diff) %>% 
      pull() %>% 
      range()
      
      # # filter(measure %!in% c("acmx", "bgmx")) %>% 
      # filter(measure %in% c("cmx", "acmx")) %>% 
      
      ggplot(aes(x = year, y = diff)) +
      geom_line() +
      scale_y_continuous(name = "Rate (per 100,000)",
                         # labels = scales::label_comma(scale = 1e5),
                         n.breaks = 3) +
      facet_grid(cohort ~ cause, scales = "free_y") +
      theme_caviz +
      theme(legend.position = "bottom",
            aspect.ratio = 1 / 3,
            strip.text.y.right = element_text(colour = "black", size = 10),
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10))
```

## Checks:

### 1) CMX vs. ACMX

** Prostate cancer, 90+

```{r}
full_join(x  = country_mort %>% select(!all_of(c("log_fitted_val", "measure"))),
          y  = country_acmx %>% select(!all_of(c("log_fitted_val", "cause", "measure"))),
          by = join_by(location, sex, year, age, pop, age_gp)) %>%
      mutate(diff = acmx - cmx) %>% 
      filter(diff < 0) %>%
      # count(cause, age)
      # select(diff) %>% 
      # pull() %>% 
      # range()
      filter(#year == sim_start_year,
             #age %in% seq(40, 70, 5),
             case_when(cause == cause2[2] ~ sex == "Male",
                       TRUE               ~ sex == "Female")) %>% 
      mutate(age = as.character(age)) %>% 
      # select(diff) %>% 
      ggplot(aes(x = year, y = diff, color = age, group = age)) +
      geom_line() +
      scale_y_continuous(n.breaks = 3) +
      # facet_wrap(. ~ cause) +
      theme_caviz +
      theme(legend.position = "bottom",
            aspect.ratio = 1 / 5,
            # strip.text.y.right = element_text(colour = "black", size = 10),
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10))
```

### 2) BSLN TPs

```{r}
bsln_tps %>% 
      mutate(cohort = year - age) %>% 
      filter(case_when(cause == cause2[2] ~ sex == "Male",
                       TRUE               ~ sex == "Female"),
             cohort == 1980) %>% 
      pivot_longer(cols = w_prc:dst_do,
                   names_to = "tp",
                   values_to = "val") %>% 
      
      filter(case_when(cause %in% cause1            ~ tp %in% c("w_lcl", "lcl_d"),
                       cause %in% c(cause2, cause3) ~ tp %in% c("w_prc", "w_lcl",
                                                                "prc_lcl",
                                                                "rgn_dst", "dst_d"))) %>% 
      
      ggplot(aes(x = year, y = val, col = tp, group = tp)) +
      geom_line() +
      scale_y_continuous(n.breaks = 3) +
      facet_wrap(. ~ cause, scales = "free_y") +
      theme_caviz +
      theme(aspect.ratio = 1 / 4,
            legend.position = "bottom",
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10))
```

```{r}
bsln_tps %>% 
      mutate(cohort = year - age) %>% 
      filter(case_when(cause == cause2[2] ~ sex == "Male",
                       TRUE               ~ sex == "Female"),
             cohort == 1980) %>% 
      pivot_longer(cols = w_prc:dst_do,
                   names_to = "tp",
                   values_to = "val") %>% 
      
      filter(case_when(cause %in% cause1            ~ tp %in% c("w_lcl", "lcl_d"),
                       cause %in% c(cause2, cause3) ~ tp %in% c("w_prc", "w_lcl",
                                                                "prc_lcl",
                                                                "rgn_dst", "dst_d",
                                                                "dst_do"))) %>% 
      
      filter(cause %in% cause3) %>% 
      
      ggplot(aes(x = year, y = val, col = tp, group = tp)) +
      geom_line() +
      scale_y_continuous(n.breaks = 3) +
      facet_wrap(. ~ cause, scales = "free_y") +
      theme_caviz +
      theme(aspect.ratio = 1 / 4,
            legend.position = "bottom")
```

```{r}
bsln_tps %>% 
      mutate(cohort = year - age) %>% 
      pivot_longer(cols = w_prc:dst_do,
                   names_to = "tp",
                   values_to = "val") %>% 
      filter(val > 1)
```

### Examining TP rowSums

```{r}
tp_rowsums <- bsln_tps %>% 
      mutate(cohort = year - age) %>% 
      group_by(age, year, cohort, sex, cause, location) %>% 
      reframe(well = rowSums(pick(starts_with("w_"))),
              prc  = rowSums(pick(starts_with("prc_"))),
              lcl  = rowSums(pick(starts_with("lcl_"))),
              rgn  = rowSums(pick(starts_with("rgn_"))),
              dst  = rowSums(pick(starts_with("dst_"))),
              d    = rowSums(pick(starts_with("d_"))),
              do   = rowSums(pick(starts_with("do_")))) %>% 
      
      pivot_longer(cols = well:do,
                   names_to = "hstate",
                   values_to = "tp_val") %>% 
      count(age, sex, cause, hstate, tp_val) 

tp_rowsums %>% 
      # filter(tp_val %!in% c(0, 1)) %>% 
      filter(cause %!in% cause3) %>% 
      filter(!near(tp_val, 0) & !near(tp_val, 1)) %>% 
      view()
```

## Examining `calc_indiv_tp_vec`

```{r}
test_hsd_stg <- full_join(x = start_hstate_dist,
                          y = bsln_stg_dist) %>% 
      filter(age == 55, sex == "Female", cause == cause1[4], year == 2000) %>% 
      mutate(category = 1,
             input_tp0 = 0.005,
             input_tp2 = 0.3,
             input_tp3 = 0.4)

test_epi <- reduce(
      .x = list(country_incd %>% select(all_of(c(uid_vec, "cix"))),
                country_prev %>% select(all_of(c(uid_vec, "cpx"))),
                country_mort %>% select(all_of(c(uid_vec, "cmx"))),
                country_acmx %>% select(all_of(c(uid_vec[-4], "acmx")))),
      .f = full_join) %>% 
      mutate(bgmx = acmx - cmx) %>% 
      filter(age == 55, sex == "Female", cause == cause1[4], year == 2000)

test_hsd_stg_epi <- test_epi[test_hsd_stg, on = .(location, sex, year, age, cause)]

t(test_hsd_stg_epi)
# test_hsd_stg <- data.table::as.data.table(test_hsd_stg)

test_indiv_tp_vec <- test_hsd_stg_epi[, c("w_w", "w_prc",   "w_lcl",   "w_rgn",   "w_dst",   "w_d",   "w_do",
                           "prc_prc", "prc_lcl",                                "prc_do",
                                      "lcl_lcl", "lcl_rgn",            "lcl_d", "lcl_do",
                                                 "rgn_rgn", "rgn_dst",          "rgn_do",
                                                            "dst_dst", "dst_d", "dst_do") :=
                  calc_indiv_tp_vec(cat       = category,
                                    s0        = precancer,
                                    s1        = local,
                                    s2        = regional,
                                    s3        = distant,
                                    w_sick    = cix,
                                    prev      = cpx,
                                    sick_d    = cmx,
                                    bgmx      = bgmx,
                                    solve_tp0 = input_tp0,
                                    solve_tp2 = input_tp2,
                                    solve_tp3 = input_tp3),
             by = .(location, cause, sex, year, age)]

test_indiv_tp_cat <- test_hsd_stg_epi[, c("w_w", "w_prc",   "w_lcl",                           "w_do",
                                          "prc_prc", "prc_lcl",                                "prc_do",
                                                     "lcl_lcl", "lcl_rgn",            "lcl_d", "lcl_do",
                                                                "rgn_rgn", "rgn_dst",          "rgn_do",
                                                                           "dst_dst", "dst_d", "dst_do") :=
                  calc_indiv_tp_cat(cat       = category,
                                    s0        = precancer,
                                    s1        = local,
                                    s2        = regional,
                                    s3        = distant,
                                    w_sick    = cix,
                                    prev      = cpx,
                                    sick_d    = cmx,
                                    bgmx      = bgmx,
                                    solve_tp0 = input_tp0,
                                    solve_tp2 = input_tp2,
                                    solve_tp3 = input_tp3),
             by = .(location, cause, sex, year, age)]

t(test_indiv_tp_cat) #%>% as.data.frame()
```

```{r}
calc_indiv_tp_cat(cat       = 1,
                  s0        = 0,
                  s1        = 1,
                  s2        = 0,
                  s3        = 0,
                  w_sick    = 0.0001057003,
                  prev      = 0.0001237225,
                  sick_d    = 0.0001051507,
                  bgmx      = 0.02762172,
                  solve_tp0 = 0.005,
                  solve_tp2 = 0.3,
                  solve_tp3 = 0.4)["w_w"] %>% class()
      unlist() %>% t() %>% t()
```

```{r}
test_indiv_tp_vec %>% 
      mutate(cohort = year - age) %>% 
      group_by(age, year, cohort, sex, cause, location) %>% 
      reframe(well = rowSums(pick(starts_with("w_"))),
              prc  = rowSums(pick(starts_with("prc_"))),
              lcl  = rowSums(pick(starts_with("lcl_"))),
              rgn  = rowSums(pick(starts_with("rgn_"))),
              dst  = rowSums(pick(starts_with("dst_"))),
              d    = rowSums(pick(starts_with("d_"))),
              do   = rowSums(pick(starts_with("do_")))) %>% 
      
      pivot_longer(cols = well:do,
                   names_to = "hstate",
                   values_to = "tp_val") %>% 
      count(age, sex, cause, hstate, tp_val)
```

```{r}
test_indiv_tp_cat %>% 
      mutate(cohort = year - age) %>% 
      group_by(age, year, cohort, sex, cause, location) %>% 
      reframe(well = rowSums(pick(starts_with("w_"))),
              prc  = rowSums(pick(starts_with("prc_"))),
              lcl  = rowSums(pick(starts_with("lcl_"))),
              rgn  = rowSums(pick(starts_with("rgn_"))),
              dst  = rowSums(pick(starts_with("dst_"))),
              d    = rowSums(pick(starts_with("d_"))),
              do   = rowSums(pick(starts_with("do_")))) %>% 
      
      pivot_longer(cols = well:do,
                   names_to = "hstate",
                   values_to = "tp_val") %>% 
      count(age, sex, cause, hstate, tp_val) 
```

