---
title: "Reducing Cardiovascular Disease in Indonesia"
subtitle: "Health Impact, Costs, and Economic Returns of Clinical and Public-Health Interventions, 2025--2050"
author: "World Bank $\\cdot$ University of Washington"
institute: "Briefing for the Ministry of Health, Republic of Indonesia"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  beamer_presentation:
    theme: "default"
    keep_tex: false
    latex_engine: xelatex
    slide_level: 2
    toc: false
    includes:
      in_header: beamer_preamble.tex
classoption: "aspectratio=169"
bibliography: references.bib
link-citations: false
---

```{r setup, include=FALSE}
## =====================================================================
##  EXECUTIVE DECK — Indonesia CVD / FAIR Choices analysis
##  World Bank · University of Washington
##
##  DATA PROVENANCE
##  Every displayed number is imported programmatically from the three
##  cost/value output workbooks written by the modeling pipeline
##  (code/cvd-fair-choices/09_cost_value.R):
##    output/indonesia_model_cost_value_formulae.xlsx                    (clinical)
##    output/indonesia_cost_value_public_health_formulae.xlsx            (public health)
##    output/indonesia_model_cost_value_clinical_public_health_formulae.xlsx (combined)
##  and from data/raw/03_asean_cvd_40q30_comparison.csv (international comparator).
##  Intervention definitions come from the workbooks' own Selected_Interventions
##  / Policy_Levers sheets. NOTHING is transcribed by hand.
##
##  Those workbooks are FORMULA workbooks. Formula cells are recalculated in a
##  TEMPORARY copy (Excel COM; the source files are never modified) so that all
##  imported values are current. If recalculation is unavailable, economic
##  metrics fall back to the combined workbook's stored values (which are
##  identical by construction — same engine, same assumptions) and a note is
##  recorded. Health-impact metrics are computed in R from the workbooks' own
##  stored per-year quantities and therefore never depend on recalculation.
## =====================================================================
library(here)
library(data.table)
library(openxlsx)
library(ggplot2)
library(ggrepel)
library(scales)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.align = "center", dev = "cairo_pdf"
)

## ---- Palette (matches beamer_preamble.tex scenario colours) ---------------
col_clinical  <- "#3B6FB6"  # clinical scenarios
col_ph        <- "#1B9E77"  # public-health scenarios
col_combined  <- "#08306B"  # combined package (darker)
col_baseline  <- "#6E6E6E"  # baseline / status quo
col_comparator<- "#BDBDBD"  # comparator countries (muted)
col_incidence <- "#2166AC"  # incidence (Well -> Sick) effects
col_casefatal <- "#B2182B"  # case-fatality (Sick -> Death) effects
col_bgmort    <- "#7F7F7F"  # background mortality
col_wbnavy    <- "#0F2A56"

theme_exec <- function(base = 11) {
  theme_minimal(base_size = base) +
    theme(
      plot.title       = element_text(face = "bold", size = base + 1, colour = col_wbnavy),
      plot.subtitle    = element_text(colour = "grey35", size = base - 2),
      plot.caption     = element_text(colour = "grey45", size = base - 4, hjust = 0),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      legend.position  = "bottom",
      legend.title     = element_blank(),
      axis.title       = element_text(size = base - 2)
    )
}

## ---- Number formatters (LaTeX-safe: escaped $ and % ; used with escape=FALSE)
fmt_int   <- function(x) formatC(round(x), format = "d", big.mark = ",")
fmt_thou  <- function(x) formatC(round(x / 1000) * 1000, format = "d", big.mark = ",")
fmt_short <- function(x, acc = 0.1) label_number(scale_cut = cut_short_scale(), accuracy = acc)(x)
fmt_pct   <- function(x, d = 1) ifelse(is.na(x), "--", paste0(formatC(x, format = "f", digits = d), "\\%"))
fmt_usd   <- function(x) ifelse(is.na(x), "--", paste0("US\\$", formatC(round(x), format = "d", big.mark = ",")))
fmt_usd2  <- function(x) ifelse(is.na(x), "--", paste0("US\\$", formatC(x, format = "f", digits = 2, big.mark = ",")))
fmt_usd_short <- function(x) ifelse(is.na(x), "--", paste0("US\\$", fmt_short(x)))
fmt_percap <- function(x) ifelse(is.na(x), "--",
  ifelse(x < 0.005, "<US\\$0.01", paste0("US\\$", formatC(x, format = "f", digits = 2, big.mark = ","))))
fmt_intl  <- function(x) ifelse(is.na(x), "--", paste0("Int\\$", fmt_short(x)))
fmt_ratio <- function(x) {
  ifelse(is.na(x), "--",
    ifelse(x >= 1000, formatC(signif(x, 3), format = "d", big.mark = ","),
    ifelse(x >= 100,  formatC(round(x), format = "d", big.mark = ","),
                      formatC(x, format = "f", digits = 1))))
}
## Two-line table headers (converts embedded "\n" to \makecell line breaks).
mkhdr <- function(v) kableExtra::linebreak(v, align = "c")
```

```{r recalc-and-read, include=FALSE}
## ---------------------------------------------------------------------------
## 1. Resolve workbook paths (project-relative) and check existence.
## ---------------------------------------------------------------------------
wb_clin <- here("output", "indonesia_model_cost_value_formulae.xlsx")
wb_ph   <- here("output", "indonesia_cost_value_public_health_formulae.xlsx")
wb_comb <- here("output", "indonesia_model_cost_value_clinical_public_health_formulae.xlsx")
csv_cmp <- here("data", "raw", "03_asean_cvd_40q30_comparison.csv")

.req_files <- c(clinical = wb_clin, public_health = wb_ph, combined = wb_comb, comparator = csv_cmp)
.missing <- .req_files[!file.exists(.req_files)]
if (length(.missing))
  stop("VALIDATION STOP — required input file(s) not found:\n  ",
       paste(sprintf("%s: %s", names(.missing), .missing), collapse = "\n  "),
       "\nRun the modeling pipeline (code/cvd-fair-choices/00_run_model_cvd_fair.R) first.",
       call. = FALSE)

## ---------------------------------------------------------------------------
## 2. Recalculate a formula workbook into a TEMP copy (Excel COM on Windows).
##    The source workbook is opened read-only and never modified.
## ---------------------------------------------------------------------------
recalc_workbook <- function(src) {
  dst <- tempfile(fileext = ".xlsx")
  ps  <- tempfile(fileext = ".ps1")
  writeLines(c(
    'param([string]$Src,[string]$Dst)',
    '$ErrorActionPreference="Stop"',
    '$xl = New-Object -ComObject Excel.Application',
    '$xl.Visible=$false; $xl.DisplayAlerts=$false; $xl.AskToUpdateLinks=$false; $xl.EnableEvents=$false',
    'try {',
    '  $wb = $xl.Workbooks.Open($Src, 0, $true)',   # ReadOnly = TRUE
    '  $xl.CalculateFullRebuild()',
    '  $wb.SaveAs($Dst, 51)',                       # 51 = xlOpenXMLWorkbook (.xlsx)
    '  $wb.Close($false)',
    '} finally {',
    '  $xl.Quit()',
    '  [System.Runtime.InteropServices.Marshal]::ReleaseComObject($xl) | Out-Null',
    '}'), ps)
  ok <- tryCatch({
    system2("powershell",
            c("-NoProfile", "-ExecutionPolicy", "Bypass", "-File", shQuote(ps),
              "-Src", shQuote(normalizePath(src)), "-Dst", shQuote(dst)),
            stdout = TRUE, stderr = TRUE)
    file.exists(dst) && file.info(dst)$size > 0
  }, error = function(e) FALSE)
  if (isTRUE(ok)) dst else NA_character_
}

recalc_ok <- TRUE
resolve_path <- function(src) {
  p <- recalc_workbook(src)
  if (is.na(p)) { recalc_ok <<- FALSE; return(src) }
  p
}
path_clin <- resolve_path(wb_clin)
path_ph   <- resolve_path(wb_ph)
path_comb <- resolve_path(wb_comb)

## ---------------------------------------------------------------------------
## 3. Read helper.
## ---------------------------------------------------------------------------
rd <- function(path, sheet) {
  as.data.table(openxlsx::read.xlsx(path, sheet = sheet))
}
sheets_of <- function(path) openxlsx::getSheetNames(path)
```

```{r build-metrics, include=FALSE}
## ---------------------------------------------------------------------------
## 4. Per-scenario metric builder. Reads a workbook's OWN sheets and computes
##    every headline metric from the model's established definitions. Health
##    metrics use stored per-year quantities (no recalculation needed); cost /
##    benefit-cost metrics use the recalculated formula cells.
## ---------------------------------------------------------------------------
HZ_START <- 2025L; HZ_END <- 2050L

build_metrics <- function(path) {
  ho <- rd(path, "Health_Outcomes")
  c4 <- rd(path, "CVD_40q30")
  ce <- rd(path, "Cost_Effectiveness")
  bc <- rd(path, "Benefit_Cost")

  ## --- Cumulative deaths averted (sum over horizon of baseline - scenario) ---
  ho[, da := baseline_deaths - modeled_deaths]
  m_da <- ho[, .(cum_deaths_averted = sum(da),
                 cum_dalys_averted  = sum(base_daly - daly),
                 life_years_gained  = sum(life_years_gained)), by = scenario]

  ## --- % reduction in deaths in 2050 ---------------------------------------
  m_d50 <- ho[year == HZ_END, .(scenario,
                pct_red_deaths_2050 = 100 * (baseline_deaths - modeled_deaths) / baseline_deaths,
                deaths_2050 = modeled_deaths, base_deaths_2050 = baseline_deaths)]

  ## --- CVD 40q30 in 2050 and % reduction (uses the R anchor column) --------
  base_q50 <- c4[scenario == "baseline" & year == HZ_END, cvd_40q30_r][1]
  base_q25 <- c4[scenario == "baseline" & year == HZ_START, cvd_40q30_r][1]
  m_q50 <- c4[year == HZ_END, .(scenario, q40q30_2050 = cvd_40q30_r)]
  m_q50[, pct_red_q40q30_2050 := 100 * (base_q50 - q40q30_2050) / base_q50]

  ## --- Cost-effectiveness (cumulative, discounted where noted) -------------
  m_ce <- ce[, .(scenario,
                 ce_deaths_averted = deaths_averted,
                 cum_cost_undisc   = incremental_cost,
                 cum_cost_disc     = disc_incremental_cost,
                 cost_per_death    = cost_per_death_averted)]

  ## --- Benefit-cost: headline = preferred VSL; secondary = preferred VSLY --
  bc_vsl  <- bc[valuation_method == "VSL"  & grepl("preferred", valuation_case),
                .(scenario, pv_benefits_vsl = pv_benefits, pv_costs = pv_costs,
                  net_benefit_vsl = pv_net_benefit, bcr_vsl = benefit_cost_ratio)]
  bc_vsly <- bc[valuation_method == "VSLY" & grepl("preferred", valuation_case),
                .(scenario, bcr_vsly = benefit_cost_ratio)]

  out <- Reduce(function(a, b) merge(a, b, by = "scenario", all = TRUE),
                list(unique(ho[, .(scenario, scenario_label)]),
                     m_da, m_d50, m_q50, m_ce, bc_vsl, bc_vsly))
  attr(out, "base_q40_2050") <- base_q50
  attr(out, "base_q40_2025") <- base_q25
  out[]
}

met_clin <- build_metrics(path_clin)
met_ph   <- build_metrics(path_ph)
met_comb <- build_metrics(path_comb)

## National population (denominator for per-capita costs) from Economic_Value.
ev_comb <- rd(path_comb, "Economic_Value")
pop_series <- unique(ev_comb[, .(year, population)])[order(year)]
mean_pop   <- mean(pop_series$population)
n_years    <- length(seq(HZ_START, HZ_END))
add_percap <- function(dt) {
  dt[, ann_cost_per_capita := cum_cost_undisc / n_years / mean_pop][]
}
met_clin <- add_percap(met_clin); met_ph <- add_percap(met_ph); met_comb <- add_percap(met_comb)

## Fallback: if live recalculation was unavailable, the clinical / public-health
## workbooks store no economic values on disk — coalesce those columns from the
## combined workbook (identical scenarios, identical engine & assumptions).
econ_cols <- c("cum_cost_undisc","cum_cost_disc","cost_per_death","ann_cost_per_capita",
               "pv_benefits_vsl","pv_costs","net_benefit_vsl","bcr_vsl","bcr_vsly")
coalesce_from_comb <- function(dt) {
  for (cc in econ_cols) {
    need <- is.na(dt[[cc]])
    if (any(need)) {
      fill <- met_comb[match(dt$scenario[need], scenario)][[cc]]
      dt[[cc]][need] <- fill
    }
  }
  dt[]
}
econ_fallback_used <- FALSE
if (!recalc_ok) {
  met_clin <- coalesce_from_comb(met_clin)
  met_ph   <- coalesce_from_comb(met_ph)
  econ_fallback_used <- TRUE
}

## Assumptions (for footnotes) — parameter_id -> value.
asm_dt <- rd(path_comb, "Calculation_Assumptions")
asm <- setNames(asm_dt$value, asm_dt$parameter_id)
gA  <- function(id, default = NA) if (id %in% names(asm)) asm[[id]] else default
disc_rate  <- as.numeric(gA("bca_discount_rate_primary", 0.03))
base_year  <- as.numeric(gA("bca_base_year", 2025))
price_year <- as.numeric(gA("bca_price_year", 2023))
vsl_elast  <- as.numeric(gA("vsl_income_elasticity_preferred", 1.5))
cur_basis  <- gA("bca_currency_basis", "PPP_international_dollars")
bca_scope  <- gA("bca_scope", "partial_mortality_benefit")
bca_stand  <- gA("bca_standing", "societal")
```

```{r validation, include=FALSE}
## ---------------------------------------------------------------------------
## 5. VALIDATION GATES — stop rendering with an informative error on failure.
## ---------------------------------------------------------------------------
val_stop <- function(cond, msg) if (!isTRUE(cond)) stop("VALIDATION STOP — ", msg, call. = FALSE)

## (a) required sheets present in every workbook
need_sheets <- c("Health_Outcomes","CVD_40q30","Cost_Effectiveness","Benefit_Cost",
                 "Economic_Value","Calculation_Assumptions")
for (nm in c("clinical","public_health","combined")) {
  p  <- switch(nm, clinical = path_clin, public_health = path_ph, combined = path_comb)
  sh <- sheets_of(p)
  val_stop(all(need_sheets %in% sh),
           sprintf("workbook '%s' is missing sheet(s): %s", nm,
                   paste(setdiff(need_sheets, sh), collapse = ", ")))
}

## (b) required aggregate scenarios present. Individual interventions are
## intentionally NOT required: the input flag can enable/disable them.
val_stop("all" %in% met_clin$scenario,
         "clinical workbook missing the aggregate 'all' scenario")
val_stop("all_public_health" %in% met_ph$scenario,
         "public-health workbook missing the aggregate 'all_public_health' scenario")
val_stop("all_clinical_public_health" %in% met_comb$scenario,
         "combined workbook missing the joint 'all_clinical_public_health' scenario")

## (c) baseline present + 2050 available in each workbook
for (nm in c("clinical","public_health","combined")) {
  p <- switch(nm, clinical = path_clin, public_health = path_ph, combined = path_comb)
  ho <- rd(p, "Health_Outcomes"); c4 <- rd(p, "CVD_40q30")
  val_stop(any(c4$scenario == "baseline"), sprintf("'%s' has no baseline in CVD_40q30", nm))
  val_stop(HZ_END %in% ho$year && HZ_END %in% c4$year,
           sprintf("'%s' has no year %d", nm, HZ_END))
}

## (d) no unexpected duplicate scenarios in Cost_Effectiveness
for (nm in c("clinical","public_health","combined")) {
  p  <- switch(nm, clinical = path_clin, public_health = path_ph, combined = path_comb)
  ce <- rd(p, "Cost_Effectiveness")
  val_stop(!any(duplicated(ce$scenario)),
           sprintf("'%s' Cost_Effectiveness has duplicate scenario id(s): %s", nm,
                   paste(unique(ce$scenario[duplicated(ce$scenario)]), collapse = ", ")))
}

## (e) headline metrics numeric
headline <- met_comb[scenario == "all_clinical_public_health"]
val_stop(is.finite(headline$cum_deaths_averted) &&
         is.finite(headline$q40q30_2050) &&
         is.finite(headline$cost_per_death) &&
         is.finite(headline$bcr_vsl),
         "combined-package headline metrics are missing/non-numeric (recalculation may have failed)")

## (f) cross-workbook baseline reconciliation (uses stored per-year quantities,
##     independent of recalculation). Tolerance = 0.1% relative.
base_deaths_2050 <- function(p) rd(p,"Health_Outcomes")[year==HZ_END, baseline_deaths][1]
base_q40_2050    <- function(p) rd(p,"CVD_40q30")[scenario=="baseline" & year==HZ_END, cvd_40q30_r][1]
bd <- c(clin = base_deaths_2050(path_clin), ph = base_deaths_2050(path_ph), comb = base_deaths_2050(path_comb))
bq <- c(clin = base_q40_2050(path_clin),    ph = base_q40_2050(path_ph),    comb = base_q40_2050(path_comb))
rel_range <- function(v) (max(v) - min(v)) / mean(v)
RECON_TOL <- 1e-3
val_stop(rel_range(bd) < RECON_TOL,
         sprintf("baseline deaths in 2050 disagree across workbooks beyond %.1f%%: %s",
                 100*RECON_TOL, paste(sprintf("%s=%.0f", names(bd), bd), collapse = ", ")))
val_stop(rel_range(bq) < RECON_TOL,
         sprintf("baseline CVD 40q30 in 2050 disagree across workbooks beyond %.1f%%: %s",
                 100*RECON_TOL, paste(sprintf("%s=%.3f", names(bq), bq), collapse = ", ")))
recon_note <- sprintf("Baseline reconciled across all three workbooks (2050 deaths agree to %.4f\\%%; 2050 CVD 40q30 to %.4f\\%%).",
                      100*rel_range(bd), 100*rel_range(bq))
```

```{r headline-scalars, include=FALSE}
## Convenience getters + headline scalars used in prose.
hv <- function(sc, col, tab = met_comb) { v <- tab[scenario == sc][[col]]; if (length(v)) v[1] else NA_real_ }

base_q_2025 <- attr(met_comb, "base_q40_2025")
base_q_2050 <- attr(met_comb, "base_q40_2050")

combo_da    <- hv("all_clinical_public_health","cum_deaths_averted")
combo_cpda  <- hv("all_clinical_public_health","cost_per_death")
combo_bcr   <- hv("all_clinical_public_health","bcr_vsl")
combo_net   <- hv("all_clinical_public_health","net_benefit_vsl")
combo_q50   <- hv("all_clinical_public_health","q40q30_2050")
combo_qred  <- hv("all_clinical_public_health","pct_red_q40q30_2050")
combo_dred  <- hv("all_clinical_public_health","pct_red_deaths_2050")
combo_daly  <- hv("all_clinical_public_health","cum_dalys_averted")

clin_da   <- hv("all","cum_deaths_averted", met_clin);  clin_bcr <- hv("all","bcr_vsl", met_clin)
clin_qred <- hv("all","pct_red_q40q30_2050", met_clin); clin_cpda <- hv("all","cost_per_death", met_clin)
ph_da     <- hv("all_public_health","cum_deaths_averted", met_ph)
ph_bcr    <- hv("all_public_health","bcr_vsl", met_ph)
ph_qred   <- hv("all_public_health","pct_red_q40q30_2050", met_ph)
ph_cpda   <- hv("all_public_health","cost_per_death", met_ph)

## Top single interventions by deaths averted (within family).
## Exec-friendly labels + display ordering.
exec_label <- c(
  I_CVD_PRIMARY="Primary CVD prevention", I_T2D_TREATMENT="Type-2 diabetes treatment",
  I_IHD_SECONDARY="Heart-attack secondary prevention", I_HF_CHRONIC="Chronic heart-failure care",
  I_HF_ACUTE="Acute heart-failure care", I_STROKE_SECONDARY="Stroke secondary prevention",
  I_RHD_SECONDARY="RHD secondary prevention", I_RHD_SURGERY="RHD cardiac surgery",
  all="All clinical (combined)",
  I_PH_TOBACCO_TAX="Tobacco tax", I_PH_TOBACCO_POLICY="Tobacco policy package",
  I_PH_SALT_POLICY="Sodium policy package", I_PH_TFA_POLICY="Trans-fat elimination",
  I_PH_SSB_TAX="Sugary-drink tax", I_PH_ALCOHOL_TAX="Alcohol tax",
  I_PH_ALCOHOL_POLICY="Alcohol advertising ban", all_public_health="All public health (combined)",
  all_clinical_public_health="Clinical + public health (combined)", baseline="Status quo (baseline)")

## Use the workbook as the source of truth for enabled interventions. Known IDs
## receive executive labels; any future IDs fall back to the workbook label.
scenario_label <- function(tab, sc) {
  if (sc %in% names(exec_label)) return(unname(exec_label[[sc]]))
  z <- tab[scenario == sc, scenario_label]
  if (length(z) && !is.na(z[1]) && nzchar(z[1])) return(z[1])
  tools::toTitleCase(gsub("_", " ", sub("^I_(PH_)?", "", sc)))
}

active_ids <- function(tab, aggregate_id) {
  ids <- tab[!scenario %in% c("baseline", aggregate_id) &
             is.finite(cum_deaths_averted), unique(scenario)]
  ids <- tab[scenario %in% ids][order(-cum_deaths_averted), unique(scenario)]
  c(ids, aggregate_id)
}
clin_order <- active_ids(met_clin, "all")
ph_order   <- active_ids(met_ph, "all_public_health")

top_clin <- met_clin[scenario %in% setdiff(clin_order, "all")][order(-cum_deaths_averted)][1]
top_ph   <- met_ph[scenario %in% setdiff(ph_order, "all_public_health")][order(-cum_deaths_averted)][1]
top_clin_label <- if (nrow(top_clin)) scenario_label(met_clin, top_clin$scenario[1]) else "the combined clinical package"
top_ph_label   <- if (nrow(top_ph)) scenario_label(met_ph, top_ph$scenario[1]) else "the combined public-health package"
```

```{r fig-40q30-data, include=FALSE}
## ---------------------------------------------------------------------------
## Data for the CVD 40q30 time-series & international-comparison figure.
##   Indonesia model scenarios (2025-2050): combined workbook CVD_40q30.
##   Observed history + comparators (1995-2050): the ASEAN comparator CSV
##   (WPP-scaled), used AS-IS with no extrapolation.
## ---------------------------------------------------------------------------
cmp <- fread(csv_cmp)
cmp[, year := as.integer(year)]
comparators <- c("Malaysia","Philippines","Singapore","Thailand","Viet Nam")

## Comparator countries — full available series (thin muted background lines).
cmp_lines <- cmp[country_label %in% comparators,
                 .(country_label, year, q = q40q30_percent)][order(country_label, year)]
cmp_lines <- unique(cmp_lines, by = c("country_label","year"))

## Indonesia observed history (solid), 1995-2023 only.
idn_obs <- cmp[country_label == "Indonesia" & series_type == "Historical WPP-scaled baseline" &
               year <= 2023, .(year, q = q40q30_percent)][order(year)]

## Indonesia MODEL scenarios (dashed), 2025-2050.
c4c <- rd(path_comb, "CVD_40q30")
idn_scen <- c4c[scenario %in% c("baseline","all","all_public_health","all_clinical_public_health") &
                year >= HZ_START,
                .(scenario, year, q = cvd_40q30_r)]
idn_scen[, scen := factor(exec_label[scenario],
           levels = c("Status quo (baseline)","All clinical (combined)",
                      "All public health (combined)","Clinical + public health (combined)"))]
scen_pal <- c("Status quo (baseline)"=col_baseline, "All clinical (combined)"=col_clinical,
              "All public health (combined)"=col_ph, "Clinical + public health (combined)"=col_combined)
```

# Reducing CVD in Indonesia

## Why this matters for Indonesia

\begin{columns}[T]
\begin{column}{0.52\textwidth}
\begin{block}{Cardiovascular disease is Indonesia's largest health burden}
\normalsize
Under current trends, a person reaching age 30 faces roughly a
\hlkey{`r fmt_pct(base_q_2025,1)` chance} of dying from cardiovascular disease
before age 70 --- the CVD ``40q30''. Most of this is \textbf{preventable or treatable}.
\end{block}
\end{column}
\begin{column}{0.46\textwidth}
\begin{itemize}\itemsep4pt
  \item Premature CVD deaths (ages 30--69) have risen for two decades and remain high.
  \item Two complementary levers:
  \begin{itemize}
    \item \textcolor{clinicalcol}{\textbf{Clinical care}} --- prevent and better treat established disease.
    \item \textcolor{phcol}{\textbf{Population prevention}} --- tobacco, salt, trans-fat, sugary drinks, alcohol.
  \end{itemize}
  \item This analysis weighs their \textbf{health impact, cost, and economic return} together.
\end{itemize}
\end{column}
\end{columns}

\vspace{3pt}
\footnotesize Source: Indonesia CVD model (calibrated to GBD 2023; UN World Population Prospects 2024). CVD 40q30 = probability of dying from cardiovascular disease between exact ages 30 and 70.

## What this analysis set out to do

\begin{block}{General aim}
\normalsize Estimate the \textbf{health and economic effects} of scaling up evidence-based
\textbf{clinical} and \textbf{public-health} cardiovascular interventions in Indonesia through \textbf{2050}.
\end{block}

\vspace{2pt}
\begin{columns}[T]
\begin{column}{0.33\textwidth}
\begin{clinicalblock}{Aim 1 --- Burden}
\footnotesize Project the CVD burden and the premature-mortality indicator (CVD 40q30) under the \textbf{status quo} and under \textbf{alternative intervention scenarios}.
\end{clinicalblock}
\end{column}
\begin{column}{0.33\textwidth}
\begin{phblock}{Aim 2 --- Value for money}
\footnotesize Estimate \textbf{deaths averted}, mortality reductions, \textbf{incremental costs}, and \textbf{cost-effectiveness} for each intervention and package.
\end{phblock}
\end{column}
\begin{column}{0.33\textwidth}
\begin{combinedblock}{Aim 3 --- Economic return}
\footnotesize Estimate \textbf{economic benefits}, \textbf{benefit--cost ratios}, and the \textbf{added value of combining} clinical and public-health action.
\end{combinedblock}
\end{column}
\end{columns}

\vspace{3pt}
\footnotesize Horizon `r HZ_START`--`r HZ_END`; the "status quo" baseline is the no-new-intervention comparator against which every scenario is measured.

## Executive summary --- the headline numbers

\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{combinedblock}{Combining clinical + public health is the big prize}
\footnotesize Together they could avert about \hlkey{`r fmt_short(combo_da)` deaths} over `r HZ_START`--`r HZ_END`,
cut the CVD 40q30 in `r HZ_END` by \hlkey{`r fmt_pct(combo_qred,0)`}, and return about
\hlkey{`r fmt_ratio(combo_bcr)` int\$} in benefits per int\$ spent.
\end{combinedblock}
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep3pt
  \item \textcolor{clinicalcol}{\textbf{All clinical interventions}}: $\approx$`r fmt_short(clin_da)` deaths averted; CVD 40q30 down `r fmt_pct(clin_qred,0)` by `r HZ_END`.
  \item \textcolor{phcol}{\textbf{All public-health policies}}: $\approx$`r fmt_short(ph_da)` deaths averted; benefit--cost ratio $\approx$`r fmt_ratio(ph_bcr)`.
  \item \textbf{Most impactful single levers}: \textcolor{clinicalcol}{`r top_clin_label`} and \textcolor{phcol}{`r top_ph_label`}.
  \item \textbf{Affordability}: the combined package costs about `r fmt_usd(combo_cpda)` per death averted.
\end{itemize}
\end{column}
\end{columns}

\vspace{2pt}
\scriptsize All figures from the model's cost/value workbooks. Deaths averted are cumulative `r HZ_START`--`r HZ_END` across modeled cardiovascular and diabetes causes; benefits valued per the Reference Case (Robinson 2019) in PPP international dollars [@Robinson2019VSL].

## How the model works --- in plain language

\begin{columns}[T]
\begin{column}{0.54\textwidth}
\footnotesize
\begin{enumerate}\itemsep3pt
  \item \textbf{People \& demography.} Indonesia's population is projected by age and sex (UN World Population Prospects 2024).
  \item \textbf{Starting health.} Disease and death rates are set to Indonesia's data (GBD 2023) and calibrated to match observed levels.
  \item \textbf{Year-by-year simulation.} Each year people can develop cardiovascular disease or diabetes, survive, or die.
  \item \textbf{Interventions.} Clinical and public-health scenarios change how many people develop disease and how many survive it.
  \item \textbf{Outcomes.} The model reports deaths, healthy years (DALYs), premature-mortality (40q30), costs, cost-effectiveness, and economic returns.
\end{enumerate}
\end{column}
\begin{column}{0.44\textwidth}
\begin{block}{One consistent engine}
\footnotesize Every scenario runs through the \textbf{same} projection, so results are directly comparable and the \textbf{combined} package is modeled as one run --- not a sum of separate estimates.
\end{block}
\vspace{2pt}
\scriptsize Calibration fits the model to GBD estimates using a standard numerical optimiser (Nelder--Mead). Life expectancy for lost-years and economic value uses UN WPP 2024.
\end{column}
\end{columns}

## How interventions change the disease pathway

\vspace{-2pt}
\begin{center}
\begin{tikzpicture}[
   >=Latex, node distance = 16mm and 30mm, font=\footnotesize,
   state/.style = {rectangle, rounded corners, draw=wbnavy, thick, fill=wbnavy!6,
                   minimum width=26mm, minimum height=11mm, align=center},
   death/.style = {rectangle, rounded corners, draw=black!55, thick, fill=black!4,
                   minimum width=26mm, minimum height=10mm, align=center}]
  \node[state] (well) {\textbf{Healthy}\\[-1pt]\scriptsize (no CVD)};
  \node[state, right=of well] (sick) {\textbf{Living with CVD}\\[-1pt]\scriptsize (or diabetes)};
  \node[death, right=of sick] (cvd)  {\textbf{CVD death}};
  \node[death, below=14mm of well, xshift=23mm] (bg) {Other-cause (background) death};

  \draw[->, very thick, incidencecol]
       (well) -- node[above, midway, text=incidencecol]{\textbf{new CVD}}
                  node[below, midway, text=incidencecol, font=\scriptsize]{incidence} (sick);
  \draw[->, very thick, casefatalcol]
       (sick) -- node[above, midway, text=casefatalcol]{\textbf{dies of CVD}}
                  node[below, midway, text=casefatalcol, font=\scriptsize]{case-fatality} (cvd);
  \draw[->, thick, bgmortcol] (well) to[bend right=12] (bg);
  \draw[->, thick, bgmortcol] (sick) to[bend left=12]  (bg);
\end{tikzpicture}
\end{center}

\vspace{-2pt}
\begin{columns}[T]
\begin{column}{0.33\textwidth}\footnotesize
\textcolor{incidencecol}{\textbf{Fewer people fall ill}}\\ \scriptsize Primary prevention \& \emph{all} public-health policies lower \textbf{incidence}.
\end{column}
\begin{column}{0.34\textwidth}\footnotesize
\textcolor{casefatalcol}{\textbf{More people survive}}\\ \scriptsize Secondary prevention, treatment \& tobacco policy lower \textbf{case-fatality}.
\end{column}
\begin{column}{0.31\textwidth}\footnotesize
\textcolor{bgmortcol}{\textbf{Background mortality}}\\ \scriptsize Other-cause deaths from demography; unchanged by these interventions.
\end{column}
\end{columns}

\vspace{3pt}
\scriptsize No recovery/remission transition is modeled: people living with CVD remain at elevated risk (a conservative assumption). Source: model engine, \texttt{06\_run\_scenarios\_indonesia\_fair.R}.

# The interventions

## Clinical interventions (FAIR Choices)

```{r tbl-clin-interventions}
clin_def <- data.table(
  Intervention = c("Primary CVD prevention","Type-2 diabetes treatment",
                   "Heart-attack secondary prevention","Stroke secondary prevention",
                   "Chronic \\& acute heart-failure care","RHD secondary prevention \\& surgery"),
  Population = c("Adults at cardiovascular risk","People with type-2 diabetes",
                 "Heart-attack (IHD) survivors","Stroke survivors",
                 "Heart-failure patients","Rheumatic heart disease patients"),
  Acts = c("\\textcolor{incidencecol}{Incidence}","\\textcolor{casefatalcol}{Case-fatality}",
           "\\textcolor{casefatalcol}{Case-fatality}","\\textcolor{casefatalcol}{Case-fatality}",
           "\\textcolor{casefatalcol}{Case-fatality}","\\textcolor{casefatalcol}{Case-fatality}"))
kbl(clin_def, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=c("Intervention","Target population","Model effect"), align="lll") |>
  kable_styling(latex_options="hold_position", font_size=8.5) |>
  column_spec(1, bold=TRUE, width="4.6cm") |>
  column_spec(2, width="6.4cm") |> column_spec(3, width="3.0cm") |>
  row_spec(0, bold=TRUE)
```

\vspace{2pt}
\footnotesize \textbf{Implementation scenario (all clinical interventions):} coverage scales up \textbf{linearly from 31\% to 80\%} of the eligible population between `r HZ_START` and `r HZ_END` (RHD surgery from a lower base). Effect sizes are drawn from the FAIR Choices intervention taxonomy; primary prevention lowers who develops CVD, while secondary prevention and treatment improve survival. \scriptsize Source: workbook \texttt{Selected\_Interventions}.

## Public-health interventions (fiscal \& regulatory policy)

```{r tbl-ph-interventions}
ph_def <- data.table(
  Policy = c("Tobacco tax","Tobacco policy package","Sodium policy package",
             "Trans-fat elimination","Sugary-drink (SSB) tax","Alcohol tax / ad ban"),
  Risk = c("Tobacco","Tobacco","Dietary salt/sodium","Industrial trans-fat",
           "Sugary drinks","Alcohol"),
  Acts = c("\\textcolor{incidencecol}{Incidence} + \\textcolor{casefatalcol}{survival}",
           "\\textcolor{incidencecol}{Incidence} + \\textcolor{casefatalcol}{survival}",
           "\\textcolor{incidencecol}{Incidence}","\\textcolor{incidencecol}{Incidence}",
           "\\textcolor{incidencecol}{Incidence}","\\textcolor{incidencecol}{Incidence}"),
  Lever = c("Raise tax to 75\\% of retail price","Clean-air law, mass media, ad ban",
            "Reformulation, labelling, media, procurement","Best-practice industrial TFA ban",
            "Price increase via excise","Price increase; advertising ban"))
kbl(ph_def, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=c("Policy","Risk factor","Model effect","Implementation lever"), align="llll") |>
  kable_styling(latex_options="hold_position", font_size=8) |>
  column_spec(1, bold=TRUE, width="3.1cm") |> column_spec(2, width="2.4cm") |>
  column_spec(3, width="3.2cm") |> column_spec(4, width="5.0cm") |>
  row_spec(0, bold=TRUE)
```

\vspace{1pt}
\footnotesize Policies reduce exposure (smoking, salt, trans-fat, sugary-drink and alcohol intake), lowering CVD \textbf{incidence}. \textbf{Tobacco} policies additionally improve \textbf{survival} after CVD. Packages are modeled as single joint runs of their component measures. \scriptsize Source: workbook \texttt{Selected\_Interventions} / \texttt{Policy\_Levers}.

## Data sources

```{r tbl-sources}
src <- data.table(
  Component = c("Population \\& demography","Cause-specific mortality \\& disease rates",
                "Disability weights (healthy-years lost)","Clinical intervention effects",
                "Public-health effect sizes","Intervention \\& policy costs",
                "Economic valuation (VSL/VSLY)","International 40q30 comparison"),
  Source = c("UN World Population Prospects","GBD (Indonesia)","GBD",
             "FAIR Choices intervention taxonomy","Published risk-factor evidence (in workbook)",
             "Programme \\& policy unit costs (in workbook)","Reference Case guidance (Robinson 2019)",
             "WPP-scaled ASEAN comparison series"),
  Years = c("2024 revision","2023","2023","current","current",
            sprintf("price year %d", price_year), "2019","1995--2050"),
  Use = c("Population, background mortality, life expectancy","Baseline burden \\& calibration targets",
          "DALYs","Incidence \\& case-fatality effects","Incidence \\& survival effects",
          "Budget impact \\& cost-effectiveness","Monetized benefits \\& benefit--cost",
          "Benchmarking Indonesia vs.\\ neighbours"))
kbl(src, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=c("Model component","Primary source","Years / version","Use in model"), align="llll") |>
  kable_styling(latex_options=c("hold_position"), font_size=7.4) |>
  column_spec(1, bold=TRUE, width="3.4cm") |> column_spec(2, width="4.2cm") |>
  column_spec(3, width="2.0cm") |> column_spec(4, width="4.3cm") |>
  row_spec(0, bold=TRUE)
```

\vspace{1pt}
\scriptsize Sources verified in the modeling pipeline and input workbooks. Specific effect-size and cost citations are documented in the input workbooks. Population: UN World Population Prospects 2024; GBD [@GBDCVD2025; @Schumacher2024GBD2021]; economic valuation [@Robinson2019VSL].

# Health impact and value for money

```{r impact-table-builders, include=FALSE}
## Builders for the impact / CE / BCA display tables.
impact_row <- function(tab, sc) {
  data.table(
    Intervention = scenario_label(tab, sc),
    da  = fmt_thou(hv(sc,"cum_deaths_averted",tab)),
    q50 = fmt_pct(hv(sc,"q40q30_2050",tab),1),
    qr  = fmt_pct(hv(sc,"pct_red_q40q30_2050",tab),1),
    dr  = fmt_pct(hv(sc,"pct_red_deaths_2050",tab),1))
}
ce_row <- function(tab, sc) {
  cbind(impact_row(tab, sc),
    data.table(cpd = fmt_usd(hv(sc,"cost_per_death",tab)),
               pcap= fmt_percap(hv(sc,"ann_cost_per_capita",tab))))
}
bca_row <- function(tab, sc) {
  data.table(
    Intervention = scenario_label(tab, sc),
    da   = fmt_thou(hv(sc,"cum_deaths_averted",tab)),
    cpd  = fmt_usd(hv(sc,"cost_per_death",tab)),
    ben  = fmt_intl(hv(sc,"pv_benefits_vsl",tab)),
    net  = fmt_intl(hv(sc,"net_benefit_vsl",tab)),
    bcr  = fmt_ratio(hv(sc,"bcr_vsl",tab)))
}
```

## Public health: impact by policy

```{r tbl-ph-impact}
tb <- rbindlist(lapply(ph_order, function(s) impact_row(met_ph, s)))
kbl(tb, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=mkhdr(c("Public-health policy","Deaths averted\n(cumulative)","CVD 40q30\n2050",
                "40q30 reduction\n2050","Deaths reduction\n2050")), align="lrrrr") |>
  kable_styling(latex_options="hold_position", font_size=8) |>
  column_spec(1, bold=TRUE, width="4.6cm") |>
  row_spec(0, bold=TRUE) |>
  row_spec(nrow(tb), bold=TRUE, background="#E7F4F1")
```

\vspace{1pt}
\footnotesize Cumulative deaths averted over `r HZ_START`--`r HZ_END` vs.\ the status-quo baseline. Packages (tobacco, sodium) are single joint runs of their component measures; individual sodium/tobacco measures are listed in the appendix. \scriptsize Source: public-health cost/value workbook.

## Clinical: impact and cost-effectiveness

```{r tbl-clin-ce}
tb <- rbindlist(lapply(clin_order, function(s) ce_row(met_clin, s)))
kbl(tb, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=mkhdr(c("Clinical intervention","Deaths averted\n(cumulative)","CVD 40q30\n2050",
                "40q30 red.\n2050","Deaths red.\n2050","Cost per\ndeath averted","Annual cost\nper capita")),
    align="lrrrrrr") |>
  kable_styling(latex_options=c("hold_position","scale_down"), font_size=8) |>
  column_spec(1, bold=TRUE) |>
  row_spec(0, bold=TRUE) |>
  row_spec(nrow(tb), bold=TRUE, background="#EAF1FB")
```

\vspace{1pt}
\footnotesize Cost per death averted uses \textbf{discounted} incremental cost (`r fmt_pct(disc_rate*100,0)`/yr). Annual cost per capita = undiscounted incremental cost averaged over `r n_years` years and Indonesia's population. Costs in market US\$ (price year `r price_year`); perspective: health system. \scriptsize Source: clinical cost/value workbook.

## Clinical: benefit--cost analysis

```{r tbl-clin-bca}
tb <- rbindlist(lapply(clin_order, function(s) bca_row(met_clin, s)))
kbl(tb, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=mkhdr(c("Clinical intervention","Deaths averted","Cost per\ndeath averted",
                "Benefits (PV)","Net benefit (PV)","Benefit--cost\nratio")), align="lrrrrr") |>
  kable_styling(latex_options=c("hold_position","scale_down"), font_size=8) |>
  column_spec(1, bold=TRUE) |>
  row_spec(0, bold=TRUE) |>
  row_spec(nrow(tb), bold=TRUE, background="#EAF1FB")
```

\vspace{1pt}
\footnotesize Benefits = value of averted mortality (preferred VSL, income elasticity `r vsl_elast`), in PPP international dollars, base year `r base_year`, discounted at `r fmt_pct(disc_rate*100,0)`/yr; standing: `r bca_stand`; scope: partial (mortality benefit only). A ratio above 1 means benefits exceed costs. \scriptsize Source: clinical cost/value workbook, preferred-VSL case.

## Public health: cost-effectiveness

```{r tbl-ph-ce}
tb <- rbindlist(lapply(ph_order, function(s) ce_row(met_ph, s)))
kbl(tb, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=mkhdr(c("Public-health policy","Deaths averted\n(cumulative)","CVD 40q30\n2050",
                "40q30 red.\n2050","Deaths red.\n2050","Cost per\ndeath averted","Annual cost\nper capita")),
    align="lrrrrrr") |>
  kable_styling(latex_options=c("hold_position","scale_down"), font_size=8) |>
  column_spec(1, bold=TRUE) |>
  row_spec(0, bold=TRUE) |>
  row_spec(nrow(tb), bold=TRUE, background="#E7F4F1")
```

\vspace{1pt}
\footnotesize Fiscal and regulatory policies reach the whole population at very low per-person cost, so cost per death averted is far lower than for clinical care. Same costing conventions as the clinical table. \scriptsize Source: public-health cost/value workbook.

## Public health: benefit--cost analysis

```{r tbl-ph-bca}
tb <- rbindlist(lapply(ph_order, function(s) bca_row(met_ph, s)))
kbl(tb, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=mkhdr(c("Public-health policy","Deaths averted","Cost per\ndeath averted",
                "Benefits (PV)","Net benefit (PV)","Benefit--cost\nratio")), align="lrrrrr") |>
  kable_styling(latex_options=c("hold_position","scale_down"), font_size=8) |>
  column_spec(1, bold=TRUE) |>
  row_spec(0, bold=TRUE) |>
  row_spec(nrow(tb), bold=TRUE, background="#E7F4F1")
```

\vspace{1pt}
\footnotesize Because fiscal/regulatory policies cost little to implement relative to the mortality they avert, their benefit--cost ratios are very large. Same valuation conventions (preferred VSL, `r gsub("_", " ", cur_basis)`, base year `r base_year`, `r fmt_pct(disc_rate*100,0)`/yr) as the clinical BCA. \scriptsize Source: public-health cost/value workbook.

# The combined package

## Clinical + public health: better together

```{r tbl-combined}
pk <- c("all","all_public_health","all_clinical_public_health")
tb <- rbindlist(lapply(pk, function(s) {
  data.table(
    Package = exec_label[[s]],
    da   = fmt_thou(hv(s,"cum_deaths_averted")),
    qr   = fmt_pct(hv(s,"pct_red_q40q30_2050"),1),
    q50  = fmt_pct(hv(s,"q40q30_2050"),1),
    cost = fmt_usd_short(hv(s,"cum_cost_disc")),
    cpd  = fmt_usd(hv(s,"cost_per_death")),
    bcr  = fmt_ratio(hv(s,"bcr_vsl")))
}))
kbl(tb, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=mkhdr(c("Package","Deaths averted\n(cumulative)","40q30 red.\n2050","CVD 40q30\n2050",
                "Total cost\n(PV)","Cost per\ndeath averted","Benefit--cost\nratio")),
    align="lrrrrrr") |>
  kable_styling(latex_options=c("hold_position","scale_down"), font_size=8.5) |>
  column_spec(1, bold=TRUE) |>
  row_spec(0, bold=TRUE) |>
  row_spec(3, bold=TRUE, background="#FCEFCB")
```

\vspace{2pt}
\begin{combinedblock}{The combined package is modeled as one run --- not a sum}
\footnotesize Because clinical and public-health effects overlap, the joint scenario is projected in a single run.
It averts about \textbf{`r fmt_short(combo_da)` deaths}, brings the `r HZ_END` CVD 40q30 down to
\textbf{`r fmt_pct(combo_q50,1)`} (from `r fmt_pct(base_q_2050,1)` at baseline), and yields a benefit--cost ratio of about
\textbf{`r fmt_ratio(combo_bcr)`} --- a net benefit of roughly \textbf{`r fmt_intl(combo_net)`}.
\end{combinedblock}

## CVD premature mortality: trajectory and neighbours

```{r fig-40q30, fig.width=9.6, fig.height=4.7, out.width="\\linewidth"}
lab_scen <- idn_scen[year == HZ_END]
lab_cmp  <- cmp_lines[, .SD[which.max(year)], by = country_label]

ggplot() +
  ## comparator countries (thin muted background)
  geom_line(data = cmp_lines, aes(year, q, group = country_label),
            colour = col_comparator, linewidth = 0.5) +
  ggrepel::geom_text_repel(data = lab_cmp, aes(year, q, label = country_label),
            colour = "grey55", size = 2.5, hjust = 0, nudge_x = 1.2, direction = "y",
            segment.colour = "grey80", segment.size = 0.2, min.segment.length = 0,
            box.padding = 0.12, max.overlaps = Inf, seed = 42) +
  ## Indonesia observed history (solid dark)
  geom_line(data = idn_obs, aes(year, q), colour = "grey20", linewidth = 1.0) +
  ## boundary between observed and modeled
  annotate("segment", x = 2024, xend = 2024, y = 10.5, yend = 20.5,
           linetype = "dotted", colour = "grey45") +
  annotate("text", x = 2024, y = 20.9, label = "Modeled scenarios →",
           hjust = 1, size = 2.7, colour = "grey35") +
  ## Indonesia model scenarios (dashed, coloured)
  geom_line(data = idn_scen, aes(year, q, colour = scen), linewidth = 1.15, linetype = "22") +
  ggrepel::geom_text_repel(data = lab_scen, aes(year, q, label = scen, colour = scen),
            size = 2.7, fontface = "bold", hjust = 0, nudge_x = 1.2, direction = "y",
            segment.colour = "grey70", segment.size = 0.2, min.segment.length = 0,
            box.padding = 0.2, max.overlaps = Inf, show.legend = FALSE, seed = 7) +
  scale_colour_manual(values = scen_pal) +
  scale_x_continuous(limits = c(1995, 2064), breaks = seq(1995, 2050, 5),
                     expand = expansion(mult = c(0.01, 0))) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(title = "CVD 40q30 in Indonesia: observed, projected, and neighbours",
       subtitle = "Probability of dying from CVD between ages 30 and 70",
       x = NULL, y = "CVD 40q30",
       caption = paste("Indonesia observed history (solid, to 2023) and ASEAN comparators are WPP-scaled reference series.",
                       "\nModeled Indonesia scenarios (dashed, 2025-2050) use the GBD-calibrated model baseline, which sits somewhat higher.")) +
  theme_exec(11) +
  theme(legend.position = "none", panel.grid.major.x = element_line(colour = "grey92", linewidth = 0.25))
```

# What it means for Indonesia

## Policy implications

\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep4pt
  \item \textbf{Biggest health gains come from doing both.} The combined package averts far more (\textasciitilde`r fmt_short(combo_da)`) than either family alone.
  \item \textbf{Public-health policies are exceptional value.} Population-wide fiscal/regulatory measures avert deaths at a fraction of clinical cost (ratios in the thousands).
  \item \textbf{Clinical scale-up delivers large, durable gains} in survival and premature mortality, at conventional cost-effectiveness.
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep4pt
  \item \textbf{Sequence for early wins.} Tobacco taxation and other prevention policies can start cutting risk quickly and cheaply, while clinical capacity scales up.
  \item \textbf{The 40q30 target is within reach.} The combined package bends Indonesia's premature-CVD-mortality curve down by about `r fmt_pct(combo_qred,0)` by `r HZ_END`.
\end{itemize}
\end{column}
\end{columns}

\vspace{2pt}
\footnotesize These are model findings; translating them into policy requires judgement on feasibility, financing and delivery. Model results quantify the \emph{potential}; they are not a delivery guarantee.

## Limitations

\begin{itemize}\footnotesize\itemsep3pt
  \item Results depend on \textbf{modeled epidemiological estimates} (GBD) and calibration; real burden may differ.
  \item \textbf{Effect sizes, future coverage, and unit costs are uncertain}; interventions are assumed to reach and sustain their targets.
  \item The disease model is \textbf{simplified} (healthy $\to$ living-with-CVD $\to$ death; no remission; no comorbidity interactions).
  \item Economic returns value \textbf{averted mortality only} (a partial benefit-cost scope) --- morbidity, productivity and downstream cost offsets are omitted, so returns are, if anything, understated.
  \item Projections are \textbf{scenarios, not forecasts or guarantees}; they show what is achievable if interventions are delivered as modeled.
\end{itemize}

## Three things to remember

\begin{columns}[T]
\begin{column}{0.33\textwidth}
\begin{combinedblock}{1. Both, together}
\footnotesize Clinical + public health averts \textbf{\textasciitilde`r fmt_short(combo_da)` deaths} by `r HZ_END` --- more than either alone.
\end{combinedblock}
\end{column}
\begin{column}{0.33\textwidth}
\begin{phblock}{2. Prevention pays}
\footnotesize Population policies return on the order of \textbf{`r fmt_ratio(ph_bcr)`\texttimes} their cost.
\end{phblock}
\end{column}
\begin{column}{0.33\textwidth}
\begin{clinicalblock}{3. A reachable target}
\footnotesize Premature CVD mortality (40q30) falls \textbf{`r fmt_pct(combo_qred,0)`} by `r HZ_END` under the combined package.
\end{clinicalblock}
\end{column}
\end{columns}

\vspace{4pt}
\begin{center}
\large Combined action against CVD could avert about \hlkey{`r fmt_short(combo_da)` deaths} and return about
\hlkey{`r fmt_ratio(combo_bcr)` int\$ per int\$ invested} through `r HZ_END`.
\end{center}

# Technical appendix

## Appendix 1 --- Scenario crosswalk

```{r tbl-crosswalk}
xw_clin <- data.table(Family = "Clinical (FAIR Choices)", ID = clin_order,
                      Label = vapply(clin_order, function(s) scenario_label(met_clin, s), character(1)))
xw_ph <- data.table(Family = "Public health", ID = ph_order,
                    Label = vapply(ph_order, function(s) scenario_label(met_ph, s), character(1)))
xw_comb <- data.table(Family = "Combined", ID = "all_clinical_public_health",
                      Label = scenario_label(met_comb, "all_clinical_public_health"))
xw <- rbindlist(list(xw_clin, xw_ph, xw_comb))
xw[, ID := gsub("_", "\\\\_", ID, fixed = TRUE)]
xw[, Family := fifelse(duplicated(Family), "", Family)]
kbl(xw, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=c("Family","Workbook scenario id","Executive label"), align="lll") |>
  kable_styling(latex_options=c("hold_position","scale_down"), font_size=6) |>
  column_spec(1, bold=TRUE) |> column_spec(2, monospace=TRUE) |>
  row_spec(0, bold=TRUE)
```

\vspace{1pt}
\tiny Packages are single joint runs of their child measures (shown in parentheses); child measures are omitted from the main tables to avoid double-counting.

## Appendix 2 --- Definitions, horizons and assumptions

\footnotesize
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\textbf{Metrics}
\begin{itemize}\scriptsize\itemsep2pt
  \item \textbf{Deaths averted (cumulative)} $=\sum_{t=`r HZ_START`}^{`r HZ_END`}(\text{baseline deaths}_t-\text{scenario deaths}_t)$, across modeled CVD \& diabetes causes.
  \item \textbf{\% reduction in deaths, `r HZ_END`} $=100\times(\text{baseline}-\text{scenario})/\text{baseline}$ in `r HZ_END`.
  \item \textbf{CVD 40q30} = probability of dying of CVD between exact ages 30 and 70 (six CVD causes, both sexes, single-year life table).
  \item \textbf{\% reduction in 40q30, `r HZ_END`}: analogous to deaths.
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\textbf{Costing \& valuation}
\begin{itemize}\scriptsize\itemsep2pt
  \item Horizon `r HZ_START`--`r HZ_END`; discount rate `r fmt_pct(disc_rate*100,0)`/yr; base year `r base_year`; price year `r price_year`.
  \item Costs in market US\$; cost perspective: health system.
  \item Benefits: averted-mortality value, \textbf{preferred VSL} (income elasticity `r vsl_elast`), in `r gsub("_"," ",cur_basis)`; standing `r bca_stand`; scope: `r gsub("_"," ",bca_scope)`.
  \item \textbf{Cost per death averted} uses discounted incremental cost.
\end{itemize}
\end{column}
\end{columns}

\vspace{3pt}
\scriptsize \textbf{Model.} Discrete-time state-transition model; single-year ages 0--95; six CVD causes (IHD, ischemic stroke, intracerebral haemorrhage, hypertensive heart disease, rheumatic heart disease, cardiomyopathy \& myocarditis) plus type-2 diabetes; calibrated to GBD 2023 (Nelder--Mead); demography and life expectancy from UN WPP 2024. \textbf{Validation:} `r recon_note` `r if (econ_fallback_used) "Economic values were sourced from the combined workbook (live recalculation unavailable); health metrics are computed directly from stored quantities." else "All economic values were recalculated live from the formula workbooks."`

## Appendix 3 --- References

\scriptsize

::: {#refs}
:::
