---
title: "Reducing Cardiovascular Disease in Indonesia"
subtitle: "Health Impact, Costs, and Cost-Effectiveness of Clinical and Public-Health Interventions, 2025--2050"
author: "World Bank $\\cdot$ University of Washington"
institute: "Briefing for the Ministry of Health, Republic of Indonesia"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  beamer_presentation:
    theme: "default"
    keep_tex: false
    latex_engine: xelatex
    slide_level: 2
    toc: false
    includes:
      in_header: beamer_preamble.tex
classoption: "aspectratio=169"
bibliography: references.bib
link-citations: false
---

```{r setup, include=FALSE}
## =====================================================================
##  EXECUTIVE DECK - Indonesia CVD / FAIR Choices analysis
##  World Bank * University of Washington
##
##  DATA PROVENANCE (BCA-free; no Excel recalculation required)
##  Every displayed number is imported programmatically from the modeling
##  pipeline's current-run outputs:
##    * output/09_deck_results.rds - the BCA-FREE results contract written by
##      code/cvd-fair-choices/09_cost_value.R. It carries, per workbook family
##      and scenario, the SAME R-engine values the workbooks store: deaths
##      averted, incremental cost, cost per death averted, 2050 death levels and
##      % reduction, and CVD 40q30 (level, baseline, % reduction). Scenario
##      identity/labels/order follow the Scenario_Catalog contract, which honours
##      the binding Intervention_Cause_Map include_flag - excluded interventions
##      never appear.
##    * the combined workbook's literal CVD_40q30 per-year sheet (cvd_40q30_r),
##      used only for the trajectory figure and the baseline 40q30 scalars.
##    * data/raw/03_asean_cvd_40q30_comparison.csv (international comparator).
##  NOTHING is transcribed by hand. The deck reads only stored/literal values,
##  so it compiles WITHOUT Excel COM and on a clean non-Windows environment. It
##  displays NO benefit-cost / VSL / VSLY content; the value-for-money metric is
##  COST PER DEATH AVERTED (in market US$). The full benefit-cost analysis lives
##  in the pipeline workbooks and is intentionally not shown here.
## =====================================================================
library(here)
library(data.table)
library(openxlsx)
library(ggplot2)
library(ggrepel)
library(scales)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.align = "center", dev = "cairo_pdf"
)

## ---- Palette (matches beamer_preamble.tex scenario colours) ---------------
col_clinical  <- "#3B6FB6"  # clinical scenarios
col_ph        <- "#1B9E77"  # public-health scenarios
col_combined  <- "#08306B"  # combined package (darker)
col_baseline  <- "#6E6E6E"  # baseline / status quo
col_comparator<- "#BDBDBD"  # comparator countries (muted)
col_incidence <- "#2166AC"  # incidence (Well -> Sick) effects
col_casefatal <- "#B2182B"  # case-fatality (Sick -> Death) effects
col_bgmort    <- "#7F7F7F"  # background mortality
col_wbnavy    <- "#0F2A56"

theme_exec <- function(base = 11) {
  theme_minimal(base_size = base) +
    theme(
      plot.title       = element_text(face = "bold", size = base + 1, colour = col_wbnavy),
      plot.subtitle    = element_text(colour = "grey35", size = base - 2),
      plot.caption     = element_text(colour = "grey45", size = base - 4, hjust = 0),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      legend.position  = "bottom",
      legend.title     = element_blank(),
      axis.title       = element_text(size = base - 2)
    )
}

## ---- Number formatters (LaTeX-safe: escaped $ and % ; used with escape=FALSE)
fmt_int   <- function(x) formatC(round(x), format = "d", big.mark = ",")
fmt_thou  <- function(x) ifelse(is.na(x), "--", formatC(round(x / 1000) * 1000, format = "d", big.mark = ","))
fmt_short <- function(x, acc = 0.1) ifelse(is.na(x), "--", label_number(scale_cut = cut_short_scale(), accuracy = acc)(x))
fmt_pct   <- function(x, d = 1) ifelse(is.na(x), "--", paste0(formatC(x, format = "f", digits = d), "\\%"))
fmt_usd   <- function(x) ifelse(is.na(x), "--", paste0("US\\$", formatC(round(x), format = "d", big.mark = ",")))
fmt_usd_short <- function(x) ifelse(is.na(x), "--", paste0("US\\$", fmt_short(x)))
## Two-line table headers (converts embedded "\n" to \makecell line breaks).
mkhdr <- function(v) kableExtra::linebreak(v, align = "c")

## Null / NA coalescing helpers (base R has neither).
`%||%` <- function(a, b) if (is.null(a) || length(a) == 0L || (length(a) == 1L && is.na(a))) b else a
`%+na%` <- function(a, b) if (is.null(a) || length(a) == 0L || is.na(a[1])) b else a[1]
```

```{r load-data, include=FALSE}
## ---------------------------------------------------------------------------
## 1. Resolve current-run inputs (project-relative) and check existence.
##    The deck reads a BCA-free R-side results contract + literal workbook
##    values only -- no Excel COM, no formula recalculation.
## ---------------------------------------------------------------------------
HZ_START <- 2025L; HZ_END <- 2050L

## Output directory: defaults to the project's output/, but can be redirected
## (e.g. for testing an alternative active-scenario subset) via EXEC_DECK_OUTDIR.
.out_dir <- Sys.getenv("EXEC_DECK_OUTDIR", unset = "")
outp <- if (nzchar(.out_dir)) function(f) file.path(.out_dir, f) else function(f) here("output", f)
deck_file <- outp("09_deck_results.rds")
wb_comb   <- outp("indonesia_model_cost_value_clinical_public_health_formulae.xlsx")
wb_clin   <- outp("indonesia_model_cost_value_formulae.xlsx")
wb_ph     <- outp("indonesia_cost_value_public_health_formulae.xlsx")
csv_cmp   <- here("data", "raw", "03_asean_cvd_40q30_comparison.csv")

.req_files <- c(results_contract = deck_file, combined_workbook = wb_comb, comparator = csv_cmp)
.missing <- .req_files[!file.exists(.req_files)]
if (length(.missing))
  stop("VALIDATION STOP -- required input file(s) not found:\n  ",
       paste(sprintf("%s: %s", names(.missing), .missing), collapse = "\n  "),
       "\nRun the modeling pipeline (code/cvd-fair-choices/00_run_model_cvd_fair.R) first.",
       call. = FALSE)

rd        <- function(path, sheet) as.data.table(openxlsx::read.xlsx(path, sheet = sheet))
sheets_of <- function(path) openxlsx::getSheetNames(path)

## ---- Results contract -----------------------------------------------------
dk   <- readRDS(deck_file)
res  <- as.data.table(dk$results)
meta <- dk$meta

met_clin <- res[workbook == "clinical"]
met_ph   <- res[workbook == "public_health"]
met_comb <- res[workbook == "combined"]

## ---- Costing/report metadata (BCA-free: cost side only) -------------------
disc_rate  <- suppressWarnings(as.numeric(meta$cost_discount_rate)); if (!is.finite(disc_rate)) disc_rate <- 0.03
price_year <- suppressWarnings(as.numeric(meta$cost_price_year));    if (!is.finite(price_year)) price_year <- 2023
ce_unit    <- meta$cost_effectiveness_unit %||% "USD per death averted"
## LaTeX-safe escaper for dynamic strings inserted into prose with escape=FALSE.
ltx <- function(s) {
  s <- as.character(s)
  s <- gsub("\\\\", "\\\\textbackslash{}", s)
  s <- gsub("([&%$#_{}])", "\\\\\\1", s)
  s <- gsub("~", "\\\\textasciitilde{}", s)
  s
}
cost_perspective <- ltx(gsub("_", " ", meta$economic_perspective %||% "health system"))
```

```{r derive-scenarios, include=FALSE}
## ---------------------------------------------------------------------------
## 2. Scenario identity is DERIVED from the contract / Scenario_Catalog, never
##    from fixed ID vectors. Optional presentation-only label overrides may be
##    supplied; any active scenario without an override falls back safely to its
##    workbook label. No excluded / hard-coded ID is ever required to exist.
## ---------------------------------------------------------------------------
exec_label_override <- c(
  I_CVD_PRIMARY = "Primary CVD prevention", I_T2D_TREATMENT = "Type-2 diabetes treatment",
  I_IHD_SECONDARY = "Heart-attack secondary prevention", I_HF_CHRONIC = "Chronic heart-failure care",
  I_HF_ACUTE = "Acute heart-failure care", I_STROKE_SECONDARY = "Stroke secondary prevention",
  I_RHD_SECONDARY = "RHD secondary prevention",
  all = "All clinical (combined)",
  I_PH_TOBACCO_TAX = "Tobacco tax", I_PH_TOBACCO_POLICY = "Tobacco policy package",
  I_PH_SALT_POLICY = "Sodium policy package", I_PH_TFA_POLICY = "Trans-fat elimination",
  all_public_health = "All public health (combined)",
  all_clinical_public_health = "Clinical + public health (combined)",
  baseline = "Status quo (baseline)")

## Family aggregate / joint scenario ids (present only if built this run).
clin_all_id <- "all"; ph_all_id <- "all_public_health"; joint_id <- "all_clinical_public_health"
has_clin  <- nrow(met_clin[scenario != "baseline"]) > 0
has_ph    <- nrow(met_ph[scenario != "baseline"])   > 0
has_joint <- joint_id %in% met_comb$scenario

## Label getter: override first, else the contract's own scenario_label.
lbl <- function(sc, tab = met_comb) {
  if (sc %in% names(exec_label_override)) return(exec_label_override[[sc]])
  v <- tab[scenario == sc, scenario_label]
  if (length(v) && !is.na(v[1]) && nzchar(v[1])) v[1] else sc
}
## Metric getter.
hv <- function(sc, col, tab = met_comb) { v <- tab[scenario == sc][[col]]; if (length(v)) v[1] else NA_real_ }

## Top-level display rows for a family (exclude baseline and package CHILDREN to
## avoid double-counting; the family aggregate is placed last). Ordered by
## deaths averted. Robust to zero/one/many active interventions.
MAX_TBL_ROWS <- 14L
top_level_rows <- function(tab, agg_id) {
  if (!nrow(tab)) return(tab[0])
  d <- tab[scenario != "baseline" &
             (is.na(scenario_role) | !(scenario_role %in% c("child")))]
  if (!nrow(d)) return(d)
  d <- d[order(-deaths_averted)]
  agg <- d[scenario == agg_id]; rest <- d[scenario != agg_id]
  rbind(rest, agg)
}
```

```{r headline-scalars, include=FALSE}
## Baseline CVD 40q30 (per-year literal from the combined workbook).
c4c <- rd(wb_comb, "CVD_40q30")
qcol <- if ("cvd_40q30_r" %in% names(c4c)) "cvd_40q30_r" else "cvd_40q30"
base_q_2025 <- c4c[scenario == "baseline" & year == HZ_START][[qcol]][1]
base_q_2050 <- c4c[scenario == "baseline" & year == HZ_END][[qcol]][1]

## Combined-package (joint) headline metrics (guarded if joint absent).
combo_da   <- if (has_joint) hv(joint_id, "deaths_averted") else NA_real_
combo_cpda <- if (has_joint) hv(joint_id, "cost_per_death_averted") else NA_real_
combo_q50  <- if (has_joint) hv(joint_id, "cvd_40q30_2050") else NA_real_
combo_qred <- if (has_joint) hv(joint_id, "cvd_40q30_pct_reduction_2050") else NA_real_
combo_dred <- if (has_joint) hv(joint_id, "pct_reduction_deaths_2050") else NA_real_

clin_da   <- hv(clin_all_id, "deaths_averted", met_clin)
clin_qred <- hv(clin_all_id, "cvd_40q30_pct_reduction_2050", met_clin)
clin_cpda <- hv(clin_all_id, "cost_per_death_averted", met_clin)
ph_da     <- hv(ph_all_id, "deaths_averted", met_ph)
ph_qred   <- hv(ph_all_id, "cvd_40q30_pct_reduction_2050", met_ph)
ph_cpda   <- hv(ph_all_id, "cost_per_death_averted", met_ph)

## Most impactful single lever per family (excludes the aggregate).
.tc <- top_level_rows(met_clin, clin_all_id); .tc <- .tc[scenario != clin_all_id]
.tp <- top_level_rows(met_ph, ph_all_id);     .tp <- .tp[scenario != ph_all_id]
top_clin_sc <- if (nrow(.tc)) .tc$scenario[1] else NA_character_
top_ph_sc   <- if (nrow(.tp)) .tp$scenario[1] else NA_character_
```

```{r validation, include=FALSE}
## ---------------------------------------------------------------------------
## 3. VALIDATION GATES - health/cost-effectiveness contract only (NO BCA).
##    Consistency is checked against Scenario_Catalog + the results contract.
## ---------------------------------------------------------------------------
val_stop <- function(cond, msg) if (!isTRUE(cond)) stop("VALIDATION STOP - ", msg, call. = FALSE)

## (a) Scenario_Catalog present in every workbook that exists (health/CE contract;
##     Benefit_Cost / Economic_Value are NOT required by the deck).
cat_sets <- list()
.wb_paths <- c(clinical = wb_clin, public_health = wb_ph, combined = wb_comb)
for (fam in names(.wb_paths)) {
  path <- .wb_paths[[fam]]
  if (!file.exists(path)) next
  sh <- sheets_of(path)
  val_stop("Scenario_Catalog" %in% sh,
           sprintf("workbook '%s' is missing the Scenario_Catalog sheet", basename(path)))
  val_stop(all(c("CVD_40q30", "Cost_Effectiveness") %in% sh),
           sprintf("workbook '%s' is missing a required health/CE sheet", basename(path)))
  cat_sets[[fam]] <- rd(path, "Scenario_Catalog")$scenario
}

## (b) Contract scenario set reconciles EXACTLY with Scenario_Catalog per family
##     (no stale extras, nothing dropped). This is how "no excluded id sneaks in"
##     is enforced without hard-coding any excluded id.
recon_family <- function(cat_key, tab) {
  cs <- cat_sets[[cat_key]]
  if (is.null(cs)) return(invisible())
  ts <- unique(tab$scenario)
  val_stop(setequal(cs, ts),
           sprintf("%s: results contract scenarios do not match Scenario_Catalog. Only in catalog: {%s}; only in results: {%s}. Re-run Model 09 on the current run.",
                   cat_key, paste(setdiff(cs, ts), collapse = ", "),
                   paste(setdiff(ts, cs), collapse = ", ")))
}
recon_family("clinical", met_clin)
recon_family("public_health", met_ph)
recon_family("combined", met_comb)

## (c) baseline available exactly once per family + 2050 present.
for (tab in list(met_clin, met_ph, met_comb)) {
  if (!nrow(tab)) next
  val_stop(sum(tab$scenario == "baseline") <= 1, "baseline appears more than once in a family")
}
val_stop(any(c4c$scenario == "baseline") && HZ_END %in% c4c$year,
         "combined workbook CVD_40q30 has no baseline / no year 2050")

## (d) unique scenario ids within each family (no duplicates).
for (tab in list(met_clin, met_ph, met_comb))
  if (nrow(tab)) val_stop(!any(duplicated(tab$scenario)),
                          "duplicate scenario id within a family results block")

## (e) headline health/CE metrics finite where the scenario exists.
if (has_joint)
  val_stop(is.finite(combo_da) && is.finite(combo_q50) && is.finite(combo_cpda),
           "combined-package headline metrics are missing/non-numeric")
val_stop(is.finite(base_q_2050), "baseline CVD 40q30 in 2050 is missing")
```

```{r fig-40q30-data, include=FALSE}
## Data for the CVD 40q30 time-series & international-comparison figure.
cmp <- fread(csv_cmp)
cmp[, year := as.integer(year)]
comparators <- c("Malaysia","Philippines","Singapore","Thailand","Viet Nam","Indonesia")

cmp_lines <- cmp[country_label %in% comparators,
                 .(country_label, year, q = q40q30_percent)][order(country_label, year)]
cmp_lines <- unique(cmp_lines, by = c("country_label","year"))

#tempatch
cmp_lines[country_label == "Indonesia", country_label := "Indonesia."]

idn_obs <- cmp[country_label == "Indonesia" & series_type == "Historical WPP-scaled baseline" &
               year <= 2023, .(year, q = q40q30_percent)][order(year)]

## Indonesia MODEL scenarios (dashed): whichever of these are BOTH in the
## current-run contract and available as a per-year series in the workbook.
fig_scen_ids <- intersect(c("baseline", clin_all_id, ph_all_id, joint_id),
                          intersect(unique(c4c$scenario), unique(met_comb$scenario)))
idn_scen <- c4c[scenario %in% fig_scen_ids & year >= HZ_START,
                .(scenario, year, q = get(qcol))]
idn_scen[, scen := vapply(scenario, lbl, character(1), USE.NAMES = FALSE)]
scen_levels <- vapply(fig_scen_ids, lbl, character(1), USE.NAMES = FALSE)
idn_scen[, scen := factor(scen, levels = scen_levels)]
scen_pal <- setNames(
  c(col_baseline, col_clinical, col_ph, col_combined)[seq_along(fig_scen_ids)],
  scen_levels)
```

```{r results-builders, include=FALSE}
## Row builders for the impact / cost-effectiveness display tables (BCA-free).
impact_row <- function(tab, sc) data.table(
  Intervention = lbl(sc, tab),
  da  = fmt_thou(hv(sc, "deaths_averted", tab)),
  q50 = fmt_pct(hv(sc, "cvd_40q30_2050", tab), 1),
  qr  = fmt_pct(hv(sc, "cvd_40q30_pct_reduction_2050", tab), 1),
  dr  = fmt_pct(hv(sc, "pct_reduction_deaths_2050", tab), 1))
ce_row <- function(tab, sc) cbind(
  impact_row(tab, sc),
  data.table(cpd = fmt_usd(hv(sc, "cost_per_death_averted", tab))))

## Build a display table for a family's top-level rows, capped for space.
## Returns list(tb, agg_row_idx, n_omitted, omitted_labels).
family_table <- function(tab, agg_id, row_fun) {
  d <- top_level_rows(tab, agg_id)
  n_omitted <- 0L; omitted <- character(0)
  if (nrow(d) > MAX_TBL_ROWS) {
    keep_agg <- d[scenario == agg_id]
    body <- d[scenario != agg_id]
    n_keep <- MAX_TBL_ROWS - nrow(keep_agg)
    omitted <- body$scenario[-seq_len(min(n_keep, nrow(body)))]
    body <- body[seq_len(min(n_keep, nrow(body)))]
    d <- rbind(body, keep_agg)
    n_omitted <- length(omitted)
    message(sprintf("  Deck: family table capped at %d rows; %d intervention(s) omitted for space: %s",
                    MAX_TBL_ROWS, n_omitted, paste(omitted, collapse = ", ")))
  }
  if (!nrow(d)) return(list(tb = data.table(), agg_idx = NA_integer_, n_omitted = 0L, omitted = character(0)))
  tb <- rbindlist(lapply(d$scenario, function(s) row_fun(tab, s)))
  list(tb = tb, agg_idx = which(d$scenario == agg_id), n_omitted = n_omitted, omitted = omitted)
}
```

# Reducing CVD in Indonesia

## Why this matters for Indonesia

\begin{columns}[T]
\begin{column}{0.52\textwidth}
\begin{block}{Cardiovascular disease is Indonesia's largest health burden}
\normalsize
Under current trends, a person reaching age 30 faces roughly a
\hlkey{`r fmt_pct(base_q_2025,1)` chance} of dying from cardiovascular disease
before age 70 --- the CVD ``40q30''. Most of this is \textbf{preventable or treatable}.
\end{block}
\end{column}
\begin{column}{0.46\textwidth}
\begin{itemize}\itemsep4pt
  \item Premature CVD deaths (ages 30--69) have risen for two decades and remain high.
  \item Two complementary levers:
  \begin{itemize}
    \item \textcolor{clinicalcol}{\textbf{Clinical care}} --- prevent and better treat established disease.
    \item \textcolor{phcol}{\textbf{Population prevention}} --- tobacco, salt, trans-fat.
  \end{itemize}
  \item This analysis weighs their \textbf{health impact and cost-effectiveness} together.
\end{itemize}
\end{column}
\end{columns}

\vspace{3pt}
\footnotesize Source: Indonesia CVD model (calibrated to GBD 2023; UN World Population Prospects 2024). CVD 40q30 = probability of dying from cardiovascular disease between exact ages 30 and 70.

## What this analysis set out to do

\begin{block}{General aim}
\normalsize Estimate the \textbf{health impact and cost-effectiveness} of scaling up evidence-based
\textbf{clinical} and \textbf{public-health} cardiovascular interventions in Indonesia through \textbf{2050}.
\end{block}

\vspace{2pt}
\begin{columns}[T]
\begin{column}{0.33\textwidth}
\begin{clinicalblock}{Aim 1 --- Burden}
\footnotesize Project the CVD burden and the premature-mortality indicator (CVD 40q30) under the \textbf{status quo} and under \textbf{alternative intervention scenarios}.
\end{clinicalblock}
\end{column}
\begin{column}{0.33\textwidth}
\begin{phblock}{Aim 2 --- Value for money}
\footnotesize Estimate \textbf{deaths averted}, mortality reductions, \textbf{incremental costs}, and the \textbf{cost per death averted} for each intervention and package.
\end{phblock}
\end{column}
\begin{column}{0.33\textwidth}
\begin{combinedblock}{Aim 3 --- Better together}
\footnotesize Quantify the \textbf{added value of combining} clinical and public-health action, modeled as a single joint run rather than a sum.
\end{combinedblock}
\end{column}
\end{columns}

\vspace{3pt}
\footnotesize Horizon `r HZ_START`--`r HZ_END`; the "status quo" baseline is the no-new-intervention comparator against which every scenario is measured.

## Executive summary --- the headline numbers

\begin{columns}[T]
\begin{column}{0.5\textwidth}
```{r exec-block, results='asis'}
if (has_joint) {
  cat(sprintf(paste0(
    "\\begin{combinedblock}{Combining clinical + public health is the big prize}\n",
    "\\footnotesize Together they could avert about \\hlkey{%s deaths} over %d--%d, ",
    "cut the CVD 40q30 in %d by \\hlkey{%s}, at about \\hlkey{%s per death averted}.\n",
    "\\end{combinedblock}\n"),
    fmt_short(combo_da), HZ_START, HZ_END, HZ_END, fmt_pct(combo_qred,0), fmt_usd(combo_cpda)))
} else {
  cat(paste0("\\begin{block}{Two complementary levers}\n\\footnotesize This run reports the ",
             "clinical and public-health families; the joint package was not part of this ",
             "configuration.\n\\end{block}\n"))
}
```
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep3pt
```{r exec-bullets, results='asis'}
if (has_clin)
  cat(sprintf("  \\item \\textcolor{clinicalcol}{\\textbf{All clinical interventions}}: $\\approx$%s deaths averted; CVD 40q30 down %s by %d.\n",
              fmt_short(clin_da), fmt_pct(clin_qred,0), HZ_END))
if (has_ph)
  cat(sprintf("  \\item \\textcolor{phcol}{\\textbf{All public-health policies}}: $\\approx$%s deaths averted at about %s per death averted.\n",
              fmt_short(ph_da), fmt_usd(ph_cpda)))
if (!is.na(top_clin_sc) && !is.na(top_ph_sc))
  cat(sprintf("  \\item \\textbf{Most impactful single levers}: \\textcolor{clinicalcol}{%s} and \\textcolor{phcol}{%s}.\n",
              lbl(top_clin_sc, met_clin), lbl(top_ph_sc, met_ph)))
if (has_joint)
  cat(sprintf("  \\item \\textbf{Affordability}: the combined package costs about %s per death averted.\n",
              fmt_usd(combo_cpda)))
```
\end{itemize}
\end{column}
\end{columns}

\vspace{2pt}
\scriptsize All figures from the model's current-run cost/value outputs. Deaths averted are cumulative `r HZ_START`--`r HZ_END` across modeled cardiovascular and diabetes causes. Cost per death averted uses discounted incremental cost in market US\$ (price year `r price_year`).

## How the model works --- in plain language

\begin{columns}[T]
\begin{column}{0.54\textwidth}
\footnotesize
\begin{enumerate}\itemsep3pt
  \item \textbf{People \& demography.} Indonesia's population is projected by age and sex (UN World Population Prospects 2024).
  \item \textbf{Starting health.} Disease and death rates are set to Indonesia's data (GBD 2023) and calibrated to match observed levels.
  \item \textbf{Year-by-year simulation.} Each year people can develop cardiovascular disease or diabetes, survive, or die.
  \item \textbf{Interventions.} Clinical and public-health scenarios change how many people develop disease and how many survive it.
  \item \textbf{Outcomes.} The model reports deaths, healthy years (DALYs), premature-mortality (40q30), costs, and cost-effectiveness.
\end{enumerate}
\end{column}
\begin{column}{0.44\textwidth}
\begin{block}{One consistent engine}
\footnotesize Every scenario runs through the \textbf{same} projection, so results are directly comparable and the \textbf{combined} package is modeled as one run --- not a sum of separate estimates.
\end{block}
\vspace{2pt}
\scriptsize Calibration fits the model to GBD estimates using a standard numerical optimiser. Demography and life expectancy use UN WPP 2024.
\end{column}
\end{columns}

## How interventions change the disease pathway

\vspace{-2pt}
\begin{center}
\begin{tikzpicture}[
   >=Latex, node distance = 16mm and 30mm, font=\footnotesize,
   state/.style = {rectangle, rounded corners, draw=wbnavy, thick, fill=wbnavy!6,
                   minimum width=26mm, minimum height=11mm, align=center},
   death/.style = {rectangle, rounded corners, draw=black!55, thick, fill=black!4,
                   minimum width=26mm, minimum height=10mm, align=center}]
  \node[state] (well) {\textbf{Healthy}\\[-1pt]\scriptsize (no CVD)};
  \node[state, right=of well] (sick) {\textbf{Living with CVD}\\[-1pt]\scriptsize (or diabetes)};
  \node[death, right=of sick] (cvd)  {\textbf{CVD death}};
  \node[death, below=14mm of well, xshift=23mm] (bg) {Other-cause (background) death};

  \draw[->, very thick, incidencecol]
       (well) -- node[above, midway, text=incidencecol]{\textbf{new CVD}}
                  node[below, midway, text=incidencecol, font=\scriptsize]{incidence} (sick);
  \draw[->, very thick, casefatalcol]
       (sick) -- node[above, midway, text=casefatalcol]{\textbf{dies of CVD}}
                  node[below, midway, text=casefatalcol, font=\scriptsize]{case-fatality} (cvd);
  \draw[->, thick, bgmortcol] (well) to[bend right=12] (bg);
  \draw[->, thick, bgmortcol] (sick) to[bend left=12]  (bg);
\end{tikzpicture}
\end{center}

\vspace{-2pt}
\begin{columns}[T]
\begin{column}{0.33\textwidth}\footnotesize
\textcolor{incidencecol}{\textbf{Fewer people fall ill}}\\ \scriptsize Primary prevention \& \emph{all} public-health policies lower \textbf{incidence}.
\end{column}
\begin{column}{0.34\textwidth}\footnotesize
\textcolor{casefatalcol}{\textbf{More people survive}}\\ \scriptsize Secondary prevention, treatment \& tobacco policy lower \textbf{case-fatality}.
\end{column}
\begin{column}{0.31\textwidth}\footnotesize
\textcolor{bgmortcol}{\textbf{Background mortality}}\\ \scriptsize Other-cause deaths from demography; unchanged by these interventions.
\end{column}
\end{columns}

\vspace{3pt}
\scriptsize No recovery/remission transition is modeled: people living with CVD remain at elevated risk (a conservative assumption). Source: model engine, \texttt{06\_run\_scenarios\_indonesia\_fair.R}.

# The interventions

## Clinical interventions (FAIR Choices)

```{r tbl-clin-interventions, results='asis'}
## Built from the ACTIVE clinical scenarios in this run (children/aggregate
## excluded); target-population / effect notes come from a small presentation map.
clin_pop <- c(
  I_CVD_PRIMARY = "Adults at cardiovascular risk", I_T2D_TREATMENT = "People with type-2 diabetes",
  I_IHD_SECONDARY = "Heart-attack (IHD) survivors", I_STROKE_SECONDARY = "Stroke survivors",
  I_HF_CHRONIC = "Chronic heart-failure patients", I_HF_ACUTE = "Acute heart-failure patients",
  I_RHD_SECONDARY = "Rheumatic heart disease patients")
clin_acts <- c(I_CVD_PRIMARY = "\\textcolor{incidencecol}{Incidence}")
clin_disp <- top_level_rows(met_clin, clin_all_id)[scenario != clin_all_id]
if (nrow(clin_disp)) {
  pluck <- function(vec, key, default) if (key %in% names(vec)) vec[[key]] else default
  cd <- data.table(
    Intervention = vapply(clin_disp$scenario, function(s) lbl(s, met_clin), character(1)),
    Population   = vapply(clin_disp$scenario, function(s) pluck(clin_pop, s, "Target patient group"), character(1)),
    Acts         = vapply(clin_disp$scenario, function(s)
                     pluck(clin_acts, s, "\\textcolor{casefatalcol}{Case-fatality}"), character(1)))
  print(kbl(cd, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=c("Intervention","Target population","Model effect"), align="lll") |>
    kable_styling(latex_options="hold_position", font_size=8.5) |>
    column_spec(1, bold=TRUE, width="4.6cm") |>
    column_spec(2, width="6.4cm") |> column_spec(3, width="3.0cm") |>
    row_spec(0, bold=TRUE))
} else cat("\\begin{center}\\large No clinical interventions selected for this run.\\end{center}")
```

\vspace{2pt}
\footnotesize \textbf{Implementation scenario (all clinical interventions):} coverage scales up \textbf{from a low base toward 80\%} of the eligible population between `r HZ_START` and `r HZ_END`. Effect sizes are drawn from the FAIR Choices intervention taxonomy; primary prevention lowers who develops CVD, while secondary prevention and treatment improve survival. \scriptsize Source: workbook \texttt{Selected\_Interventions}.

## Public-health interventions (fiscal \& regulatory policy)

```{r tbl-ph-interventions, results='asis'}
ph_meta <- data.table(
  scenario = c("I_PH_TOBACCO_TAX","I_PH_TOBACCO_POLICY","I_PH_SALT_POLICY","I_PH_TFA_POLICY",
               "I_PH_SALT_REFORM","I_PH_SALT_FOPL","I_PH_SALT_MEDIA","I_PH_SALT_ENV",
               "I_PH_TOB_CLEAN_AIR","I_PH_TOB_MEDIA","I_PH_TOB_AD_BAN"),
  Risk  = c("Tobacco","Tobacco","Dietary salt/sodium","Industrial trans-fat",
            "Dietary salt/sodium","Dietary salt/sodium","Dietary salt/sodium","Dietary salt/sodium",
            "Tobacco","Tobacco","Tobacco"),
  Acts  = c("\\textcolor{incidencecol}{Incidence} + \\textcolor{casefatalcol}{survival}",
            "\\textcolor{incidencecol}{Incidence} + \\textcolor{casefatalcol}{survival}",
            "\\textcolor{incidencecol}{Incidence}","\\textcolor{incidencecol}{Incidence}",
            "\\textcolor{incidencecol}{Incidence}","\\textcolor{incidencecol}{Incidence}",
            "\\textcolor{incidencecol}{Incidence}","\\textcolor{incidencecol}{Incidence}",
            "\\textcolor{incidencecol}{Incidence} + \\textcolor{casefatalcol}{survival}",
            "\\textcolor{incidencecol}{Incidence} + \\textcolor{casefatalcol}{survival}",
            "\\textcolor{incidencecol}{Incidence} + \\textcolor{casefatalcol}{survival}"),
  Lever = c("Raise tax to 75\\% of retail price","Clean-air law, mass media, ad ban",
            "Reformulation, labelling, media, procurement","Best-practice industrial TFA ban",
            "Mandatory reformulation","Front-of-pack labelling","Mass-media campaign","Public-institution environment",
            "Smoke-free / clean-air law","Mass-media campaign","Advertising ban"))
ph_disp <- top_level_rows(met_ph, ph_all_id)[scenario != ph_all_id]
if (nrow(ph_disp)) {
  pd <- merge(ph_disp[, .(scenario)], ph_meta, by = "scenario", all.x = TRUE, sort = FALSE)
  pd[, Policy := vapply(scenario, function(s) lbl(s, met_ph), character(1))]
  pd[is.na(Risk),  Risk  := "Behavioural risk factor"]
  pd[is.na(Acts),  Acts  := "\\textcolor{incidencecol}{Incidence}"]
  pd[is.na(Lever), Lever := "Fiscal / regulatory measure"]
  print(kbl(pd[, .(Policy, Risk, Acts, Lever)], format="latex", booktabs=TRUE, escape=FALSE,
      col.names=c("Policy","Risk factor","Model effect","Implementation lever"), align="llll") |>
    kable_styling(latex_options="hold_position", font_size=8) |>
    column_spec(1, bold=TRUE, width="3.1cm") |> column_spec(2, width="2.4cm") |>
    column_spec(3, width="3.2cm") |> column_spec(4, width="5.0cm") |>
    row_spec(0, bold=TRUE))
} else cat("\\begin{center}\\large No public-health policies selected for this run.\\end{center}")
```

\vspace{1pt}
\footnotesize Policies reduce exposure (smoking, salt, trans-fat intake), lowering CVD \textbf{incidence}. \textbf{Tobacco} policies additionally improve \textbf{survival} after CVD. Packages are modeled as single joint runs of their component measures. \scriptsize Source: workbook \texttt{Selected\_Interventions} / \texttt{Policy\_Levers}.

## Data sources

```{r tbl-sources}
src <- data.table(
  Component = c("Population \\& demography","Cause-specific mortality \\& disease rates",
                "Disability weights (healthy-years lost)","Clinical intervention effects",
                "Public-health effect sizes","Intervention \\& policy costs",
                "International 40q30 comparison"),
  Source = c("UN World Population Prospects","GBD (Indonesia)","GBD",
             "FAIR Choices intervention taxonomy","Published risk-factor evidence (in workbook)",
             "Programme \\& policy unit costs (in workbook)",
             "WPP-scaled ASEAN comparison series"),
  Years = c("2024 revision","2023","2023","current","current",
            sprintf("price year %d", price_year), "1995--2050"),
  Use = c("Population, background mortality, life expectancy","Baseline burden \\& calibration targets",
          "DALYs","Incidence \\& case-fatality effects","Incidence \\& survival effects",
          "Budget impact \\& cost-effectiveness",
          "Benchmarking Indonesia vs.\\ neighbours"))
kbl(src, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=c("Model component","Primary source","Years / version","Use in model"), align="llll") |>
  kable_styling(latex_options=c("hold_position"), font_size=7.6) |>
  column_spec(1, bold=TRUE, width="3.4cm") |> column_spec(2, width="4.2cm") |>
  column_spec(3, width="2.0cm") |> column_spec(4, width="4.3cm") |>
  row_spec(0, bold=TRUE)
```

\vspace{1pt}
\scriptsize Sources verified in the modeling pipeline and input workbooks. Specific effect-size and cost citations are documented in the input workbooks. Population: UN World Population Prospects 2024; GBD [@GBDCVD2025; @Schumacher2024GBD2021].

# Health impact and value for money

## Clinical: impact and cost-effectiveness

```{r tbl-clin-ce, results='asis'}
ft <- family_table(met_clin, clin_all_id, ce_row)
if (nrow(ft$tb)) {
  k <- kbl(ft$tb, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=mkhdr(c("Clinical intervention","Deaths averted\n(cumulative)","CVD 40q30\n2050",
                  "40q30 red.\n2050","Deaths red.\n2050","Cost per\ndeath averted")),
      align="lrrrrr") |>
    kable_styling(latex_options=c("hold_position","scale_down"), font_size=8.5) |>
    column_spec(1, bold=TRUE) |> row_spec(0, bold=TRUE)
  if (!is.na(ft$agg_idx)) k <- k |> row_spec(ft$agg_idx, bold=TRUE, background="#EAF1FB")
  print(k)
  if (ft$n_omitted > 0)
    cat(sprintf("\n\n\\vspace{1pt}\\scriptsize %d additional clinical intervention(s) omitted for space; see the appendix crosswalk and the workbook.\n", ft$n_omitted))
} else cat("\\begin{center}\\large No clinical interventions selected for this run.\\end{center}")
```

<!-- \vspace{1pt} -->
<!-- \footnotesize Cost per death averted uses \textbf{discounted} incremental cost (`r fmt_pct(disc_rate*100,0)`/yr). Costs in market US\$ (price year `r price_year`); perspective: `r cost_perspective`. \scriptsize Source: clinical cost/value workbook. -->

\vspace{1pt}
\footnotesize Cost per death averted is based on \textbf{discounted} incremental costs (`r fmt_pct(disc_rate*100,0)` per year) and assumes a linear scale-up from baseline coverage in `r HZ_START` to 80\% coverage by `r HZ_END`. Costs are reported in market US\$ (price year `r price_year`) from the `r cost_perspective` perspective.

## Public health: impact and cost-effectiveness

```{r tbl-ph-ce, results='asis'}
ft <- family_table(met_ph, ph_all_id, ce_row)
if (nrow(ft$tb)) {
  k <- kbl(ft$tb, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=mkhdr(c("Public-health policy","Deaths averted\n(cumulative)","CVD 40q30\n2050",
                  "40q30 red.\n2050","Deaths red.\n2050","Cost per\ndeath averted")),
      align="lrrrrr") |>
    kable_styling(latex_options=c("hold_position","scale_down"), font_size=8.5) |>
    column_spec(1, bold=TRUE) |> row_spec(0, bold=TRUE)
  if (!is.na(ft$agg_idx)) k <- k |> row_spec(ft$agg_idx, bold=TRUE, background="#E7F4F1")
  print(k)
  if (ft$n_omitted > 0)
    cat(sprintf("\n\n\\vspace{1pt}\\scriptsize %d additional public-health intervention(s) omitted for space; see the appendix crosswalk and the workbook.\n", ft$n_omitted))
} else cat("\\begin{center}\\large No public-health policies selected for this run.\\end{center}")
```

\vspace{1pt}
\footnotesize Cumulative deaths averted during `r HZ_START`--`r HZ_END`, relative to the status-quo baseline. The tobacco package combines smoke-free policies, health warnings and mass-media campaigns, advertising bans, and cessation support. The sodium package combines food reformulation, front-of-pack nutrition labelling, healthy public food-procurement policies, and behavior-change communication. Each package is modeled as a single joint scenario.

<!-- ; detailed component definitions are provided in the appendix. \scriptsize Source: public-health cost/value workbook. -->


# The combined package

## Clinical + public health: better together

```{r tbl-combined, results='asis'}
pk <- intersect(c(clin_all_id, ph_all_id, joint_id), met_comb$scenario)
if (length(pk)) {
  tb <- rbindlist(lapply(pk, function(s) data.table(
    Package = lbl(s),
    da   = fmt_thou(hv(s, "deaths_averted")),
    qr   = fmt_pct(hv(s, "cvd_40q30_pct_reduction_2050"), 1),
    q50  = fmt_pct(hv(s, "cvd_40q30_2050"), 1),
    cost = fmt_usd_short(hv(s, "disc_incremental_cost")),
    cpd  = fmt_usd(hv(s, "cost_per_death_averted")))))
  k <- kbl(tb, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=mkhdr(c("Package","Deaths averted\n(cumulative)","40q30 red.\n2050","CVD 40q30\n2050",
                  "Total cost\n(PV, discounted)","Cost per\ndeath averted")),
      align="lrrrrr") |>
    kable_styling(latex_options=c("hold_position","scale_down"), font_size=9) |>
    column_spec(1, bold=TRUE) |> row_spec(0, bold=TRUE)
  jidx <- which(pk == joint_id)
  if (length(jidx)) k <- k |> row_spec(jidx, bold=TRUE, background="#FCEFCB")
  print(k)
} else cat("\\begin{center}\\large No package scenarios available for this run.\\end{center}")
```

\vspace{2pt}
```{r combined-block, results='asis'}
if (has_joint) {
  cat(sprintf(paste0(
    "\\begin{combinedblock}{The combined package is modeled as one run --- not a sum}\n",
    "\\footnotesize Because clinical and public-health effects overlap, the joint scenario is ",
    "projected in a single run. It averts about \\textbf{%s deaths}, brings the %d CVD 40q30 down to ",
    "\\textbf{%s} (from %s at baseline), at about \\textbf{%s per death averted}.\n",
    "\\end{combinedblock}\n"),
    fmt_short(combo_da), HZ_END, fmt_pct(combo_q50,1), fmt_pct(base_q_2050,1), fmt_usd(combo_cpda)))
} else {
  cat(paste0("\\begin{block}{Single-family run}\n\\footnotesize The joint clinical + ",
             "public-health scenario was not built in this configuration; family results are ",
             "shown above.\n\\end{block}\n"))
}
```

## CVD premature mortality: trajectory and neighbours

```{r fig-40q30, fig.width=9.6, fig.height=4.7, out.width="\\linewidth"}
# Remove the modeled status-quo baseline
idn_scen_plot <- idn_scen[scenario != "baseline"]
lab_scen <- idn_scen_plot[year == HZ_END]

lab_cmp <- cmp_lines[, .SD[which.max(year)], by = country_label]

ggplot() +
  geom_line(
    data = cmp_lines,
    aes(year, q, group = country_label),
    colour = col_comparator,
    linewidth = 0.5
  ) +
  ggrepel::geom_text_repel(
    data = lab_cmp,
    aes(year, q, label = country_label),
    colour = "grey55", size = 2.5, hjust = 0,
    nudge_x = 1.2, direction = "y",
    segment.colour = "grey80", segment.size = 0.2,
    min.segment.length = 0, box.padding = 0.12,
    max.overlaps = Inf, seed = 42
  ) +
  geom_line(
    data = idn_obs,
    aes(year, q),
    colour = "grey20",
    linewidth = 1.0
  ) +
  annotate(
    "segment",
    x = 2024, xend = 2024,
    y = 10.5, yend = 20.5,
    linetype = "dotted",
    colour = "grey45"
  ) +
  annotate(
    "text",
    x = 2024, y = 20.9,
    label = "Modeled scenarios →",
    hjust = 1, size = 2.7,
    colour = "grey35"
  ) +
  geom_line(
    data = idn_scen_plot,  # filtered data
    aes(year, q, colour = scen),
    linewidth = 1.15,
    linetype = "22"
  ) +
  ggrepel::geom_text_repel(
    data = lab_scen,
    aes(year, q, label = scen, colour = scen),
    size = 2.7, fontface = "bold", hjust = 0,
    nudge_x = 1.2, direction = "y",
    segment.colour = "grey70", segment.size = 0.2,
    min.segment.length = 0, box.padding = 0.2,
    max.overlaps = Inf, show.legend = FALSE, seed = 7
  ) +
  scale_colour_manual(values = scen_pal) +
  scale_x_continuous(
    limits = c(1995, 2064),
    breaks = seq(1995, 2050, 5),
    expand = expansion(mult = c(0.01, 0))
  ) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "CVD 40q30 in Indonesia: observed, projected, and neighbours",
    subtitle = "Probability of dying from CVD between ages 30 and 70",
    x = NULL,
    y = "CVD 40q30",
    caption = paste(
      "Indonesia observed history (solid, to 2023) and ASEAN comparators are WPP-scaled reference series.",
      "\nModeled intervention scenarios are dashed for 2025–2050.",
      "\nASEAN comparator trends use linear extrapolation for illustrative reference."
    )
  ) +
  theme_exec(11) +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_line(
      colour = "grey92",
      linewidth = 0.25
    )
  )
```

# What it means for Indonesia

## Policy implications

\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep4pt
```{r impl-left, results='asis'}
if (has_joint)
  cat(sprintf("  \\item \\textbf{Biggest health gains come from doing both.} The combined package averts far more (\\textasciitilde%s) than either family alone.\n", fmt_short(combo_da)))
cat("  \\item \\textbf{Public-health policies are exceptional value.} Population-wide fiscal/regulatory measures avert deaths at a fraction of clinical cost per death averted.\n")
cat("  \\item \\textbf{Clinical scale-up delivers large, durable gains} in survival and premature mortality.\n")
```
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep4pt
  \item \textbf{Sequence for early wins.} Tobacco taxation and other prevention policies can start cutting risk quickly and cheaply, while clinical capacity scales up.
```{r impl-right, results='asis'}
if (has_joint)
  cat(sprintf("  \\item \\textbf{The 40q30 target is within reach.} The combined package bends Indonesia's premature-CVD-mortality curve down by about %s by %d.\n", fmt_pct(combo_qred,0), HZ_END))
```
\end{itemize}
\end{column}
\end{columns}

\vspace{2pt}
\footnotesize These are model findings; translating them into policy requires judgement on feasibility, financing and delivery. Model results quantify the \emph{potential}; they are not a delivery guarantee.

## Limitations

\begin{itemize}\footnotesize\itemsep3pt
  \item Results depend on \textbf{modeled epidemiological estimates} (GBD) and calibration; real burden may differ.
  \item \textbf{Effect sizes, future coverage, and unit costs are uncertain}; interventions are assumed to reach and sustain their targets.
  \item The disease model is \textbf{simplified} (healthy $\to$ living-with-CVD $\to$ death; no remission; no comorbidity interactions).
  \item Cost-effectiveness is expressed as \textbf{cost per death averted}; morbidity and productivity effects are not monetised here.
  \item Projections are \textbf{scenarios, not forecasts or guarantees}; they show what is achievable if interventions are delivered as modeled.
\end{itemize}

## Three things to remember

\begin{columns}[T]
\begin{column}{0.33\textwidth}
```{r remember-1, results='asis'}
if (has_joint) {
  cat(sprintf("\\begin{combinedblock}{1. Both, together}\n\\footnotesize Clinical + public health averts \\textbf{\\textasciitilde%s deaths} by %d --- more than either alone.\n\\end{combinedblock}\n", fmt_short(combo_da), HZ_END))
} else {
  cat("\\begin{combinedblock}{1. Two levers}\n\\footnotesize Clinical care and population prevention each cut premature CVD mortality.\n\\end{combinedblock}\n")
}
```
\end{column}
\begin{column}{0.33\textwidth}
```{r remember-2, results='asis'}
cat(sprintf("\\begin{phblock}{2. Prevention is affordable}\n\\footnotesize Population policies avert deaths at about \\textbf{%s} each.\n\\end{phblock}\n", fmt_usd(ph_cpda %+na% NA_real_)))
```
\end{column}
\begin{column}{0.33\textwidth}
```{r remember-3, results='asis'}
if (has_joint) {
  cat(sprintf("\\begin{clinicalblock}{3. A reachable target}\n\\footnotesize Premature CVD mortality (40q30) falls \\textbf{%s} by %d under the combined package.\n\\end{clinicalblock}\n", fmt_pct(combo_qred,0), HZ_END))
} else {
  cat(sprintf("\\begin{clinicalblock}{3. A reachable target}\n\\footnotesize All clinical interventions cut premature CVD mortality (40q30) by \\textbf{%s} by %d.\n\\end{clinicalblock}\n", fmt_pct(clin_qred,0), HZ_END))
}
```
\end{column}
\end{columns}

\vspace{4pt}
```{r closing-banner, results='asis'}
if (has_joint) {
  cat(sprintf(paste0("\\begin{center}\n\\large Combined action against CVD could avert about \\hlkey{%s deaths} ",
                     "at about \\hlkey{%s per death averted} through %d.\n\\end{center}\n"),
              fmt_short(combo_da), fmt_usd(combo_cpda), HZ_END))
} else {
  cat(sprintf(paste0("\\begin{center}\n\\large Scaling up CVD interventions could avert hundreds of thousands ",
                     "of deaths through %d at a modest cost per death averted.\n\\end{center}\n"), HZ_END))
}
```

# Technical appendix

## Appendix 1 --- Scenario crosswalk

```{r tbl-crosswalk}
## Built dynamically from the current run's combined-workbook scenarios.
fam_order <- c(clinical = 1, public_health = 2, clinical_public_health = 3)
xwd <- copy(met_comb)[, .(scenario, intervention_family, scenario_label, n_interventions)]
xwd[, ford := fam_order[intervention_family]]
xwd[is.na(ford), ford := 9]
setorder(xwd, ford, -n_interventions)
fam_disp <- c(clinical = "Clinical (FAIR Choices)", public_health = "Public health",
              clinical_public_health = "Combined")
xwd[, Family := fam_disp[intervention_family]]
xwd[is.na(Family), Family := intervention_family]
## Only show the family label on its first row (visual grouping).
xwd[, Family := ifelse(duplicated(Family), "", Family)]
xwd[, IDesc := gsub("_", "\\\\_", scenario)]
xw <- xwd[, .(Family, ID = IDesc, Label = scenario_label)]
kbl(xw, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=c("Family","Workbook scenario id","Executive label"), align="lll") |>
  kable_styling(latex_options=c("hold_position","scale_down"), font_size=6) |>
  column_spec(1, bold=TRUE) |> column_spec(2, monospace=TRUE) |>
  row_spec(0, bold=TRUE)
```

\vspace{1pt}
\tiny Packages are single joint runs of their child measures; child measures are omitted from the main tables to avoid double-counting. This crosswalk lists exactly the scenarios produced in the current run.

## Appendix 2 --- Definitions, horizons and assumptions

\footnotesize
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\textbf{Metrics}
\begin{itemize}\scriptsize\itemsep2pt
  \item \textbf{Deaths averted (cumulative)} $=\sum_{t=`r HZ_START`}^{`r HZ_END`}(\text{baseline deaths}_t-\text{scenario deaths}_t)$, across modeled CVD \& diabetes causes.
  \item \textbf{\% reduction in deaths, `r HZ_END`} $=100\times(\text{baseline}-\text{scenario})/\text{baseline}$ in `r HZ_END`.
  \item \textbf{CVD 40q30} = probability of dying of CVD between exact ages 30 and 70 (six CVD causes, both sexes, single-year life table).
  \item \textbf{\% reduction in 40q30, `r HZ_END`}: analogous to deaths.
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\textbf{Costing \& cost-effectiveness}
\begin{itemize}\scriptsize\itemsep2pt
  \item Horizon `r HZ_START`--`r HZ_END`; cost discount rate `r fmt_pct(disc_rate*100,0)`/yr; price year `r price_year`.
  \item Costs in \textbf{market US\$}; cost perspective: `r cost_perspective`.
  \item \textbf{Cost per death averted} $=$ cumulative \textbf{discounted} incremental cost $\div$ cumulative \textbf{undiscounted} deaths averted.
  \item A ``Dominant'' scenario averts deaths while lowering cost; ``not defined'' means no deaths averted.
\end{itemize}
\end{column}
\end{columns}

\vspace{3pt}
\scriptsize \textbf{Model.} Discrete-time state-transition model; single-year ages 0--95; six CVD causes (IHD, ischemic stroke, intracerebral haemorrhage, hypertensive heart disease, rheumatic heart disease, cardiomyopathy \& myocarditis) plus type-2 diabetes; calibrated to GBD 2023; demography and life expectancy from UN WPP 2024. \textbf{Data provenance:} every number is read from the current-run results contract (\texttt{output/09\_deck\_results.rds}) and the workbooks' literal CVD 40q30 series; scenario identity follows the \texttt{Scenario\_Catalog} contract (binding \texttt{include\_flag}). No Excel recalculation is used.

## Appendix 3 --- References

\scriptsize

::: {#refs}
:::
