---
title: "Reducing Cardiovascular Disease in Indonesia"
subtitle: "Health Impact, Costs, and Cost-Effectiveness of Clinical and Public-Health Interventions, 2025--2050"
author: "Disease Control Priorities"
institute: "Briefing for the Ministry of Health, Republic of Indonesia"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  beamer_presentation:
    theme: "default"
    keep_tex: false
    latex_engine: xelatex
    slide_level: 2
    toc: false
    includes:
      in_header: beamer_preamble.tex
classoption: "aspectratio=169"
bibliography: references.bib
link-citations: false
header-includes:
  - '\definecolor{srcgrey}{HTML}{5B6B7B}'
  - '\newcommand{\srcmethod}[1]{{\par\vspace{1pt}\scriptsize\itshape\color{srcgrey}#1\par}}'
  - '\newcommand{\costcaveat}{\textbf{Preliminary costing results:} unit costs, service use, and delivery assumptions require further validation against BPJS and Ministry of Health data before policy use.}'
---

```{r setup, include=FALSE}
## =====================================================================
##  EXECUTIVE DECK - Indonesia CVD / FAIR Choices analysis
##  World Bank * University of Washington
##
##  DATA PROVENANCE (BCA-free; no Excel recalculation required)
##  Every displayed number is imported programmatically from the modeling
##  pipeline's current-run outputs:
##    * output/09_deck_results.rds - the BCA-FREE results contract written by
##      code/cvd-fair-choices/09_cost_value.R. It carries, per workbook family
##      and scenario, the SAME R-engine values the workbooks store: deaths
##      averted, incremental cost, cost per death averted, 2050 death levels and
##      % reduction, and CVD 40q30 (level, baseline, % reduction). Scenario
##      identity/labels/order follow the Scenario_Catalog contract, which honours
##      the binding Intervention_Cause_Map include_flag - excluded interventions
##      never appear.
##    * the combined workbook's literal CVD_40q30 per-year sheet (cvd_40q30_r),
##      used only for the trajectory figure and the baseline 40q30 scalars.
##    * data/raw/03_asean_cvd_40q30_comparison.csv (international comparator).
##  NOTHING is transcribed by hand. The deck reads only stored/literal values,
##  so it compiles WITHOUT Excel COM and on a clean non-Windows environment. It
##  displays NO benefit-cost / VSL / VSLY content; the value-for-money metric is
##  COST PER DEATH AVERTED (in market US$). The full benefit-cost analysis lives
##  in the pipeline workbooks and is intentionally not shown here.
##
##  2026 REFRESH (executive/plain-language pass):
##    * Every content-slide title is a plain-language takeaway; the metric
##      formerly labelled "40q30" is written as "Probability of death from CVD
##      before age 70" in ALL rendered text (internal object/sheet/column names
##      such as CVD_40q30 / cvd_40q30_r are kept in code only).
##    * Every analytical slide carries a muted italic Source/Method footnote
##      (\srcmethod macro, defined in header-includes); cost/CE slides also carry
##      the fixed BPJS/MOH validation caveat (\costcaveat).
##    * Each of the three CE slides shows TWO discounted per-capita cost measures,
##      each drawn from / reconstructed from that family's OWN cost workbook
##      (Budget_Impact population is authoritative; the combined workbook is never
##      a shortcut for the clinical or public-health tables):
##        (1) Average annual additional cost per capita, 2026-2050
##              = cumulative DISCOUNTED additional cost 2026-2050
##                / (25 x 2026 population)              [parentheses essential]
##        (2) Additional cost per capita in 2050 at target coverage
##              = discounted additional cost in 2050 / 2050 population.
##      The clinical workbook is Excel-recalculated (cached) and read directly; the
##      public-health and combined formula workbooks are UNcached, so their annual
##      discounted additional cost is reproduced in R from the SAME Annual_Cost
##      inputs the workbook formulae use and reconciled to the run's contract total.
##      No Excel COM is required.
##    * The national cascade sub-section additionally reports the year-specific
##      annual additional cost per capita (cascade workbook, Budget_Impact).
##    * A two-panel stacked clinical budget-impact figure (2025-2050) decomposes
##      cumulative discounted total and per-capita cost by clinical intervention,
##      read from the clinical workbook Annual_Cost and QA-reconciled by year.
##    * The isolated 70-30-30 -> 70-70-70 cascade strategy is read ONLY from its
##      cascade workbook (national under output/70_30_30_to_70_70_70/; SUBNATIONAL
##      38-province under output/70_30_30_to_70_70_70_subnational/); never merged
##      into the ordinary results contract.
##    * A high-level DCP costing-methods slide + an appendix of the adapted
##      equations cite Watkins et al. (2020).
## =====================================================================
library(here)
library(data.table)
library(openxlsx)
library(ggplot2)
library(ggrepel)
library(scales)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.align = "center", dev = "cairo_pdf"
)

## ---- Palette (matches beamer_preamble.tex scenario colours) ---------------
col_clinical  <- "#3B6FB6"  # clinical scenarios
col_ph        <- "#1B9E77"  # public-health scenarios
col_combined  <- "#08306B"  # combined package (darker)
col_baseline  <- "#6E6E6E"  # baseline / status quo
col_comparator<- "#BDBDBD"  # comparator countries (muted)
col_incidence <- "#2166AC"  # incidence (Well -> Sick) effects
col_casefatal <- "#B2182B"  # case-fatality (Sick -> Death) effects
col_bgmort    <- "#7F7F7F"  # background mortality
col_wbnavy    <- "#0F2A56"

theme_exec <- function(base = 11) {
  theme_minimal(base_size = base) +
    theme(
      plot.title       = element_text(face = "bold", size = base + 1, colour = col_wbnavy),
      plot.subtitle    = element_text(colour = "grey35", size = base - 2),
      plot.caption     = element_text(colour = "grey45", size = base - 4, hjust = 0),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      legend.position  = "bottom",
      legend.title     = element_blank(),
      axis.title       = element_text(size = base - 2)
    )
}

## ---- Number formatters (LaTeX-safe: escaped $ and % ; used with escape=FALSE)
fmt_int   <- function(x) formatC(round(x), format = "d", big.mark = ",")
fmt_thou  <- function(x) ifelse(is.na(x), "--", formatC(round(x / 1000) * 1000, format = "d", big.mark = ","))
fmt_short <- function(x, acc = 0.1) ifelse(is.na(x), "--", label_number(scale_cut = cut_short_scale(), accuracy = acc)(x))
fmt_pct   <- function(x, d = 1) ifelse(is.na(x), "--", paste0(formatC(x, format = "f", digits = d), "\\%"))
fmt_usd   <- function(x) ifelse(is.na(x), "--", paste0("US\\$", formatC(round(x), format = "d", big.mark = ",")))
fmt_usd_short <- function(x) ifelse(is.na(x), "--", paste0("US\\$", fmt_short(x)))
## Per-capita US$ to 2 decimals (LaTeX-safe: escaped $).
fmt_usd_pc <- function(x) ifelse(is.na(x), "--", paste0("US\\$", formatC(x, format = "f", digits = 2, big.mark = ",")))
## Per-capita US$ with ADAPTIVE precision so small public-health costs never show as
## $0.00: >=1 -> 2dp; >=0.1 -> 2dp; >=0.01 -> 3dp; else 4dp. Keeps big values clean
## while preserving 2+ significant figures for cents-scale policy costs.
fmt_usd_pc2 <- function(x) {
  vapply(x, function(v) {
    if (is.na(v)) return("--")
    a <- abs(v)
    d <- if (a >= 0.1 || a == 0) 2L else if (a >= 0.01) 3L else 4L
    paste0("US\\$", formatC(v, format = "f", digits = d, big.mark = ","))
  }, character(1))
}
## Title-safe (plain-text, NO % or $ so it survives pandoc -> \frametitle):
## express a count in the most readable unit ("3.5 million", "1.13 million").
fmt_title_num <- function(x, acc = 0.1) {
  if (is.na(x)) return("--")
  if (abs(x) >= 1e6) return(paste0(formatC(x / 1e6, format = "f", digits = 1), " million"))
  if (abs(x) >= 1e3) return(paste0(formatC(round(x / 1e3), format = "d", big.mark = ","), " thousand"))
  formatC(round(x), format = "d", big.mark = ",")
}
## USD in billions/millions as plain words (no $) for prose reconciliation lines.
fmt_bn <- function(x, d = 2) ifelse(is.na(x), "--", paste0(formatC(x / 1e9, format = "f", digits = d), " billion"))
## Two-line table headers (converts embedded "\n" to \makecell line breaks).
mkhdr <- function(v) kableExtra::linebreak(v, align = "c")

## Null / NA coalescing helpers (base R has neither).
`%||%` <- function(a, b) if (is.null(a) || length(a) == 0L || (length(a) == 1L && is.na(a))) b else a
`%+na%` <- function(a, b) if (is.null(a) || length(a) == 0L || is.na(a[1])) b else a[1]
```

```{r xlsx-companion-setup, include=FALSE}
## =====================================================================
##  COMPANION EXCEL WORKBOOK  (reports/executive_slides.xlsx)
##  --------------------------------------------------------------------
##  A companion workbook is built programmatically AS THIS DECK KNITS.
##  Every figure/plot and every table that appears on a slide is captured
##  -- in slide order -- via reg_fig()/reg_tbl() at the exact point the
##  slide builds it, so the workbook reuses the SAME data objects and the
##  SAME plots the slides render. The workbook is assembled and written by
##  the final `build-xlsx-workbook` chunk at the end of the document.
##
##  Registration is INERT for the rendered deck:
##    * reg_tbl() returns its data invisibly -> emits nothing on the slide
##      (safe inside results='asis' chunks);
##    * reg_fig() saves the plot to PNG and RETURNS the ggplot, so the
##      figure chunk still auto-prints the identical figure to the deck.
##  A failure inside either helper is swallowed (the deck must still build).
## =====================================================================
.xlsx_enabled <- TRUE
.xlsx_items   <- list()
.xlsx_fig_dir <- file.path(tempdir(), "executive_slides_xlsx_figs")
dir.create(.xlsx_fig_dir, showWarnings = FALSE, recursive = TRUE)

## Register a table. `data` is the SAME object passed to kbl(); `col_names`
## the SAME display headers used on the slide (raw -- may carry LaTeX markup
## or embedded "\n"; both are cleaned to plain text at write time).
reg_tbl <- function(id, title, data, col_names = NULL, note = NULL) {
  if (!isTRUE(.xlsx_enabled)) return(invisible(data))
  try({
    .xlsx_items[[length(.xlsx_items) + 1L]] <<- list(
      kind = "table", id = id, title = as.character(title),
      data = as.data.frame(data, stringsAsFactors = FALSE),
      col_names = if (is.null(col_names)) NULL else as.character(col_names),
      note = note)
  }, silent = TRUE)
  invisible(data)
}

## Register a figure. Saves the ggplot to PNG at the slide's own dimensions
## and RETURNS the plot so the chunk auto-prints it into the deck unchanged.
reg_fig <- function(plot, id, title, width, height, dpi = 200, note = NULL) {
  if (!isTRUE(.xlsx_enabled)) return(plot)
  path <- file.path(.xlsx_fig_dir, paste0(id, ".png"))
  ok <- tryCatch({
    suppressWarnings(suppressMessages(
      ggplot2::ggsave(path, plot = plot, width = width, height = height,
                      units = "in", dpi = dpi, device = "png", bg = "white")))
    TRUE
  }, error = function(e) FALSE)
  try({
    .xlsx_items[[length(.xlsx_items) + 1L]] <<- list(
      kind = "figure", id = id, title = as.character(title),
      path = if (isTRUE(ok)) path else NA_character_,
      width = width, height = height, note = note)
  }, silent = TRUE)
  plot
}
```

```{r load-data, include=FALSE}
## ---------------------------------------------------------------------------
## 1. Resolve current-run inputs (project-relative) and check existence.
##    The deck reads a BCA-free R-side results contract + literal workbook
##    values only -- no Excel COM, no formula recalculation.
## ---------------------------------------------------------------------------
HZ_START <- 2025L; HZ_END <- 2050L

## Output directory: defaults to the project's output/, but can be redirected
## (e.g. for testing an alternative active-scenario subset) via EXEC_DECK_OUTDIR.
.out_dir <- Sys.getenv("EXEC_DECK_OUTDIR", unset = "")
outp <- if (nzchar(.out_dir)) function(f) file.path(.out_dir, f) else function(f) here("output", f)
deck_file <- outp("09_deck_results.rds")
wb_comb   <- outp("indonesia_model_cost_value_clinical_public_health_formulae.xlsx")
wb_clin   <- outp("indonesia_model_cost_value_formulae.xlsx")
wb_ph     <- outp("indonesia_cost_value_public_health_formulae.xlsx")
## Cached R-value companion to the clinical formula workbook: holds the SAME Model-09
## values as literals. Used for the clinical Budget_Impact when the formula workbook is
## written uncached (openxlsx default) so its computed cost cells read back as NA.
wb_clin_cached <- outp("indonesia_model_cost_value.xlsx")
csv_cmp   <- here("data", "raw", "03_asean_cvd_40q30_comparison.csv")

## Isolated 70-30-30 -> 70-70-70 cascade run (its OWN output dir; never merged with the
## ordinary results contract). Overridable for testing, defaults to the standard path.
.casc_dir <- Sys.getenv("EXEC_DECK_CASCADE_DIR", unset = "")
casc_wb <- if (nzchar(.casc_dir)) file.path(.casc_dir, "indonesia_70_30_30_to_70_70_70_cost_value_formulae.xlsx") else
  here("output", "70_30_30_to_70_70_70", "indonesia_70_30_30_to_70_70_70_cost_value_formulae.xlsx")

## Joint cascade + public-health run. Its CVD_40q30 sheet contains, among other
## scenarios, the hypertension/cholesterol + diabetes cascade (S_70_30_30_TO_70_70_70)
## and the cascade combined with the tobacco tax ONLY
## (S_70_30_30_TO_70_70_70_I_PH_TOBACCO_TAX) -- not the full tobacco regulatory
## package. Used by fig-40q30-cascade only.
.casc_ph_wb <- Sys.getenv("EXEC_DECK_CASCADE_PH_WB", unset = "")
casc_ph_wb <- if (nzchar(.casc_ph_wb)) .casc_ph_wb else
  outp("indonesia_model_cost_value_70_30_30_to_70_70_70_public_health_formulae.xlsx")

## Isolated SUBNATIONAL 70-30-30 -> 70-70-70 cascade run (38 provinces; its OWN output dir).
## The expected deliverable is the province-level formula workbook; overridable for testing.
.subnat_dir <- Sys.getenv("EXEC_DECK_SUBNAT_DIR", unset = "")
subnat_wb <- if (nzchar(.subnat_dir))
  file.path(.subnat_dir, "indonesia_70_30_30_to_70_70_70_cost_value_formulae_subnational.xlsx") else
  here("output", "70_30_30_to_70_70_70_subnational",
       "indonesia_70_30_30_to_70_70_70_cost_value_formulae_subnational.xlsx")

.req_files <- c(results_contract = deck_file, clinical_workbook = wb_clin,
                clinical_workbook_cached = wb_clin_cached,
                public_health_workbook = wb_ph, combined_workbook = wb_comb,
                comparator = csv_cmp, cascade_workbook = casc_wb,
                cascade_public_health_workbook = casc_ph_wb,
                subnational_workbook = subnat_wb)
.missing <- .req_files[!file.exists(.req_files)]
if (length(.missing))
  stop("VALIDATION STOP -- required input file(s) not found:\n  ",
       paste(sprintf("%s: %s", names(.missing), .missing), collapse = "\n  "),
       "\nRun the modeling pipeline (code/cvd-fair-choices/00_run_model_cvd_fair.R) first.",
       call. = FALSE)

rd        <- function(path, sheet) as.data.table(openxlsx::read.xlsx(path, sheet = sheet))
sheets_of <- function(path) openxlsx::getSheetNames(path)

## ---- Results contract -----------------------------------------------------
dk   <- readRDS(deck_file)
res  <- as.data.table(dk$results)
meta <- dk$meta

met_clin <- res[workbook == "clinical"]
met_ph   <- res[workbook == "public_health"]
met_comb <- res[workbook == "combined"]

## ---- Costing/report metadata (BCA-free: cost side only) -------------------
disc_rate  <- suppressWarnings(as.numeric(meta$cost_discount_rate)); if (!is.finite(disc_rate)) disc_rate <- 0.03
price_year <- suppressWarnings(as.numeric(meta$cost_price_year));    if (!is.finite(price_year)) price_year <- 2023
ce_unit    <- meta$cost_effectiveness_unit %||% "USD per death averted"
## LaTeX-safe escaper for dynamic strings inserted into prose with escape=FALSE.
ltx <- function(s) {
  s <- as.character(s)
  s <- gsub("\\\\", "\\\\textbackslash{}", s)
  s <- gsub("([&%$#_{}])", "\\\\\\1", s)
  s <- gsub("~", "\\\\textasciitilde{}", s)
  s
}
cost_perspective <- ltx(gsub("_", " ", meta$economic_perspective %||% "health system"))
```

```{r derive-scenarios, include=FALSE}
## ---------------------------------------------------------------------------
## 2. Scenario identity is DERIVED from the contract / Scenario_Catalog, never
##    from fixed ID vectors. Optional presentation-only label overrides may be
##    supplied; any active scenario without an override falls back safely to its
##    workbook label. No excluded / hard-coded ID is ever required to exist.
## ---------------------------------------------------------------------------
exec_label_override <- c(
  I_CVD_PRIMARY = "Primary CVD prevention", I_T2D_TREATMENT = "Type-2 diabetes treatment",
  I_IHD_SECONDARY = "Heart-attack secondary prevention", I_HF_CHRONIC = "Chronic heart-failure care",
  I_HF_ACUTE = "Acute heart-failure care", I_STROKE_SECONDARY = "Stroke secondary prevention",
  I_RHD_SECONDARY = "RHD secondary prevention", I_RHD_SURGERY = "RHD surgery",
  all = "All clinical (combined)",
  I_PH_TOBACCO_TAX = "Tobacco tax", I_PH_TOBACCO_POLICY = "Tobacco policy package",
  I_PH_SALT_POLICY = "Sodium policy package", I_PH_TFA_POLICY = "Trans-fat elimination",
  all_public_health = "All public health (combined)",
  all_clinical_public_health = "Clinical + public health (combined)",
  baseline = "Status quo (baseline)")

## Family aggregate / joint scenario ids (present only if built this run).
clin_all_id <- "all"; ph_all_id <- "all_public_health"; joint_id <- "all_clinical_public_health"
has_clin  <- nrow(met_clin[scenario != "baseline"]) > 0
has_ph    <- nrow(met_ph[scenario != "baseline"])   > 0
has_joint <- joint_id %in% met_comb$scenario

## Label getter: override first, else the contract's own scenario_label.
lbl <- function(sc, tab = met_comb) {
  if (sc %in% names(exec_label_override)) return(exec_label_override[[sc]])
  v <- tab[scenario == sc, scenario_label]
  if (length(v) && !is.na(v[1]) && nzchar(v[1])) v[1] else sc
}
## Metric getter.
hv <- function(sc, col, tab = met_comb) { v <- tab[scenario == sc][[col]]; if (length(v)) v[1] else NA_real_ }

## Top-level display rows for a family (exclude baseline and package CHILDREN to
## avoid double-counting; the family aggregate is placed last). Ordered by
## deaths averted. Robust to zero/one/many active interventions.
MAX_TBL_ROWS <- 14L
top_level_rows <- function(tab, agg_id) {
  if (!nrow(tab)) return(tab[0])
  d <- tab[scenario != "baseline" &
             (is.na(scenario_role) | !(scenario_role %in% c("child")))]
  if (!nrow(d)) return(d)
  d <- d[order(-deaths_averted)]
  agg <- d[scenario == agg_id]; rest <- d[scenario != agg_id]
  rbind(rest, agg)
}
```

```{r headline-scalars, include=FALSE}
## Baseline CVD 40q30 (per-year literal from the combined workbook).
c4c <- rd(wb_comb, "CVD_40q30")
qcol <- if ("cvd_40q30_r" %in% names(c4c)) "cvd_40q30_r" else "cvd_40q30"
base_q_2025 <- c4c[scenario == "baseline" & year == HZ_START][[qcol]][1]
base_q_2050 <- c4c[scenario == "baseline" & year == HZ_END][[qcol]][1]

## Combined-package (joint) headline metrics (guarded if joint absent).
combo_da   <- if (has_joint) hv(joint_id, "deaths_averted") else NA_real_
combo_cpda <- if (has_joint) hv(joint_id, "cost_per_death_averted") else NA_real_
combo_q50  <- if (has_joint) hv(joint_id, "cvd_40q30_2050") else NA_real_
combo_qred <- if (has_joint) hv(joint_id, "cvd_40q30_pct_reduction_2050") else NA_real_
combo_dred <- if (has_joint) hv(joint_id, "pct_reduction_deaths_2050") else NA_real_

clin_da   <- hv(clin_all_id, "deaths_averted", met_clin)
clin_qred <- hv(clin_all_id, "cvd_40q30_pct_reduction_2050", met_clin)
clin_cpda <- hv(clin_all_id, "cost_per_death_averted", met_clin)
ph_da     <- hv(ph_all_id, "deaths_averted", met_ph)
ph_qred   <- hv(ph_all_id, "cvd_40q30_pct_reduction_2050", met_ph)
ph_cpda   <- hv(ph_all_id, "cost_per_death_averted", met_ph)

## Most impactful single lever per family (excludes the aggregate).
.tc <- top_level_rows(met_clin, clin_all_id); .tc <- .tc[scenario != clin_all_id]
.tp <- top_level_rows(met_ph, ph_all_id);     .tp <- .tp[scenario != ph_all_id]
top_clin_sc <- if (nrow(.tc)) .tc$scenario[1] else NA_character_
top_ph_sc   <- if (nrow(.tp)) .tp$scenario[1] else NA_character_
```

```{r validation, include=FALSE}
## ---------------------------------------------------------------------------
## 3. VALIDATION GATES - health/cost-effectiveness contract only (NO BCA).
##    Consistency is checked against Scenario_Catalog + the results contract.
## ---------------------------------------------------------------------------
val_stop <- function(cond, msg) if (!isTRUE(cond)) stop("VALIDATION STOP - ", msg, call. = FALSE)

## (a) Scenario_Catalog present in every workbook that exists (health/CE contract;
##     Benefit_Cost / Economic_Value are NOT required by the deck).
cat_sets <- list()
.wb_paths <- c(clinical = wb_clin, public_health = wb_ph, combined = wb_comb)
for (fam in names(.wb_paths)) {
  path <- .wb_paths[[fam]]
  if (!file.exists(path)) next
  sh <- sheets_of(path)
  val_stop("Scenario_Catalog" %in% sh,
           sprintf("workbook '%s' is missing the Scenario_Catalog sheet", basename(path)))
  val_stop(all(c("CVD_40q30", "Cost_Effectiveness") %in% sh),
           sprintf("workbook '%s' is missing a required health/CE sheet", basename(path)))
  cat_sets[[fam]] <- rd(path, "Scenario_Catalog")$scenario
}

## (b) Contract scenario set reconciles EXACTLY with Scenario_Catalog per family
##     (no stale extras, nothing dropped). This is how "no excluded id sneaks in"
##     is enforced without hard-coding any excluded id.
recon_family <- function(cat_key, tab) {
  cs <- cat_sets[[cat_key]]
  if (is.null(cs)) return(invisible())
  ts <- unique(tab$scenario)
  val_stop(setequal(cs, ts),
           sprintf("%s: results contract scenarios do not match Scenario_Catalog. Only in catalog: {%s}; only in results: {%s}. Re-run Model 09 on the current run.",
                   cat_key, paste(setdiff(cs, ts), collapse = ", "),
                   paste(setdiff(ts, cs), collapse = ", ")))
}
recon_family("clinical", met_clin)
recon_family("public_health", met_ph)
recon_family("combined", met_comb)

## (c) baseline available exactly once per family + 2050 present.
for (tab in list(met_clin, met_ph, met_comb)) {
  if (!nrow(tab)) next
  val_stop(sum(tab$scenario == "baseline") <= 1, "baseline appears more than once in a family")
}
val_stop(any(c4c$scenario == "baseline") && HZ_END %in% c4c$year,
         "combined workbook CVD_40q30 has no baseline / no year 2050")

## (d) unique scenario ids within each family (no duplicates).
for (tab in list(met_clin, met_ph, met_comb))
  if (nrow(tab)) val_stop(!any(duplicated(tab$scenario)),
                          "duplicate scenario id within a family results block")

## (e) headline health/CE metrics finite where the scenario exists.
if (has_joint)
  val_stop(is.finite(combo_da) && is.finite(combo_q50) && is.finite(combo_cpda),
           "combined-package headline metrics are missing/non-numeric")
val_stop(is.finite(base_q_2050), "baseline CVD 40q30 in 2050 is missing")
```

```{r fig-40q30-data, include=FALSE}
## Data for the CVD 40q30 time-series & international-comparison figure.
cmp <- fread(csv_cmp)
cmp[, year := as.integer(year)]
comparators <- c("Malaysia","Philippines","Singapore","Thailand","Viet Nam","Indonesia")

cmp_lines <- cmp[country_label %in% comparators,
                 .(country_label, year, q = q40q30_percent)][order(country_label, year)]
cmp_lines <- unique(cmp_lines, by = c("country_label","year"))

#tempatch
cmp_lines[country_label == "Indonesia", country_label := "Indonesia."]

idn_obs <- cmp[country_label == "Indonesia" & series_type == "Historical WPP-scaled baseline" &
               year <= 2023, .(year, q = q40q30_percent)][order(year)]

## Indonesia MODEL scenarios (dashed): whichever of these are BOTH in the
## current-run contract and available as a per-year series in the workbook.
fig_scen_ids <- intersect(c("baseline", clin_all_id, ph_all_id, joint_id),
                          intersect(unique(c4c$scenario), unique(met_comb$scenario)))
idn_scen <- c4c[scenario %in% fig_scen_ids & year >= HZ_START,
                .(scenario, year, q = get(qcol))]
idn_scen[, scen := vapply(scenario, lbl, character(1), USE.NAMES = FALSE)]
scen_levels <- vapply(fig_scen_ids, lbl, character(1), USE.NAMES = FALSE)
idn_scen[, scen := factor(scen, levels = scen_levels)]
scen_pal <- setNames(
  c(col_baseline, col_clinical, col_ph, col_combined)[seq_along(fig_scen_ids)],
  scen_levels)
```

```{r national-cost-recon, include=FALSE}
## ---------------------------------------------------------------------------
## National per-capita cost measures - reconstructed from EACH family's OWN cost
## workbook, keyed by scenario (never the combined workbook as a shortcut).
##
## The two discounted measures shown in every national CE table are:
##   (1) Average annual additional cost per capita, 2026-2050 (discounted)
##         = [ cumulative DISCOUNTED additional cost over 2026-2050 ]
##           / ( 25 x 2026 population )            <-- parentheses are essential
##   (2) Additional cost per capita in 2050 at target coverage (discounted)
##         = discounted additional cost in 2050 / 2050 population
## Population is the workbook's own national_population (Budget_Impact), identical
## across scenarios by year; 2026 and 2050 are authoritative.
##
## Because openxlsx writes formula cells UNcached, the clinical workbook's
## Budget_Impact is fully Excel-recalculated (cached) and read directly, while the
## public-health and combined formula workbooks are uncached: their annual
## discounted additional cost is reproduced from the SAME authoritative Annual_Cost
## inputs the workbook formulae use, then reconciled to the run's contract total
## (disc_incremental_cost) to a tight relative tolerance. No Excel COM is required.
## ---------------------------------------------------------------------------
.ca_num <- function(path, id) {
  ca <- rd(path, "Calculation_Assumptions")
  v  <- suppressWarnings(as.numeric(ca[parameter_id == id, value][1])); v
}

## Cached clinical Budget_Impact resolver. The clinical formula workbook's Budget_Impact
## holds computed (formula) cost cells that are populated only when the workbook has been
## Excel-recalculated; when it is written uncached (openxlsx default) they read back as NA.
## In that case fall back to the companion R-value workbook, which stores the SAME Model-09
## R-engine values as literals, so the clinical family's discounted per-capita measures
## remain available without Excel COM.
read_clin_bi <- function() {
  bi <- rd(wb_clin, "Budget_Impact")
  if (!"disc_incremental_cost" %in% names(bi) || all(is.na(bi$disc_incremental_cost)))
    bi <- rd(wb_clin_cached, "Budget_Impact")
  bi
}
## Cached clinical Annual_Cost resolver (same fallback rationale as read_clin_bi). The
## R-value companion carries de-duplicated rows and no shared_duplicate_count column, so
## a missing/uncached count is treated as 1 (each component counted once, unweighted).
read_clin_ac <- function() {
  ac <- rd(wb_clin, "Annual_Cost")
  if (!"disc_cost_incremental" %in% names(ac) || all(is.na(ac$disc_cost_incremental)))
    ac <- rd(wb_clin_cached, "Annual_Cost")
  if (!"shared_duplicate_count" %in% names(ac) || all(is.na(ac$shared_duplicate_count)))
    ac[, shared_duplicate_count := 1]
  ac
}

## Annual discounted additional cost + national population, by (scenario, year).
build_annual_cost <- function(fam) {
  if (fam == "clinical") {
    ## Cached clinical Budget_Impact (formula workbook if recalculated, else R-value companion).
    bi <- read_clin_bi()
    need <- c("scenario","year","disc_incremental_cost","national_population","cumulative_disc_incremental_cost")
    val_stop(all(need %in% names(bi)), "clinical Budget_Impact missing required column(s)")
    ## Reconcile the annual series to the workbook's cumulative field, and confirm the
    ## 2026-2050 numerator equals cumulative[2050] - disc[2025] (2025 IS in the cumulative).
    .rc <- bi[, .(a = cumsum(disc_incremental_cost),
                  c = cumulative_disc_incremental_cost), by = scenario]
    val_stop(max(abs(.rc$a - .rc$c)) < 1,
             "clinical annual discounted cost does not reconcile to the cumulative field")
    .n1 <- bi[year %in% (HZ_START + 1L):HZ_END, .(num = sum(disc_incremental_cost)), by = scenario]
    .n2 <- bi[year == HZ_END, .(scenario, cum = cumulative_disc_incremental_cost)][
             bi[year == HZ_START, .(scenario, d25 = disc_incremental_cost)], on = "scenario"]
    .nn <- merge(.n1, .n2, by = "scenario")
    val_stop(max(abs(.nn$num - (.nn$cum - .nn$d25))) < 1,
             "clinical 2026-2050 numerator does not equal cumulative[2050] - 2025 value")
    out <- bi[, .(scenario, year, disc = disc_incremental_cost, pop = national_population)]
  } else if (fam == "public_health") {
    ## Uncached: reproduce annual policy-implementation cost from Annual_Cost inputs.
    r  <- .ca_num(wb_ph, "cost_discount_rate");  y0 <- .ca_num(wb_ph, "analysis_start_year")
    ps <- .ca_num(wb_ph, "policy_start_year");   rp <- .ca_num(wb_ph, "policy_cost_ramp_years")
    b0 <- .ca_num(wb_ph, "baseline_implementation_fraction")
    val_stop(all(is.finite(c(r, y0, ps, rp))), "public-health cost assumptions (rate/start/ramp) missing")
    if (!is.finite(b0)) b0 <- 0
    ac <- rd(wb_ph, "Annual_Cost")
    need <- c("scenario","year","pop_scenario","population_in_need_fraction","frequency_per_year","unit_cost_usd")
    val_stop(all(need %in% names(ac)), "public-health Annual_Cost missing required input column(s)")
    imf <- function(t) { f <- (t - (ps - 1)) / rp; f[t < ps] <- 0; pmin(pmax(f, 0), 1) }
    ac[, disc := pop_scenario * population_in_need_fraction * frequency_per_year * unit_cost_usd *
                 (imf(year) - b0) * (1 / (1 + r)^(year - y0))]
    a  <- ac[, .(disc = sum(disc)), by = .(scenario, year)]
    bi <- rd(wb_ph, "Budget_Impact")
    val_stop("national_population" %in% names(bi), "public-health Budget_Impact missing national_population")
    out <- merge(a, bi[, .(scenario, year, pop = national_population)], by = c("scenario","year"))
  } else if (fam == "combined") {
    ## Uncached: reproduce annual additional cost from Annual_Cost inputs.
    r  <- .ca_num(wb_comb, "cost_discount_rate"); y0 <- .ca_num(wb_comb, "analysis_start_year")
    val_stop(all(is.finite(c(r, y0))), "combined cost assumptions (rate/start) missing")
    ac <- rd(wb_comb, "Annual_Cost")
    need <- c("scenario","year","pin_fraction","frequency_per_year","unit_cost_usd",
              "q_scenario","q_baseline","covimpl_scenario","covimpl_baseline")
    val_stop(all(need %in% names(ac)), "combined Annual_Cost missing required input column(s)")
    ac[, disc := pin_fraction * frequency_per_year * unit_cost_usd *
                 (q_scenario * covimpl_scenario - q_baseline * covimpl_baseline) *
                 (1 / (1 + r)^(year - y0))]
    a  <- ac[, .(disc = sum(disc)), by = .(scenario, year)]
    bi <- rd(wb_comb, "Budget_Impact")
    val_stop("national_population" %in% names(bi), "combined Budget_Impact missing national_population")
    out <- merge(a, bi[, .(scenario, year, pop = national_population)], by = c("scenario","year"))
  } else stop("unknown family: ", fam)
  setorder(out, scenario, year)
  ## --- Structural validation ---------------------------------------------
  yrs <- sort(unique(out$year))
  val_stop(setequal(yrs, HZ_START:HZ_END),
           sprintf("%s cost workbook year coverage is not %d-%d", fam, HZ_START, HZ_END))
  ## national population identical across scenarios each year (single denominator)
  val_stop(nrow(out[, .(nu = uniqueN(pop)), by = year][nu > 1]) == 0,
           sprintf("%s national_population differs across scenarios within a year", fam))
  ## reconcile reconstructed full-horizon total to the run's contract total
  ct <- res[workbook == fam, .(scenario, contract = disc_incremental_cost)]
  rc <- merge(out[, .(tot = sum(disc)), by = scenario], ct, by = "scenario")
  rc[, rel := abs(tot - contract) / pmax(abs(contract), 1)]
  val_stop(nrow(rc) > 0 && max(rc$rel) < 1e-4,
           sprintf("%s reconstructed discounted cost does not reconcile to the contract total (max rel.diff %.2g)",
                   fam, if (nrow(rc)) max(rc$rel) else NA_real_))
  out[]
}
annual_cost_tabs <- list(
  clinical      = build_annual_cost("clinical"),
  public_health = build_annual_cost("public_health"),
  combined      = build_annual_cost("combined"))

## The two discounted per-capita measures for one (family, scenario).
DISPLAY_YEARS <- (HZ_START + 1L):HZ_END                 # 2026..2050
val_stop(length(DISPLAY_YEARS) == 25L,
         sprintf("expected exactly 25 years 2026-%d; got %d", HZ_END, length(DISPLAY_YEARS)))
pc_avg_annual <- function(fam, sc) {                    # measure (1)
  d <- annual_cost_tabs[[fam]][scenario == sc]
  if (!nrow(d)) return(NA_real_)
  num   <- d[year %in% DISPLAY_YEARS, sum(disc)]
  pop26 <- d[year == HZ_START + 1L, pop][1]
  if (!is.finite(num) || !is.finite(pop26) || pop26 <= 0) return(NA_real_)
  num / (25 * pop26)
}
pc_year2050 <- function(fam, sc) {                      # measure (2)
  d <- annual_cost_tabs[[fam]][scenario == sc]
  if (!nrow(d)) return(NA_real_)
  disc50 <- d[year == HZ_END, disc][1]; pop50 <- d[year == HZ_END, pop][1]
  if (!is.finite(disc50) || !is.finite(pop50) || pop50 <= 0) return(NA_real_)
  disc50 / pop50
}
## Independent validation: measure (2) for the clinical family must equal the
## clinical workbook's own cached disc_incremental_cost_per_capita in 2050.
{ .bi <- read_clin_bi()
  if ("disc_incremental_cost_per_capita" %in% names(.bi)) {
    .chk <- .bi[year == HZ_END, .(scenario, stored = disc_incremental_cost_per_capita)]
    .cmp <- merge(.chk, data.table(scenario = .chk$scenario,
                                   rebuilt = vapply(.chk$scenario, function(s) pc_year2050("clinical", s), numeric(1))),
                  by = "scenario")
    val_stop(max(abs(.cmp$stored - .cmp$rebuilt)) < 1e-6,
             "clinical 2050 per-capita reconstruction disagrees with the workbook's stored value") }
}
## Map contract workbook key -> the annual-cost family key (same names here).
fam_of <- function(tab) tab[1, workbook]
```

```{r results-builders, include=FALSE}
## Row builders for the impact / cost-effectiveness display tables (BCA-free).
impact_row <- function(tab, sc) data.table(
  Intervention = lbl(sc, tab),
  da  = fmt_thou(hv(sc, "deaths_averted", tab)),
  q50 = fmt_pct(hv(sc, "cvd_40q30_2050", tab), 1),
  qr  = fmt_pct(hv(sc, "cvd_40q30_pct_reduction_2050", tab), 1),
  dr  = fmt_pct(hv(sc, "pct_reduction_deaths_2050", tab), 1))
## The two DISCOUNTED per-capita measures for one scenario, reconstructed from the
## family's own cost workbook (see the national-cost-recon chunk):
##   pc1 = average annual additional cost per capita, 2026-2050
##       = cumulative discounted additional cost 2026-2050 / (25 x 2026 population)
##   pc2 = discounted additional cost per capita in 2050 / 2050 population
## Adaptive precision keeps cents-scale public-health costs from rounding to zero.
pc1 <- function(tab, sc) fmt_usd_pc2(pc_avg_annual(fam_of(tab), sc))
pc2 <- function(tab, sc) fmt_usd_pc2(pc_year2050(fam_of(tab), sc))
ce_row <- function(tab, sc) cbind(
  impact_row(tab, sc),
  data.table(cpd  = fmt_usd(hv(sc, "cost_per_death_averted", tab)),
             pc1c = pc1(tab, sc),
             pc2c = pc2(tab, sc)))

## Build a display table for a family's top-level rows, capped for space.
## Returns list(tb, agg_row_idx, n_omitted, omitted_labels).
family_table <- function(tab, agg_id, row_fun) {
  d <- top_level_rows(tab, agg_id)
  n_omitted <- 0L; omitted <- character(0)
  if (nrow(d) > MAX_TBL_ROWS) {
    keep_agg <- d[scenario == agg_id]
    body <- d[scenario != agg_id]
    n_keep <- MAX_TBL_ROWS - nrow(keep_agg)
    omitted <- body$scenario[-seq_len(min(n_keep, nrow(body)))]
    body <- body[seq_len(min(n_keep, nrow(body)))]
    d <- rbind(body, keep_agg)
    n_omitted <- length(omitted)
    message(sprintf("  Deck: family table capped at %d rows; %d intervention(s) omitted for space: %s",
                    MAX_TBL_ROWS, n_omitted, paste(omitted, collapse = ", ")))
  }
  if (!nrow(d)) return(list(tb = data.table(), agg_idx = NA_integer_, n_omitted = 0L, omitted = character(0)))
  tb <- rbindlist(lapply(d$scenario, function(s) row_fun(tab, s)))
  list(tb = tb, agg_idx = which(d$scenario == agg_id), n_omitted = n_omitted, omitted = omitted)
}
```

```{r national-cost-recon-undisc, include=FALSE}
## ---------------------------------------------------------------------------
## UNDISCOUNTED national cost measures - an exact parallel to national-cost-recon,
## reconstructed from EACH family's OWN cost workbook, keyed by scenario. This
## block leaves every discounted helper/object untouched and only ADDS the
## undiscounted analogues (suffix _undisc). Every workbook-specific rule is
## preserved: the same population denominators, coverage/implementation
## quantities, scenario identity, and shared-component de-duplication as the
## discounted path -- the ONLY change is that the per-year discount factor is
## omitted from the annual cost. Because the formula workbooks write ALL computed
## cost cells (discounted AND undiscounted) uncached, the reconstruction reads the
## same authoritative INPUT columns the workbook formulae use:
##   * clinical      -> the cached companion Budget_Impact stores incremental_cost
##                      (annual undiscounted) + cumulative_incremental_cost as literals.
##   * public-health -> annual undiscounted policy-implementation cost from Annual_Cost
##                      inputs (drop the 1/(1+r)^(t-y0) factor from the discounted build).
##   * combined      -> annual undiscounted additional cost from Annual_Cost inputs
##                      (same drop-the-discount-factor reconstruction).
## Each reconstruction is reconciled to the run's contract UNDISCOUNTED total
## (res$incremental_cost) to a tight relative tolerance -- the undiscounted twin of
## the discounted reconciliation. No Excel COM is required.
## ---------------------------------------------------------------------------
build_annual_cost_undisc <- function(fam) {
  if (fam == "clinical") {
    ## Cached clinical Budget_Impact stores undiscounted incremental_cost as literals.
    bi <- read_clin_bi()
    need <- c("scenario","year","incremental_cost","national_population","cumulative_incremental_cost")
    val_stop(all(need %in% names(bi)), "clinical Budget_Impact missing required undiscounted column(s)")
    .rc <- bi[, .(a = cumsum(incremental_cost), c = cumulative_incremental_cost), by = scenario]
    val_stop(max(abs(.rc$a - .rc$c)) < 1,
             "clinical annual undiscounted cost does not reconcile to the cumulative field")
    out <- bi[, .(scenario, year, undisc = incremental_cost, pop = national_population)]
  } else if (fam == "public_health") {
    ## Uncached: reproduce annual UNDISCOUNTED policy-implementation cost from Annual_Cost inputs.
    ps <- .ca_num(wb_ph, "policy_start_year");   rp <- .ca_num(wb_ph, "policy_cost_ramp_years")
    b0 <- .ca_num(wb_ph, "baseline_implementation_fraction")
    val_stop(all(is.finite(c(ps, rp))), "public-health cost assumptions (start/ramp) missing")
    if (!is.finite(b0)) b0 <- 0
    ac <- rd(wb_ph, "Annual_Cost")
    need <- c("scenario","year","pop_scenario","population_in_need_fraction","frequency_per_year","unit_cost_usd")
    val_stop(all(need %in% names(ac)), "public-health Annual_Cost missing required input column(s)")
    imf <- function(t) { f <- (t - (ps - 1)) / rp; f[t < ps] <- 0; pmin(pmax(f, 0), 1) }
    ac[, undisc := pop_scenario * population_in_need_fraction * frequency_per_year * unit_cost_usd *
                   (imf(year) - b0)]
    a  <- ac[, .(undisc = sum(undisc)), by = .(scenario, year)]
    bi <- rd(wb_ph, "Budget_Impact")
    val_stop("national_population" %in% names(bi), "public-health Budget_Impact missing national_population")
    out <- merge(a, bi[, .(scenario, year, pop = national_population)], by = c("scenario","year"))
  } else if (fam == "combined") {
    ## Uncached: reproduce annual UNDISCOUNTED additional cost from Annual_Cost inputs.
    ac <- rd(wb_comb, "Annual_Cost")
    need <- c("scenario","year","pin_fraction","frequency_per_year","unit_cost_usd",
              "q_scenario","q_baseline","covimpl_scenario","covimpl_baseline")
    val_stop(all(need %in% names(ac)), "combined Annual_Cost missing required input column(s)")
    ac[, undisc := pin_fraction * frequency_per_year * unit_cost_usd *
                   (q_scenario * covimpl_scenario - q_baseline * covimpl_baseline)]
    a  <- ac[, .(undisc = sum(undisc)), by = .(scenario, year)]
    bi <- rd(wb_comb, "Budget_Impact")
    val_stop("national_population" %in% names(bi), "combined Budget_Impact missing national_population")
    out <- merge(a, bi[, .(scenario, year, pop = national_population)], by = c("scenario","year"))
  } else stop("unknown family: ", fam)
  setorder(out, scenario, year)
  ## --- Structural validation (mirrors the discounted build) ---------------
  yrs <- sort(unique(out$year))
  val_stop(setequal(yrs, HZ_START:HZ_END),
           sprintf("%s undiscounted cost workbook year coverage is not %d-%d", fam, HZ_START, HZ_END))
  val_stop(nrow(out[, .(nu = uniqueN(pop)), by = year][nu > 1]) == 0,
           sprintf("%s national_population differs across scenarios within a year (undiscounted)", fam))
  ## reconcile reconstructed full-horizon total to the run's contract UNDISCOUNTED total
  ct <- res[workbook == fam, .(scenario, contract = incremental_cost)][is.finite(contract)]
  rc <- merge(out[, .(tot = sum(undisc)), by = scenario], ct, by = "scenario")
  rc[, rel := abs(tot - contract) / pmax(abs(contract), 1)]
  val_stop(nrow(rc) > 0 && max(rc$rel) < 1e-4,
           sprintf("%s reconstructed undiscounted cost does not reconcile to the contract total (max rel.diff %.2g)",
                   fam, if (nrow(rc)) max(rc$rel) else NA_real_))
  out[]
}
annual_cost_tabs_undisc <- list(
  clinical      = build_annual_cost_undisc("clinical"),
  public_health = build_annual_cost_undisc("public_health"),
  combined      = build_annual_cost_undisc("combined"))

## The two UNDISCOUNTED per-capita measures for one (family, scenario).
pc_avg_annual_undisc <- function(fam, sc) {                 # measure (1), undiscounted
  d <- annual_cost_tabs_undisc[[fam]][scenario == sc]
  if (!nrow(d)) return(NA_real_)
  num   <- d[year %in% DISPLAY_YEARS, sum(undisc)]
  pop26 <- d[year == HZ_START + 1L, pop][1]
  if (!is.finite(num) || !is.finite(pop26) || pop26 <= 0) return(NA_real_)
  num / (25 * pop26)
}
pc_year2050_undisc <- function(fam, sc) {                   # measure (2), undiscounted
  d <- annual_cost_tabs_undisc[[fam]][scenario == sc]
  if (!nrow(d)) return(NA_real_)
  u50 <- d[year == HZ_END, undisc][1]; pop50 <- d[year == HZ_END, pop][1]
  if (!is.finite(u50) || !is.finite(pop50) || pop50 <= 0) return(NA_real_)
  u50 / pop50
}
## Independent validation: measure (2) UNDISCOUNTED for the clinical family must equal
## the clinical workbook's own cached incremental_cost_per_capita in 2050.
{ .bi <- read_clin_bi()
  if ("incremental_cost_per_capita" %in% names(.bi)) {
    .chk <- .bi[year == HZ_END, .(scenario, stored = incremental_cost_per_capita)][is.finite(stored)]
    if (nrow(.chk)) {
      .cmp <- data.table(scenario = .chk$scenario, stored = .chk$stored,
                         rebuilt = vapply(.chk$scenario, function(s) pc_year2050_undisc("clinical", s), numeric(1)))
      val_stop(max(abs(.cmp$stored - .cmp$rebuilt)) < 1e-6,
               "clinical 2050 undiscounted per-capita reconstruction disagrees with the workbook's stored value") }
  }
}
```

```{r results-builders-undisc, include=FALSE}
## UNDISCOUNTED row builders / scalars -- exact twins of the discounted ones in
## results-builders, differing ONLY in the monetary measures (health is identical):
##   * cost per death averted (undiscounted) = UNDISCOUNTED additional cost
##       (res$incremental_cost) / the SAME cumulative undiscounted deaths averted.
##   * pc1u = average annual UNDISCOUNTED additional cost per capita, 2026-2050
##       = cumulative undiscounted additional cost 2026-2050 / (25 x 2026 population)
##   * pc2u = undiscounted additional cost per capita in 2050 / 2050 population.
pc1u <- function(tab, sc) fmt_usd_pc2(pc_avg_annual_undisc(fam_of(tab), sc))
pc2u <- function(tab, sc) fmt_usd_pc2(pc_year2050_undisc(fam_of(tab), sc))
cpda_undisc <- function(sc, tab) {
  ic <- hv(sc, "incremental_cost", tab); da <- hv(sc, "deaths_averted", tab)
  if (is.finite(ic) && is.finite(da) && da > 0) ic / da else NA_real_
}
ce_row_undisc <- function(tab, sc) cbind(
  impact_row(tab, sc),
  data.table(cpd  = fmt_usd(cpda_undisc(sc, tab)),
             pc1c = pc1u(tab, sc),
             pc2c = pc2u(tab, sc)))

## Undiscounted headline scalars (health scalars are unchanged and reused as-is).
combo_cpda_undisc <- if (has_joint) cpda_undisc(joint_id, met_comb) else NA_real_
combo_cost_undisc <- if (has_joint) hv(joint_id, "incremental_cost", met_comb) else NA_real_
clin_cpda_undisc  <- cpda_undisc(clin_all_id, met_clin)
ph_cpda_undisc    <- cpda_undisc(ph_all_id, met_ph)
```

```{r cascade-data, include=FALSE}
## ---------------------------------------------------------------------------
## 5. ISOLATED 70-30-30 -> 70-70-70 cascade strategy analysis. Read ONLY from
##    the cascade workbook's stored/literal values; never merged with the
##    ordinary results contract. Fail loudly on any structural surprise.
## ---------------------------------------------------------------------------
csh   <- openxlsx::getSheetNames(casc_wb)
.creq <- c("Run_Metadata","Scenario_Catalog","Cascade_Assumptions","Cascade_Trajectory",
           "Annual_Mortality","CVD_40q30","Annual_Cost","Budget_Impact","Cost_Effectiveness")
val_stop(all(.creq %in% csh),
         sprintf("cascade workbook missing sheet(s): %s", paste(setdiff(.creq, csh), collapse = ", ")))
cread <- function(s) as.data.table(openxlsx::read.xlsx(casc_wb, sheet = s))
need_cols <- function(dt, cols, sheet)
  val_stop(all(cols %in% names(dt)),
           sprintf("cascade sheet '%s' missing column(s): %s", sheet, paste(setdiff(cols, names(dt)), collapse = ", ")))

## Cached-value fallback for UNcached formula cells. The cascade FORMULA workbook is the
## structural + methodological source of truth, but openxlsx writes formula cells UNcached:
## its computed cost / cost-effectiveness columns read back as NA (Excel would populate them
## only on open). When a required numeric column is entirely non-finite, fall back to the
## companion literal-value workbook (indonesia_70_30_30_to_70_70_70_cost_value.xlsx) for ONLY
## that sheet's numbers; every structural sheet is still read from the formula workbook. The
## companion is used only after both workbooks are validated to describe the SAME run.
casc_wb_cached <- sub("_formulae\\.xlsx$", ".xlsx", casc_wb)
.casc_same_run <- FALSE
ensure_casc_same_run <- function() {
  if (.casc_same_run) return(invisible(TRUE))
  val_stop(file.exists(casc_wb_cached),
           sprintf("cascade formula workbook has uncached cost cells and its literal-value companion is missing: %s",
                   basename(casc_wb_cached)))
  ## (i) identical Annual_Cost row keys => same scenarios / interventions / years / cost keys.
  keyc <- c("scenario","year","cost_record_id","cost_component_key","cost_join_key")
  kf <- cread("Annual_Cost"); kc <- as.data.table(openxlsx::read.xlsx(casc_wb_cached, sheet = "Annual_Cost"))
  val_stop(all(keyc %in% names(kf)) && all(keyc %in% names(kc)),
           "cannot verify cascade formula/cached same-run: an Annual_Cost key column is missing")
  kfk <- sort(do.call(paste, c(kf[, ..keyc], sep = "")))
  kck <- sort(do.call(paste, c(kc[, ..keyc], sep = "")))
  val_stop(length(kfk) == length(kck) && all(kfk == kck),
           "cascade cached-value workbook does not match the formula workbook (different scenarios / interventions / years / cost keys)")
  ## (ii) matching run provenance metadata.
  mf <- cread("Run_Metadata"); mc <- as.data.table(openxlsx::read.xlsx(casc_wb_cached, sheet = "Run_Metadata"))
  for (it in c("Input workbook","Scenarios costed","Model / pipeline","Location")) {
    vf <- mf[item == it, value]; vc <- mc[item == it, value]
    val_stop(length(vf) && length(vc) && identical(vf[1], vc[1]),
             sprintf("cascade formula vs cached workbook disagree on Run_Metadata '%s' (different run)", it))
  }
  .casc_same_run <<- TRUE
  invisible(TRUE)
}
## Read a cascade sheet preferring the formula workbook; if any required numeric column is
## absent or entirely non-finite there (uncached formulas), take the sheet from the validated
## literal-value companion instead.
cread_num <- function(sheet, num_cols) {
  d <- cread(sheet)
  ok <- all(num_cols %in% names(d)) &&
        all(vapply(num_cols, function(cc) any(is.finite(d[[cc]])), logical(1)))
  if (ok) return(d)
  ensure_casc_same_run()
  as.data.table(openxlsx::read.xlsx(casc_wb_cached, sheet = sheet))
}

## (a) Exactly the two expected scenarios (baseline + the cascade scenario).
CASC_SCEN <- "S_70_30_30_TO_70_70_70"
c_cat <- cread("Scenario_Catalog"); need_cols(c_cat, "scenario", "Scenario_Catalog")
val_stop(setequal(c_cat$scenario, c("baseline", CASC_SCEN)),
         sprintf("cascade Scenario_Catalog must be exactly {baseline, %s}; got {%s}",
                 CASC_SCEN, paste(c_cat$scenario, collapse = ", ")))

## (b) Assumptions -> parameter lookup + metadata.
c_as <- cread("Cascade_Assumptions"); need_cols(c_as, c("parameter_id","value"), "Cascade_Assumptions")
getp  <- function(id) { v <- c_as[parameter_id == id, value][1]; val_stop(!is.na(v), sprintf("cascade assumption '%s' missing", id)); v }
getpn <- function(id) suppressWarnings(as.numeric(getp(id)))
casc_disc  <- getpn("cost_discount_rate")
casc_price <- getpn("cost_price_year")
casc_curr  <- getp("currency")
casc_diag  <- getpn("target_diagnosis_share")
casc_t30   <- getpn("first_target_treatment_conditional")
casc_c30   <- getpn("first_target_control_conditional")
casc_t40   <- getpn("final_target_treatment_conditional")
casc_c40   <- getpn("final_target_control_conditional")
casc_half  <- getpn("treated_uncontrolled_effect_fraction")
casc_eff30 <- getpn("first_target_effective_coverage_exact")
casc_eff40 <- getpn("final_target_effective_coverage_exact")
casc_y0    <- as.integer(getpn("analysis_start_year"))
casc_y1    <- as.integer(getpn("analysis_end_year"))
casc_ytar1 <- as.integer(getpn("first_target_year"))
casc_ytar2 <- as.integer(getpn("final_target_year"))
c_meta <- cread("Run_Metadata")
casc_persp <- { v <- c_meta[item == "Economic perspective", value]; ltx(gsub("_", " ", if (length(v)) v[1] else "health system")) }
val_stop(is.finite(casc_disc) && is.finite(casc_price), "cascade discount rate / price year missing")
## Nested-share + effective-coverage reconciliation (fail loud if the workbook drifts).
val_stop(abs(casc_diag * casc_t30 * casc_c30 - 0.063) < 1e-6, "cascade 2030 controlled share does not reconcile to 6.3%")
val_stop(abs(casc_diag * casc_t40 * casc_c40 - 0.343) < 1e-6, "cascade 2040 controlled share does not reconcile to 34.3%")
.eff_cov <- function(diag, t, ctrl, half) diag * t * ctrl + half * (diag * t - diag * t * ctrl)
val_stop(abs(.eff_cov(casc_diag, casc_t30, casc_c30, casc_half) - casc_eff30) < 1e-6,
         "cascade 2030 effective coverage does not reconcile to the stored exact value")
val_stop(abs(.eff_cov(casc_diag, casc_t40, casc_c40, casc_half) - casc_eff40) < 1e-6,
         "cascade 2040 effective coverage does not reconcile to the stored exact value")

## (c) Hypertension cascade shares (sex-averaged) for the bar figure.
c_tr <- cread("Cascade_Trajectory")
need_cols(c_tr, c("intervention_id","risk_factor_id","sex","year","controlled_share","treated_uncontrolled_share"), "Cascade_Trajectory")
htn_shares <- function(iv, rf) {
  d <- c_tr[intervention_id == iv & risk_factor_id == rf & year %in% c(HZ_START, casc_ytar1, casc_ytar2) &
              !is.na(controlled_share)]
  d[, treated := controlled_share + treated_uncontrolled_share]
  d[, .(controlled = mean(controlled_share), treated = mean(treated)), by = year][order(year)]
}
htn_sh <- htn_shares("I_CVD_PRIMARY", "RF_HTN")
val_stop(nrow(htn_sh) == 3L, "hypertension cascade needs baseline + two milestone years of shares")
## diagnosed row: not modelled at baseline (leave NA); = target at milestone years
htn_sh[, diagnosed := ifelse(year == HZ_START, NA_real_, casc_diag)]
dm_base <- { d <- c_tr[intervention_id == "I_T2D_TREATMENT" & year == HZ_START & !is.na(controlled_share)]
             d[, treated := controlled_share + treated_uncontrolled_share]
             list(treated = mean(d$treated), controlled = mean(d$controlled_share)) }

## (d) Probability-of-death-before-70 series (R-reconciled literal values).
c_q <- cread("CVD_40q30"); need_cols(c_q, c("scenario","year","cvd_40q30_r"), "CVD_40q30")
casc_q <- c_q[, .(scenario, year, q = cvd_40q30_r)][order(scenario, year)]
val_stop(all(is.finite(casc_q$q)) && setequal(casc_q$year, casc_y0:casc_y1),
         "cascade probability-of-death series has gaps or non-finite values")

## (e) Cumulative DISCOUNTED cost levels (baseline vs intervention), de-duplicated.
c_ac <- cread_num("Annual_Cost", c("disc_cost_baseline","disc_cost_scenario","disc_cost_incremental"))
need_cols(c_ac, c("year","disc_cost_baseline","disc_cost_scenario","disc_cost_incremental"), "Annual_Cost")
## Count each shared component once. The formula workbook carries an explicit
## shared_duplicate_count; the literal-value companion omits it (its rows are already the
## de-duplicated set -- each component appears exactly once per year), so a missing/NA count
## is treated as 1. Any residual double-count is still caught by the Budget_Impact
## reconciliation guard below.
if (!"shared_duplicate_count" %in% names(c_ac) || all(is.na(c_ac$shared_duplicate_count)))
  c_ac[, shared_duplicate_count := 1]
c_ac[, w := 1 / pmax(shared_duplicate_count, 1)]              # count shared components once
casc_cost <- c_ac[, .(base = sum(disc_cost_baseline * w), scen = sum(disc_cost_scenario * w),
                      inc  = sum(disc_cost_incremental * w)), by = year][order(year)]
casc_cost[, `:=`(cum_base = cumsum(base), cum_scen = cumsum(scen), cum_inc = cumsum(inc))]
c_bi <- cread_num("Budget_Impact", c("cumulative_disc_incremental_cost","disc_incremental_cost"))
need_cols(c_bi, c("year","cumulative_disc_incremental_cost","disc_incremental_cost","national_population"), "Budget_Impact")
casc_cost <- merge(casc_cost, c_bi[, .(year, bi_cum = cumulative_disc_incremental_cost,
                                       bi_disc_annual = disc_incremental_cost,
                                       bi_pop = national_population)], by = "year")
## ANNUAL (year-specific, NOT cumulative) additional cost per capita = that year's
## discounted additional cost / that year's national population (both authoritative
## Budget_Impact fields). Distinct from the 2026-2050 average shown in the CE tables.
casc_cost[, pc_annual := bi_disc_annual / bi_pop]
val_stop(all(is.finite(casc_cost$pc_annual)), "cascade annual per-capita cost is non-finite")
## Guard: our de-duplicated level difference must reconcile to the authoritative budget sheet.
val_stop(max(abs((casc_cost$cum_scen - casc_cost$cum_base) - casc_cost$bi_cum)) < 1,
         "cascade cost levels (Annual_Cost) do not reconcile to Budget_Impact cumulative additional cost")

## (f) Cumulative deaths (six CVD causes + type-2 diabetes), baseline vs intervention.
c_am <- cread("Annual_Mortality"); need_cols(c_am, c("scenario","year","cause_deaths","base_deaths","deaths_averted"), "Annual_Mortality")
casc_base_d <- c_am[scenario == "baseline",  .(base = sum(cause_deaths)), by = year][order(year)]
casc_scen_d <- c_am[scenario == CASC_SCEN,   .(scen = sum(cause_deaths)), by = year][order(year)]
casc_deaths <- merge(casc_base_d, casc_scen_d, by = "year")
casc_deaths[, `:=`(cum_base = cumsum(base), cum_scen = cumsum(scen))]
casc_deaths[, cum_averted := cum_base - cum_scen]

## (g) Headline metrics from the workbook Cost_Effectiveness (single source of truth).
c_ce <- cread_num("Cost_Effectiveness", c("deaths_averted","disc_incremental_cost","cost_per_death_averted")); need_cols(c_ce, c("scenario","deaths_averted","disc_incremental_cost","cost_per_death_averted"), "Cost_Effectiveness")
c_ce <- c_ce[scenario == CASC_SCEN]
val_stop(nrow(c_ce) == 1L, "cascade Cost_Effectiveness must have exactly the cascade scenario")
casc_da   <- c_ce$deaths_averted[1]
casc_cost_tot <- c_ce$disc_incremental_cost[1]
casc_cpda <- c_ce$cost_per_death_averted[1]
## Reconcile CE against the annual sheets we plotted (fail loud on drift).
val_stop(abs(casc_deaths[year == casc_y1, cum_averted] - casc_da) < 1,
         "cascade deaths averted (Annual_Mortality) do not reconcile to Cost_Effectiveness")
val_stop(abs(casc_cost[year == casc_y1, bi_cum] - casc_cost_tot) < 1e4,
         "cascade cumulative additional cost (Budget_Impact) does not reconcile to Cost_Effectiveness")

## Cumulative snapshots for the results table (2030 / 2040 / 2050).
casc_tab_years <- c(casc_ytar1, casc_ytar2, casc_y1)
casc_snapshot <- function(y) list(
  d_base = casc_deaths[year == y, cum_base], d_scen = casc_deaths[year == y, cum_scen],
  d_av   = casc_deaths[year == y, cum_averted],
  c_base = casc_cost[year == y, cum_base],   c_scen = casc_cost[year == y, cum_scen],
  c_add  = casc_cost[year == y, bi_cum],
  pc_ann = casc_cost[year == y, pc_annual])   # year-specific annual additional cost per capita
```

```{r cascade-data-undisc, include=FALSE}
## ---------------------------------------------------------------------------
## UNDISCOUNTED cascade cost levels -- an exact parallel to the (e)/(g) blocks of
## cascade-data, ADDED alongside the discounted objects (never replacing them).
## The cascade FORMULA workbook is the structural source of truth but writes its
## computed cost cells uncached, so cread_num() transparently takes the numbers
## from the validated literal-value companion (same-run guarded). The undiscounted
## build reuses the SAME de-duplication rule (count each shared component once) and
## the SAME authoritative Budget_Impact fields as the discounted build; the ONLY
## difference is that it sums the undiscounted annual_cost_* columns instead of the
## discounted disc_cost_* columns, and reconciles to the workbook's UNDISCOUNTED
## cumulative_incremental_cost / Cost_Effectiveness incremental_cost.
c_ac_u <- cread_num("Annual_Cost", c("annual_cost_baseline","annual_cost_scenario","annual_cost_incremental"))
need_cols(c_ac_u, c("year","annual_cost_baseline","annual_cost_scenario","annual_cost_incremental"), "Annual_Cost")
if (!"shared_duplicate_count" %in% names(c_ac_u) || all(is.na(c_ac_u$shared_duplicate_count)))
  c_ac_u[, shared_duplicate_count := 1]
c_ac_u[, w := 1 / pmax(shared_duplicate_count, 1)]           # count shared components once
casc_cost_u <- c_ac_u[, .(base = sum(annual_cost_baseline * w), scen = sum(annual_cost_scenario * w),
                          inc  = sum(annual_cost_incremental * w)), by = year][order(year)]
casc_cost_u[, `:=`(cum_base = cumsum(base), cum_scen = cumsum(scen), cum_inc = cumsum(inc))]
c_bi_u <- cread_num("Budget_Impact", c("cumulative_incremental_cost","incremental_cost"))
need_cols(c_bi_u, c("year","cumulative_incremental_cost","incremental_cost","national_population"), "Budget_Impact")
casc_cost_u <- merge(casc_cost_u, c_bi_u[, .(year, bi_cum = cumulative_incremental_cost,
                                             bi_ann = incremental_cost, bi_pop = national_population)], by = "year")
## ANNUAL (year-specific, NOT cumulative) undiscounted additional cost per capita.
casc_cost_u[, pc_annual := bi_ann / bi_pop]
val_stop(all(is.finite(casc_cost_u$pc_annual)), "cascade annual undiscounted per-capita cost is non-finite")
## de-duplicated level difference must reconcile to the authoritative UNDISCOUNTED budget sheet.
val_stop(max(abs((casc_cost_u$cum_scen - casc_cost_u$cum_base) - casc_cost_u$bi_cum)) < 1,
         "cascade undiscounted cost levels do not reconcile to Budget_Impact cumulative additional cost")

## Undiscounted headline metrics from the workbook Cost_Effectiveness (single source of truth).
c_ce_u <- cread_num("Cost_Effectiveness", c("deaths_averted","incremental_cost"))
c_ce_u <- c_ce_u[scenario == CASC_SCEN]
val_stop(nrow(c_ce_u) == 1L, "cascade undiscounted Cost_Effectiveness must have exactly the cascade scenario")
casc_cost_tot_undisc <- c_ce_u$incremental_cost[1]
casc_cpda_undisc <- if (is.finite(casc_da) && casc_da > 0) casc_cost_tot_undisc / casc_da else NA_real_
val_stop(abs(casc_cost_u[year == casc_y1, bi_cum] - casc_cost_tot_undisc) < 1e4,
         "cascade cumulative undiscounted additional cost does not reconcile to Cost_Effectiveness")

## Undiscounted cumulative snapshots for the results table (2030 / 2040 / 2050).
casc_snapshot_undisc <- function(y) list(
  d_base = casc_deaths[year == y, cum_base], d_scen = casc_deaths[year == y, cum_scen],
  d_av   = casc_deaths[year == y, cum_averted],
  c_base = casc_cost_u[year == y, cum_base], c_scen = casc_cost_u[year == y, cum_scen],
  c_add  = casc_cost_u[year == y, bi_cum],
  pc_ann = casc_cost_u[year == y, pc_annual])   # year-specific annual undiscounted additional cost per capita
```

```{r deck-titles, include=FALSE}
## Dynamic, plain-text (no %/$) takeaway titles derived from current outputs.
title_exec <- if (has_joint)
  sprintf("Together, clinical care and prevention could avert about %s deaths by %d",
          fmt_title_num(combo_da), HZ_END) else
  "Clinical care and population prevention each cut premature cardiovascular deaths"
title_cascade_results <- sprintf(
  "A stronger cascade could prevent about %s deaths by %d",
  fmt_title_num(casc_da), casc_y1)

## UNDISCOUNTED-counterpart takeaway titles (plain text; no %/$). Each states the
## finding and makes the undiscounted basis explicit; monetary detail lives in the
## slide body/labels, never in the frame title.
title_exec_undisc <- if (has_joint)
  sprintf("Without discounting, combining care and prevention still averts about %s deaths by %d",
          fmt_title_num(combo_da), HZ_END) else
  "Without discounting, clinical care and prevention each still cut premature cardiovascular deaths"
title_cascade_results_undisc <- sprintf(
  "On an undiscounted basis, a stronger cascade still prevents about %s deaths by %d",
  fmt_title_num(casc_da), casc_y1)
```

```{r subnat-data, include=FALSE}
## ===========================================================================
## ISOLATED SUBNATIONAL (38-province) 70-30-30 -> 70-70-70 cascade analysis.
## Read ONLY from the province-level cascade workbook; never merged with any other
## results. Formula-only fields are reproduced from cached authoritative inputs
## (validated on the fully-cached national cascade workbook); population is the
## workbook's own province_population.
## ===========================================================================
sread <- function(s) as.data.table(openxlsx::read.xlsx(subnat_wb, sheet = s))
.ssh  <- openxlsx::getSheetNames(subnat_wb)
.sreq <- c("Run_Metadata","Cascade_Assumptions","Scenario_Catalog","Cascade_Trajectory",
           "Annual_Mortality","Health_Outcomes","CVD_40q30","Annual_Cost","Budget_Impact",
           "Cost_Effectiveness","Province_Reconciliation")
val_stop(all(.sreq %in% .ssh),
         sprintf("subnational workbook missing sheet(s): %s", paste(setdiff(.sreq, .ssh), collapse = ", ")))
sneed <- function(dt, cols, sheet)
  val_stop(all(cols %in% names(dt)),
           sprintf("subnational sheet '%s' missing column(s): %s", sheet, paste(setdiff(cols, names(dt)), collapse = ", ")))

## Exactly the baseline + cascade scenario.
SUB_SCEN <- "S_70_30_30_TO_70_70_70"
s_cat <- sread("Scenario_Catalog"); sneed(s_cat, "scenario", "Scenario_Catalog")
val_stop(setequal(s_cat$scenario, c("baseline", SUB_SCEN)),
         sprintf("subnational Scenario_Catalog must be exactly {baseline, %s}", SUB_SCEN))

## Assumptions + metadata.
s_as <- sread("Cascade_Assumptions"); sneed(s_as, c("parameter_id","value"), "Cascade_Assumptions")
sgetn <- function(id) suppressWarnings(as.numeric(s_as[parameter_id == id, value][1]))
sub_disc  <- sgetn("cost_discount_rate"); sub_y0 <- as.integer(sgetn("analysis_start_year"))
sub_price <- as.integer(sgetn("cost_price_year"))
sub_ytar1 <- as.integer(sgetn("first_target_year")); sub_ytar2 <- as.integer(sgetn("final_target_year"))
sub_eff30 <- sgetn("first_target_effective_coverage_exact")
sub_eff40 <- sgetn("final_target_effective_coverage_exact")
val_stop(all(is.finite(c(sub_disc, sub_y0, sub_ytar1, sub_ytar2))), "subnational assumptions missing")
s_rmd <- sread("Run_Metadata")
sub_persp <- { v <- s_rmd[item == "Economic perspective", value]; ltx(gsub("_", " ", if (length(v)) v[1] else "health system")) }

## ---- Province roster + BPS(province_code) -> ISO-3166-2 crosswalk -----------
## The offline boundary set (rnaturalearth admin-1) predates the 5 post-2012
## provinces (North Kalimantan; the four new Papua provinces); each is mapped onto
## its PARENT polygon, of which it is an exact geographic subset. Counts are summed
## and rates are population-weighted onto the parent, so the parent region is
## reconstituted exactly. All 38 provinces are shown individually in the Pareto bar.
s_ce <- sread("Cost_Effectiveness"); sneed(s_ce, c("location","province_code"), "Cost_Effectiveness")
prov_meta <- unique(s_ce[, .(location, province_code = as.character(province_code))])
val_stop(uniqueN(prov_meta$location) == 38L && nrow(prov_meta) == 38L,
         sprintf("expected exactly 38 unique provinces; got %d rows / %d locations",
                 nrow(prov_meta), uniqueN(prov_meta$location)))
bps2iso <- c(
  "11"="ID-AC","12"="ID-SU","13"="ID-SB","14"="ID-RI","15"="ID-JA","16"="ID-SS","17"="ID-BE",
  "18"="ID-LA","19"="ID-BB","21"="ID-KR","31"="ID-JK","32"="ID-JB","33"="ID-JT","34"="ID-YO",
  "35"="ID-JI","36"="ID-BT","51"="ID-BA","52"="ID-NB","53"="ID-NT","61"="ID-KB","62"="ID-KT",
  "63"="ID-KS","64"="ID-KI","65"="ID-KI","71"="ID-SA","72"="ID-ST","73"="ID-SN","74"="ID-SG",
  "75"="ID-GO","76"="ID-SR","81"="ID-MA","82"="ID-MU","91"="ID-PA","92"="ID-PB","93"="ID-PA",
  "94"="ID-PA","95"="ID-PA","96"="ID-PB")
prov_meta[, iso := bps2iso[province_code]]
val_stop(!any(is.na(prov_meta$iso)),
         sprintf("province(s) without a boundary mapping: %s", paste(prov_meta[is.na(iso), location], collapse = ", ")))
n_merged_prov <- nrow(prov_meta[, .N, by = iso][N > 1][, .(iso)])   # parent polygons receiving >1 province
n_new_prov    <- sum(prov_meta$province_code %in% c("65","93","94","95","96"))

## Province population by year (authoritative; cached) for weighting/denominators.
s_bi <- sread("Budget_Impact"); sneed(s_bi, c("location","year","province_population"), "Budget_Impact")
prov_pop <- s_bi[, .(location, year, pop = province_population)]

## ---- (5a) Cumulative deaths averted by province, 2025-2050 (6 CVD + T2D) ----
s_ho <- sread("Health_Outcomes"); sneed(s_ho, c("location","scenario","year","modeled_deaths","baseline_deaths"), "Health_Outcomes")
da_prov <- s_ho[scenario == SUB_SCEN, .(da = sum(baseline_deaths - modeled_deaths)), by = location]
val_stop(nrow(da_prov) == 38L && all(is.finite(da_prov$da)), "subnational deaths averted incomplete")
setorder(da_prov, -da)
da_prov[, cum_share := cumsum(da) / sum(da) * 100]
da_prov[, location := factor(location, levels = location)]        # sorted order for the Pareto
sub_da_total <- sum(da_prov$da)
## Reconcile baseline-death aggregation to the workbook's Province_Reconciliation.
s_pr <- sread("Province_Reconciliation"); sneed(s_pr, c("metric","province_sum","national","rel_diff","status"), "Province_Reconciliation")
.prd <- s_pr[grepl("baseline_modeled_deaths", metric)]
sub_recon_reldiff <- if (nrow(.prd)) suppressWarnings(as.numeric(.prd$rel_diff[1])) * 100 else NA_real_
sub_recon_status  <- if (nrow(.prd)) .prd$status[1] else NA_character_

## ---- (5b) CVD primary-care effective coverage, 2025 & 2040 (sex-pooled) -----
s_tr <- sread("Cascade_Trajectory")
sneed(s_tr, c("location","intervention_id","sex","year","scenario_effective_coverage"), "Cascade_Trajectory")
cvp <- s_tr[intervention_id == "I_CVD_PRIMARY" & year %in% c(HZ_START, sub_ytar2)]
## Verify the two sexes are identical at the 70-70-70 target year (so the pooled mean is exact there).
.sxchk <- dcast(cvp[year == sub_ytar2], location ~ sex, value.var = "scenario_effective_coverage")
val_stop(max(abs(.sxchk[[2]] - .sxchk[[3]])) < 1e-9,
         "subnational CVD coverage: the two sexes are not identical at the target year")
cov_prov <- cvp[, .(cov = mean(scenario_effective_coverage)), by = .(location, year)]  # equal adult-population weights

## ---- (5c) % reduction in CVD deaths (six CVD causes; exclude T2D), 2030/2040 -
s_am <- sread("Annual_Mortality"); sneed(s_am, c("location","scenario","year","cause","cause_deaths","base_deaths"), "Annual_Mortality")
CVD_CAUSES <- c("cmd","hhd","hstroke","ihd","istroke","rhd")   # six modelled CVD causes (no dm2)
val_stop(all(CVD_CAUSES %in% unique(s_am$cause)), "subnational Annual_Mortality missing a CVD cause code")
cvd_red <- s_am[scenario == SUB_SCEN & cause %in% CVD_CAUSES & year %in% c(sub_ytar1, sub_ytar2),
                .(base = sum(base_deaths), scen = sum(cause_deaths)), by = .(location, year)]
cvd_red[, pctred := (base - scen) / base * 100]

## ---- (5d) % reduction in probability of CVD death before age 70, 2030/2040 --
s_q <- sread("CVD_40q30"); sneed(s_q, c("location","scenario","year","cvd_40q30_r"), "CVD_40q30")
qb <- s_q[scenario == "baseline",  .(location, year, base_r = cvd_40q30_r)]
qs <- s_q[scenario == SUB_SCEN,    .(location, year, scen_r = cvd_40q30_r)]
q40 <- merge(qb, qs, by = c("location","year"))
val_stop(all(is.finite(q40$base_r)) && all(is.finite(q40$scen_r)) && all(q40$base_r > 0),
         "subnational probability-of-death series has gaps or non-positive baseline")
q40[, pctred := (base_r - scen_r) / base_r * 100]
q40 <- q40[year %in% c(sub_ytar1, sub_ytar2)]

## ---- (5e) Annual DISCOUNTED additional cost per capita, 2030 & 2040 ---------
## Reproduce annual discounted additional cost from cached Annual_Cost inputs
## (formula validated to machine precision on the national cascade workbook).
s_ac <- sread("Annual_Cost")
sneed(s_ac, c("location","scenario","year","frequency_per_year","unit_cost_usd","population_in_need_fraction",
              "r_quantity_scenario","r_quantity_baseline"), "Annual_Cost")
s_ac[, disc := frequency_per_year * unit_cost_usd * population_in_need_fraction *
       (r_quantity_scenario * coverage_scenario - r_quantity_baseline * coverage_baseline) *
       (1 / (1 + sub_disc)^(year - sub_y0))]
cost_prov <- s_ac[, .(disc = sum(disc)), by = .(location, year)]
cost_prov <- merge(cost_prov, prov_pop, by = c("location","year"))
cost_prov[, pc := disc / pop]
cost_prov <- cost_prov[year %in% c(sub_ytar1, sub_ytar2)]
val_stop(all(is.finite(cost_prov$pc)), "subnational per-capita cost has non-finite values")

## Each map-year dataset must cover all 38 provinces exactly once before plotting.
chk38 <- function(dt, yrs, what) for (y in yrs) {
  n <- uniqueN(dt[year == y, location])
  val_stop(n == 38L && !any(duplicated(dt[year == y, location])),
           sprintf("%s: year %d has %d unique provinces (expected 38)", what, y, n))
}
chk38(cov_prov,  c(HZ_START, sub_ytar2), "coverage map")
chk38(cvd_red,   c(sub_ytar1, sub_ytar2), "CVD-death reduction map")
chk38(q40,       c(sub_ytar1, sub_ytar2), "probability-of-death reduction map")
chk38(cost_prov, c(sub_ytar1, sub_ytar2), "per-capita cost map")

## ---- Boundary geometry (offline: rnaturalearth admin-1) + join validation ---
val_stop(requireNamespace("sf", quietly = TRUE) && requireNamespace("rnaturalearth", quietly = TRUE),
         "packages 'sf' and 'rnaturalearth' are required for the subnational maps")
suppressWarnings(suppressMessages({ library(sf); library(rnaturalearth) }))
idn_sf <- rnaturalearth::ne_states(country = "Indonesia", returnclass = "sf")
idn_sf <- idn_sf[, c("iso_3166_2", "name")]; names(idn_sf)[1] <- "iso"
val_stop(all(prov_meta$iso %in% idn_sf$iso),
         sprintf("boundary geometry is missing polygon(s) for iso: %s",
                 paste(setdiff(unique(prov_meta$iso), idn_sf$iso), collapse = ", ")))
n_polygons <- nrow(idn_sf)

## Aggregation helpers province -> boundary polygon (iso).
sum_to_iso <- function(dt, valcol) {           # counts: sum
  d <- merge(dt, prov_meta[, .(location, iso)], by = "location")
  d[, .(value = sum(get(valcol))), by = iso]
}
popw_to_iso <- function(dt, valcol, yr) {      # rates: population-weighted mean at year yr
  w <- prov_pop[year == yr, .(location, pop)]
  d <- merge(dt, prov_meta[, .(location, iso)], by = "location")
  d <- merge(d, w, by = "location")
  d[, .(value = sum(get(valcol) * pop) / sum(pop)), by = iso]
}
ratio_to_iso <- function(dt) {                 # per-capita cost: sum(cost)/sum(pop)
  d <- merge(dt, prov_meta[, .(location, iso)], by = "location")
  d[, .(value = sum(disc) / sum(pop)), by = iso]
}
## Attach values (one row per iso) to geometry, returning an sf for one panel/year.
map_sf <- function(iso_dt) {
  iso_df <- as.data.frame(iso_dt)                     # plain frame -> use merge.sf, not data.table
  m <- merge(idn_sf, iso_df, by = "iso", all.x = TRUE)
  val_stop(!any(is.na(m$value)), "a boundary polygon received no value in a subnational map")
  m
}
## Build a faceted two-year sf: bind per-year sf with a 'panel' label.
maps_two_year <- function(dt_by_year, years, panel_labs, agg = c("sum","popw","ratio"), valcol = "value") {
  agg <- match.arg(agg)
  out <- lapply(seq_along(years), function(i) {
    yr <- years[i]
    dyr <- dt_by_year[year == yr]
    iso_dt <- switch(agg,
                     sum   = sum_to_iso(dyr, valcol),
                     popw  = popw_to_iso(dyr, valcol, yr),
                     ratio = ratio_to_iso(dyr))
    s <- map_sf(iso_dt); s$panel <- panel_labs[i]; s
  })
  do.call(rbind, out)
}
## Common map theme.
theme_map_exec <- function(base = 10) {
  theme_void(base_size = base) +
    theme(legend.position = "bottom",
          legend.title = element_text(size = base - 2, colour = col_wbnavy),
          legend.text  = element_text(size = base - 3),
          legend.key.height = unit(0.28, "cm"), legend.key.width = unit(0.95, "cm"),
          strip.text = element_text(face = "bold", colour = col_wbnavy, size = base),
          plot.margin = margin(2, 2, 2, 2))
}
## Sequential ramps (colour-blind friendly).
ramp_blue <- c("#F7FBFF","#DEEBF7","#9ECAE1","#4292C6","#08519C","#08306B")
ramp_grn  <- c("#F7FCF5","#C7E9C0","#74C476","#31A354","#006D2C")
ramp_org  <- c("#FFF5EB","#FEE6CE","#FDAE6B","#F16913","#D94801","#7F2704")

## Standard caveat pieces reused on every subnational slide (built once).
sub_caveat <- paste0(
  "\\textbf{Preliminary---validate before use:} provincial baseline cascade estimates and ",
  "province-adjusted costs require validation by Indonesia's Ministry of Health and BPJS.")
sub_hetero <- paste0(
  "Provincial CVD treatment coverage is a crude anchor from observed CVD mortality and the modelled ",
  "coverage-effect relationship (extreme values capped); provincial diabetes coverage is scaled in proportion.")
sub_geo_note <- sprintf(
  "Boundaries: offline admin-1 polygons; the %d post-2012 provinces appear within their parent region (an exact geographic subset).",
  n_new_prov)

## ---- Dynamic plain-text takeaway titles (no % or $ for \frametitle) ---------
sub_n50 <- which(da_prov$cum_share >= 50)[1]                 # provinces to reach half the gains
title_sub_pareto <- sprintf(
  "About half of averted deaths come from just %d of Indonesia's 38 provinces", sub_n50)
title_sub_cov <- sprintf(
  "CVD primary-care coverage rises toward a common target in every province by %d", sub_ytar2)
.q40max <- q40[year == sub_ytar2][which.max(pctred)]
title_sub_40q30 <- sprintf(
  "Premature cardiovascular mortality falls in every province by %d", sub_ytar2)
title_sub_cvddeath <- "Cardiovascular deaths fall most where the baseline gaps and burden are largest"
.costmax <- cost_prov[year == sub_ytar2][which.max(pc)]
title_sub_cost <- "The provincial financing requirement grows as coverage scales up"

## ---- Common positive-anchored scale limits for the subnational choropleths -----
## Continuous colour bars must start at the minimum strictly-positive displayed value
## (no detached 0%/US$0 tick). Zeros/negatives are squished to the low end (see each
## map's oob=scales::squish) rather than dropped to NA/grey. Guards the degenerate
## all-non-positive / single-value cases. Applied to ALL subnational 70-70-70 maps.
pos_scale_lims <- function(v) {
  f <- v[is.finite(v)]
  if (!length(f)) return(c(0, 1))
  p  <- f[f > 0]
  lo <- if (length(p)) min(p) else min(f)
  hi <- max(f)
  if (!is.finite(hi) || hi <= lo) hi <- lo + abs(lo) * 1e-6 + 1e-9
  c(lo, hi)
}
```

```{r subnat-data-undisc, include=FALSE}
## ---------------------------------------------------------------------------
## UNDISCOUNTED subnational annual additional cost per capita -- an exact parallel
## to block (5e) of subnat-data, ADDED alongside the discounted objects. The
## discounted per-row cost s_ac$disc is already reproduced from the cached
## Annual_Cost inputs (formula validated to machine precision on the national
## cascade workbook); the undiscounted twin is the SAME reconstruction with the
## 1/(1+r)^(t-y0) discount factor omitted -- nothing else changes (same population
## denominator province_population, same coverage/quantity inputs, same de-dup).
s_ac[, undisc := frequency_per_year * unit_cost_usd * population_in_need_fraction *
       (r_quantity_scenario * coverage_scenario - r_quantity_baseline * coverage_baseline)]
## Internal identity: the undiscounted series is exactly the discounted one without
## the discount factor (fail loud if the reconstruction drifted).
.nz_u <- s_ac[abs(undisc) > 1e-9]
val_stop(max(abs(.nz_u$disc / .nz_u$undisc - 1 / (1 + sub_disc)^(.nz_u$year - sub_y0))) < 1e-6,
         "subnational undiscounted cost is not the discounted series without the discount factor")
cost_prov_u <- s_ac[, .(undisc = sum(undisc)), by = .(location, year)]
cost_prov_u <- merge(cost_prov_u, prov_pop, by = c("location", "year"))
cost_prov_u[, pc := undisc / pop]
cost_prov_u <- cost_prov_u[year %in% c(sub_ytar1, sub_ytar2)]
val_stop(all(is.finite(cost_prov_u$pc)), "subnational undiscounted per-capita cost has non-finite values")
chk38(cost_prov_u, c(sub_ytar1, sub_ytar2), "undiscounted per-capita cost map")
title_sub_cost_undisc <- "The undiscounted provincial financing requirement grows as coverage scales up"
```

# The challenge, and the opportunity

## Most premature cardiovascular deaths in Indonesia can be prevented or treated

\begin{columns}[T]
\begin{column}{0.52\textwidth}
\begin{block}{Cardiovascular disease is Indonesia's largest health burden}
\normalsize
Under current trends, a person reaching age 30 faces roughly a
\hlkey{`r fmt_pct(base_q_2025,1)` chance} of dying from cardiovascular disease
before age 70. Most of this is \textbf{preventable or treatable}.
\end{block}
\end{column}
\begin{column}{0.46\textwidth}
\begin{itemize}\itemsep4pt
  \item Premature CVD deaths (ages 30--69) have risen for two decades and remain high.
  \item Two complementary levers:
  \begin{itemize}
    \item \textcolor{clinicalcol}{\textbf{Clinical care}} --- prevent and better treat established disease.
    \item \textcolor{phcol}{\textbf{Population prevention}} --- tobacco, salt, trans-fat.
  \end{itemize}
  \item This analysis weighs their \textbf{health impact and value for money} together.
\end{itemize}
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} Indonesia CVD model, calibrated to GBD 2023 and UN World Population Prospects 2024; combined results workbook (annual probability-of-death sheet). \textbf{Method:} the probability of death from CVD before age 70 is the chance of dying from the six modelled cardiovascular diseases between exact ages 30 and 70, read from a single-year life table in the `r HZ_START` baseline year.}

## We measured the health gains and the value for money of proven interventions

\begin{block}{General aim}
\normalsize Estimate the \textbf{health impact and value for money} of scaling up evidence-based
\textbf{clinical} and \textbf{public-health} cardiovascular interventions in Indonesia through \textbf{2050}.
\end{block}

\vspace{2pt}
\begin{columns}[T]
\begin{column}{0.33\textwidth}
\begin{clinicalblock}{Aim 1 --- Burden}
\footnotesize Project the CVD burden and the probability of death from CVD before age 70 under the \textbf{current course} and under \textbf{alternative intervention scenarios}.
\end{clinicalblock}
\end{column}
\begin{column}{0.33\textwidth}
\begin{phblock}{Aim 2 --- Value for money}
\footnotesize Estimate \textbf{deaths averted}, mortality reductions, \textbf{additional costs}, and the \textbf{cost per death averted} for each intervention and package.
\end{phblock}
\end{column}
\begin{column}{0.33\textwidth}
\begin{combinedblock}{Aim 3 --- Better together}
\footnotesize Quantify the \textbf{added value of combining} clinical and public-health action, modelled as a single joint run rather than a sum.
\end{combinedblock}
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} study design; Indonesia CVD model. \textbf{Method:} every scenario runs through one projection over `r HZ_START`--`r HZ_END`; the baseline (current course) is the no-new-intervention comparator against which each scenario is measured.}

## `r title_exec`

\begin{columns}[T]
\begin{column}{0.5\textwidth}
```{r exec-block, results='asis'}
if (has_joint) {
  cat(sprintf(paste0(
    "\\begin{combinedblock}{Combining clinical + public health is the big prize}\n",
    "\\footnotesize Together they could avert about \\hlkey{%s deaths} over %d--%d, ",
    "cut the probability of death from CVD before age 70 in %d by \\hlkey{%s}, at about \\hlkey{%s per death averted}.\n",
    "\\end{combinedblock}\n"),
    fmt_short(combo_da), HZ_START, HZ_END, HZ_END, fmt_pct(combo_qred,0), fmt_usd(combo_cpda)))
} else {
  cat(paste0("\\begin{block}{Two complementary levers}\n\\footnotesize This run reports the ",
             "clinical and public-health families; the joint package was not part of this ",
             "configuration.\n\\end{block}\n"))
}
```
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep3pt
```{r exec-bullets, results='asis'}
if (has_clin)
  cat(sprintf("  \\item \\textcolor{clinicalcol}{\\textbf{All clinical interventions}}: $\\approx$%s deaths averted; probability of death from CVD before age 70 down %s by %d.\n",
              fmt_short(clin_da), fmt_pct(clin_qred,0), HZ_END))
if (has_ph)
  cat(sprintf("  \\item \\textcolor{phcol}{\\textbf{All public-health policies}}: $\\approx$%s deaths averted at about %s per death averted.\n",
              fmt_short(ph_da), fmt_usd(ph_cpda)))
if (!is.na(top_clin_sc) && !is.na(top_ph_sc))
  cat(sprintf("  \\item \\textbf{Most impactful single levers}: \\textcolor{clinicalcol}{%s} and \\textcolor{phcol}{%s}.\n",
              lbl(top_clin_sc, met_clin), lbl(top_ph_sc, met_ph)))
if (has_joint)
  cat(sprintf("  \\item \\textbf{Affordability}: the combined package costs about %s per death averted.\n",
              fmt_usd(combo_cpda)))
```
\end{itemize}
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} results contract (\texttt{09\_deck\_results.rds}) + combined workbook. \textbf{Method:} deaths averted cumulative `r HZ_START`--`r HZ_END` (six CVD causes $+$ type-2 diabetes); cost per death averted $=$ discounted additional cost $\div$ deaths averted (`r price_year` US\$). \costcaveat}

## `r title_exec_undisc`

\begin{columns}[T]
\begin{column}{0.5\textwidth}
```{r exec-block-undisc, results='asis'}
if (has_joint) {
  cat(sprintf(paste0(
    "\\begin{combinedblock}{Combining clinical + public health, valued undiscounted}\n",
    "\\footnotesize Together they could avert about \\hlkey{%s deaths} over %d--%d, ",
    "cut the probability of death from CVD before age 70 in %d by \\hlkey{%s}, at about \\hlkey{%s per death averted (undiscounted)}.\n",
    "\\end{combinedblock}\n"),
    fmt_short(combo_da), HZ_START, HZ_END, HZ_END, fmt_pct(combo_qred,0), fmt_usd(combo_cpda_undisc)))
} else {
  cat(paste0("\\begin{block}{Two complementary levers}\n\\footnotesize This run reports the ",
             "clinical and public-health families; the joint package was not part of this ",
             "configuration.\n\\end{block}\n"))
}
```
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep3pt
```{r exec-bullets-undisc, results='asis'}
if (has_clin)
  cat(sprintf("  \\item \\textcolor{clinicalcol}{\\textbf{All clinical interventions}}: $\\approx$%s deaths averted; probability of death from CVD before age 70 down %s by %d.\n",
              fmt_short(clin_da), fmt_pct(clin_qred,0), HZ_END))
if (has_ph)
  cat(sprintf("  \\item \\textcolor{phcol}{\\textbf{All public-health policies}}: $\\approx$%s deaths averted at about %s per death averted (undiscounted).\n",
              fmt_short(ph_da), fmt_usd(ph_cpda_undisc)))
if (!is.na(top_clin_sc) && !is.na(top_ph_sc))
  cat(sprintf("  \\item \\textbf{Most impactful single levers}: \\textcolor{clinicalcol}{%s} and \\textcolor{phcol}{%s}.\n",
              lbl(top_clin_sc, met_clin), lbl(top_ph_sc, met_ph)))
if (has_joint)
  cat(sprintf("  \\item \\textbf{Affordability (undiscounted)}: the combined package costs about %s per death averted.\n",
              fmt_usd(combo_cpda_undisc)))
```
\end{itemize}
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} results contract (\texttt{09\_deck\_results.rds}) + combined workbook. \textbf{Method:} deaths averted cumulative `r HZ_START`--`r HZ_END` (six CVD causes $+$ type-2 diabetes); \textbf{undiscounted} cost per death averted $=$ \textbf{undiscounted} additional cost $\div$ deaths averted (`r price_year` US\$, `r cost_perspective`). Health results are identical to the discounted slide. \costcaveat}

## The model follows how prevention and treatment change disease and survival

\begin{columns}[T]
\begin{column}{0.54\textwidth}
\footnotesize
\begin{enumerate}\itemsep3pt
  \item \textbf{People \& demography.} Indonesia's population is projected by age and sex (UN World Population Prospects 2024).
  \item \textbf{Starting health.} Disease and death rates are set to Indonesia's data (GBD 2023) and calibrated to match observed levels.
  \item \textbf{Year-by-year simulation.} Each year people can develop cardiovascular disease or diabetes, survive, or die.
  \item \textbf{Interventions.} Clinical and public-health scenarios change how many people develop disease (new cases) and how many survive it.
  \item \textbf{Outcomes.} The model reports deaths, healthy years lost or gained (DALYs), the probability of death from CVD before age 70, costs, and value for money.
\end{enumerate}
\end{column}
\begin{column}{0.44\textwidth}
\begin{block}{One consistent engine}
\footnotesize Every scenario runs through the \textbf{same} projection, so results are directly comparable and the \textbf{combined} package is modelled as one run --- not a sum of separate estimates.
\end{block}
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} Indonesia CVD model engine; GBD 2023 (starting rates); UN World Population Prospects 2024 (demography, life expectancy). \textbf{Method:} a discrete-time state-transition simulation by single-year age and sex, calibrated to match observed GBD levels using a standard numerical optimiser.}

## Interventions work by preventing new disease and by helping people survive it

\vspace{-2pt}
\begin{center}
\begin{tikzpicture}[
   >=Latex, node distance = 16mm and 30mm, font=\footnotesize,
   state/.style = {rectangle, rounded corners, draw=wbnavy, thick, fill=wbnavy!6,
                   minimum width=26mm, minimum height=11mm, align=center},
   death/.style = {rectangle, rounded corners, draw=black!55, thick, fill=black!4,
                   minimum width=26mm, minimum height=10mm, align=center}]
  \node[state] (well) {\textbf{Healthy}\\[-1pt]\scriptsize (no CVD)};
  \node[state, right=of well] (sick) {\textbf{Living with CVD}\\[-1pt]\scriptsize (or diabetes)};
  \node[death, right=of sick] (cvd)  {\textbf{CVD death}};
  \node[death, below=14mm of well, xshift=23mm] (bg) {Other-cause (background) death};

  \draw[->, very thick, incidencecol]
       (well) -- node[above, midway, text=incidencecol]{\textbf{new CVD}}
                  node[below, midway, text=incidencecol, font=\scriptsize]{new cases} (sick);
  \draw[->, very thick, casefatalcol]
       (sick) -- node[above, midway, text=casefatalcol]{\textbf{dies of CVD}}
                  node[below, midway, text=casefatalcol, font=\scriptsize]{risk of dying with CVD} (cvd);
  \draw[->, thick, bgmortcol] (well) to[bend right=12] (bg);
  \draw[->, thick, bgmortcol] (sick) to[bend left=12]  (bg);
\end{tikzpicture}
\end{center}

\vspace{-2pt}
\begin{columns}[T]
\begin{column}{0.33\textwidth}\footnotesize
\textcolor{incidencecol}{\textbf{Fewer people fall ill}}\\ \scriptsize Primary prevention \& \emph{all} public-health policies reduce \textbf{new cases}.
\end{column}
\begin{column}{0.34\textwidth}\footnotesize
\textcolor{casefatalcol}{\textbf{More people survive}}\\ \scriptsize Secondary prevention, treatment \& tobacco policy lower the \textbf{risk of dying after developing CVD}.
\end{column}
\begin{column}{0.31\textwidth}\footnotesize
\textcolor{bgmortcol}{\textbf{Background mortality}}\\ \scriptsize Other-cause deaths from demography; unchanged by these interventions.
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} Indonesia CVD model engine (\texttt{06\_run\_scenarios\_indonesia\_fair.R}). \textbf{Method:} interventions act on two transitions --- new cases (Healthy $\rightarrow$ Living with CVD) and the risk of dying after developing CVD (Living with CVD $\rightarrow$ death); no recovery/remission is modelled, a conservative assumption.}

# The interventions we tested

## Clinical care prevents disease and helps people survive it

```{r tbl-clin-interventions, results='asis'}
## Built from the ACTIVE clinical scenarios in this run (children/aggregate
## excluded); target-population / effect notes come from a small presentation map.
clin_pop <- c(
  I_CVD_PRIMARY = "Adults at cardiovascular risk", I_T2D_TREATMENT = "People with type-2 diabetes",
  I_IHD_SECONDARY = "Heart-attack (IHD) survivors", I_STROKE_SECONDARY = "Stroke survivors",
  I_HF_CHRONIC = "Chronic heart-failure patients", I_HF_ACUTE = "Acute heart-failure patients",
  I_RHD_SECONDARY = "Rheumatic heart disease patients")
clin_acts <- c(I_CVD_PRIMARY = "\\textcolor{incidencecol}{Fewer new cases}")
clin_disp <- top_level_rows(met_clin, clin_all_id)[scenario != clin_all_id]
if (nrow(clin_disp)) {
  pluck <- function(vec, key, default) if (key %in% names(vec)) vec[[key]] else default
  cd <- data.table(
    Intervention = vapply(clin_disp$scenario, function(s) lbl(s, met_clin), character(1)),
    Population   = vapply(clin_disp$scenario, function(s) pluck(clin_pop, s, "Target patient group"), character(1)),
    Acts         = vapply(clin_disp$scenario, function(s)
                     pluck(clin_acts, s, "\\textcolor{casefatalcol}{Better survival}"), character(1)))
  print(kbl(cd, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=c("Intervention","Target population","Model effect"), align="lll") |>
    kable_styling(latex_options="hold_position", font_size=8.5) |>
    column_spec(1, bold=TRUE, width="4.6cm") |>
    column_spec(2, width="6.4cm") |> column_spec(3, width="3.0cm") |>
    row_spec(0, bold=TRUE))
  reg_tbl("tbl-clin-interventions",
          "Clinical care prevents disease and helps people survive it",
          cd, c("Intervention", "Target population", "Model effect"))
} else cat("\\begin{center}\\large No clinical interventions selected for this run.\\end{center}")
```

\vspace{2pt}
\footnotesize \textbf{Implementation scenario (all clinical interventions):} coverage scales up \textbf{from a low base toward 80\%} of the people who need each service between `r HZ_START` and `r HZ_END`. Primary prevention lowers who develops CVD (new cases); secondary prevention and treatment improve survival.

\srcmethod{\textbf{Source:} clinical results workbook, sheets \texttt{Selected\_Interventions} and \texttt{Scenario\_Catalog}. \textbf{Method:} the table lists the clinical interventions active in this run (excluded interventions never appear); effect sizes come from the FAIR Choices intervention taxonomy.}

## Population policies cut tobacco, salt, and trans-fat risk across the whole population

```{r tbl-ph-interventions, results='asis'}
ph_meta <- data.table(
  scenario = c("I_PH_TOBACCO_TAX","I_PH_TOBACCO_POLICY","I_PH_SALT_POLICY","I_PH_TFA_POLICY",
               "I_PH_SALT_REFORM","I_PH_SALT_FOPL","I_PH_SALT_MEDIA","I_PH_SALT_ENV",
               "I_PH_TOB_CLEAN_AIR","I_PH_TOB_MEDIA","I_PH_TOB_AD_BAN"),
  Risk  = c("Tobacco","Tobacco","Dietary salt/sodium","Industrial trans-fat",
            "Dietary salt/sodium","Dietary salt/sodium","Dietary salt/sodium","Dietary salt/sodium",
            "Tobacco","Tobacco","Tobacco"),
  Acts  = c("\\textcolor{incidencecol}{New cases} + \\textcolor{casefatalcol}{survival}",
            "\\textcolor{incidencecol}{New cases} + \\textcolor{casefatalcol}{survival}",
            "\\textcolor{incidencecol}{New cases}","\\textcolor{incidencecol}{New cases}",
            "\\textcolor{incidencecol}{New cases}","\\textcolor{incidencecol}{New cases}",
            "\\textcolor{incidencecol}{New cases}","\\textcolor{incidencecol}{New cases}",
            "\\textcolor{incidencecol}{New cases} + \\textcolor{casefatalcol}{survival}",
            "\\textcolor{incidencecol}{New cases} + \\textcolor{casefatalcol}{survival}",
            "\\textcolor{incidencecol}{New cases} + \\textcolor{casefatalcol}{survival}"),
  Lever = c("Raise tax to 75\\% of retail price","Clean-air law, mass media, ad ban",
            "Reformulation, labelling, media, procurement","Best-practice industrial TFA ban",
            "Mandatory reformulation","Front-of-pack labelling","Mass-media campaign","Public-institution environment",
            "Smoke-free / clean-air law","Mass-media campaign","Advertising ban"))
ph_disp <- top_level_rows(met_ph, ph_all_id)[scenario != ph_all_id]
if (nrow(ph_disp)) {
  pd <- merge(ph_disp[, .(scenario)], ph_meta, by = "scenario", all.x = TRUE, sort = FALSE)
  pd[, Policy := vapply(scenario, function(s) lbl(s, met_ph), character(1))]
  pd[is.na(Risk),  Risk  := "Behavioural risk factor"]
  pd[is.na(Acts),  Acts  := "\\textcolor{incidencecol}{New cases}"]
  pd[is.na(Lever), Lever := "Fiscal / regulatory measure"]
  print(kbl(pd[, .(Policy, Risk, Acts, Lever)], format="latex", booktabs=TRUE, escape=FALSE,
      col.names=c("Policy","Risk factor","Model effect","Implementation lever"), align="llll") |>
    kable_styling(latex_options="hold_position", font_size=7.8) |>
    column_spec(1, bold=TRUE, width="2.9cm") |> column_spec(2, width="2.3cm") |>
    column_spec(3, width="3.1cm") |> column_spec(4, width="5.7cm") |>
    row_spec(0, bold=TRUE))
  reg_tbl("tbl-ph-interventions",
          "Population policies cut tobacco, salt, and trans-fat risk across the whole population",
          pd[, .(Policy, Risk, Acts, Lever)],
          c("Policy", "Risk factor", "Model effect", "Implementation lever"))
} else cat("\\begin{center}\\large No public-health policies selected for this run.\\end{center}")
```

\vspace{1pt}
\footnotesize Policies reduce exposure (smoking, salt, trans-fat intake), lowering the number of \textbf{new CVD cases}. \textbf{Tobacco} policies additionally improve \textbf{survival} after CVD. Packages are modelled as single joint runs of their component measures.

\srcmethod{\textbf{Source:} public-health results workbook, sheets \texttt{Selected\_Interventions} / \texttt{Policy\_Levers} and \texttt{Scenario\_Catalog}. \textbf{Method:} the table lists the policies active in this run; effect sizes come from published risk-factor evidence stored in the workbook.}

## Every input is traceable to a recognised data source

```{r tbl-sources}
src <- data.table(
  Component = c("Population \\& demography","Cause-specific mortality \\& disease rates",
                "Disability weights (healthy years lost)","Clinical intervention effects",
                "Public-health effect sizes","Intervention \\& policy costs",
                "International CVD mortality comparison"),
  Source = c("UN World Population Prospects","GBD (Indonesia)","GBD",
             "FAIR Choices intervention taxonomy","Published risk-factor evidence (in workbook)",
             "Programme \\& policy unit costs (in workbook)",
             "WPP-scaled ASEAN comparison series"),
  Years = c("2024 revision","2023","2023","current","current",
            sprintf("price year %d", price_year), "1995--2050"),
  Use = c("Population, background mortality, life expectancy","Baseline burden \\& calibration targets",
          "Healthy years lost or gained (DALYs)","New-case \\& survival effects","New-case \\& survival effects",
          "Budget impact \\& value for money",
          "Benchmarking Indonesia vs.\\ neighbours"))
reg_tbl("tbl-sources", "Every input is traceable to a recognised data source",
        src, c("Model component", "Primary source", "Years / version", "Use in model"))
kbl(src, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=c("Model component","Primary source","Years / version","Use in model"), align="llll") |>
  kable_styling(latex_options=c("hold_position"), font_size=6.6) |>
  column_spec(1, bold=TRUE, width="3.2cm") |> column_spec(2, width="4.0cm") |>
  column_spec(3, width="1.9cm") |> column_spec(4, width="4.7cm") |>
  row_spec(0, bold=TRUE)
```

\vspace{1pt}
\scriptsize Population: UN World Population Prospects 2024; epidemiology: GBD 2023 [@GBDCVD2025; @Schumacher2024GBD2021].

\srcmethod{\textbf{Source:} modelling pipeline and input workbooks. \textbf{Method:} these inputs drive the demographic projection, calibration targets, and effect sizes; effect-size and cost citations are in the workbooks.}

## Costs depend on who needs care, how many receive it, and local prices

\footnotesize We estimate the \textbf{additional health-system resources required to expand effective coverage}, using a costing approach adapted from Disease Control Priorities (DCP) [@Watkins2020ResourceRequirements].

<!-- \footnotesize We estimate costs with an intervention-based approach adapted from the Disease Control Priorities (DCP) cost model of @Watkins2020ResourceRequirements, applied here from a \textbf{health-system perspective} with \textbf{annual} projections. Each service is costed as: \emph{who needs it} $\times$ \emph{how much of that need is met} $\times$ \emph{services per person per year} $\times$ \emph{Indonesia-adjusted unit cost}. -->

\vspace{2pt}
\begin{center}
\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[
   >=Latex, node distance = 5mm, font=\scriptsize,
   costbox/.style = {rectangle, rounded corners, draw=wbnavy, thick, fill=wbnavy!6,
                  text width=22mm, minimum height=14mm, align=center},
   costout/.style = {rectangle, rounded corners, draw=clinicalcol, very thick, fill=clinicalcol!8,
                  text width=27mm, minimum height=14mm, align=center}]
  \node[costbox] (pin)  {\textbf{People who need}\\ the service};
  \node[costbox, right=of pin] (cov) {\textbf{Current coverage}\\ \& scale-up target};
  \node[costbox, right=of cov] (freq){\textbf{Services}\\ per person / year};
  \node[costbox, right=of freq] (uc) {\textbf{Indonesia unit}\\ cost};
  \node[costbox, right=of uc] (ann) {\textbf{Annual cost}\\ baseline \& scenario};
  \node[costout,  right=of ann] (res){\textbf{Additional \& discounted cost;} cost per person; cost per death prevented};
  \draw[->,thick] (pin)--(cov); \draw[->,thick] (cov)--(freq); \draw[->,thick] (freq)--(uc);
  \draw[->,thick] (uc)--(ann);  \draw[->,thick, clinicalcol] (ann)--(res);
\end{tikzpicture}}
\end{center}

\vspace{2pt}
<!-- \footnotesize The DCP paper supplies the \textbf{conceptual framework} --- unit costs, demographic/epidemiological estimates of the people in need, baseline coverage, and health-system delivery costs. This project \textbf{adapts} that framework to Indonesia and does not claim to reproduce the paper exactly. -->

\footnotesize\textbf{What this tells policymakers:} the analysis estimates how much additional health-system spending is required to move from today's coverage toward universal, high-quality access—and relates that investment to the health gains achieved.

\srcmethod{\textbf{Source:} intervention-based costing adapted from Watkins et al. (2020); unit costs and coverage from the input workbooks; demography from UN WPP 2024. \textbf{Method:} annual cost = people in need $\times$ coverage $\times$ annual service frequency $\times$ Indonesia-adjusted unit cost; additional cost = scenario minus baseline; equations in the appendix. \costcaveat}

# What the interventions deliver, and at what cost

## Scaling up clinical care sharply cuts premature cardiovascular deaths

```{r tbl-clin-ce, results='asis'}
ft <- family_table(met_clin, clin_all_id, ce_row)
if (nrow(ft$tb)) {
  k <- kbl(ft$tb, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=mkhdr(c("Clinical intervention","Deaths averted\n(cumulative)","Prob. of CVD death\nbefore age 70, 2050",
                  "Reduction in that\nprobability, 2050","Deaths reduction\n2050","Cost per\ndeath averted",
                  "Avg. annual add'l\ncost/capita, 2026-50 (disc.)","Add'l cost/capita\nin 2050 (disc.)")),
      align="lrrrrrrr") |>
    kable_styling(latex_options=c("hold_position","scale_down"), font_size=8.5) |>
    column_spec(1, bold=TRUE) |> row_spec(0, bold=TRUE)
  if (!is.na(ft$agg_idx)) k <- k |> row_spec(ft$agg_idx, bold=TRUE, background="#EAF1FB")
  print(k)
  reg_tbl("tbl-clin-ce",
          "Scaling up clinical care sharply cuts premature cardiovascular deaths",
          ft$tb,
          c("Clinical intervention","Deaths averted\n(cumulative)","Prob. of CVD death\nbefore age 70, 2050",
            "Reduction in that\nprobability, 2050","Deaths reduction\n2050","Cost per\ndeath averted",
            "Avg. annual add'l\ncost/capita, 2026-50 (disc.)","Add'l cost/capita\nin 2050 (disc.)"))
  if (ft$n_omitted > 0)
    cat(sprintf("\n\n\\vspace{1pt}\\scriptsize %d additional clinical intervention(s) omitted for space; see the appendix crosswalk and the workbook.\n", ft$n_omitted))
} else cat("\\begin{center}\\large No clinical interventions selected for this run.\\end{center}")
```

\srcmethod{\textbf{Source:} clinical cost workbook (\texttt{indonesia\_model\_cost\_value\_formulae.xlsx}, sheet \texttt{Budget\_Impact}); population is the workbook's own \texttt{national\_population}. \textbf{Method:} cost per death averted $=$ discounted additional cost $\div$ deaths averted (`r fmt_pct(disc_rate*100,0)`/yr, `r price_year` US\$, `r cost_perspective`). \textbf{Avg. annual additional cost per capita, 2026--`r HZ_END`} $=$ cumulative discounted additional cost over 2026--`r HZ_END` $\div$ (25 $\times$ `r HZ_START+1L` population). \textbf{Additional cost per capita in `r HZ_END`} $=$ discounted additional cost in `r HZ_END` $\div$ `r HZ_END` population. \costcaveat}

## Undiscounted costs are higher, yet clinical care still sharply cuts premature deaths

```{r tbl-clin-ce-undisc, results='asis'}
## Undiscounted counterpart of tbl-clin-ce: identical health columns; every monetary
## column is undiscounted (ce_row_undisc). Costs from the clinical workbook's own
## undiscounted incremental_cost / incremental_cost_per_capita.
ftu <- family_table(met_clin, clin_all_id, ce_row_undisc)
if (nrow(ftu$tb)) {
  k <- kbl(ftu$tb, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=mkhdr(c("Clinical intervention","Deaths averted\n(cumulative)","Prob. of CVD death\nbefore age 70, 2050",
                  "Reduction in that\nprobability, 2050","Deaths reduction\n2050","Cost per death\naverted (undisc.)",
                  "Avg. annual add'l\ncost/capita, 2026-50 (undisc.)","Add'l cost/capita\nin 2050 (undisc.)")),
      align="lrrrrrrr") |>
    kable_styling(latex_options=c("hold_position","scale_down"), font_size=8.5) |>
    column_spec(1, bold=TRUE) |> row_spec(0, bold=TRUE)
  if (!is.na(ftu$agg_idx)) k <- k |> row_spec(ftu$agg_idx, bold=TRUE, background="#EAF1FB")
  print(k)
  reg_tbl("tbl-clin-ce-undisc",
          "Undiscounted costs are higher, yet clinical care still sharply cuts premature deaths",
          ftu$tb,
          c("Clinical intervention","Deaths averted\n(cumulative)","Prob. of CVD death\nbefore age 70, 2050",
            "Reduction in that\nprobability, 2050","Deaths reduction\n2050","Cost per death\naverted (undisc.)",
            "Avg. annual add'l\ncost/capita, 2026-50 (undisc.)","Add'l cost/capita\nin 2050 (undisc.)"))
  if (ftu$n_omitted > 0)
    cat(sprintf("\n\n\\vspace{1pt}\\scriptsize %d additional clinical intervention(s) omitted for space; see the appendix crosswalk and the workbook.\n", ftu$n_omitted))
} else cat("\\begin{center}\\large No clinical interventions selected for this run.\\end{center}")
```

\srcmethod{\textbf{Source:} clinical cost workbook (\texttt{indonesia\_model\_cost\_value\_formulae.xlsx}, sheet \texttt{Budget\_Impact}); population is the workbook's own \texttt{national\_population}. \textbf{Method (undiscounted):} cost per death averted $=$ \textbf{undiscounted} additional cost $\div$ deaths averted (`r price_year` US\$, `r cost_perspective`). \textbf{Avg. annual additional cost per capita, 2026--`r HZ_END`} $=$ cumulative \textbf{undiscounted} additional cost over 2026--`r HZ_END` $\div$ (25 $\times$ `r HZ_START+1L` population). \textbf{Additional cost per capita in `r HZ_END`} $=$ \textbf{undiscounted} additional cost in `r HZ_END` $\div$ `r HZ_END` population. Health results are identical to the discounted slide. \costcaveat}

```{r clin-budget-data, include=FALSE}
## ---------------------------------------------------------------------------
## Clinical budget impact over time, decomposed by clinical intervention.
## Source of truth: the CLINICAL cost workbook (Annual_Cost) - the all-clinical
## scenario's discounted additional cost, keyed by (year, intervention_id), with
## the workbook's shared-cost rule (count each shared component once). Population
## is the workbook's own national_population. Two cumulative measures:
##   left  = cumulative DISCOUNTED total additional cost through year t
##   right = that cumulative total / national_population(t)  (per capita through t)
## QA: intervention-level totals must reconcile, by year, to the all-clinical
## Budget_Impact totals (annual and cumulative).
## ---------------------------------------------------------------------------
.cb_ac <- read_clin_ac()
.cb_bi <- read_clin_bi()
need_ac <- c("scenario","year","intervention_id","disc_cost_incremental","shared_duplicate_count")
val_stop(all(need_ac %in% names(.cb_ac)), "clinical Annual_Cost missing a column needed for the budget figure")
val_stop("all" %in% .cb_ac$scenario, "clinical Annual_Cost has no all-clinical scenario for the budget figure")
.cb_ac[, w := 1 / pmax(shared_duplicate_count, 1)]     # shared components counted once
clin_bud <- .cb_ac[scenario == "all",
                   .(disc = sum(disc_cost_incremental * w)), by = .(intervention_id, year)]
setorder(clin_bud, intervention_id, year)
clin_bud[, cum := cumsum(disc), by = intervention_id]
clin_pop <- .cb_bi[scenario == "all", .(year, pop = national_population)]
clin_bud <- merge(clin_bud, clin_pop, by = "year")
clin_bud[, cum_pc := cum / pop]                        # cumulative discounted cost per capita through t

## QA reconciliation to the all-clinical Budget_Impact (annual + cumulative).
.qa_ann <- merge(clin_bud[, .(s = sum(disc)), by = year],
                 .cb_bi[scenario == "all", .(year, bi = disc_incremental_cost)], by = "year")
val_stop(max(abs(.qa_ann$s - .qa_ann$bi)) < 1,
         "clinical budget figure: intervention sums do not reconcile to all-clinical annual cost")
.qa_cum <- merge(clin_bud[, .(s = sum(cum)), by = year],
                 .cb_bi[scenario == "all", .(year, bic = cumulative_disc_incremental_cost)], by = "year")
val_stop(max(abs(.qa_cum$s - .qa_cum$bic)) < 1,
         "clinical budget figure: intervention sums do not reconcile to all-clinical cumulative cost")
val_stop(setequal(clin_bud$year, HZ_START:HZ_END), "clinical budget figure: year coverage is not 2025-2050")

## Plain-language labels + stacking order (largest cumulative contribution first).
.ord <- clin_bud[year == HZ_END, .(cum), by = intervention_id][order(-cum)]
.ord[, lab := vapply(intervention_id, function(s) lbl(s, met_clin), character(1))]
lev <- .ord$lab
clin_bud[, ivlab := vapply(intervention_id, function(s) lbl(s, met_clin), character(1))]
clin_bud[, ivlab := factor(ivlab, levels = rev(lev))]  # largest at bottom of the stack
## Accessible clinical palette (blue-anchored), one colour per intervention, shared across panels.
clin_bud_pal <- setNames(
  c("#08306B","#2171B5","#6BAED6","#B2182B","#EF8A62","#00897B","#66C2A5","#B39DDB")[seq_len(nrow(.ord))],
  .ord$lab)
## Long form: two panels.
cb_long <- rbind(
  clin_bud[, .(year, ivlab, value = cum / 1e9,
               panel = "Cumulative discounted additional cost (US$ billions)")],
  clin_bud[, .(year, ivlab, value = cum_pc,
               panel = "Cumulative discounted additional cost per capita (US$)")])
cb_long[, panel := factor(panel, levels = c("Cumulative discounted additional cost (US$ billions)",
                                            "Cumulative discounted additional cost per capita (US$)"))]
## Takeaway title (plain text; no % or $ so it survives \frametitle).
clin_top_lab   <- .ord$lab[1]
clin_top_share <- .ord$cum[1] / sum(.ord$cum) * 100
title_clin_budget <- sprintf(
  "%s drives most of Indonesia's clinical cardiovascular budget through %d",
  clin_top_lab, HZ_END)
```

```{r clin-budget-data-undisc, include=FALSE}
## ---------------------------------------------------------------------------
## UNDISCOUNTED clinical budget impact over time -- an exact twin of clin-budget-data
## using the workbook's UNDISCOUNTED annual_cost_incremental (shared components counted
## once via the same w). Reuses the discounted figure's palette (clin_bud_pal) and
## stacking order (lev) so the two figures are visually comparable; only the monetary
## magnitudes differ. QA: intervention-level totals reconcile, by year, to the
## all-clinical Budget_Impact UNDISCOUNTED annual + cumulative totals.
## ---------------------------------------------------------------------------
need_ac_u <- c("scenario","year","intervention_id","annual_cost_incremental")
val_stop(all(need_ac_u %in% names(.cb_ac)), "clinical Annual_Cost missing a column needed for the undiscounted budget figure")
clin_bud_u <- .cb_ac[scenario == "all",
                     .(undisc = sum(annual_cost_incremental * w)), by = .(intervention_id, year)]
setorder(clin_bud_u, intervention_id, year)
clin_bud_u[, cum := cumsum(undisc), by = intervention_id]
clin_bud_u <- merge(clin_bud_u, clin_pop, by = "year")     # clin_pop (national_population) from clin-budget-data
clin_bud_u[, cum_pc := cum / pop]                          # cumulative undiscounted cost per capita through t
## QA reconciliation to the all-clinical Budget_Impact UNDISCOUNTED annual + cumulative.
.qa_ann_u <- merge(clin_bud_u[, .(s = sum(undisc)), by = year],
                   .cb_bi[scenario == "all", .(year, bi = incremental_cost)], by = "year")
val_stop(max(abs(.qa_ann_u$s - .qa_ann_u$bi)) < 1,
         "undiscounted clinical budget figure: intervention sums do not reconcile to all-clinical annual cost")
.qa_cum_u <- merge(clin_bud_u[, .(s = sum(cum)), by = year],
                   .cb_bi[scenario == "all", .(year, bic = cumulative_incremental_cost)], by = "year")
val_stop(max(abs(.qa_cum_u$s - .qa_cum_u$bic)) < 1,
         "undiscounted clinical budget figure: intervention sums do not reconcile to all-clinical cumulative cost")
val_stop(setequal(clin_bud_u$year, HZ_START:HZ_END), "undiscounted clinical budget figure: year coverage is not 2025-2050")
clin_bud_u[, ivlab := vapply(intervention_id, function(s) lbl(s, met_clin), character(1))]
clin_bud_u[, ivlab := factor(ivlab, levels = rev(lev))]    # same stacking order + palette as the discounted figure
cb_long_u <- rbind(
  clin_bud_u[, .(year, ivlab, value = cum / 1e9,
                 panel = "Cumulative additional cost (US$ billions)")],
  clin_bud_u[, .(year, ivlab, value = cum_pc,
                 panel = "Cumulative additional cost per capita (US$)")])
cb_long_u[, panel := factor(panel, levels = c("Cumulative additional cost (US$ billions)",
                                              "Cumulative additional cost per capita (US$)"))]
.ord_u <- clin_bud_u[year == HZ_END, .(cum), by = intervention_id][order(-cum)]
.ord_u[, lab := vapply(intervention_id, function(s) lbl(s, met_clin), character(1))]
title_clin_budget_undisc <- sprintf(
  "%s still drives most of the undiscounted clinical cardiovascular budget through %d",
  .ord_u$lab[1], HZ_END)
```

## `r title_clin_budget`

```{r fig-clin-budget, fig.width=9.6, fig.height=3.2, out.width="\\linewidth"}
reg_fig(
ggplot(cb_long, aes(year, value, fill = ivlab)) +
  geom_area(colour = "white", linewidth = 0.15) +
  facet_wrap(~panel, scales = "free_y") +
  scale_fill_manual(values = clin_bud_pal, breaks = lev, drop = FALSE) +
  scale_x_continuous(limits = c(HZ_START, HZ_END), breaks = seq(HZ_START, HZ_END, 5),
                     expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.03))) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  labs(title = NULL, x = "Year", y = NULL) +
  theme_exec(11) +
  theme(legend.position = "bottom", legend.text = element_text(size = 7.4),
        legend.key.size = unit(0.32, "cm"),
        strip.text = element_text(face = "bold", colour = col_wbnavy, size = 9.5),
        panel.grid.major.x = element_line(colour = "grey92", linewidth = 0.25)),
"fig-clin-budget", title_clin_budget, 9.6, 3.2)
```

\srcmethod{\textbf{Source:} clinical cost workbook (\texttt{indonesia\_model\_cost\_value\_formulae.xlsx}, sheet \texttt{Annual\_Cost}); all-clinical scenario, discounted additional cost by intervention (shared components counted once); population is the workbook's own \texttt{national\_population}. \textbf{Method:} `r HZ_START` is the discount base year and coverage scales up over `r HZ_START`--`r HZ_END`; left is cumulative discounted additional cost through each year, right is that total $\div$ that year's population; intervention totals reconcile to the all-clinical totals by year. \costcaveat}

## `r title_clin_budget_undisc`

```{r fig-clin-budget-undisc, fig.width=9.6, fig.height=3.2, out.width="\\linewidth"}
## Undiscounted counterpart of fig-clin-budget: same interventions, colours, stacking
## order and layout; both panels show UNDISCOUNTED cumulative additional cost.
reg_fig(
ggplot(cb_long_u, aes(year, value, fill = ivlab)) +
  geom_area(colour = "white", linewidth = 0.15) +
  facet_wrap(~panel, scales = "free_y") +
  scale_fill_manual(values = clin_bud_pal, breaks = lev, drop = FALSE) +
  scale_x_continuous(limits = c(HZ_START, HZ_END), breaks = seq(HZ_START, HZ_END, 5),
                     expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.03))) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  labs(title = NULL, x = "Year", y = NULL) +
  theme_exec(11) +
  theme(legend.position = "bottom", legend.text = element_text(size = 7.4),
        legend.key.size = unit(0.32, "cm"),
        strip.text = element_text(face = "bold", colour = col_wbnavy, size = 9.5),
        panel.grid.major.x = element_line(colour = "grey92", linewidth = 0.25)),
"fig-clin-budget-undisc", title_clin_budget_undisc, 9.6, 3.2)
```

\srcmethod{\textbf{Source:} clinical cost workbook (\texttt{indonesia\_model\_cost\_value\_formulae.xlsx}, sheet \texttt{Annual\_Cost}); all-clinical scenario, \textbf{undiscounted} additional cost by intervention (shared components counted once); population is the workbook's own \texttt{national\_population}. \textbf{Method (undiscounted):} coverage scales up over `r HZ_START`--`r HZ_END`; left is cumulative \textbf{undiscounted} additional cost through each year, right is that total $\div$ that year's population (`r price_year` US\$); intervention totals reconcile to the all-clinical undiscounted totals by year. \costcaveat}

## Population policies avert deaths at very low cost

```{r tbl-ph-ce, results='asis'}
ft <- family_table(met_ph, ph_all_id, ce_row)
if (nrow(ft$tb)) {
  k <- kbl(ft$tb, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=mkhdr(c("Public-health policy","Deaths averted\n(cumulative)","Prob. of CVD death\nbefore age 70, 2050",
                  "Reduction in that\nprobability, 2050","Deaths reduction\n2050","Cost per\ndeath averted",
                  "Avg. annual add'l\ncost/capita, 2026-50 (disc.)","Add'l cost/capita\nin 2050 (disc.)")),
      align="lrrrrrrr") |>
    kable_styling(latex_options=c("hold_position","scale_down"), font_size=8.5) |>
    column_spec(1, bold=TRUE) |> row_spec(0, bold=TRUE)
  if (!is.na(ft$agg_idx)) k <- k |> row_spec(ft$agg_idx, bold=TRUE, background="#E7F4F1")
  print(k)
  reg_tbl("tbl-ph-ce",
          "Population policies avert deaths at very low cost",
          ft$tb,
          c("Public-health policy","Deaths averted\n(cumulative)","Prob. of CVD death\nbefore age 70, 2050",
            "Reduction in that\nprobability, 2050","Deaths reduction\n2050","Cost per\ndeath averted",
            "Avg. annual add'l\ncost/capita, 2026-50 (disc.)","Add'l cost/capita\nin 2050 (disc.)"))
  if (ft$n_omitted > 0)
    cat(sprintf("\n\n\\vspace{1pt}\\scriptsize %d additional public-health intervention(s) omitted for space; see the appendix crosswalk and the workbook.\n", ft$n_omitted))
} else cat("\\begin{center}\\large No public-health policies selected for this run.\\end{center}")
```

\srcmethod{\textbf{Source:} public-health cost workbook (\texttt{indonesia\_cost\_value\_public\_health\_formulae.xlsx}); annual additional cost reproduced from its \texttt{Annual\_Cost} inputs and reconciled to the run total; population is the workbook's own \texttt{national\_population}. \textbf{Method:} cumulative deaths averted `r HZ_START`--`r HZ_END` vs baseline; tobacco and sodium packages each modelled as one joint run. \textbf{Avg. annual additional cost per capita, 2026--`r HZ_END`} $=$ cumulative discounted additional cost over 2026--`r HZ_END` $\div$ (25 $\times$ `r HZ_START+1L` population); \textbf{cost per capita in `r HZ_END`} $=$ discounted additional cost in `r HZ_END` $\div$ `r HZ_END` population. \costcaveat}

## Even undiscounted, population policies avert deaths at very low cost

```{r tbl-ph-ce-undisc, results='asis'}
## Undiscounted counterpart of tbl-ph-ce: identical health columns; every monetary
## column undiscounted (annual undiscounted cost reproduced from the PH workbook's own
## Annual_Cost inputs and reconciled to the run's undiscounted total).
ftu <- family_table(met_ph, ph_all_id, ce_row_undisc)
if (nrow(ftu$tb)) {
  k <- kbl(ftu$tb, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=mkhdr(c("Public-health policy","Deaths averted\n(cumulative)","Prob. of CVD death\nbefore age 70, 2050",
                  "Reduction in that\nprobability, 2050","Deaths reduction\n2050","Cost per death\naverted (undisc.)",
                  "Avg. annual add'l\ncost/capita, 2026-50 (undisc.)","Add'l cost/capita\nin 2050 (undisc.)")),
      align="lrrrrrrr") |>
    kable_styling(latex_options=c("hold_position","scale_down"), font_size=8.5) |>
    column_spec(1, bold=TRUE) |> row_spec(0, bold=TRUE)
  if (!is.na(ftu$agg_idx)) k <- k |> row_spec(ftu$agg_idx, bold=TRUE, background="#E7F4F1")
  print(k)
  reg_tbl("tbl-ph-ce-undisc",
          "Even undiscounted, population policies avert deaths at very low cost",
          ftu$tb,
          c("Public-health policy","Deaths averted\n(cumulative)","Prob. of CVD death\nbefore age 70, 2050",
            "Reduction in that\nprobability, 2050","Deaths reduction\n2050","Cost per death\naverted (undisc.)",
            "Avg. annual add'l\ncost/capita, 2026-50 (undisc.)","Add'l cost/capita\nin 2050 (undisc.)"))
  if (ftu$n_omitted > 0)
    cat(sprintf("\n\n\\vspace{1pt}\\scriptsize %d additional public-health intervention(s) omitted for space; see the appendix crosswalk and the workbook.\n", ftu$n_omitted))
} else cat("\\begin{center}\\large No public-health policies selected for this run.\\end{center}")
```

\srcmethod{\textbf{Source:} public-health cost workbook (\texttt{indonesia\_cost\_value\_public\_health\_formulae.xlsx}); annual additional cost reproduced from its \texttt{Annual\_Cost} inputs and reconciled to the run's \textbf{undiscounted} total; population is the workbook's own \texttt{national\_population}. \textbf{Method (undiscounted):} cumulative deaths averted `r HZ_START`--`r HZ_END` vs baseline; tobacco and sodium packages each modelled as one joint run. \textbf{Avg. annual additional cost per capita, 2026--`r HZ_END`} $=$ cumulative \textbf{undiscounted} additional cost over 2026--`r HZ_END` $\div$ (25 $\times$ `r HZ_START+1L` population); \textbf{cost per capita in `r HZ_END`} $=$ \textbf{undiscounted} additional cost in `r HZ_END` $\div$ `r HZ_END` population (`r price_year` US\$). Health results are identical to the discounted slide. \costcaveat}


# Clinical and population action, combined

## Clinical care and prevention achieve more together than alone

```{r tbl-combined, results='asis'}
pk <- intersect(c(clin_all_id, ph_all_id, joint_id), met_comb$scenario)
if (length(pk)) {
  tb <- rbindlist(lapply(pk, function(s) data.table(
    Package = lbl(s),
    da   = fmt_thou(hv(s, "deaths_averted")),
    qr   = fmt_pct(hv(s, "cvd_40q30_pct_reduction_2050"), 1),
    q50  = fmt_pct(hv(s, "cvd_40q30_2050"), 1),
    cost = fmt_usd_short(hv(s, "disc_incremental_cost")),
    cpd  = fmt_usd(hv(s, "cost_per_death_averted")),
    pc1c = pc1(met_comb, s),
    pc2c = pc2(met_comb, s))))
  k <- kbl(tb, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=mkhdr(c("Package","Deaths averted\n(cumulative)","Reduction in prob.\nof CVD death <70, 2050","Prob. of CVD death\nbefore age 70, 2050",
                  "Discounted total\ncost","Cost per\ndeath averted","Avg. annual add'l\ncost/capita, 2026-50 (disc.)","Add'l cost/capita\nin 2050 (disc.)")),
      align="lrrrrrrr") |>
    kable_styling(latex_options=c("hold_position","scale_down"), font_size=9) |>
    column_spec(1, bold=TRUE) |> row_spec(0, bold=TRUE)
  jidx <- which(pk == joint_id)
  if (length(jidx)) k <- k |> row_spec(jidx, bold=TRUE, background="#FCEFCB")
  print(k)
  reg_tbl("tbl-combined",
          "Clinical care and prevention achieve more together than alone",
          tb,
          c("Package","Deaths averted\n(cumulative)","Reduction in prob.\nof CVD death <70, 2050","Prob. of CVD death\nbefore age 70, 2050",
            "Discounted total\ncost","Cost per\ndeath averted","Avg. annual add'l\ncost/capita, 2026-50 (disc.)","Add'l cost/capita\nin 2050 (disc.)"))
} else cat("\\begin{center}\\large No package scenarios available for this run.\\end{center}")
```

```{r combined-block, results='asis'}
if (has_joint) {
  cat(sprintf(paste0(
    "\\begin{combinedblock}{The combined package is modelled as one run --- not a sum}\n",
    "\\footnotesize Because the two families' effects overlap, the joint scenario is projected in a single run: about \\textbf{%s deaths} averted and the %d probability of CVD death before age 70 cut to \\textbf{%s} (from %s), at about \\textbf{%s per death averted}.\n",
    "\\end{combinedblock}\n"),
    fmt_short(combo_da), HZ_END, fmt_pct(combo_q50,1), fmt_pct(base_q_2050,1), fmt_usd(combo_cpda)))
} else {
  cat(paste0("\\begin{block}{Single-family run}\n\\footnotesize The joint clinical + ",
             "public-health scenario was not built in this configuration; family results are ",
             "shown above.\n\\end{block}\n"))
}
```

\srcmethod{\textbf{Source:} combined cost workbook (\texttt{indonesia\_model\_cost\_value\_clinical\_public\_health\_formulae.xlsx}); annual additional cost reproduced from its \texttt{Annual\_Cost} inputs and reconciled to the run total; population is the workbook's own \texttt{national\_population}. \textbf{Method:} joint package $=$ one projection applying both families (not their sum). \textbf{Avg. annual additional cost per capita, 2026--`r HZ_END`} $=$ cumulative discounted additional cost over 2026--`r HZ_END` $\div$ (25 $\times$ `r HZ_START+1L` population); \textbf{cost per capita in `r HZ_END`} $=$ discounted additional cost in `r HZ_END` $\div$ `r HZ_END` population (`r fmt_pct(disc_rate*100,0)`/yr, `r price_year` US\$). \costcaveat}

## Undiscounted, clinical care and prevention still achieve more together

```{r tbl-combined-undisc, results='asis'}
## Undiscounted counterpart of tbl-combined: identical health columns; total cost, cost
## per death averted, and both per-capita measures are undiscounted (reproduced from the
## combined workbook's own Annual_Cost inputs and reconciled to the run's undiscounted total).
pku <- intersect(c(clin_all_id, ph_all_id, joint_id), met_comb$scenario)
if (length(pku)) {
  tbu <- rbindlist(lapply(pku, function(s) data.table(
    Package = lbl(s),
    da   = fmt_thou(hv(s, "deaths_averted")),
    qr   = fmt_pct(hv(s, "cvd_40q30_pct_reduction_2050"), 1),
    q50  = fmt_pct(hv(s, "cvd_40q30_2050"), 1),
    cost = fmt_usd_short(hv(s, "incremental_cost")),
    cpd  = fmt_usd(cpda_undisc(s, met_comb)),
    pc1c = pc1u(met_comb, s),
    pc2c = pc2u(met_comb, s))))
  k <- kbl(tbu, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=mkhdr(c("Package","Deaths averted\n(cumulative)","Reduction in prob.\nof CVD death <70, 2050","Prob. of CVD death\nbefore age 70, 2050",
                  "Undiscounted total\ncost","Cost per death\naverted (undisc.)","Avg. annual add'l\ncost/capita, 2026-50 (undisc.)","Add'l cost/capita\nin 2050 (undisc.)")),
      align="lrrrrrrr") |>
    kable_styling(latex_options=c("hold_position","scale_down"), font_size=9) |>
    column_spec(1, bold=TRUE) |> row_spec(0, bold=TRUE)
  jidxu <- which(pku == joint_id)
  if (length(jidxu)) k <- k |> row_spec(jidxu, bold=TRUE, background="#FCEFCB")
  print(k)
  reg_tbl("tbl-combined-undisc",
          "Undiscounted, clinical care and prevention still achieve more together",
          tbu,
          c("Package","Deaths averted\n(cumulative)","Reduction in prob.\nof CVD death <70, 2050","Prob. of CVD death\nbefore age 70, 2050",
            "Undiscounted total\ncost","Cost per death\naverted (undisc.)","Avg. annual add'l\ncost/capita, 2026-50 (undisc.)","Add'l cost/capita\nin 2050 (undisc.)"))
} else cat("\\begin{center}\\large No package scenarios available for this run.\\end{center}")
```

```{r combined-block-undisc, results='asis'}
if (has_joint) {
  cat(sprintf(paste0(
    "\\begin{combinedblock}{The combined package is one run --- valued undiscounted}\n",
    "\\footnotesize Because the two families' effects overlap, the joint scenario is projected in a single run: about \\textbf{%s deaths} averted and the %d probability of CVD death before age 70 cut to \\textbf{%s} (from %s), at about \\textbf{%s per death averted (undiscounted)}.\n",
    "\\end{combinedblock}\n"),
    fmt_short(combo_da), HZ_END, fmt_pct(combo_q50,1), fmt_pct(base_q_2050,1), fmt_usd(combo_cpda_undisc)))
} else {
  cat(paste0("\\begin{block}{Single-family run}\n\\footnotesize The joint clinical + ",
             "public-health scenario was not built in this configuration; family results are ",
             "shown above.\n\\end{block}\n"))
}
```

\srcmethod{\textbf{Source:} combined cost workbook (\texttt{indonesia\_model\_cost\_value\_clinical\_public\_health\_formulae.xlsx}); annual additional cost reproduced from its \texttt{Annual\_Cost} inputs and reconciled to the run's \textbf{undiscounted} total; population is the workbook's own \texttt{national\_population}. \textbf{Method (undiscounted):} joint package $=$ one projection applying both families (not their sum). \textbf{Avg. annual additional cost per capita, 2026--`r HZ_END`} $=$ cumulative \textbf{undiscounted} additional cost over 2026--`r HZ_END` $\div$ (25 $\times$ `r HZ_START+1L` population); \textbf{cost per capita in `r HZ_END`} $=$ \textbf{undiscounted} additional cost in `r HZ_END` $\div$ `r HZ_END` population (`r price_year` US\$). Health results are identical to the discounted slide. \costcaveat}

## Indonesia's premature cardiovascular mortality can fall toward its neighbours

```{r fig-40q30, fig.width=9.6, fig.height=4.2, out.width="\\linewidth"}
# Remove the modeled status-quo baseline
idn_scen_plot <- idn_scen[scenario != "baseline"]
lab_scen <- idn_scen_plot[year == HZ_END]

lab_cmp <- cmp_lines[, .SD[which.max(year)], by = country_label]

p_fig_40q30 <- ggplot() +
  geom_line(
    data = cmp_lines,
    aes(year, q, group = country_label),
    colour = col_comparator,
    linewidth = 0.5
  ) +
  ggrepel::geom_text_repel(
    data = lab_cmp,
    aes(year, q, label = country_label),
    colour = "grey55", size = 2.5, hjust = 0,
    nudge_x = 1.2, direction = "y",
    segment.colour = "grey80", segment.size = 0.2,
    min.segment.length = 0, box.padding = 0.12,
    max.overlaps = Inf, seed = 42
  ) +
  geom_line(
    data = idn_obs,
    aes(year, q),
    colour = "grey20",
    linewidth = 1.0
  ) +
  annotate(
    "segment",
    x = 2024, xend = 2024,
    y = 10.5, yend = 20.5,
    linetype = "dotted",
    colour = "grey45"
  ) +
  annotate(
    "text",
    x = 2024, y = 20.9,
    label = "Modeled scenarios →",
    hjust = 1, size = 2.7,
    colour = "grey35"
  ) +
  geom_line(
    data = idn_scen_plot,  # filtered data
    aes(year, q, colour = scen),
    linewidth = 1.15,
    linetype = "22"
  ) +
  ggrepel::geom_text_repel(
    data = lab_scen,
    aes(year, q, label = scen, colour = scen),
    size = 2.7, fontface = "bold", hjust = 0,
    nudge_x = 1.2, direction = "y",
    segment.colour = "grey70", segment.size = 0.2,
    min.segment.length = 0, box.padding = 0.2,
    max.overlaps = Inf, show.legend = FALSE, seed = 7
  ) +
  scale_colour_manual(values = scen_pal) +
  scale_x_continuous(
    limits = c(1995, 2064),
    breaks = seq(1995, 2050, 5),
    expand = expansion(mult = c(0.01, 0))
  ) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "Probability of death from CVD before age 70: observed, projected, and neighbours",
    subtitle = "Chance of dying from CVD between exact ages 30 and 70",
    x = NULL,
    y = "Probability of CVD death before age 70 (%)",
    caption = paste(
      "Indonesia observed history (solid, to 2023) and ASEAN comparators are WPP-scaled reference series.",
      "\nModelled intervention scenarios are dashed for 2025–2050.",
      "\nASEAN comparator trends use linear extrapolation for illustrative reference."
    )
  ) +
  theme_exec(11) +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_line(
      colour = "grey92",
      linewidth = 0.25
    )
  )
reg_fig(p_fig_40q30, "fig-40q30",
        "Indonesia's premature cardiovascular mortality can fall toward its neighbours",
        9.6, 4.2)
```

\srcmethod{\textbf{Source:} combined results workbook (annual probability-of-death sheet, R-reconciled values); ASEAN comparator series in \texttt{data/raw/}. \textbf{Method:} model scenarios (dashed) run `r HZ_START`--`r HZ_END` for the six modelled cardiovascular causes over exact ages 30--70; comparators are WPP-scaled reference series shown for context only.}

# A stronger care cascade: the 70-30-30 to 70-70-70 strategy

## The cascade scale-up is ambitious but staged

\footnotesize Hypertension and type-2 diabetes are modelled with \textbf{diagnosed $\rightarrow$ treated $\rightarrow$ controlled} care cascades; \textbf{cholesterol} treatment coverage follows hypertension coverage. Coverage rises in two stages, then holds.

\vspace{1pt}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{clinicalblock}{Two staged milestones}
\footnotesize
\begin{itemize}\itemsep1pt
  \item \textbf{`r casc_ytar1` --- 70-30-30:} `r fmt_pct(casc_diag*100,0)` diagnosed, `r fmt_pct(casc_t30*100,0)` of those treated, `r fmt_pct(casc_c30*100,0)` of those controlled.
  \item \textbf{`r casc_ytar2` --- 70-70-70:} `r fmt_pct(casc_diag*100,0)` diagnosed, `r fmt_pct(casc_t40*100,0)` treated, `r fmt_pct(casc_c40*100,0)` controlled; held through `r casc_y1`.
  \item Linear scale-up from the observed baseline; coverage already above a milestone is \textbf{not scaled down}.
\end{itemize}
\end{clinicalblock}
\end{column}
\begin{column}{0.48\textwidth}
\begin{combinedblock}{Shares of all people with the condition}
\footnotesize
\begin{tabular}{lcc}
 & \textbf{`r casc_ytar1`} & \textbf{`r casc_ytar2`}\\
\midrule
Diagnosed  & `r fmt_pct(casc_diag*100,0)` & `r fmt_pct(casc_diag*100,0)`\\
Treated    & `r fmt_pct(casc_diag*casc_t30*100,0)` & `r fmt_pct(casc_diag*casc_t40*100,0)`\\
Controlled & `r fmt_pct(casc_diag*casc_t30*casc_c30*100,1)` & `r fmt_pct(casc_diag*casc_t40*casc_c40*100,1)`\\
\end{tabular}\\[3pt]
{\scriptsize treated $=70\%\times\{30,70\}\%$; controlled $=$ treated $\times\{30,70\}\%$.}
\end{combinedblock}
\end{column}
\end{columns}

\vspace{1pt}
\footnotesize \textbf{Effective modelled coverage} $=$ controlled $+\tfrac{1}{2}\times$ treated-but-uncontrolled $=$ \textbf{`r fmt_pct(casc_eff30*100,2)`} (`r casc_ytar1`) and \textbf{`r fmt_pct(casc_eff40*100,2)`} (`r casc_ytar2`): full effect to controlled patients, half to treated-but-uncontrolled.

\srcmethod{\textbf{Source:} cascade workbook (\texttt{Cascade\_Assumptions}, \texttt{Cascade\_Trajectory}); isolated runner; NCD-RisC coverage inputs. \textbf{Method:} nested targets on the diagnosed/treated/controlled shares; piecewise-linear scale-up, no backsliding.}

## Indonesia's hypertension cascade can progress in two achievable stages

```{r fig-htn-cascade, fig.width=9.4, fig.height=3.7, out.width="0.94\\linewidth"}
#Patch: manual update to Farizals value automatize later

htn_sh[year == 2025, `:=`(
  controlled = 0.04,
  treated    = 0.22,
  diagnosed  = 0.27
)]

htn_sh[year == 2030, `:=`(
  controlled = 0.063,
  treated    = 0.21,
  diagnosed  = 0.700
)]

htn_sh[year == 2040, `:=`(
  controlled = 0.343,
  treated    = 0.490,
  diagnosed  = 0.700
)]


htn_long <- melt(htn_sh[, .(year, Diagnosed = diagnosed, Treated = treated, Controlled = controlled)],
                 id.vars = "year", variable.name = "stage", value.name = "share")
htn_long <- htn_long[!is.na(share)]
htn_long[, stage := factor(stage, levels = c("Diagnosed","Treated","Controlled"))]
htn_long[, xlab := factor(year, levels = c(HZ_START, casc_ytar1, casc_ytar2),
                          labels = c(paste0("Baseline\n(", HZ_START, ")"),
                                     paste0(casc_ytar1, "\n(70-30-30)"),
                                     paste0(casc_ytar2, "\n(70-70-70)")))]
casc_blues <- c(Diagnosed = "#9ECAE1", Treated = "#4292C6", Controlled = "#08519C")
reg_fig(
ggplot(htn_long, aes(xlab, share, fill = stage)) +
  geom_col(position = position_dodge2(width = 0.8, preserve = "single"), width = 0.8) +
  geom_text(aes(label = paste0(formatC(share * 100, format = "f", digits = 1), "%")),
            position = position_dodge2(width = 0.8, preserve = "single"),
            vjust = -0.4, size = 3.0, fontface = "bold") +
  scale_fill_manual(values = casc_blues) +
  scale_y_continuous(labels = function(x) paste0(x * 100, "%"), limits = c(0, 0.80),
                     expand = expansion(mult = c(0, 0.02))) +
  labs(title = "Hypertension care cascade under the 70-30-30 to 70-70-70 scale-up",
       subtitle = "Diagnosed, treated, and controlled shares of all people with hypertension (intervention scenario)",
       x = NULL, y = "% of people with hypertension") +
  theme_exec(11) +
  theme(legend.position = "top", panel.grid.major.x = element_blank()),
"fig-htn-cascade", "Indonesia's hypertension cascade can progress in two achievable stages", 9.4, 3.7)
```

\srcmethod{\textbf{Source:} cascade \texttt{Cascade\_Trajectory} (sex-averaged) and \texttt{Cascade\_Assumptions}; scale-up begins `r HZ_START+1L`. \textbf{Method:} baseline (`r HZ_START`) shows only measured treated/controlled --- diagnosis is not modelled at baseline, so no diagnosed bar. Type-2 diabetes follows the same targets (baseline treated $\approx$ `r fmt_pct(dm_base$treated*100,1)`, controlled $\approx$ `r fmt_pct(dm_base$controlled*100,1)`).}

## Stronger prevention lowers premature deaths at a rising but bounded cost

```{r fig-casc-2panel, fig.width=9.6, fig.height=3.7, out.width="\\linewidth"}
lab_q  <- c(baseline = "Baseline (current course)", cascade = "Intervention")
qdat <- copy(casc_q)[, series := ifelse(scenario == "baseline", "Baseline (current course)", "Intervention")]
qdat[, panel := "Probability of death from CVD before age 70 (%)"]
setnames(qdat, "q", "value")
cdat <- rbind(
  casc_cost[, .(year, series = "Baseline (current course)", value = cum_base / 1e9)],
  casc_cost[, .(year, series = "Intervention",              value = cum_scen / 1e9)])
cdat[, panel := "Cumulative discounted health-system cost (US$ billions)"]
two <- rbind(qdat[, .(year, series, value, panel)], cdat)
two[, panel := factor(panel, levels = c("Probability of death from CVD before age 70 (%)",
                                         "Cumulative discounted health-system cost (US$ billions)"))]
casc_pal <- c("Baseline (current course)" = col_baseline, "Intervention" = "#D55E00")
ends <- two[, .SD[year == max(year)], by = .(series, panel)]
## Readable endpoint labels: short text, visible leader segments, and ample right-side
## room so labels never sit on the lines or clip. Deterministic (fixed seed + offsets).
ends[, lab := ifelse(series == "Intervention", "Intervention", "Baseline")]
reg_fig(
ggplot(two, aes(year, value, colour = series)) +
  geom_line(linewidth = 1.15) +
  ggrepel::geom_text_repel(data = ends, aes(label = lab), size = 2.7, fontface = "bold",
                           hjust = 0, nudge_x = 1.6, direction = "y",
                           segment.colour = "grey55", segment.size = 0.3,
                           min.segment.length = 0, box.padding = 0.5, point.padding = 0.3,
                           max.overlaps = Inf, seed = 3, show.legend = FALSE) +
  facet_wrap(~panel, scales = "free_y") +
  scale_colour_manual(values = casc_pal) +
  scale_x_continuous(limits = c(HZ_START, casc_y1 + 11), breaks = seq(HZ_START, casc_y1, 5)) +
  labs(title = NULL, x = "Year", y = NULL) +
  theme_exec(11) +
  theme(legend.position = "none",
        strip.text = element_text(face = "bold", colour = col_wbnavy, size = 10),
        panel.grid.major.x = element_line(colour = "grey92", linewidth = 0.25)),
"fig-casc-2panel", "Stronger prevention lowers premature deaths at a rising but bounded cost", 9.6, 3.7)
```

\srcmethod{\textbf{Source:} cascade workbook --- annual probability-of-death sheet (R-reconciled) and \texttt{Annual\_Cost}, cumulatively summed and reconciled to \texttt{Budget\_Impact}. \textbf{Method:} probability over the six CVD causes, ages 30--70; cumulative discounted health-system cost (`r fmt_pct(casc_disc*100,0)`/yr to `r HZ_START`, `r casc_price` US\$) of the CVD primary-care and type-2 diabetes cascades. \costcaveat}

## Undiscounted, stronger prevention still lowers premature deaths at a rising cost

```{r fig-casc-2panel-undisc, fig.width=9.6, fig.height=3.7, out.width="\\linewidth"}
## Undiscounted counterpart of fig-casc-2panel: the health panel is unchanged; the cost
## panel shows cumulative UNDISCOUNTED baseline vs intervention health-system cost.
cdat_u <- rbind(
  casc_cost_u[, .(year, series = "Baseline (current course)", value = cum_base / 1e9)],
  casc_cost_u[, .(year, series = "Intervention",              value = cum_scen / 1e9)])
cdat_u[, panel := "Cumulative health-system cost (US$ billions, undiscounted)"]
two_u <- rbind(qdat[, .(year, series, value, panel)], cdat_u)
two_u[, panel := factor(panel, levels = c("Probability of death from CVD before age 70 (%)",
                                          "Cumulative health-system cost (US$ billions, undiscounted)"))]
ends_u <- two_u[, .SD[year == max(year)], by = .(series, panel)]
ends_u[, lab := ifelse(series == "Intervention", "Intervention", "Baseline")]
reg_fig(
ggplot(two_u, aes(year, value, colour = series)) +
  geom_line(linewidth = 1.15) +
  ggrepel::geom_text_repel(data = ends_u, aes(label = lab), size = 2.7, fontface = "bold",
                           hjust = 0, nudge_x = 1.6, direction = "y",
                           segment.colour = "grey55", segment.size = 0.3,
                           min.segment.length = 0, box.padding = 0.5, point.padding = 0.3,
                           max.overlaps = Inf, seed = 3, show.legend = FALSE) +
  facet_wrap(~panel, scales = "free_y") +
  scale_colour_manual(values = casc_pal) +
  scale_x_continuous(limits = c(HZ_START, casc_y1 + 11), breaks = seq(HZ_START, casc_y1, 5)) +
  labs(title = NULL, x = "Year", y = NULL) +
  theme_exec(11) +
  theme(legend.position = "none",
        strip.text = element_text(face = "bold", colour = col_wbnavy, size = 10),
        panel.grid.major.x = element_line(colour = "grey92", linewidth = 0.25)),
"fig-casc-2panel-undisc", "Undiscounted, stronger prevention still lowers premature deaths at a rising cost", 9.6, 3.7)
```

\srcmethod{\textbf{Source:} cascade workbook --- annual probability-of-death sheet (R-reconciled) and \texttt{Annual\_Cost}, cumulatively summed and reconciled to \texttt{Budget\_Impact} \textbf{undiscounted} cumulative additional cost. \textbf{Method:} probability over the six CVD causes, ages 30--70 (identical to the discounted slide); cumulative \textbf{undiscounted} health-system cost (`r casc_price` US\$) of the CVD primary-care and type-2 diabetes cascades. \costcaveat}

## `r title_cascade_results`

```{r tbl-cascade, results='asis'}
snp <- lapply(casc_tab_years, casc_snapshot); names(snp) <- as.character(casc_tab_years)
mm  <- function(x) formatC(x / 1e6, format = "f", digits = 3, big.mark = ",")   # deaths, millions
bn  <- function(x) formatC(x / 1e9, format = "f", digits = 2)                    # cost, US$ billions
pcf <- function(x) paste0("US\\$", formatC(x, format = "f", digits = 2, big.mark = ",")) # US$/person
casc_tb <- data.table(
  Metric = c("\\quad Baseline (current course)", "\\quad Intervention",
             "\\quad Baseline (current course)", "\\quad Intervention",
             "Additional cost (US\\$ billion, discounted)", "Deaths prevented (million)",
             "Additional cost per capita in that year (US\\$, discounted)"),
  y1 = c(mm(snp[[1]]$d_base), mm(snp[[1]]$d_scen), bn(snp[[1]]$c_base), bn(snp[[1]]$c_scen), bn(snp[[1]]$c_add), mm(snp[[1]]$d_av), pcf(snp[[1]]$pc_ann)),
  y2 = c(mm(snp[[2]]$d_base), mm(snp[[2]]$d_scen), bn(snp[[2]]$c_base), bn(snp[[2]]$c_scen), bn(snp[[2]]$c_add), mm(snp[[2]]$d_av), pcf(snp[[2]]$pc_ann)),
  y3 = c(mm(snp[[3]]$d_base), mm(snp[[3]]$d_scen), bn(snp[[3]]$c_base), bn(snp[[3]]$c_scen), bn(snp[[3]]$c_add), mm(snp[[3]]$d_av), pcf(snp[[3]]$pc_ann)))
reg_tbl("tbl-cascade", title_cascade_results, casc_tb,
        c("Cumulative from 2025 through year (unless noted)", as.character(casc_tab_years)))
kbl(casc_tb, format = "latex", booktabs = TRUE, escape = FALSE,
    col.names = c("Cumulative from 2025 through year (unless noted)", as.character(casc_tab_years)),
    align = "lrrr") |>
  kable_styling(latex_options = c("hold_position"), font_size = 8) |>
  column_spec(1, width = "6.6cm") |> column_spec(2:4, width = "2.2cm") |>
  row_spec(0, bold = TRUE) |>
  pack_rows("Projected deaths --- CVD (six causes) + type-2 diabetes (million)", 1, 2, bold = TRUE, latex_gap_space = "0.2em", escape = FALSE) |>
  pack_rows("Projected health-system cost (US\\$ billion, discounted)", 3, 4, bold = TRUE, latex_gap_space = "0.2em", escape = FALSE) |>
  pack_rows("Net effect of the strategy", 5, 6, bold = TRUE, latex_gap_space = "0.2em") |>
  pack_rows("Annual financing intensity (that year only, not cumulative)", 7, 7, bold = TRUE, latex_gap_space = "0.2em", escape = FALSE)
```

\srcmethod{\textbf{Source:} cascade \texttt{Annual\_Mortality}, \texttt{Annual\_Cost}/\texttt{Budget\_Impact}. \textbf{Method:} deaths/cost columns cumulative from `r HZ_START` through that year (six CVD causes $+$ type-2 diabetes; discounted `r fmt_pct(casc_disc*100,0)` to `r HZ_START`, `r casc_price` US\$); the per-capita row is that year only, not cumulative. \costcaveat}

## `r title_cascade_results_undisc`

```{r tbl-cascade-undisc, results='asis'}
## Undiscounted counterpart of tbl-cascade: deaths rows are unchanged (deaths are never
## discounted); baseline cost, intervention cost, additional cost, and the annual
## additional cost per capita are all UNDISCOUNTED (from casc_snapshot_undisc).
snpu <- lapply(casc_tab_years, casc_snapshot_undisc); names(snpu) <- as.character(casc_tab_years)
mm  <- function(x) formatC(x / 1e6, format = "f", digits = 3, big.mark = ",")   # deaths, millions
bn  <- function(x) formatC(x / 1e9, format = "f", digits = 2)                    # cost, US$ billions
pcf <- function(x) paste0("US\\$", formatC(x, format = "f", digits = 2, big.mark = ",")) # US$/person
casc_tb_u <- data.table(
  Metric = c("\\quad Baseline (current course)", "\\quad Intervention",
             "\\quad Baseline (current course)", "\\quad Intervention",
             "Additional cost (US\\$ billion, undiscounted)", "Deaths prevented (million)",
             "Additional cost per capita in that year (US\\$, undiscounted)"),
  y1 = c(mm(snpu[[1]]$d_base), mm(snpu[[1]]$d_scen), bn(snpu[[1]]$c_base), bn(snpu[[1]]$c_scen), bn(snpu[[1]]$c_add), mm(snpu[[1]]$d_av), pcf(snpu[[1]]$pc_ann)),
  y2 = c(mm(snpu[[2]]$d_base), mm(snpu[[2]]$d_scen), bn(snpu[[2]]$c_base), bn(snpu[[2]]$c_scen), bn(snpu[[2]]$c_add), mm(snpu[[2]]$d_av), pcf(snpu[[2]]$pc_ann)),
  y3 = c(mm(snpu[[3]]$d_base), mm(snpu[[3]]$d_scen), bn(snpu[[3]]$c_base), bn(snpu[[3]]$c_scen), bn(snpu[[3]]$c_add), mm(snpu[[3]]$d_av), pcf(snpu[[3]]$pc_ann)))
reg_tbl("tbl-cascade-undisc", title_cascade_results_undisc, casc_tb_u,
        c("Cumulative from 2025 through year (unless noted)", as.character(casc_tab_years)))
kbl(casc_tb_u, format = "latex", booktabs = TRUE, escape = FALSE,
    col.names = c("Cumulative from 2025 through year (unless noted)", as.character(casc_tab_years)),
    align = "lrrr") |>
  kable_styling(latex_options = c("hold_position"), font_size = 8) |>
  column_spec(1, width = "6.6cm") |> column_spec(2:4, width = "2.2cm") |>
  row_spec(0, bold = TRUE) |>
  pack_rows("Projected deaths --- CVD (six causes) + type-2 diabetes (million)", 1, 2, bold = TRUE, latex_gap_space = "0.2em", escape = FALSE) |>
  pack_rows("Projected health-system cost (US\\$ billion, undiscounted)", 3, 4, bold = TRUE, latex_gap_space = "0.2em", escape = FALSE) |>
  pack_rows("Net effect of the strategy", 5, 6, bold = TRUE, latex_gap_space = "0.2em") |>
  pack_rows("Annual financing intensity (that year only, not cumulative)", 7, 7, bold = TRUE, latex_gap_space = "0.2em", escape = FALSE)
```

\srcmethod{\textbf{Source:} cascade \texttt{Annual\_Mortality}, \texttt{Annual\_Cost}/\texttt{Budget\_Impact}. \textbf{Method (undiscounted):} deaths/cost columns cumulative from `r HZ_START` through that year (six CVD causes $+$ type-2 diabetes; \textbf{undiscounted}, `r casc_price` US\$); the per-capita row is that year only, not cumulative. Deaths are identical to the discounted slide. \costcaveat}

## The care cascade and tobacco policies jointly move Indonesia further toward its neighbours

```{r fig-40q30-cascade, fig.width=8.0, fig.height=6.0, out.width="\\linewidth"}
## ---------------------------------------------------------------------------
## FIVE EXPLICITLY-MODELLED annual probability-of-death-before-70 trajectories,
## each read as the stored R-reconciled literal (cvd_40q30_r) from its OWN
## workbook's CVD_40q30 sheet -- never summed, multiplied, averaged, or
## interpolated. All source workbooks' baselines reconcile EXACTLY (asserted below),
## so the six-cause / exact-ages-30-70 definition is identical and the lines are
## directly comparable:
## * Hypertension/cholesterol + diabetes 70-70-70 cascade
##     -> S_70_30_30_TO_70_70_70 (joint cascade + public-health wb)
## * 70-70-70 cascade + tobacco tax ONLY (targeted interaction; the tobacco tax
##   is the ONLY public-health intervention added -- NOT clean-air, mass-media or
##   the advertising ban, and NOT the full all_cascade_public_health package)
##     -> S_70_30_30_TO_70_70_70_I_PH_TOBACCO_TAX (joint cascade + public-health wb)
## * Tobacco tax -> I_PH_TOBACCO_TAX (public-health wb)
## * All public-health interventions -> ph_all_id (all_public_health, public-health wb)
## * All clinical + public health -> all_clinical_public_health (combined wb)
## The all-public-health line is read as the public-health workbook's OWN
## all-public-health scenario (its stored annual cvd_40q30_r) -- NOT reused from the
## combined workbook, NOT summed from individual policies, NOT the tobacco-tax line.
## Reuses the SAME observed history (idn_obs) and ASEAN comparators (cmp_lines)
## as the earlier neighbours slide; the modelled status-quo baseline is NOT drawn.
## ---------------------------------------------------------------------------
scen_specs <- list(
  list(id = "S_70_30_30_TO_70_70_70", wb = casc_ph_wb,
       label = "70-70-70 target achieved",
       col = col_clinical),
  list(id = "S_70_30_30_TO_70_70_70_I_PH_TOBACCO_TAX", wb = casc_ph_wb,
       label = "70-70-70 target achieved + tobacco tax",
       col = "#7B3294"),
  list(id = "all_clinical_public_health", wb = wb_comb, label = "All clinical + tobacco tax & regulatory package", col = col_combined)
)

read_scen_q3 <- function(path, sid) {
  d <- rd(path, "CVD_40q30")
  if ("location" %in% names(d)) d <- d[location == "Indonesia"]
  qc <- if ("cvd_40q30_r" %in% names(d)) "cvd_40q30_r" else "cvd_40q30"
  d[scenario == sid & year >= HZ_START & !is.na(get(qc)), .(year, q = get(qc))][order(year)]
}

idn_scen5 <- rbindlist(lapply(scen_specs, function(s) {
  x <- read_scen_q3(s$wb, s$id)
  x[, `:=`(scenario = s$id, scen = s$label)]
  x[]
}))

## Fail loudly if any requested scenario lacks a full HZ_START..HZ_END trajectory.
for (s in scen_specs) {
  ys <- idn_scen5[scenario == s$id, year]
  if (!length(ys) || max(ys) < HZ_END || min(ys) > HZ_START)
    stop(sprintf("VALIDATION STOP -- scenario '%s' has no full %d-%d trajectory in %s",
                 s$id, HZ_START, HZ_END, basename(s$wb)), call. = FALSE)
}

## Baseline reconciliation: every plotted line's own workbook must share the identical
## six-cause / ages-30-70 baseline probability series across HZ_START..HZ_END, so the
## five lines are directly comparable.
.base_ref <- read_scen_q3(wb_comb, "baseline")
for (s in scen_specs) {
  .b <- read_scen_q3(s$wb, "baseline")
  .bm <- merge(.base_ref, .b, by = "year")
  if (nrow(.bm) < length(HZ_START:HZ_END) || max(abs(.bm$q.x - .bm$q.y)) > 1e-6)
    stop(sprintf("VALIDATION STOP -- baseline probability series in %s does not reconcile with the combined workbook baseline",
                 basename(s$wb)), call. = FALSE)
}

scen5_levels <- vapply(scen_specs, `[[`, character(1), "label")
idn_scen5[, scen := factor(scen, levels = scen5_levels)]
scen5_pal <- setNames(vapply(scen_specs, `[[`, character(1), "col"), scen5_levels)

## Endpoint labels ONLY for ASEAN comparators (right-hand side).
## Scenario (intervention) labels are deliberately omitted from the RHS;
## they appear only in the legend.
cmp_end <- cmp_lines[, .SD[which.max(year)], by = country_label][
  , .(year, q, country_label)]

p_fig_40q30_cascade <- ggplot() +
  geom_line(
    data = cmp_lines,
    aes(year, q, group = country_label),
    colour = col_comparator,
    linewidth = 0.5
  ) +
  geom_line(
    data = idn_obs,
    aes(year, q),
    colour = "grey20",
    linewidth = 1.0
  ) +
  annotate(
    "segment",
    x = 2024, xend = 2024, y = 10.5, yend = 20.5,
    linetype = "dotted", colour = "grey45"
  ) +
  annotate(
    "text",
    x = 2024, y = 20.9,
    label = "Modeled scenarios →",
    hjust = 1, size = 2.7, colour = "grey35"
  ) +
  geom_line(
    data = idn_scen5,
    aes(year, q, colour = scen),
    linewidth = 1.15,
    linetype = "22"
  ) +
  ## Country names only (larger, grey) on the right-hand side
  ggrepel::geom_text_repel(
    data = cmp_end,
    aes(year, q, label = country_label),
    colour = "grey40",
    size = 3.4,
    hjust = 0,
    nudge_x = 1.3,
    direction = "y",
    segment.colour = "grey65",
    segment.size = 0.25,
    min.segment.length = 0,
    box.padding = 0.35,
    point.padding = 0.2,
    max.overlaps = Inf,
    show.legend = FALSE,
    seed = 11
  ) +
  scale_colour_manual(
    values = scen5_pal,
    name = NULL
  ) +
  scale_x_continuous(
    limits = c(1995, 2068),
    breaks = seq(1995, 2050, 5),
    expand = expansion(mult = c(0.01, 0))
  ) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "Cascade, tobacco policies, and combined packages",
    subtitle = "Chance of dying from CVD between exact ages 30 and 70",
    x = NULL,
    y = "Probability of CVD death before age 70 (%)"
    ) +
  theme_exec(11) +
  theme(
    legend.position = c(0.02, 0.02),   # lower-left
    legend.justification = c(0, 0),
    legend.background = element_rect(fill = alpha("white", 0.85), colour = "grey80", linewidth = 0.3),
    legend.key.width = unit(1.1, "cm"),
    legend.key.height = unit(0.40, "cm"),
    legend.text = element_text(size = 7.8),
    panel.grid.major.x = element_line(colour = "grey92", linewidth = 0.25)
  )

reg_fig(p_fig_40q30_cascade,
        "fig-40q30-cascade",
        "The care cascade and tobacco policies jointly move Indonesia further toward its neighbours",
        8.0, 6.0)
```

\srcmethod{\textbf{Source:} each modelled line is the stored R-reconciled literal from the annual probability-of-death-before-70 sheet of its \emph{own} workbook --- \textbf{70-70-70 hypertension/cholesterol + diabetes cascade} and \textbf{cascade + tobacco tax:} joint cascade/public-health workbook \texttt{indonesia\_model\_cost\_value\_70\_30\_30\_to\_70\_70\_70\_public\_health\_formulae.xlsx} (scenarios \texttt{S\_70\_30\_30\_TO\_70\_70\_70} and \texttt{S\_70\_30\_30\_TO\_70\_70\_70\_I\_PH\_TOBACCO\_TAX}); \textbf{Tobacco tax:} public-health workbook \texttt{indonesia\_cost\_value\_public\_health\_formulae.xlsx} (scenario \texttt{I\_PH\_TOBACCO\_TAX}); \textbf{All public-health interventions:} the SAME public-health workbook (scenario \texttt{all\_public\_health}); \textbf{All clinical and public-health interventions:} combined workbook \texttt{indonesia\_model\_cost\_value\_clinical\_public\_health\_formulae.xlsx} (scenario \texttt{all\_clinical\_public\_health}). ASEAN comparators in \texttt{data/raw/}. \textbf{Method:} every intervention line is a \emph{single joint model run} applying its listed interventions simultaneously --- never a sum, product, or average of separately modelled effects; all source workbooks share an identical baseline (asserted), so every line is the probability of death from the six modelled cardiovascular causes over exact ages 30--70. The plotted purple ``cascade + tobacco tax'' line (scenario \texttt{S\_70\_30\_30\_TO\_70\_70\_70\_I\_PH\_TOBACCO\_TAX}) adds ONLY the tobacco tax to the cascade --- it does NOT include the clean-air law, mass-media campaign, or advertising ban, and is NOT the broader \texttt{all\_cascade\_public\_health} package. Comparators are WPP-scaled reference series shown for context only.}

# Where the gains fall: a province-level view of the cascade

## `r title_sub_pareto`

```{r fig-sub-pareto, fig.width=9.7, fig.height=3.5, out.width="\\linewidth"}
sf_scale <- max(da_prov$da)
reg_fig(
ggplot(da_prov, aes(x = location)) +
  geom_col(aes(y = da), fill = col_clinical, width = 0.72) +
  geom_line(aes(y = cum_share / 100 * sf_scale, group = 1), colour = col_casefatal, linewidth = 0.7) +
  geom_point(aes(y = cum_share / 100 * sf_scale), colour = col_casefatal, size = 0.9) +
  scale_y_continuous(
    name = "Cumulative deaths averted, 2025-2050 (thousands)",
    labels = function(x) formatC(x / 1e3, format = "d", big.mark = ","),
    expand = expansion(mult = c(0, 0.02)),
    sec.axis = sec_axis(~ . / sf_scale * 100, name = "Cumulative share of national total (%)",
                        labels = function(x) paste0(x, "%"))) +
  labs(x = NULL) +
  theme_exec(10) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 5.6),
        axis.title.y.left  = element_text(colour = col_clinical, size = 8),
        axis.title.y.right = element_text(colour = col_casefatal, size = 8),
        panel.grid.major.x = element_blank()),
"fig-sub-pareto", title_sub_pareto, 9.7, 3.5)
```

\srcmethod{\textbf{Source:} isolated subnational cascade workbook (\texttt{...\_formulae\_subnational.xlsx}), sheet \texttt{Health\_Outcomes}. \textbf{Method:} bars are cumulative deaths averted `r HZ_START`--`r HZ_END` per province (official cascade scope: six CVD causes $+$ type-2 diabetes); the line is the running share of the national total. Province-summed baseline deaths reconcile to \texttt{Province\_Reconciliation} (`r sub_recon_status`; residual `r fmt_pct(sub_recon_reldiff,2)`). `r sub_caveat` `r sub_hetero`}

## `r title_sub_cov`

```{r fig-sub-cov, fig.width=9.4, fig.height=3.0, out.width="\\linewidth"}
cov_map <- maps_two_year(cov_prov, c(HZ_START, sub_ytar2), c(as.character(HZ_START), as.character(sub_ytar2)),
                         agg = "popw", valcol = "cov")
cov_lims <- pos_scale_lims(cov_map$value)   # bar starts at the min positive displayed value
reg_fig(
ggplot(cov_map) +
  geom_sf(aes(fill = value), colour = "white", linewidth = 0.12) +
  facet_wrap(~panel) +
  scale_fill_gradientn(colours = ramp_blue, limits = cov_lims, oob = scales::squish,
                       labels = function(x) paste0(formatC(x * 100, format = "f", digits = 0), "%"),
                       name = "CVD primary-care effective coverage used by the model") +
  coord_sf(expand = FALSE) +
  theme_map_exec(10),
"fig-sub-cov", title_sub_cov, 9.4, 3.0)
```

\srcmethod{\textbf{Source:} subnational cascade workbook, sheet \texttt{Cascade\_Trajectory}. \textbf{Method:} CVD primary-care effective coverage (the model's own effect/cost driver) for `r HZ_START` and `r sub_ytar2`; the workbook provides no sex-population split, so the two sexes are pooled with equal weight (their modelled coverage is identical from the `r sub_ytar2` target onward); one common scale. `r sub_geo_note` `r sub_caveat` `r sub_hetero`}

## `r title_sub_cvddeath`

```{r fig-sub-cvddeath, fig.width=9.4, fig.height=3.0, out.width="\\linewidth"}
## % reduction in CVD deaths: aggregate summed baseline & scenario CVD deaths to the
## boundary polygon, then recompute the reduction (never average percentages).
cvd_map <- do.call(rbind, lapply(c(sub_ytar1, sub_ytar2), function(yr) {
  d <- merge(cvd_red[year == yr], prov_meta[, .(location, iso)], by = "location")
  a <- d[, .(base = sum(base), scen = sum(scen)), by = iso]
  a[, value := (base - scen) / base * 100]
  s <- map_sf(a[, .(iso, value)]); s$panel <- as.character(yr); s
}))
cvd_lims <- pos_scale_lims(cvd_map$value)   # bar starts at the min positive displayed value
reg_fig(
ggplot(cvd_map) +
  geom_sf(aes(fill = value), colour = "white", linewidth = 0.12) +
  facet_wrap(~panel) +
  scale_fill_gradientn(colours = ramp_grn, limits = cvd_lims, oob = scales::squish,
                       labels = function(x) paste0(formatC(x, format = "f", digits = 1), "%"),
                       name = "Reduction in CVD deaths vs baseline (%)") +
  coord_sf(expand = FALSE) +
  theme_map_exec(10),
"fig-sub-cvddeath", title_sub_cvddeath, 9.4, 3.0)
```

\srcmethod{\textbf{Source:} subnational cascade workbook, sheet \texttt{Annual\_Mortality}. \textbf{Method:} percentage reduction in deaths from the six modelled CVD causes (type-2 diabetes excluded) $=$ (baseline $-$ scenario) $\div$ baseline, for `r sub_ytar1` and `r sub_ytar2`; one common scale. `r sub_geo_note` `r sub_caveat` `r sub_hetero`}

## `r title_sub_40q30`

```{r fig-sub-40q30, fig.width=9.4, fig.height=3.0, out.width="\\linewidth"}
q40_map <- maps_two_year(q40, c(sub_ytar1, sub_ytar2), c(as.character(sub_ytar1), as.character(sub_ytar2)),
                         agg = "popw", valcol = "pctred")
q40_lims <- pos_scale_lims(q40_map$value)   # bar starts at the min positive displayed value
reg_fig(
ggplot(q40_map) +
  geom_sf(aes(fill = value), colour = "white", linewidth = 0.12) +
  facet_wrap(~panel) +
  scale_fill_gradientn(colours = ramp_grn, limits = q40_lims, oob = scales::squish,
                       labels = function(x) paste0(formatC(x, format = "f", digits = 1), "%"),
                       name = "Reduction in the probability of death from CVD before age 70 (%)") +
  coord_sf(expand = FALSE) +
  theme_map_exec(10),
"fig-sub-40q30", title_sub_40q30, 9.4, 3.0)
```

\srcmethod{\textbf{Source:} subnational cascade workbook, probability-of-death-before-70 sheet (R-reconciled literal probabilities). \textbf{Method:} percentage reduction in the probability of death from CVD before age 70 $=$ (baseline $-$ scenario) $\div$ baseline, for `r sub_ytar1` and `r sub_ytar2`, pooled to the boundary polygon with population weights; one common scale. `r sub_geo_note` `r sub_caveat` `r sub_hetero`}

## `r title_sub_cost`

```{r fig-sub-cost, fig.width=9.4, fig.height=3.0, out.width="\\linewidth"}
cost_map <- maps_two_year(cost_prov, c(sub_ytar1, sub_ytar2), c(as.character(sub_ytar1), as.character(sub_ytar2)),
                          agg = "ratio")
cost_lims <- pos_scale_lims(cost_map$value)   # bar starts at the min positive displayed value
reg_fig(
ggplot(cost_map) +
  geom_sf(aes(fill = value), colour = "white", linewidth = 0.12) +
  facet_wrap(~panel) +
  scale_fill_gradientn(colours = ramp_org, limits = cost_lims, oob = scales::squish,
                       breaks = scales::breaks_pretty(n = 4),
                       labels = function(x) formatC(x, format = "f", digits = 2),
                       name = "Annual additional cost per capita (US$ per person, discounted)") +
  coord_sf(expand = FALSE) +
  theme_map_exec(10),
"fig-sub-cost", title_sub_cost, 9.4, 3.0)
```

\srcmethod{\textbf{Source:} subnational cascade workbook, sheets \texttt{Annual\_Cost} / \texttt{Budget\_Impact}. \textbf{Method:} \emph{annual} (not cumulative) discounted additional cost in each province divided by that province's population, for `r sub_ytar1` and `r sub_ytar2` (discounted `r fmt_pct(sub_disc*100,0)`/yr to `r HZ_START`, `r sub_price` US\$); one common scale. `r sub_geo_note` `r sub_caveat` `r sub_hetero`}

## `r title_sub_cost_undisc`

```{r fig-sub-cost-undisc, fig.width=9.4, fig.height=3.0, out.width="\\linewidth"}
## Undiscounted counterpart of fig-sub-cost: each province's ANNUAL UNDISCOUNTED
## additional cost / that province's population (sum(cost)/sum(pop) per boundary
## polygon), for the two milestone years. Same geometry, same common positive-anchored
## legend rule (zeros squished to the low colour, no detached US$0 tick).
cost_map_u <- do.call(rbind, lapply(c(sub_ytar1, sub_ytar2), function(yr) {
  d <- merge(cost_prov_u[year == yr], prov_meta[, .(location, iso)], by = "location")
  a <- d[, .(value = sum(undisc) / sum(pop)), by = iso]
  s <- map_sf(a); s$panel <- as.character(yr); s
}))
cost_lims_u <- pos_scale_lims(cost_map_u$value)
reg_fig(
ggplot(cost_map_u) +
  geom_sf(aes(fill = value), colour = "white", linewidth = 0.12) +
  facet_wrap(~panel) +
  scale_fill_gradientn(colours = ramp_org, limits = cost_lims_u, oob = scales::squish,
                       breaks = scales::breaks_pretty(n = 4),
                       labels = function(x) formatC(x, format = "f", digits = 2),
                       name = "Annual additional cost per capita (US$ per person, undiscounted)") +
  coord_sf(expand = FALSE) +
  theme_map_exec(10),
"fig-sub-cost-undisc", title_sub_cost_undisc, 9.4, 3.0)
```

\srcmethod{\textbf{Source:} subnational cascade workbook, sheets \texttt{Annual\_Cost} / \texttt{Budget\_Impact}. \textbf{Method (undiscounted):} \emph{annual} (not cumulative) \textbf{undiscounted} additional cost in each province divided by that province's population, for `r sub_ytar1` and `r sub_ytar2` (`r sub_price` US\$); one common scale. `r sub_geo_note` `r sub_caveat` `r sub_hetero`}

# What this means for policy

## Prevention and clinical care together offer the biggest, most affordable gains

\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep4pt
```{r impl-left, results='asis'}
if (has_joint)
  cat(sprintf("  \\item \\textbf{Biggest health gains come from doing both.} The combined package averts far more (\\textasciitilde%s) than either family alone.\n", fmt_short(combo_da)))
cat("  \\item \\textbf{Public-health policies are exceptional value.} Population-wide fiscal/regulatory measures avert deaths at a fraction of clinical cost per death averted.\n")
cat("  \\item \\textbf{Clinical scale-up delivers large, durable gains} in survival and premature mortality.\n")
```
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep4pt
  \item \textbf{Sequence for early wins.} Tobacco taxation and other prevention policies can start cutting risk quickly and cheaply, while clinical capacity scales up.
```{r impl-right, results='asis'}
if (has_joint)
  cat(sprintf("  \\item \\textbf{The premature-mortality target is within reach.} The combined package bends Indonesia's probability of death from CVD before age 70 down by about %s by %d.\n", fmt_pct(combo_qred,0), HZ_END))
```
\end{itemize}
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} results contract (\texttt{09\_deck\_results.rds}); cascade workbook. \textbf{Method:} interpretation of modelled deaths averted, premature-mortality reduction, and cost per death averted; policy translation needs judgement on feasibility, financing, and delivery. \costcaveat}

## What these results can and cannot tell us

\begin{itemize}\footnotesize\itemsep3pt
  \item Results depend on \textbf{modelled epidemiological estimates} (GBD) and calibration; the real burden may differ.
  \item \textbf{Effect sizes, future coverage, and unit costs are uncertain}; interventions are assumed to reach and sustain their targets.
  \item The disease model is \textbf{simplified} (healthy $\to$ living with CVD $\to$ death; no recovery; no comorbidity interactions).
  \item Value for money is expressed as \textbf{cost per death averted}; effects on illness (morbidity) and productivity are not valued in money here.
  \item \textbf{Costing results are preliminary.} Unit costs, service use, and delivery assumptions require further validation against \textbf{BPJS and Ministry of Health} data before policy use.
  \item Projections are \textbf{scenarios, not forecasts or guarantees}; they show what is achievable if interventions are delivered as modelled.
\end{itemize}

\srcmethod{\textbf{Source:} Indonesia CVD model and its input workbooks (GBD 2023; UN WPP 2024; FAIR Choices effects; programme and policy unit costs). \textbf{Method:} qualitative statement of the assumptions and uncertainties that bound the quantitative results above.}

## The case for action: two levers, affordable, and a reachable target

\begin{columns}[T]
\begin{column}{0.33\textwidth}
```{r remember-1, results='asis'}
if (has_joint) {
  cat(sprintf("\\begin{combinedblock}{1. Both, together}\n\\footnotesize Clinical + public health averts \\textbf{\\textasciitilde%s deaths} by %d --- more than either alone.\n\\end{combinedblock}\n", fmt_short(combo_da), HZ_END))
} else {
  cat("\\begin{combinedblock}{1. Two levers}\n\\footnotesize Clinical care and population prevention each cut premature CVD mortality.\n\\end{combinedblock}\n")
}
```
\end{column}
\begin{column}{0.33\textwidth}
```{r remember-2, results='asis'}
cat(sprintf("\\begin{phblock}{2. Prevention is affordable}\n\\footnotesize Population policies avert deaths at about \\textbf{%s} each.\n\\end{phblock}\n", fmt_usd(ph_cpda %+na% NA_real_)))
```
\end{column}
\begin{column}{0.33\textwidth}
```{r remember-3, results='asis'}
if (has_joint) {
  cat(sprintf("\\begin{clinicalblock}{3. A reachable target}\n\\footnotesize The probability of death from CVD before age 70 falls \\textbf{%s} by %d under the combined package.\n\\end{clinicalblock}\n", fmt_pct(combo_qred,0), HZ_END))
} else {
  cat(sprintf("\\begin{clinicalblock}{3. A reachable target}\n\\footnotesize All clinical interventions cut the probability of death from CVD before age 70 by \\textbf{%s} by %d.\n\\end{clinicalblock}\n", fmt_pct(clin_qred,0), HZ_END))
}
```
\end{column}
\end{columns}

\vspace{4pt}
```{r closing-banner, results='asis'}
if (has_joint) {
  cat(sprintf(paste0("\\begin{center}\n\\large Combined action against CVD could avert about \\hlkey{%s deaths} ",
                     "at about \\hlkey{%s per death averted} through %d.\n\\end{center}\n"),
              fmt_short(combo_da), fmt_usd(combo_cpda), HZ_END))
} else {
  cat(sprintf(paste0("\\begin{center}\n\\large Scaling up CVD interventions could avert hundreds of thousands ",
                     "of deaths through %d at a modest cost per death averted.\n\\end{center}\n"), HZ_END))
}
```

\srcmethod{\textbf{Source:} results contract (\texttt{09\_deck\_results.rds}). \textbf{Method:} deaths averted cumulative `r HZ_START`--`r HZ_END`; reduction in the probability of death from CVD before age 70 in `r HZ_END`; cost per death averted uses discounted additional cost. \costcaveat}

## Undiscounted, the case still holds: two levers, affordable, and a reachable target

\begin{columns}[T]
\begin{column}{0.33\textwidth}
```{r remember-1-undisc, results='asis'}
if (has_joint) {
  cat(sprintf("\\begin{combinedblock}{1. Both, together}\n\\footnotesize Clinical + public health averts \\textbf{\\textasciitilde%s deaths} by %d --- more than either alone.\n\\end{combinedblock}\n", fmt_short(combo_da), HZ_END))
} else {
  cat("\\begin{combinedblock}{1. Two levers}\n\\footnotesize Clinical care and population prevention each cut premature CVD mortality.\n\\end{combinedblock}\n")
}
```
\end{column}
\begin{column}{0.33\textwidth}
```{r remember-2-undisc, results='asis'}
cat(sprintf("\\begin{phblock}{2. Prevention is affordable}\n\\footnotesize Population policies avert deaths at about \\textbf{%s} each (undiscounted).\n\\end{phblock}\n", fmt_usd(ph_cpda_undisc %+na% NA_real_)))
```
\end{column}
\begin{column}{0.33\textwidth}
```{r remember-3-undisc, results='asis'}
if (has_joint) {
  cat(sprintf("\\begin{clinicalblock}{3. A reachable target}\n\\footnotesize The probability of death from CVD before age 70 falls \\textbf{%s} by %d under the combined package.\n\\end{clinicalblock}\n", fmt_pct(combo_qred,0), HZ_END))
} else {
  cat(sprintf("\\begin{clinicalblock}{3. A reachable target}\n\\footnotesize All clinical interventions cut the probability of death from CVD before age 70 by \\textbf{%s} by %d.\n\\end{clinicalblock}\n", fmt_pct(clin_qred,0), HZ_END))
}
```
\end{column}
\end{columns}

\vspace{4pt}
```{r closing-banner-undisc, results='asis'}
if (has_joint) {
  cat(sprintf(paste0("\\begin{center}\n\\large Combined action against CVD could avert about \\hlkey{%s deaths} ",
                     "at about \\hlkey{%s per death averted (undiscounted)} through %d.\n\\end{center}\n"),
              fmt_short(combo_da), fmt_usd(combo_cpda_undisc), HZ_END))
} else {
  cat(sprintf(paste0("\\begin{center}\n\\large Scaling up CVD interventions could avert hundreds of thousands ",
                     "of deaths through %d at a modest undiscounted cost per death averted.\n\\end{center}\n"), HZ_END))
}
```

\srcmethod{\textbf{Source:} results contract (\texttt{09\_deck\_results.rds}). \textbf{Method (undiscounted):} deaths averted cumulative `r HZ_START`--`r HZ_END`; reduction in the probability of death from CVD before age 70 in `r HZ_END`; cost per death averted uses \textbf{undiscounted} additional cost (`r price_year` US\$). Health results are identical to the discounted slide. \costcaveat}

# Technical appendix

## Appendix 1 --- Scenario crosswalk

```{r tbl-crosswalk}
## Built dynamically from the current run's combined-workbook scenarios.
fam_order <- c(clinical = 1, public_health = 2, clinical_public_health = 3)
xwd <- copy(met_comb)[, .(scenario, intervention_family, scenario_label, n_interventions)]
xwd[, ford := fam_order[intervention_family]]
xwd[is.na(ford), ford := 9]
setorder(xwd, ford, -n_interventions)
fam_disp <- c(clinical = "Clinical (FAIR Choices)", public_health = "Public health",
              clinical_public_health = "Combined")
xwd[, Family := fam_disp[intervention_family]]
xwd[is.na(Family), Family := intervention_family]
## Only show the family label on its first row (visual grouping).
xwd[, Family := ifelse(duplicated(Family), "", Family)]
xwd[, IDesc := gsub("_", "\\\\_", scenario)]
xw <- xwd[, .(Family, ID = IDesc, Label = scenario_label)]
reg_tbl("tbl-crosswalk", "Appendix 1 - Scenario crosswalk", xw,
        c("Family", "Workbook scenario id", "Executive label"))
kbl(xw, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=c("Family","Workbook scenario id","Executive label"), align="lll") |>
  kable_styling(latex_options=c("hold_position"), font_size=5.2) |>
  column_spec(1, bold=TRUE, width="3.0cm") |> column_spec(2, monospace=TRUE, width="4.0cm") |>
  column_spec(3, width="6.6cm") |> row_spec(0, bold=TRUE)
```

\srcmethod{\textbf{Source:} combined workbook \texttt{Scenario\_Catalog} (binding active-scenario contract). \textbf{Method:} identity/labels/family read from the current run; excluded interventions never appear; packages are single joint runs of their children (omitted from main tables to avoid double-counting).}

## Appendix 2 --- Definitions, horizons and assumptions

\footnotesize
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\textbf{Metrics}
\begin{itemize}\scriptsize\itemsep2pt
  \item \textbf{Deaths averted (cumulative)} $=\sum_{t=`r HZ_START`}^{`r HZ_END`}(\text{baseline deaths}_t-\text{scenario deaths}_t)$, across modelled CVD \& diabetes causes.
  \item \textbf{\% reduction in deaths, `r HZ_END`} $=100\times(\text{baseline}-\text{scenario})/\text{baseline}$ in `r HZ_END`.
  \item \textbf{Probability of death from CVD before age 70} $=$ probability of dying of CVD between exact ages 30 and 70 (six CVD causes, both sexes, single-year life table).
  \item \textbf{\% reduction in that probability, `r HZ_END`}: analogous to deaths.
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\textbf{Costing \& value for money}
\begin{itemize}\scriptsize\itemsep2pt
  \item Horizon `r HZ_START`--`r HZ_END`; cost discount rate `r fmt_pct(disc_rate*100,0)`/yr; price year `r price_year`.
  \item Costs in \textbf{market US\$}; cost perspective: `r cost_perspective`.
  \item \textbf{Cost per death averted} $=$ cumulative \textbf{discounted} additional (incremental) cost $\div$ cumulative \textbf{undiscounted} deaths averted.
  \item A ``Dominant'' scenario averts deaths while lowering cost; ``not defined'' means no deaths averted.
\end{itemize}
\end{column}
\end{columns}

\vspace{3pt}
\scriptsize \textbf{Model.} Discrete-time state-transition model; single-year ages 0--95; six CVD causes (IHD, ischemic stroke, intracerebral haemorrhage, hypertensive heart disease, rheumatic heart disease, cardiomyopathy \& myocarditis) plus type-2 diabetes; calibrated to GBD 2023; demography and life expectancy from UN WPP 2024. \textbf{Data provenance:} every ordinary-run number is read from the current-run results contract (\texttt{output/09\_deck\_results.rds}) and the workbooks' literal probability-of-death series; the cascade strategy is read from its isolated workbook; scenario identity follows the \texttt{Scenario\_Catalog} contract (binding \texttt{include\_flag}). No Excel recalculation is used.

## Appendix 3 --- How costs are calculated

\footnotesize Intervention-based costing adapted from the DCP Cost Model of @Watkins2020ResourceRequirements, from a `r cost_perspective` perspective with annual projections. For intervention $i$, sex/age group $s$, year $t$:

\vspace{-2pt}
\begin{columns}[T]
\begin{column}{0.52\textwidth}
{\scriptsize
\begin{align*}
C_{i,s,t} &= PIN_{i,s,t}\times cov_{i,s,t}\times f_i\times uc_i\\[2pt]
\Delta C_{s,t} &= C_{s,t}-C_{0,t}\\[2pt]
PV(\Delta C_s) &= \sum_{t=`r HZ_START+1L`}^{`r HZ_END`}\frac{\Delta C_{s,t}}{(1+r)^{\,t-`r HZ_START`}}\\[2pt]
\text{Cost per death prevented}_s &= \frac{PV(\Delta C_s)}{\sum_{t=`r HZ_START+1L`}^{`r HZ_END`}(D_{0,t}-D_{s,t})}
\end{align*}}
\end{column}
\begin{column}{0.46\textwidth}
\scriptsize
$PIN$ = people in need (population $\times$ epidemiological need); $cov$ = coverage (baseline or scenario); $f_i$ = services per person per year; $uc_i$ = Indonesia-adjusted unit cost; subscript $0$ = baseline (current course), $s$ = scenario; $r$ = `r fmt_pct(disc_rate*100,0)`/yr discount rate; price year `r price_year`; currency market US\$; perspective `r cost_perspective`. Deaths $D$ remain \textbf{undiscounted} in the cost-per-death calculation. Downstream cost offsets are \textbf{excluded}.
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} adapted from Watkins et al. (2020); unit costs, coverage, and people-in-need from the input workbooks and \texttt{Annual\_Cost}; demography from UN WPP 2024. \textbf{Method:} annual costs by intervention $\times$ group $\times$ year, differenced against baseline, discounted to the `r HZ_START` base year. \costcaveat}

## Appendix 4 --- References

\scriptsize

::: {#refs}
:::

```{r build-xlsx-workbook, include=FALSE}
## =====================================================================
##  ASSEMBLE + WRITE THE COMPANION WORKBOOK: reports/executive_slides.xlsx
##  Runs after every slide chunk, so .xlsx_items holds every figure and
##  table captured (in slide order) by reg_fig()/reg_tbl(). Each item gets
##  its own logically-named worksheet (<=31 chars) labelled with the slide
##  title and the chunk/figure/table id; tables are written as real,
##  editable cells; figures are embedded as PNG images at readable
##  resolution. A leading Contents sheet indexes everything.
##
##  Wrapped in tryCatch so a workbook hiccup can never break the deck:
##  the slide PDF is the primary artifact and its rendering is untouched.
## =====================================================================
if (isTRUE(.xlsx_enabled) && length(.xlsx_items)) tryCatch({

  xlsx_out <- here::here("reports", "executive_slides.xlsx")

  ## ---- Strip LaTeX markup from a string so cells/labels are plain text ----
  latex_clean <- function(x) {
    x <- as.character(x); x[is.na(x)] <- ""
    ## \textcolor{colour}{text} -> text (two braced args; keep the second)
    x <- gsub("\\\\textcolor\\{[^{}]*\\}\\{([^{}]*)\\}", "\\1", x)
    x <- gsub("\\\\makecell\\[[^]]*\\]\\{([^{}]*)\\}", "\\1", x)
    ## remaining single-argument commands \cmd{...} -> ... (iterate for nesting)
    repeat { y <- gsub("\\\\[A-Za-z@]+\\*?\\{([^{}]*)\\}", "\\1", x); if (identical(y, x)) break; x <- y }
    ## escaped specials
    x <- gsub("US\\\\\\$", "US$", x)
    x <- gsub("\\\\%", "%", x); x <- gsub("\\\\\\$", "$", x)
    x <- gsub("\\\\_", "_", x); x <- gsub("\\\\&", "&", x); x <- gsub("\\\\#", "#", x)
    x <- gsub("\\$\\\\approx\\$", "~", x)
    x <- gsub("\\\\textasciitilde\\{\\}", "~", x)
    ## any leftover backslash commands (no braces) then stray braces
    x <- gsub("\\\\[A-Za-z@]+", "", x); x <- gsub("[{}]", "", x)
    ## header line breaks -> spaces; collapse whitespace
    x <- gsub("[\r\n]+", " ", x); x <- gsub("[ \t]+", " ", x)
    trimws(x)
  }
  ## ---- Sheet-name helper: Excel forbids \ / ? * [ ] : and caps at 31 chars --
  sanitize_sheet <- function(s) substr(gsub("[\\\\/?*\\[\\]:]", "-", s), 1, 31)

  ## openxlsx 4.2.x registers a drawing + vmlDrawing relationship (and a drawing
  ## Content_Types override) for EVERY worksheet, but writes drawing<i>.xml only for
  ## sheets that hold an image and never writes the vmlDrawing parts -- leaving
  ## dangling part references that Excel tolerates but strict readers (openpyxl /
  ## LibreOffice / Sheets) reject. Remove ONLY the phantom references, keeping the
  ## real embedded-image drawings intact. Guarded no-op on a future openxlsx layout.
  strip_phantom_drawings <- function(wb) {
    tryCatch({
      has_img <- vapply(seq_along(wb$worksheets),
                        function(i) !is.null(wb$drawings[[i]]) && length(wb$drawings[[i]]) > 0, logical(1))
      if (!is.null(wb$Content_Types)) {
        is_draw  <- grepl("/xl/drawings/drawing[0-9]+\\.xml", wb$Content_Types)
        keep_pat <- if (any(has_img))
          paste0("PartName=\"/xl/drawings/drawing", which(has_img), "\\.xml\"", collapse = "|") else NA_character_
        keep_img <- if (is.na(keep_pat)) rep(FALSE, length(wb$Content_Types)) else grepl(keep_pat, wb$Content_Types)
        wb$Content_Types <- wb$Content_Types[!is_draw | keep_img]
      }
      if (!is.null(wb$worksheets_rels))
        for (i in seq_along(wb$worksheets_rels)) {
          r <- wb$worksheets_rels[[i]]; if (!length(r)) next
          r <- r[!grepl("/vmlDrawing", r)]                       # vmlDrawing parts are never written
          if (!has_img[i]) r <- r[!grepl("/drawings/drawing[0-9]+\\.xml", r)]
          wb$worksheets_rels[[i]] <- r
        }
    }, error = function(e) message("  (phantom-drawing strip skipped: ", conditionMessage(e), ")"))
    invisible(wb)
  }

  wb <- openxlsx::createWorkbook()
  st_title <- openxlsx::createStyle(fontSize = 13, textDecoration = "bold", fontColour = "#0F2A56")
  st_sub   <- openxlsx::createStyle(fontSize = 9,  textDecoration = "italic", fontColour = "#5B6B7B")
  st_link  <- openxlsx::createStyle(fontColour = "#3B6FB6", textDecoration = "underline")
  st_hdr   <- openxlsx::createStyle(textDecoration = "bold", fgFill = "#EAF1FB",
                                    halign = "center", valign = "center", wrapText = TRUE,
                                    border = "TopBottom", borderColour = "#B7C7DA")
  st_ihdr  <- openxlsx::createStyle(textDecoration = "bold", fgFill = "#0F2A56", fontColour = "white")

  ## Pre-compute one worksheet name per item.
  item_sheets <- vapply(seq_along(.xlsx_items),
                        function(i) sanitize_sheet(sprintf("%02d_%s", i, .xlsx_items[[i]]$id)), character(1))

  ## ---- Contents / index sheet (added first => leftmost tab) ---------------
  openxlsx::addWorksheet(wb, "Contents")
  openxlsx::writeData(wb, "Contents", "Executive deck - companion workbook", startCol = 1, startRow = 1)
  openxlsx::addStyle(wb, "Contents", st_title, rows = 1, cols = 1)
  openxlsx::writeData(wb, "Contents",
    sprintf("Every figure and table from executive_slides.Rmd, in slide order (generated %s).",
            format(Sys.time(), "%Y-%m-%d %H:%M")),
    startCol = 1, startRow = 2)
  openxlsx::addStyle(wb, "Contents", st_sub, rows = 2, cols = 1)
  contents <- data.frame(
    `#`            = seq_along(.xlsx_items),
    Type           = vapply(.xlsx_items, function(z) z$kind, character(1)),
    `Slide title`  = vapply(.xlsx_items, function(z) latex_clean(z$title), character(1)),
    Identifier     = vapply(.xlsx_items, function(z) z$id, character(1)),
    Worksheet      = item_sheets,
    check.names = FALSE, stringsAsFactors = FALSE)
  openxlsx::writeData(wb, "Contents", contents, startRow = 4, startCol = 1, colNames = TRUE)
  openxlsx::addStyle(wb, "Contents", st_ihdr, rows = 4, cols = seq_len(ncol(contents)), gridExpand = TRUE)
  openxlsx::setColWidths(wb, "Contents", cols = 1:5, widths = c(5, 9, 62, 26, 24))
  openxlsx::freezePane(wb, "Contents", firstActiveRow = 5)

  ## ---- One worksheet per item --------------------------------------------
  for (i in seq_along(.xlsx_items)) {
    it <- .xlsx_items[[i]]; sh <- item_sheets[i]
    openxlsx::addWorksheet(wb, sh)
    openxlsx::writeData(wb, sh, latex_clean(it$title), startRow = 1, startCol = 1)
    openxlsx::addStyle(wb, sh, st_title, rows = 1, cols = 1)
    openxlsx::writeData(wb, sh, sprintf("%s  |  %s", toupper(it$kind), it$id), startRow = 2, startCol = 1)
    openxlsx::addStyle(wb, sh, st_sub, rows = 2, cols = 1)
    try({
      openxlsx::writeFormula(wb, sh,
        x = openxlsx::makeHyperlinkString(sheet = "Contents", row = 1, col = 1, text = "<- Contents"),
        startRow = 3, startCol = 1)
      openxlsx::addStyle(wb, sh, st_link, rows = 3, cols = 1)
    }, silent = TRUE)

    r0 <- 5L
    if (identical(it$kind, "table")) {
      df <- it$data
      ## Cleaned display headers: prefer the slide's own col_names, else the object's names.
      hdr <- if (!is.null(it$col_names) && length(it$col_names) == ncol(df))
               latex_clean(it$col_names) else latex_clean(names(df))
      names(df) <- make.unique(hdr)
      ## Clean every character/factor cell to plain text (LaTeX markup -> readable).
      for (j in seq_along(df))
        if (is.character(df[[j]]) || is.factor(df[[j]])) df[[j]] <- latex_clean(as.character(df[[j]]))
      openxlsx::writeData(wb, sh, df, startRow = r0, startCol = 1, colNames = TRUE, headerStyle = st_hdr)
      ## Readable, capped column widths.
      widths <- vapply(seq_along(df), function(j) {
        w <- max(nchar(names(df)[j]), max(nchar(as.character(df[[j]])), 0L, na.rm = TRUE))
        min(60, max(12, w + 2))
      }, numeric(1))
      openxlsx::setColWidths(wb, sh, cols = seq_along(df), widths = widths)
      openxlsx::freezePane(wb, sh, firstActiveRow = r0 + 1)
    } else {
      if (!is.na(it$path) && file.exists(it$path)) {
        openxlsx::insertImage(wb, sh, file = it$path, startRow = r0, startCol = 1,
                              width = it$width, height = it$height, units = "in", dpi = 200)
      } else {
        openxlsx::writeData(wb, sh, "[figure image could not be generated for this run]",
                            startRow = r0, startCol = 1)
      }
    }
  }

  strip_phantom_drawings(wb)
  openxlsx::saveWorkbook(wb, xlsx_out, overwrite = TRUE)
  message(sprintf("Companion workbook written: %s  (%d items: %d figures, %d tables)",
                  xlsx_out, length(.xlsx_items),
                  sum(vapply(.xlsx_items, function(z) z$kind == "figure", logical(1))),
                  sum(vapply(.xlsx_items, function(z) z$kind == "table",  logical(1)))))
}, error = function(e)
  warning("Companion workbook (executive_slides.xlsx) was NOT written: ", conditionMessage(e), call. = FALSE))
```
