---
title: "Reducing Cardiovascular Disease in Indonesia"
subtitle: "Health Impact, Costs, and Cost-Effectiveness of Clinical and Public-Health Interventions, 2025--2050"
author: "World Bank $\\cdot$ University of Washington"
institute: "Briefing for the Ministry of Health, Republic of Indonesia"
date: "`r format(Sys.Date(), '%B %Y')`"
output:
  beamer_presentation:
    theme: "default"
    keep_tex: false
    latex_engine: xelatex
    slide_level: 2
    toc: false
    includes:
      in_header: beamer_preamble.tex
classoption: "aspectratio=169"
bibliography: references.bib
link-citations: false
header-includes:
  - '\definecolor{srcgrey}{HTML}{5B6B7B}'
  - '\newcommand{\srcmethod}[1]{{\par\vspace{1pt}\scriptsize\itshape\color{srcgrey}#1\par}}'
  - '\newcommand{\costcaveat}{\textbf{Preliminary costing results:} unit costs, service use, and delivery assumptions require further validation against BPJS and Ministry of Health data before policy use.}'
---

```{r setup, include=FALSE}
## =====================================================================
##  EXECUTIVE DECK - Indonesia CVD / FAIR Choices analysis
##  World Bank * University of Washington
##
##  DATA PROVENANCE (BCA-free; no Excel recalculation required)
##  Every displayed number is imported programmatically from the modeling
##  pipeline's current-run outputs:
##    * output/09_deck_results.rds - the BCA-FREE results contract written by
##      code/cvd-fair-choices/09_cost_value.R. It carries, per workbook family
##      and scenario, the SAME R-engine values the workbooks store: deaths
##      averted, incremental cost, cost per death averted, 2050 death levels and
##      % reduction, and CVD 40q30 (level, baseline, % reduction). Scenario
##      identity/labels/order follow the Scenario_Catalog contract, which honours
##      the binding Intervention_Cause_Map include_flag - excluded interventions
##      never appear.
##    * the combined workbook's literal CVD_40q30 per-year sheet (cvd_40q30_r),
##      used only for the trajectory figure and the baseline 40q30 scalars.
##    * data/raw/03_asean_cvd_40q30_comparison.csv (international comparator).
##  NOTHING is transcribed by hand. The deck reads only stored/literal values,
##  so it compiles WITHOUT Excel COM and on a clean non-Windows environment. It
##  displays NO benefit-cost / VSL / VSLY content; the value-for-money metric is
##  COST PER DEATH AVERTED (in market US$). The full benefit-cost analysis lives
##  in the pipeline workbooks and is intentionally not shown here.
##
##  2026 REFRESH (executive/plain-language pass):
##    * Every content-slide title is a plain-language takeaway; the metric
##      formerly labelled "40q30" is written as "Probability of death from CVD
##      before age 70" in ALL rendered text (internal object/sheet/column names
##      such as CVD_40q30 / cvd_40q30_r are kept in code only).
##    * Every analytical slide carries a muted italic Source/Method footnote
##      (\srcmethod macro, defined in header-includes); cost/CE slides also carry
##      the fixed BPJS/MOH validation caveat (\costcaveat).
##    * Average annual additional cost per capita (undiscounted & discounted) is
##      added to the three CE slides: numerator = the contract's cumulative
##      additional cost (= workbook Budget_Impact); denominator = 25 x the 2026
##      national population, counted ONCE per location x year x age x sex from the
##      baseline (current-course) demography in dt_output_dalys.rds (UN WPP 2024).
##    * A separate 70-30-30 -> 70-70-70 cascade strategy sub-section (4 slides)
##      reads ONLY the ISOLATED cascade workbook under output/70_30_30_to_70_70_70/
##      (validated to contain exactly {baseline, S_70_30_30_TO_70_70_70}); it is
##      NEVER merged into the ordinary results contract.
##    * A high-level DCP costing-methods slide + an appendix of the adapted
##      equations cite Watkins et al. (2020).
## =====================================================================
library(here)
library(data.table)
library(openxlsx)
library(ggplot2)
library(ggrepel)
library(scales)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.align = "center", dev = "cairo_pdf"
)

## ---- Palette (matches beamer_preamble.tex scenario colours) ---------------
col_clinical  <- "#3B6FB6"  # clinical scenarios
col_ph        <- "#1B9E77"  # public-health scenarios
col_combined  <- "#08306B"  # combined package (darker)
col_baseline  <- "#6E6E6E"  # baseline / status quo
col_comparator<- "#BDBDBD"  # comparator countries (muted)
col_incidence <- "#2166AC"  # incidence (Well -> Sick) effects
col_casefatal <- "#B2182B"  # case-fatality (Sick -> Death) effects
col_bgmort    <- "#7F7F7F"  # background mortality
col_wbnavy    <- "#0F2A56"

theme_exec <- function(base = 11) {
  theme_minimal(base_size = base) +
    theme(
      plot.title       = element_text(face = "bold", size = base + 1, colour = col_wbnavy),
      plot.subtitle    = element_text(colour = "grey35", size = base - 2),
      plot.caption     = element_text(colour = "grey45", size = base - 4, hjust = 0),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      legend.position  = "bottom",
      legend.title     = element_blank(),
      axis.title       = element_text(size = base - 2)
    )
}

## ---- Number formatters (LaTeX-safe: escaped $ and % ; used with escape=FALSE)
fmt_int   <- function(x) formatC(round(x), format = "d", big.mark = ",")
fmt_thou  <- function(x) ifelse(is.na(x), "--", formatC(round(x / 1000) * 1000, format = "d", big.mark = ","))
fmt_short <- function(x, acc = 0.1) ifelse(is.na(x), "--", label_number(scale_cut = cut_short_scale(), accuracy = acc)(x))
fmt_pct   <- function(x, d = 1) ifelse(is.na(x), "--", paste0(formatC(x, format = "f", digits = d), "\\%"))
fmt_usd   <- function(x) ifelse(is.na(x), "--", paste0("US\\$", formatC(round(x), format = "d", big.mark = ",")))
fmt_usd_short <- function(x) ifelse(is.na(x), "--", paste0("US\\$", fmt_short(x)))
## Per-capita US$ to 2 decimals (LaTeX-safe: escaped $).
fmt_usd_pc <- function(x) ifelse(is.na(x), "--", paste0("US\\$", formatC(x, format = "f", digits = 2, big.mark = ",")))
## Title-safe (plain-text, NO % or $ so it survives pandoc -> \frametitle):
## express a count in the most readable unit ("3.5 million", "1.13 million").
fmt_title_num <- function(x, acc = 0.1) {
  if (is.na(x)) return("--")
  if (abs(x) >= 1e6) return(paste0(formatC(x / 1e6, format = "f", digits = 1), " million"))
  if (abs(x) >= 1e3) return(paste0(formatC(round(x / 1e3), format = "d", big.mark = ","), " thousand"))
  formatC(round(x), format = "d", big.mark = ",")
}
## USD in billions/millions as plain words (no $) for prose reconciliation lines.
fmt_bn <- function(x, d = 2) ifelse(is.na(x), "--", paste0(formatC(x / 1e9, format = "f", digits = d), " billion"))
## Two-line table headers (converts embedded "\n" to \makecell line breaks).
mkhdr <- function(v) kableExtra::linebreak(v, align = "c")

## Null / NA coalescing helpers (base R has neither).
`%||%` <- function(a, b) if (is.null(a) || length(a) == 0L || (length(a) == 1L && is.na(a))) b else a
`%+na%` <- function(a, b) if (is.null(a) || length(a) == 0L || is.na(a[1])) b else a[1]
```

```{r load-data, include=FALSE}
## ---------------------------------------------------------------------------
## 1. Resolve current-run inputs (project-relative) and check existence.
##    The deck reads a BCA-free R-side results contract + literal workbook
##    values only -- no Excel COM, no formula recalculation.
## ---------------------------------------------------------------------------
HZ_START <- 2025L; HZ_END <- 2050L

## Output directory: defaults to the project's output/, but can be redirected
## (e.g. for testing an alternative active-scenario subset) via EXEC_DECK_OUTDIR.
.out_dir <- Sys.getenv("EXEC_DECK_OUTDIR", unset = "")
outp <- if (nzchar(.out_dir)) function(f) file.path(.out_dir, f) else function(f) here("output", f)
deck_file <- outp("09_deck_results.rds")
wb_comb   <- outp("indonesia_model_cost_value_clinical_public_health_formulae.xlsx")
wb_clin   <- outp("indonesia_model_cost_value_formulae.xlsx")
wb_ph     <- outp("indonesia_cost_value_public_health_formulae.xlsx")
dalys_file<- outp("dt_output_dalys.rds")   # authoritative model demography (per-capita denominator)
csv_cmp   <- here("data", "raw", "03_asean_cvd_40q30_comparison.csv")

## Isolated 70-30-30 -> 70-70-70 cascade run (its OWN output dir; never merged with the
## ordinary results contract). Overridable for testing, defaults to the standard path.
.casc_dir <- Sys.getenv("EXEC_DECK_CASCADE_DIR", unset = "")
casc_wb <- if (nzchar(.casc_dir)) file.path(.casc_dir, "indonesia_70_30_30_to_70_70_70_cost_value_formulae.xlsx") else
  here("output", "70_30_30_to_70_70_70", "indonesia_70_30_30_to_70_70_70_cost_value_formulae.xlsx")

.req_files <- c(results_contract = deck_file, combined_workbook = wb_comb,
                model_demography = dalys_file, comparator = csv_cmp, cascade_workbook = casc_wb)
.missing <- .req_files[!file.exists(.req_files)]
if (length(.missing))
  stop("VALIDATION STOP -- required input file(s) not found:\n  ",
       paste(sprintf("%s: %s", names(.missing), .missing), collapse = "\n  "),
       "\nRun the modeling pipeline (code/cvd-fair-choices/00_run_model_cvd_fair.R) first.",
       call. = FALSE)

rd        <- function(path, sheet) as.data.table(openxlsx::read.xlsx(path, sheet = sheet))
sheets_of <- function(path) openxlsx::getSheetNames(path)

## ---- Results contract -----------------------------------------------------
dk   <- readRDS(deck_file)
res  <- as.data.table(dk$results)
meta <- dk$meta

met_clin <- res[workbook == "clinical"]
met_ph   <- res[workbook == "public_health"]
met_comb <- res[workbook == "combined"]

## ---- Costing/report metadata (BCA-free: cost side only) -------------------
disc_rate  <- suppressWarnings(as.numeric(meta$cost_discount_rate)); if (!is.finite(disc_rate)) disc_rate <- 0.03
price_year <- suppressWarnings(as.numeric(meta$cost_price_year));    if (!is.finite(price_year)) price_year <- 2023
ce_unit    <- meta$cost_effectiveness_unit %||% "USD per death averted"
## LaTeX-safe escaper for dynamic strings inserted into prose with escape=FALSE.
ltx <- function(s) {
  s <- as.character(s)
  s <- gsub("\\\\", "\\\\textbackslash{}", s)
  s <- gsub("([&%$#_{}])", "\\\\\\1", s)
  s <- gsub("~", "\\\\textasciitilde{}", s)
  s
}
cost_perspective <- ltx(gsub("_", " ", meta$economic_perspective %||% "health system"))
```

```{r derive-scenarios, include=FALSE}
## ---------------------------------------------------------------------------
## 2. Scenario identity is DERIVED from the contract / Scenario_Catalog, never
##    from fixed ID vectors. Optional presentation-only label overrides may be
##    supplied; any active scenario without an override falls back safely to its
##    workbook label. No excluded / hard-coded ID is ever required to exist.
## ---------------------------------------------------------------------------
exec_label_override <- c(
  I_CVD_PRIMARY = "Primary CVD prevention", I_T2D_TREATMENT = "Type-2 diabetes treatment",
  I_IHD_SECONDARY = "Heart-attack secondary prevention", I_HF_CHRONIC = "Chronic heart-failure care",
  I_HF_ACUTE = "Acute heart-failure care", I_STROKE_SECONDARY = "Stroke secondary prevention",
  I_RHD_SECONDARY = "RHD secondary prevention",
  all = "All clinical (combined)",
  I_PH_TOBACCO_TAX = "Tobacco tax", I_PH_TOBACCO_POLICY = "Tobacco policy package",
  I_PH_SALT_POLICY = "Sodium policy package", I_PH_TFA_POLICY = "Trans-fat elimination",
  all_public_health = "All public health (combined)",
  all_clinical_public_health = "Clinical + public health (combined)",
  baseline = "Status quo (baseline)")

## Family aggregate / joint scenario ids (present only if built this run).
clin_all_id <- "all"; ph_all_id <- "all_public_health"; joint_id <- "all_clinical_public_health"
has_clin  <- nrow(met_clin[scenario != "baseline"]) > 0
has_ph    <- nrow(met_ph[scenario != "baseline"])   > 0
has_joint <- joint_id %in% met_comb$scenario

## Label getter: override first, else the contract's own scenario_label.
lbl <- function(sc, tab = met_comb) {
  if (sc %in% names(exec_label_override)) return(exec_label_override[[sc]])
  v <- tab[scenario == sc, scenario_label]
  if (length(v) && !is.na(v[1]) && nzchar(v[1])) v[1] else sc
}
## Metric getter.
hv <- function(sc, col, tab = met_comb) { v <- tab[scenario == sc][[col]]; if (length(v)) v[1] else NA_real_ }

## Top-level display rows for a family (exclude baseline and package CHILDREN to
## avoid double-counting; the family aggregate is placed last). Ordered by
## deaths averted. Robust to zero/one/many active interventions.
MAX_TBL_ROWS <- 14L
top_level_rows <- function(tab, agg_id) {
  if (!nrow(tab)) return(tab[0])
  d <- tab[scenario != "baseline" &
             (is.na(scenario_role) | !(scenario_role %in% c("child")))]
  if (!nrow(d)) return(d)
  d <- d[order(-deaths_averted)]
  agg <- d[scenario == agg_id]; rest <- d[scenario != agg_id]
  rbind(rest, agg)
}
```

```{r headline-scalars, include=FALSE}
## Baseline CVD 40q30 (per-year literal from the combined workbook).
c4c <- rd(wb_comb, "CVD_40q30")
qcol <- if ("cvd_40q30_r" %in% names(c4c)) "cvd_40q30_r" else "cvd_40q30"
base_q_2025 <- c4c[scenario == "baseline" & year == HZ_START][[qcol]][1]
base_q_2050 <- c4c[scenario == "baseline" & year == HZ_END][[qcol]][1]

## Combined-package (joint) headline metrics (guarded if joint absent).
combo_da   <- if (has_joint) hv(joint_id, "deaths_averted") else NA_real_
combo_cpda <- if (has_joint) hv(joint_id, "cost_per_death_averted") else NA_real_
combo_q50  <- if (has_joint) hv(joint_id, "cvd_40q30_2050") else NA_real_
combo_qred <- if (has_joint) hv(joint_id, "cvd_40q30_pct_reduction_2050") else NA_real_
combo_dred <- if (has_joint) hv(joint_id, "pct_reduction_deaths_2050") else NA_real_

clin_da   <- hv(clin_all_id, "deaths_averted", met_clin)
clin_qred <- hv(clin_all_id, "cvd_40q30_pct_reduction_2050", met_clin)
clin_cpda <- hv(clin_all_id, "cost_per_death_averted", met_clin)
ph_da     <- hv(ph_all_id, "deaths_averted", met_ph)
ph_qred   <- hv(ph_all_id, "cvd_40q30_pct_reduction_2050", met_ph)
ph_cpda   <- hv(ph_all_id, "cost_per_death_averted", met_ph)

## Most impactful single lever per family (excludes the aggregate).
.tc <- top_level_rows(met_clin, clin_all_id); .tc <- .tc[scenario != clin_all_id]
.tp <- top_level_rows(met_ph, ph_all_id);     .tp <- .tp[scenario != ph_all_id]
top_clin_sc <- if (nrow(.tc)) .tc$scenario[1] else NA_character_
top_ph_sc   <- if (nrow(.tp)) .tp$scenario[1] else NA_character_
```

```{r validation, include=FALSE}
## ---------------------------------------------------------------------------
## 3. VALIDATION GATES - health/cost-effectiveness contract only (NO BCA).
##    Consistency is checked against Scenario_Catalog + the results contract.
## ---------------------------------------------------------------------------
val_stop <- function(cond, msg) if (!isTRUE(cond)) stop("VALIDATION STOP - ", msg, call. = FALSE)

## (a) Scenario_Catalog present in every workbook that exists (health/CE contract;
##     Benefit_Cost / Economic_Value are NOT required by the deck).
cat_sets <- list()
.wb_paths <- c(clinical = wb_clin, public_health = wb_ph, combined = wb_comb)
for (fam in names(.wb_paths)) {
  path <- .wb_paths[[fam]]
  if (!file.exists(path)) next
  sh <- sheets_of(path)
  val_stop("Scenario_Catalog" %in% sh,
           sprintf("workbook '%s' is missing the Scenario_Catalog sheet", basename(path)))
  val_stop(all(c("CVD_40q30", "Cost_Effectiveness") %in% sh),
           sprintf("workbook '%s' is missing a required health/CE sheet", basename(path)))
  cat_sets[[fam]] <- rd(path, "Scenario_Catalog")$scenario
}

## (b) Contract scenario set reconciles EXACTLY with Scenario_Catalog per family
##     (no stale extras, nothing dropped). This is how "no excluded id sneaks in"
##     is enforced without hard-coding any excluded id.
recon_family <- function(cat_key, tab) {
  cs <- cat_sets[[cat_key]]
  if (is.null(cs)) return(invisible())
  ts <- unique(tab$scenario)
  val_stop(setequal(cs, ts),
           sprintf("%s: results contract scenarios do not match Scenario_Catalog. Only in catalog: {%s}; only in results: {%s}. Re-run Model 09 on the current run.",
                   cat_key, paste(setdiff(cs, ts), collapse = ", "),
                   paste(setdiff(ts, cs), collapse = ", ")))
}
recon_family("clinical", met_clin)
recon_family("public_health", met_ph)
recon_family("combined", met_comb)

## (c) baseline available exactly once per family + 2050 present.
for (tab in list(met_clin, met_ph, met_comb)) {
  if (!nrow(tab)) next
  val_stop(sum(tab$scenario == "baseline") <= 1, "baseline appears more than once in a family")
}
val_stop(any(c4c$scenario == "baseline") && HZ_END %in% c4c$year,
         "combined workbook CVD_40q30 has no baseline / no year 2050")

## (d) unique scenario ids within each family (no duplicates).
for (tab in list(met_clin, met_ph, met_comb))
  if (nrow(tab)) val_stop(!any(duplicated(tab$scenario)),
                          "duplicate scenario id within a family results block")

## (e) headline health/CE metrics finite where the scenario exists.
if (has_joint)
  val_stop(is.finite(combo_da) && is.finite(combo_q50) && is.finite(combo_cpda),
           "combined-package headline metrics are missing/non-numeric")
val_stop(is.finite(base_q_2050), "baseline CVD 40q30 in 2050 is missing")
```

```{r fig-40q30-data, include=FALSE}
## Data for the CVD 40q30 time-series & international-comparison figure.
cmp <- fread(csv_cmp)
cmp[, year := as.integer(year)]
comparators <- c("Malaysia","Philippines","Singapore","Thailand","Viet Nam","Indonesia")

cmp_lines <- cmp[country_label %in% comparators,
                 .(country_label, year, q = q40q30_percent)][order(country_label, year)]
cmp_lines <- unique(cmp_lines, by = c("country_label","year"))

#tempatch
cmp_lines[country_label == "Indonesia", country_label := "Indonesia."]

idn_obs <- cmp[country_label == "Indonesia" & series_type == "Historical WPP-scaled baseline" &
               year <= 2023, .(year, q = q40q30_percent)][order(year)]

## Indonesia MODEL scenarios (dashed): whichever of these are BOTH in the
## current-run contract and available as a per-year series in the workbook.
fig_scen_ids <- intersect(c("baseline", clin_all_id, ph_all_id, joint_id),
                          intersect(unique(c4c$scenario), unique(met_comb$scenario)))
idn_scen <- c4c[scenario %in% fig_scen_ids & year >= HZ_START,
                .(scenario, year, q = get(qcol))]
idn_scen[, scen := vapply(scenario, lbl, character(1), USE.NAMES = FALSE)]
scen_levels <- vapply(fig_scen_ids, lbl, character(1), USE.NAMES = FALSE)
idn_scen[, scen := factor(scen, levels = scen_levels)]
scen_pal <- setNames(
  c(col_baseline, col_clinical, col_ph, col_combined)[seq_along(fig_scen_ids)],
  scen_levels)
```

```{r results-builders, include=FALSE}
## Row builders for the impact / cost-effectiveness display tables (BCA-free).
impact_row <- function(tab, sc) data.table(
  Intervention = lbl(sc, tab),
  da  = fmt_thou(hv(sc, "deaths_averted", tab)),
  q50 = fmt_pct(hv(sc, "cvd_40q30_2050", tab), 1),
  qr  = fmt_pct(hv(sc, "cvd_40q30_pct_reduction_2050", tab), 1),
  dr  = fmt_pct(hv(sc, "pct_reduction_deaths_2050", tab), 1))
## Average annual additional cost per capita for one scenario (undisc / disc):
## cumulative additional cost / (25 x 2026 national population). `percap_den` is
## built in the percapita chunk (evaluated before any table chunk runs).
pc_undisc <- function(tab, sc) fmt_usd_pc(hv(sc, "incremental_cost", tab)      / percap_den)
pc_disc   <- function(tab, sc) fmt_usd_pc(hv(sc, "disc_incremental_cost", tab) / percap_den)
ce_row <- function(tab, sc) cbind(
  impact_row(tab, sc),
  data.table(cpd = fmt_usd(hv(sc, "cost_per_death_averted", tab)),
             pcu = pc_undisc(tab, sc),
             pcd = pc_disc(tab, sc)))

## Build a display table for a family's top-level rows, capped for space.
## Returns list(tb, agg_row_idx, n_omitted, omitted_labels).
family_table <- function(tab, agg_id, row_fun) {
  d <- top_level_rows(tab, agg_id)
  n_omitted <- 0L; omitted <- character(0)
  if (nrow(d) > MAX_TBL_ROWS) {
    keep_agg <- d[scenario == agg_id]
    body <- d[scenario != agg_id]
    n_keep <- MAX_TBL_ROWS - nrow(keep_agg)
    omitted <- body$scenario[-seq_len(min(n_keep, nrow(body)))]
    body <- body[seq_len(min(n_keep, nrow(body)))]
    d <- rbind(body, keep_agg)
    n_omitted <- length(omitted)
    message(sprintf("  Deck: family table capped at %d rows; %d intervention(s) omitted for space: %s",
                    MAX_TBL_ROWS, n_omitted, paste(omitted, collapse = ", ")))
  }
  if (!nrow(d)) return(list(tb = data.table(), agg_idx = NA_integer_, n_omitted = 0L, omitted = character(0)))
  tb <- rbindlist(lapply(d$scenario, function(s) row_fun(tab, s)))
  list(tb = tb, agg_idx = which(d$scenario == agg_id), n_omitted = n_omitted, omitted = omitted)
}
```

```{r percapita, include=FALSE}
## ---------------------------------------------------------------------------
## 4. Average annual ADDITIONAL cost per capita (team request), for the three
##    aggregate scenarios. Numerator = cumulative additional cost over the
##    scale-up horizon (contract value = workbook Budget_Impact cumulative
##    additional cost; 2025 additional cost is 0 by construction, so the
##    horizon total equals the 2026-2050 sum). Denominator = 25 x 2026 national
##    population, counted ONCE per location x year x age x sex from the model's
##    authoritative demography, NOT the eligible/treated population.
## ---------------------------------------------------------------------------
percap_years <- (HZ_START + 1L):HZ_END                       # 2026..2050
val_stop(length(percap_years) == 25L,
         sprintf("expected 25 projection years (2026-%d); got %d", HZ_END, length(percap_years)))

## Single defensible national population series = the baseline (current-course)
## demographic projection, which is cause-invariant. Guard that claim.
.dd     <- as.data.table(readRDS(dalys_file))
.pk     <- c("location", "year", "age", "sex")
.base26 <- .dd[scenario == "baseline" & location == "Indonesia" & year == (HZ_START + 1L)]
val_stop(nrow(.base26) > 0, "model demography has no baseline Indonesia row for the 2026 population denominator")
.spread <- .base26[, .(sp = max(population) - min(population)), by = .pk]
val_stop(all(.spread$sp < 1),
         "no single defensible 2026 national-population series: baseline population differs by >=1 person across causes within a location/year/age/sex cell")
.pop_cells <- .base26[, .(population = mean(population)), by = .pk]   # collapse fp noise across causes
pop_2026   <- .pop_cells[, sum(population)]
val_stop(is.finite(pop_2026) && pop_2026 > 1e8, "2026 national population is missing or implausible")
percap_den <- length(percap_years) * pop_2026
rm(.dd)  # free the large demography table

## Formatted "undiscounted / discounted" per-capita callout for one aggregate scenario.
percap_vals <- function(tab, sc) {
  undisc <- hv(sc, "incremental_cost", tab) / percap_den
  disc   <- hv(sc, "disc_incremental_cost", tab) / percap_den
  list(undisc = undisc, disc = disc,
       txt = sprintf("\\textbf{Average annual additional cost per capita} (simple average over the %d-year scale-up, %d population denominator): \\textbf{%s} undiscounted, \\textbf{%s} discounted.",
                     length(percap_years), (HZ_START + 1L), fmt_usd_pc(undisc), fmt_usd_pc(disc)))
}
## Emit the per-capita callout as a compact tinted single line under a cost table.
percap_callout <- function(tab, sc, famcol = "wbnavy") {
  v <- percap_vals(tab, sc)
  if (!is.finite(v$undisc) || !is.finite(v$disc)) return(invisible())
  cat(sprintf("\n\n\\vspace{2pt}\\noindent\\colorbox{%s!8}{\\parbox{\\dimexpr\\linewidth-2\\fboxsep}{\\footnotesize %s}}\n",
              famcol, v$txt))
}
```

```{r cascade-data, include=FALSE}
## ---------------------------------------------------------------------------
## 5. ISOLATED 70-30-30 -> 70-70-70 cascade strategy analysis. Read ONLY from
##    the cascade workbook's stored/literal values; never merged with the
##    ordinary results contract. Fail loudly on any structural surprise.
## ---------------------------------------------------------------------------
csh   <- openxlsx::getSheetNames(casc_wb)
.creq <- c("Run_Metadata","Scenario_Catalog","Cascade_Assumptions","Cascade_Trajectory",
           "Annual_Mortality","CVD_40q30","Annual_Cost","Budget_Impact","Cost_Effectiveness")
val_stop(all(.creq %in% csh),
         sprintf("cascade workbook missing sheet(s): %s", paste(setdiff(.creq, csh), collapse = ", ")))
cread <- function(s) as.data.table(openxlsx::read.xlsx(casc_wb, sheet = s))
need_cols <- function(dt, cols, sheet)
  val_stop(all(cols %in% names(dt)),
           sprintf("cascade sheet '%s' missing column(s): %s", sheet, paste(setdiff(cols, names(dt)), collapse = ", ")))

## (a) Exactly the two expected scenarios (baseline + the cascade scenario).
CASC_SCEN <- "S_70_30_30_TO_70_70_70"
c_cat <- cread("Scenario_Catalog"); need_cols(c_cat, "scenario", "Scenario_Catalog")
val_stop(setequal(c_cat$scenario, c("baseline", CASC_SCEN)),
         sprintf("cascade Scenario_Catalog must be exactly {baseline, %s}; got {%s}",
                 CASC_SCEN, paste(c_cat$scenario, collapse = ", ")))

## (b) Assumptions -> parameter lookup + metadata.
c_as <- cread("Cascade_Assumptions"); need_cols(c_as, c("parameter_id","value"), "Cascade_Assumptions")
getp  <- function(id) { v <- c_as[parameter_id == id, value][1]; val_stop(!is.na(v), sprintf("cascade assumption '%s' missing", id)); v }
getpn <- function(id) suppressWarnings(as.numeric(getp(id)))
casc_disc  <- getpn("cost_discount_rate")
casc_price <- getpn("cost_price_year")
casc_curr  <- getp("currency")
casc_diag  <- getpn("target_diagnosis_share")
casc_t30   <- getpn("first_target_treatment_conditional")
casc_c30   <- getpn("first_target_control_conditional")
casc_t40   <- getpn("final_target_treatment_conditional")
casc_c40   <- getpn("final_target_control_conditional")
casc_half  <- getpn("treated_uncontrolled_effect_fraction")
casc_eff30 <- getpn("first_target_effective_coverage_exact")
casc_eff40 <- getpn("final_target_effective_coverage_exact")
casc_y0    <- as.integer(getpn("analysis_start_year"))
casc_y1    <- as.integer(getpn("analysis_end_year"))
casc_ytar1 <- as.integer(getpn("first_target_year"))
casc_ytar2 <- as.integer(getpn("final_target_year"))
c_meta <- cread("Run_Metadata")
casc_persp <- { v <- c_meta[item == "Economic perspective", value]; ltx(gsub("_", " ", if (length(v)) v[1] else "health system")) }
val_stop(is.finite(casc_disc) && is.finite(casc_price), "cascade discount rate / price year missing")
## Nested-share + effective-coverage reconciliation (fail loud if the workbook drifts).
val_stop(abs(casc_diag * casc_t30 * casc_c30 - 0.063) < 1e-6, "cascade 2030 controlled share does not reconcile to 6.3%")
val_stop(abs(casc_diag * casc_t40 * casc_c40 - 0.343) < 1e-6, "cascade 2040 controlled share does not reconcile to 34.3%")
.eff_cov <- function(diag, t, ctrl, half) diag * t * ctrl + half * (diag * t - diag * t * ctrl)
val_stop(abs(.eff_cov(casc_diag, casc_t30, casc_c30, casc_half) - casc_eff30) < 1e-6,
         "cascade 2030 effective coverage does not reconcile to the stored exact value")
val_stop(abs(.eff_cov(casc_diag, casc_t40, casc_c40, casc_half) - casc_eff40) < 1e-6,
         "cascade 2040 effective coverage does not reconcile to the stored exact value")

## (c) Hypertension cascade shares (sex-averaged) for the bar figure.
c_tr <- cread("Cascade_Trajectory")
need_cols(c_tr, c("intervention_id","risk_factor_id","sex","year","controlled_share","treated_uncontrolled_share"), "Cascade_Trajectory")
htn_shares <- function(iv, rf) {
  d <- c_tr[intervention_id == iv & risk_factor_id == rf & year %in% c(HZ_START, casc_ytar1, casc_ytar2) &
              !is.na(controlled_share)]
  d[, treated := controlled_share + treated_uncontrolled_share]
  d[, .(controlled = mean(controlled_share), treated = mean(treated)), by = year][order(year)]
}
htn_sh <- htn_shares("I_CVD_PRIMARY", "RF_HTN")
val_stop(nrow(htn_sh) == 3L, "hypertension cascade needs baseline + two milestone years of shares")
## diagnosed row: not modelled at baseline (leave NA); = target at milestone years
htn_sh[, diagnosed := ifelse(year == HZ_START, NA_real_, casc_diag)]
dm_base <- { d <- c_tr[intervention_id == "I_T2D_TREATMENT" & year == HZ_START & !is.na(controlled_share)]
             d[, treated := controlled_share + treated_uncontrolled_share]
             list(treated = mean(d$treated), controlled = mean(d$controlled_share)) }

## (d) Probability-of-death-before-70 series (R-reconciled literal values).
c_q <- cread("CVD_40q30"); need_cols(c_q, c("scenario","year","cvd_40q30_r"), "CVD_40q30")
casc_q <- c_q[, .(scenario, year, q = cvd_40q30_r)][order(scenario, year)]
val_stop(all(is.finite(casc_q$q)) && setequal(casc_q$year, casc_y0:casc_y1),
         "cascade probability-of-death series has gaps or non-finite values")

## (e) Cumulative DISCOUNTED cost levels (baseline vs intervention), de-duplicated.
c_ac <- cread("Annual_Cost")
need_cols(c_ac, c("year","disc_cost_baseline","disc_cost_scenario","disc_cost_incremental","shared_duplicate_count"), "Annual_Cost")
c_ac[, w := 1 / pmax(shared_duplicate_count, 1)]              # count shared components once
casc_cost <- c_ac[, .(base = sum(disc_cost_baseline * w), scen = sum(disc_cost_scenario * w),
                      inc  = sum(disc_cost_incremental * w)), by = year][order(year)]
casc_cost[, `:=`(cum_base = cumsum(base), cum_scen = cumsum(scen), cum_inc = cumsum(inc))]
c_bi <- cread("Budget_Impact"); need_cols(c_bi, c("year","cumulative_disc_incremental_cost"), "Budget_Impact")
casc_cost <- merge(casc_cost, c_bi[, .(year, bi_cum = cumulative_disc_incremental_cost)], by = "year")
## Guard: our de-duplicated level difference must reconcile to the authoritative budget sheet.
val_stop(max(abs((casc_cost$cum_scen - casc_cost$cum_base) - casc_cost$bi_cum)) < 1,
         "cascade cost levels (Annual_Cost) do not reconcile to Budget_Impact cumulative additional cost")

## (f) Cumulative deaths (six CVD causes + type-2 diabetes), baseline vs intervention.
c_am <- cread("Annual_Mortality"); need_cols(c_am, c("scenario","year","cause_deaths","base_deaths","deaths_averted"), "Annual_Mortality")
casc_base_d <- c_am[scenario == "baseline",  .(base = sum(cause_deaths)), by = year][order(year)]
casc_scen_d <- c_am[scenario == CASC_SCEN,   .(scen = sum(cause_deaths)), by = year][order(year)]
casc_deaths <- merge(casc_base_d, casc_scen_d, by = "year")
casc_deaths[, `:=`(cum_base = cumsum(base), cum_scen = cumsum(scen))]
casc_deaths[, cum_averted := cum_base - cum_scen]

## (g) Headline metrics from the workbook Cost_Effectiveness (single source of truth).
c_ce <- cread("Cost_Effectiveness"); need_cols(c_ce, c("scenario","deaths_averted","disc_incremental_cost","cost_per_death_averted"), "Cost_Effectiveness")
c_ce <- c_ce[scenario == CASC_SCEN]
val_stop(nrow(c_ce) == 1L, "cascade Cost_Effectiveness must have exactly the cascade scenario")
casc_da   <- c_ce$deaths_averted[1]
casc_cost_tot <- c_ce$disc_incremental_cost[1]
casc_cpda <- c_ce$cost_per_death_averted[1]
## Reconcile CE against the annual sheets we plotted (fail loud on drift).
val_stop(abs(casc_deaths[year == casc_y1, cum_averted] - casc_da) < 1,
         "cascade deaths averted (Annual_Mortality) do not reconcile to Cost_Effectiveness")
val_stop(abs(casc_cost[year == casc_y1, bi_cum] - casc_cost_tot) < 1e4,
         "cascade cumulative additional cost (Budget_Impact) does not reconcile to Cost_Effectiveness")

## Cumulative snapshots for the results table (2030 / 2040 / 2050).
casc_tab_years <- c(casc_ytar1, casc_ytar2, casc_y1)
casc_snapshot <- function(y) list(
  d_base = casc_deaths[year == y, cum_base], d_scen = casc_deaths[year == y, cum_scen],
  d_av   = casc_deaths[year == y, cum_averted],
  c_base = casc_cost[year == y, cum_base],   c_scen = casc_cost[year == y, cum_scen],
  c_add  = casc_cost[year == y, bi_cum])
```

```{r deck-titles, include=FALSE}
## Dynamic, plain-text (no %/$) takeaway titles derived from current outputs.
title_exec <- if (has_joint)
  sprintf("Together, clinical care and prevention could avert about %s deaths by %d",
          fmt_title_num(combo_da), HZ_END) else
  "Clinical care and population prevention each cut premature cardiovascular deaths"
title_cascade_results <- sprintf(
  "A stronger cascade could prevent about %s deaths by %d",
  fmt_title_num(casc_da), casc_y1)
```

# The challenge, and the opportunity

## Most premature cardiovascular deaths in Indonesia can be prevented or treated

\begin{columns}[T]
\begin{column}{0.52\textwidth}
\begin{block}{Cardiovascular disease is Indonesia's largest health burden}
\normalsize
Under current trends, a person reaching age 30 faces roughly a
\hlkey{`r fmt_pct(base_q_2025,1)` chance} of dying from cardiovascular disease
before age 70. Most of this is \textbf{preventable or treatable}.
\end{block}
\end{column}
\begin{column}{0.46\textwidth}
\begin{itemize}\itemsep4pt
  \item Premature CVD deaths (ages 30--69) have risen for two decades and remain high.
  \item Two complementary levers:
  \begin{itemize}
    \item \textcolor{clinicalcol}{\textbf{Clinical care}} --- prevent and better treat established disease.
    \item \textcolor{phcol}{\textbf{Population prevention}} --- tobacco, salt, trans-fat.
  \end{itemize}
  \item This analysis weighs their \textbf{health impact and value for money} together.
\end{itemize}
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} Indonesia CVD model, calibrated to GBD 2023 and UN World Population Prospects 2024; combined results workbook (annual probability-of-death sheet). \textbf{Method:} the probability of death from CVD before age 70 is the chance of dying from the six modelled cardiovascular diseases between exact ages 30 and 70, read from a single-year life table in the `r HZ_START` baseline year.}

## We measured the health gains and the value for money of proven interventions

\begin{block}{General aim}
\normalsize Estimate the \textbf{health impact and value for money} of scaling up evidence-based
\textbf{clinical} and \textbf{public-health} cardiovascular interventions in Indonesia through \textbf{2050}.
\end{block}

\vspace{2pt}
\begin{columns}[T]
\begin{column}{0.33\textwidth}
\begin{clinicalblock}{Aim 1 --- Burden}
\footnotesize Project the CVD burden and the probability of death from CVD before age 70 under the \textbf{current course} and under \textbf{alternative intervention scenarios}.
\end{clinicalblock}
\end{column}
\begin{column}{0.33\textwidth}
\begin{phblock}{Aim 2 --- Value for money}
\footnotesize Estimate \textbf{deaths averted}, mortality reductions, \textbf{additional costs}, and the \textbf{cost per death averted} for each intervention and package.
\end{phblock}
\end{column}
\begin{column}{0.33\textwidth}
\begin{combinedblock}{Aim 3 --- Better together}
\footnotesize Quantify the \textbf{added value of combining} clinical and public-health action, modelled as a single joint run rather than a sum.
\end{combinedblock}
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} study design; Indonesia CVD model. \textbf{Method:} every scenario runs through one projection over `r HZ_START`--`r HZ_END`; the baseline (current course) is the no-new-intervention comparator against which each scenario is measured.}

## `r title_exec`

\begin{columns}[T]
\begin{column}{0.5\textwidth}
```{r exec-block, results='asis'}
if (has_joint) {
  cat(sprintf(paste0(
    "\\begin{combinedblock}{Combining clinical + public health is the big prize}\n",
    "\\footnotesize Together they could avert about \\hlkey{%s deaths} over %d--%d, ",
    "cut the probability of death from CVD before age 70 in %d by \\hlkey{%s}, at about \\hlkey{%s per death averted}.\n",
    "\\end{combinedblock}\n"),
    fmt_short(combo_da), HZ_START, HZ_END, HZ_END, fmt_pct(combo_qred,0), fmt_usd(combo_cpda)))
} else {
  cat(paste0("\\begin{block}{Two complementary levers}\n\\footnotesize This run reports the ",
             "clinical and public-health families; the joint package was not part of this ",
             "configuration.\n\\end{block}\n"))
}
```
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep3pt
```{r exec-bullets, results='asis'}
if (has_clin)
  cat(sprintf("  \\item \\textcolor{clinicalcol}{\\textbf{All clinical interventions}}: $\\approx$%s deaths averted; probability of death from CVD before age 70 down %s by %d.\n",
              fmt_short(clin_da), fmt_pct(clin_qred,0), HZ_END))
if (has_ph)
  cat(sprintf("  \\item \\textcolor{phcol}{\\textbf{All public-health policies}}: $\\approx$%s deaths averted at about %s per death averted.\n",
              fmt_short(ph_da), fmt_usd(ph_cpda)))
if (!is.na(top_clin_sc) && !is.na(top_ph_sc))
  cat(sprintf("  \\item \\textbf{Most impactful single levers}: \\textcolor{clinicalcol}{%s} and \\textcolor{phcol}{%s}.\n",
              lbl(top_clin_sc, met_clin), lbl(top_ph_sc, met_ph)))
if (has_joint)
  cat(sprintf("  \\item \\textbf{Affordability}: the combined package costs about %s per death averted.\n",
              fmt_usd(combo_cpda)))
```
\end{itemize}
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} results contract (\texttt{09\_deck\_results.rds}) + combined workbook. \textbf{Method:} deaths averted cumulative `r HZ_START`--`r HZ_END` (six CVD causes $+$ type-2 diabetes); cost per death averted $=$ discounted additional cost $\div$ deaths averted (`r price_year` US\$). \costcaveat}

## The model follows how prevention and treatment change disease and survival

\begin{columns}[T]
\begin{column}{0.54\textwidth}
\footnotesize
\begin{enumerate}\itemsep3pt
  \item \textbf{People \& demography.} Indonesia's population is projected by age and sex (UN World Population Prospects 2024).
  \item \textbf{Starting health.} Disease and death rates are set to Indonesia's data (GBD 2023) and calibrated to match observed levels.
  \item \textbf{Year-by-year simulation.} Each year people can develop cardiovascular disease or diabetes, survive, or die.
  \item \textbf{Interventions.} Clinical and public-health scenarios change how many people develop disease (new cases) and how many survive it.
  \item \textbf{Outcomes.} The model reports deaths, healthy years lost or gained (DALYs), the probability of death from CVD before age 70, costs, and value for money.
\end{enumerate}
\end{column}
\begin{column}{0.44\textwidth}
\begin{block}{One consistent engine}
\footnotesize Every scenario runs through the \textbf{same} projection, so results are directly comparable and the \textbf{combined} package is modelled as one run --- not a sum of separate estimates.
\end{block}
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} Indonesia CVD model engine; GBD 2023 (starting rates); UN World Population Prospects 2024 (demography, life expectancy). \textbf{Method:} a discrete-time state-transition simulation by single-year age and sex, calibrated to match observed GBD levels using a standard numerical optimiser.}

## Interventions work by preventing new disease and by helping people survive it

\vspace{-2pt}
\begin{center}
\begin{tikzpicture}[
   >=Latex, node distance = 16mm and 30mm, font=\footnotesize,
   state/.style = {rectangle, rounded corners, draw=wbnavy, thick, fill=wbnavy!6,
                   minimum width=26mm, minimum height=11mm, align=center},
   death/.style = {rectangle, rounded corners, draw=black!55, thick, fill=black!4,
                   minimum width=26mm, minimum height=10mm, align=center}]
  \node[state] (well) {\textbf{Healthy}\\[-1pt]\scriptsize (no CVD)};
  \node[state, right=of well] (sick) {\textbf{Living with CVD}\\[-1pt]\scriptsize (or diabetes)};
  \node[death, right=of sick] (cvd)  {\textbf{CVD death}};
  \node[death, below=14mm of well, xshift=23mm] (bg) {Other-cause (background) death};

  \draw[->, very thick, incidencecol]
       (well) -- node[above, midway, text=incidencecol]{\textbf{new CVD}}
                  node[below, midway, text=incidencecol, font=\scriptsize]{new cases} (sick);
  \draw[->, very thick, casefatalcol]
       (sick) -- node[above, midway, text=casefatalcol]{\textbf{dies of CVD}}
                  node[below, midway, text=casefatalcol, font=\scriptsize]{risk of dying with CVD} (cvd);
  \draw[->, thick, bgmortcol] (well) to[bend right=12] (bg);
  \draw[->, thick, bgmortcol] (sick) to[bend left=12]  (bg);
\end{tikzpicture}
\end{center}

\vspace{-2pt}
\begin{columns}[T]
\begin{column}{0.33\textwidth}\footnotesize
\textcolor{incidencecol}{\textbf{Fewer people fall ill}}\\ \scriptsize Primary prevention \& \emph{all} public-health policies reduce \textbf{new cases}.
\end{column}
\begin{column}{0.34\textwidth}\footnotesize
\textcolor{casefatalcol}{\textbf{More people survive}}\\ \scriptsize Secondary prevention, treatment \& tobacco policy lower the \textbf{risk of dying after developing CVD}.
\end{column}
\begin{column}{0.31\textwidth}\footnotesize
\textcolor{bgmortcol}{\textbf{Background mortality}}\\ \scriptsize Other-cause deaths from demography; unchanged by these interventions.
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} Indonesia CVD model engine (\texttt{06\_run\_scenarios\_indonesia\_fair.R}). \textbf{Method:} interventions act on two transitions --- new cases (Healthy $\rightarrow$ Living with CVD) and the risk of dying after developing CVD (Living with CVD $\rightarrow$ death); no recovery/remission is modelled, a conservative assumption.}

# The interventions we tested

## Clinical care prevents disease and helps people survive it

```{r tbl-clin-interventions, results='asis'}
## Built from the ACTIVE clinical scenarios in this run (children/aggregate
## excluded); target-population / effect notes come from a small presentation map.
clin_pop <- c(
  I_CVD_PRIMARY = "Adults at cardiovascular risk", I_T2D_TREATMENT = "People with type-2 diabetes",
  I_IHD_SECONDARY = "Heart-attack (IHD) survivors", I_STROKE_SECONDARY = "Stroke survivors",
  I_HF_CHRONIC = "Chronic heart-failure patients", I_HF_ACUTE = "Acute heart-failure patients",
  I_RHD_SECONDARY = "Rheumatic heart disease patients")
clin_acts <- c(I_CVD_PRIMARY = "\\textcolor{incidencecol}{Fewer new cases}")
clin_disp <- top_level_rows(met_clin, clin_all_id)[scenario != clin_all_id]
if (nrow(clin_disp)) {
  pluck <- function(vec, key, default) if (key %in% names(vec)) vec[[key]] else default
  cd <- data.table(
    Intervention = vapply(clin_disp$scenario, function(s) lbl(s, met_clin), character(1)),
    Population   = vapply(clin_disp$scenario, function(s) pluck(clin_pop, s, "Target patient group"), character(1)),
    Acts         = vapply(clin_disp$scenario, function(s)
                     pluck(clin_acts, s, "\\textcolor{casefatalcol}{Better survival}"), character(1)))
  print(kbl(cd, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=c("Intervention","Target population","Model effect"), align="lll") |>
    kable_styling(latex_options="hold_position", font_size=8.5) |>
    column_spec(1, bold=TRUE, width="4.6cm") |>
    column_spec(2, width="6.4cm") |> column_spec(3, width="3.0cm") |>
    row_spec(0, bold=TRUE))
} else cat("\\begin{center}\\large No clinical interventions selected for this run.\\end{center}")
```

\vspace{2pt}
\footnotesize \textbf{Implementation scenario (all clinical interventions):} coverage scales up \textbf{from a low base toward 80\%} of the people who need each service between `r HZ_START` and `r HZ_END`. Primary prevention lowers who develops CVD (new cases); secondary prevention and treatment improve survival.

\srcmethod{\textbf{Source:} clinical results workbook, sheets \texttt{Selected\_Interventions} and \texttt{Scenario\_Catalog}. \textbf{Method:} the table lists the clinical interventions active in this run (excluded interventions never appear); effect sizes come from the FAIR Choices intervention taxonomy.}

## Population policies cut tobacco, salt, and trans-fat risk across the whole population

```{r tbl-ph-interventions, results='asis'}
ph_meta <- data.table(
  scenario = c("I_PH_TOBACCO_TAX","I_PH_TOBACCO_POLICY","I_PH_SALT_POLICY","I_PH_TFA_POLICY",
               "I_PH_SALT_REFORM","I_PH_SALT_FOPL","I_PH_SALT_MEDIA","I_PH_SALT_ENV",
               "I_PH_TOB_CLEAN_AIR","I_PH_TOB_MEDIA","I_PH_TOB_AD_BAN"),
  Risk  = c("Tobacco","Tobacco","Dietary salt/sodium","Industrial trans-fat",
            "Dietary salt/sodium","Dietary salt/sodium","Dietary salt/sodium","Dietary salt/sodium",
            "Tobacco","Tobacco","Tobacco"),
  Acts  = c("\\textcolor{incidencecol}{New cases} + \\textcolor{casefatalcol}{survival}",
            "\\textcolor{incidencecol}{New cases} + \\textcolor{casefatalcol}{survival}",
            "\\textcolor{incidencecol}{New cases}","\\textcolor{incidencecol}{New cases}",
            "\\textcolor{incidencecol}{New cases}","\\textcolor{incidencecol}{New cases}",
            "\\textcolor{incidencecol}{New cases}","\\textcolor{incidencecol}{New cases}",
            "\\textcolor{incidencecol}{New cases} + \\textcolor{casefatalcol}{survival}",
            "\\textcolor{incidencecol}{New cases} + \\textcolor{casefatalcol}{survival}",
            "\\textcolor{incidencecol}{New cases} + \\textcolor{casefatalcol}{survival}"),
  Lever = c("Raise tax to 75\\% of retail price","Clean-air law, mass media, ad ban",
            "Reformulation, labelling, media, procurement","Best-practice industrial TFA ban",
            "Mandatory reformulation","Front-of-pack labelling","Mass-media campaign","Public-institution environment",
            "Smoke-free / clean-air law","Mass-media campaign","Advertising ban"))
ph_disp <- top_level_rows(met_ph, ph_all_id)[scenario != ph_all_id]
if (nrow(ph_disp)) {
  pd <- merge(ph_disp[, .(scenario)], ph_meta, by = "scenario", all.x = TRUE, sort = FALSE)
  pd[, Policy := vapply(scenario, function(s) lbl(s, met_ph), character(1))]
  pd[is.na(Risk),  Risk  := "Behavioural risk factor"]
  pd[is.na(Acts),  Acts  := "\\textcolor{incidencecol}{New cases}"]
  pd[is.na(Lever), Lever := "Fiscal / regulatory measure"]
  print(kbl(pd[, .(Policy, Risk, Acts, Lever)], format="latex", booktabs=TRUE, escape=FALSE,
      col.names=c("Policy","Risk factor","Model effect","Implementation lever"), align="llll") |>
    kable_styling(latex_options="hold_position", font_size=7.8) |>
    column_spec(1, bold=TRUE, width="2.9cm") |> column_spec(2, width="2.3cm") |>
    column_spec(3, width="3.1cm") |> column_spec(4, width="5.7cm") |>
    row_spec(0, bold=TRUE))
} else cat("\\begin{center}\\large No public-health policies selected for this run.\\end{center}")
```

\vspace{1pt}
\footnotesize Policies reduce exposure (smoking, salt, trans-fat intake), lowering the number of \textbf{new CVD cases}. \textbf{Tobacco} policies additionally improve \textbf{survival} after CVD. Packages are modelled as single joint runs of their component measures.

\srcmethod{\textbf{Source:} public-health results workbook, sheets \texttt{Selected\_Interventions} / \texttt{Policy\_Levers} and \texttt{Scenario\_Catalog}. \textbf{Method:} the table lists the policies active in this run; effect sizes come from published risk-factor evidence stored in the workbook.}

## Every input is traceable to a recognised data source

```{r tbl-sources}
src <- data.table(
  Component = c("Population \\& demography","Cause-specific mortality \\& disease rates",
                "Disability weights (healthy years lost)","Clinical intervention effects",
                "Public-health effect sizes","Intervention \\& policy costs",
                "International CVD mortality comparison"),
  Source = c("UN World Population Prospects","GBD (Indonesia)","GBD",
             "FAIR Choices intervention taxonomy","Published risk-factor evidence (in workbook)",
             "Programme \\& policy unit costs (in workbook)",
             "WPP-scaled ASEAN comparison series"),
  Years = c("2024 revision","2023","2023","current","current",
            sprintf("price year %d", price_year), "1995--2050"),
  Use = c("Population, background mortality, life expectancy","Baseline burden \\& calibration targets",
          "Healthy years lost or gained (DALYs)","New-case \\& survival effects","New-case \\& survival effects",
          "Budget impact \\& value for money",
          "Benchmarking Indonesia vs.\\ neighbours"))
kbl(src, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=c("Model component","Primary source","Years / version","Use in model"), align="llll") |>
  kable_styling(latex_options=c("hold_position"), font_size=6.6) |>
  column_spec(1, bold=TRUE, width="3.2cm") |> column_spec(2, width="4.0cm") |>
  column_spec(3, width="1.9cm") |> column_spec(4, width="4.7cm") |>
  row_spec(0, bold=TRUE)
```

\vspace{1pt}
\scriptsize Population: UN World Population Prospects 2024; epidemiology: GBD 2023 [@GBDCVD2025; @Schumacher2024GBD2021].

\srcmethod{\textbf{Source:} modelling pipeline and input workbooks. \textbf{Method:} these inputs drive the demographic projection, calibration targets, and effect sizes; effect-size and cost citations are in the workbooks.}

## Costs depend on who needs care, how many receive it, and local prices

\footnotesize We estimate the \textbf{additional health-system resources required to expand effective coverage}, using a costing approach adapted from Disease Control Priorities (DCP) [@Watkins2020ResourceRequirements].

<!-- \footnotesize We estimate costs with an intervention-based approach adapted from the Disease Control Priorities (DCP) cost model of @Watkins2020ResourceRequirements, applied here from a \textbf{health-system perspective} with \textbf{annual} projections. Each service is costed as: \emph{who needs it} $\times$ \emph{how much of that need is met} $\times$ \emph{services per person per year} $\times$ \emph{Indonesia-adjusted unit cost}. -->

\vspace{2pt}
\begin{center}
\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[
   >=Latex, node distance = 5mm, font=\scriptsize,
   costbox/.style = {rectangle, rounded corners, draw=wbnavy, thick, fill=wbnavy!6,
                  text width=22mm, minimum height=14mm, align=center},
   costout/.style = {rectangle, rounded corners, draw=clinicalcol, very thick, fill=clinicalcol!8,
                  text width=27mm, minimum height=14mm, align=center}]
  \node[costbox] (pin)  {\textbf{People who need}\\ the service};
  \node[costbox, right=of pin] (cov) {\textbf{Current coverage}\\ \& scale-up target};
  \node[costbox, right=of cov] (freq){\textbf{Services}\\ per person / year};
  \node[costbox, right=of freq] (uc) {\textbf{Indonesia unit}\\ cost};
  \node[costbox, right=of uc] (ann) {\textbf{Annual cost}\\ baseline \& scenario};
  \node[costout,  right=of ann] (res){\textbf{Additional \& discounted cost;} cost per person; cost per death prevented};
  \draw[->,thick] (pin)--(cov); \draw[->,thick] (cov)--(freq); \draw[->,thick] (freq)--(uc);
  \draw[->,thick] (uc)--(ann);  \draw[->,thick, clinicalcol] (ann)--(res);
\end{tikzpicture}}
\end{center}

\vspace{2pt}
<!-- \footnotesize The DCP paper supplies the \textbf{conceptual framework} --- unit costs, demographic/epidemiological estimates of the people in need, baseline coverage, and health-system delivery costs. This project \textbf{adapts} that framework to Indonesia and does not claim to reproduce the paper exactly. -->

\footnotesize\textbf{What this tells policymakers:} the analysis estimates how much additional health-system spending is required to move from today's coverage toward universal, high-quality access—and relates that investment to the health gains achieved.

\srcmethod{\textbf{Source:} intervention-based costing adapted from Watkins et al. (2020); unit costs and coverage from the input workbooks; demography from UN WPP 2024. \textbf{Method:} annual cost = people in need $\times$ coverage $\times$ annual service frequency $\times$ Indonesia-adjusted unit cost; additional cost = scenario minus baseline; equations in the appendix. \costcaveat}

# What the interventions deliver, and at what cost

## Scaling up clinical care sharply cuts premature cardiovascular deaths

```{r tbl-clin-ce, results='asis'}
ft <- family_table(met_clin, clin_all_id, ce_row)
if (nrow(ft$tb)) {
  k <- kbl(ft$tb, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=mkhdr(c("Clinical intervention","Deaths averted\n(cumulative)","Prob. of CVD death\nbefore age 70, 2050",
                  "Reduction in that\nprobability, 2050","Deaths reduction\n2050","Cost per\ndeath averted",
                  "Add'l cost/capita\nper yr, undisc.","Add'l cost/capita\nper yr, disc.")),
      align="lrrrrrrr") |>
    kable_styling(latex_options=c("hold_position","scale_down"), font_size=8.5) |>
    column_spec(1, bold=TRUE) |> row_spec(0, bold=TRUE)
  if (!is.na(ft$agg_idx)) k <- k |> row_spec(ft$agg_idx, bold=TRUE, background="#EAF1FB")
  print(k)
  if (ft$n_omitted > 0)
    cat(sprintf("\n\n\\vspace{1pt}\\scriptsize %d additional clinical intervention(s) omitted for space; see the appendix crosswalk and the workbook.\n", ft$n_omitted))
} else cat("\\begin{center}\\large No clinical interventions selected for this run.\\end{center}")
```

\srcmethod{\textbf{Source:} results contract (\texttt{09\_deck\_results.rds}) + clinical workbook; `r (HZ_START+1L)` population from model demography (UN WPP 2024). \textbf{Method:} cost per death averted $=$ discounted additional cost $\div$ deaths averted (`r fmt_pct(disc_rate*100,0)`/yr, `r price_year` US\$, `r cost_perspective`); the two per-capita columns $=$ each row's cumulative additional cost (undiscounted / discounted) $\div$ (25 $\times$ `r (HZ_START+1L)` national population), a simple average over the 25-year horizon. \costcaveat}

## Population policies avert deaths at very low cost

```{r tbl-ph-ce, results='asis'}
ft <- family_table(met_ph, ph_all_id, ce_row)
if (nrow(ft$tb)) {
  k <- kbl(ft$tb, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=mkhdr(c("Public-health policy","Deaths averted\n(cumulative)","Prob. of CVD death\nbefore age 70, 2050",
                  "Reduction in that\nprobability, 2050","Deaths reduction\n2050","Cost per\ndeath averted",
                  "Add'l cost/capita\nper yr, undisc.","Add'l cost/capita\nper yr, disc.")),
      align="lrrrrrrr") |>
    kable_styling(latex_options=c("hold_position","scale_down"), font_size=8.5) |>
    column_spec(1, bold=TRUE) |> row_spec(0, bold=TRUE)
  if (!is.na(ft$agg_idx)) k <- k |> row_spec(ft$agg_idx, bold=TRUE, background="#E7F4F1")
  print(k)
  if (ft$n_omitted > 0)
    cat(sprintf("\n\n\\vspace{1pt}\\scriptsize %d additional public-health intervention(s) omitted for space; see the appendix crosswalk and the workbook.\n", ft$n_omitted))
} else cat("\\begin{center}\\large No public-health policies selected for this run.\\end{center}")
```

\srcmethod{\textbf{Source:} results contract (\texttt{09\_deck\_results.rds}) + public-health workbook; `r (HZ_START+1L)` population from model demography (UN WPP 2024). \textbf{Method:} cumulative deaths averted `r HZ_START`--`r HZ_END` vs baseline; tobacco and sodium packages each modelled as one joint run; the two per-capita columns $=$ each row's cumulative additional cost (undiscounted / discounted) $\div$ (25 $\times$ `r (HZ_START+1L)` national population). \costcaveat}


# Clinical and population action, combined

## Clinical care and prevention achieve more together than alone

```{r tbl-combined, results='asis'}
pk <- intersect(c(clin_all_id, ph_all_id, joint_id), met_comb$scenario)
if (length(pk)) {
  tb <- rbindlist(lapply(pk, function(s) data.table(
    Package = lbl(s),
    da   = fmt_thou(hv(s, "deaths_averted")),
    qr   = fmt_pct(hv(s, "cvd_40q30_pct_reduction_2050"), 1),
    q50  = fmt_pct(hv(s, "cvd_40q30_2050"), 1),
    cost = fmt_usd_short(hv(s, "disc_incremental_cost")),
    cpd  = fmt_usd(hv(s, "cost_per_death_averted")),
    pcu  = pc_undisc(met_comb, s),
    pcd  = pc_disc(met_comb, s))))
  k <- kbl(tb, format="latex", booktabs=TRUE, escape=FALSE,
      col.names=mkhdr(c("Package","Deaths averted\n(cumulative)","Reduction in prob.\nof CVD death <70, 2050","Prob. of CVD death\nbefore age 70, 2050",
                  "Discounted total\ncost","Cost per\ndeath averted","Add'l cost/capita\nper yr, undisc.","Add'l cost/capita\nper yr, disc.")),
      align="lrrrrrrr") |>
    kable_styling(latex_options=c("hold_position","scale_down"), font_size=9) |>
    column_spec(1, bold=TRUE) |> row_spec(0, bold=TRUE)
  jidx <- which(pk == joint_id)
  if (length(jidx)) k <- k |> row_spec(jidx, bold=TRUE, background="#FCEFCB")
  print(k)
} else cat("\\begin{center}\\large No package scenarios available for this run.\\end{center}")
```

```{r combined-block, results='asis'}
if (has_joint) {
  cat(sprintf(paste0(
    "\\begin{combinedblock}{The combined package is modelled as one run --- not a sum}\n",
    "\\footnotesize Because the two families' effects overlap, the joint scenario is projected in a single run: about \\textbf{%s deaths} averted and the %d probability of CVD death before age 70 cut to \\textbf{%s} (from %s), at about \\textbf{%s per death averted}.\n",
    "\\end{combinedblock}\n"),
    fmt_short(combo_da), HZ_END, fmt_pct(combo_q50,1), fmt_pct(base_q_2050,1), fmt_usd(combo_cpda)))
} else {
  cat(paste0("\\begin{block}{Single-family run}\n\\footnotesize The joint clinical + ",
             "public-health scenario was not built in this configuration; family results are ",
             "shown above.\n\\end{block}\n"))
}
```

\srcmethod{\textbf{Source:} results contract (\texttt{09\_deck\_results.rds}) + combined workbook; `r (HZ_START+1L)` population from model demography (UN WPP 2024). \textbf{Method:} joint package $=$ one projection applying both families (not their sum); the two per-capita columns $=$ each row's cumulative additional cost (undiscounted / discounted, `r fmt_pct(disc_rate*100,0)`/yr, `r price_year` US\$) $\div$ (25 $\times$ `r (HZ_START+1L)` national population). \costcaveat}

## Indonesia's premature cardiovascular mortality can fall toward its neighbours

```{r fig-40q30, fig.width=9.6, fig.height=4.2, out.width="\\linewidth"}
# Remove the modeled status-quo baseline
idn_scen_plot <- idn_scen[scenario != "baseline"]
lab_scen <- idn_scen_plot[year == HZ_END]

lab_cmp <- cmp_lines[, .SD[which.max(year)], by = country_label]

ggplot() +
  geom_line(
    data = cmp_lines,
    aes(year, q, group = country_label),
    colour = col_comparator,
    linewidth = 0.5
  ) +
  ggrepel::geom_text_repel(
    data = lab_cmp,
    aes(year, q, label = country_label),
    colour = "grey55", size = 2.5, hjust = 0,
    nudge_x = 1.2, direction = "y",
    segment.colour = "grey80", segment.size = 0.2,
    min.segment.length = 0, box.padding = 0.12,
    max.overlaps = Inf, seed = 42
  ) +
  geom_line(
    data = idn_obs,
    aes(year, q),
    colour = "grey20",
    linewidth = 1.0
  ) +
  annotate(
    "segment",
    x = 2024, xend = 2024,
    y = 10.5, yend = 20.5,
    linetype = "dotted",
    colour = "grey45"
  ) +
  annotate(
    "text",
    x = 2024, y = 20.9,
    label = "Modeled scenarios →",
    hjust = 1, size = 2.7,
    colour = "grey35"
  ) +
  geom_line(
    data = idn_scen_plot,  # filtered data
    aes(year, q, colour = scen),
    linewidth = 1.15,
    linetype = "22"
  ) +
  ggrepel::geom_text_repel(
    data = lab_scen,
    aes(year, q, label = scen, colour = scen),
    size = 2.7, fontface = "bold", hjust = 0,
    nudge_x = 1.2, direction = "y",
    segment.colour = "grey70", segment.size = 0.2,
    min.segment.length = 0, box.padding = 0.2,
    max.overlaps = Inf, show.legend = FALSE, seed = 7
  ) +
  scale_colour_manual(values = scen_pal) +
  scale_x_continuous(
    limits = c(1995, 2064),
    breaks = seq(1995, 2050, 5),
    expand = expansion(mult = c(0.01, 0))
  ) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "Probability of death from CVD before age 70: observed, projected, and neighbours",
    subtitle = "Chance of dying from CVD between exact ages 30 and 70",
    x = NULL,
    y = "Probability of CVD death before age 70 (%)",
    caption = paste(
      "Indonesia observed history (solid, to 2023) and ASEAN comparators are WPP-scaled reference series.",
      "\nModelled intervention scenarios are dashed for 2025–2050.",
      "\nASEAN comparator trends use linear extrapolation for illustrative reference."
    )
  ) +
  theme_exec(11) +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_line(
      colour = "grey92",
      linewidth = 0.25
    )
  )
```

\srcmethod{\textbf{Source:} combined results workbook (annual probability-of-death sheet, R-reconciled values); ASEAN comparator series in \texttt{data/raw/}. \textbf{Method:} model scenarios (dashed) run `r HZ_START`--`r HZ_END` for the six modelled cardiovascular causes over exact ages 30--70; comparators are WPP-scaled reference series shown for context only.}

# A stronger care cascade: the 70-30-30 to 70-70-70 strategy

## The cascade scale-up is ambitious but staged

\footnotesize Hypertension and type-2 diabetes are modelled with \textbf{diagnosed $\rightarrow$ treated $\rightarrow$ controlled} care cascades; \textbf{cholesterol} treatment coverage follows hypertension coverage. Coverage rises in two stages, then holds.

\vspace{1pt}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{clinicalblock}{Two staged milestones}
\footnotesize
\begin{itemize}\itemsep1pt
  \item \textbf{`r casc_ytar1` --- 70-30-30:} `r fmt_pct(casc_diag*100,0)` diagnosed, `r fmt_pct(casc_t30*100,0)` of those treated, `r fmt_pct(casc_c30*100,0)` of those controlled.
  \item \textbf{`r casc_ytar2` --- 70-70-70:} `r fmt_pct(casc_diag*100,0)` diagnosed, `r fmt_pct(casc_t40*100,0)` treated, `r fmt_pct(casc_c40*100,0)` controlled; held through `r casc_y1`.
  \item Linear scale-up from the observed baseline; coverage already above a milestone is \textbf{not scaled down}.
\end{itemize}
\end{clinicalblock}
\end{column}
\begin{column}{0.48\textwidth}
\begin{combinedblock}{Shares of all people with the condition}
\footnotesize
\begin{tabular}{lcc}
 & \textbf{`r casc_ytar1`} & \textbf{`r casc_ytar2`}\\
\midrule
Diagnosed  & `r fmt_pct(casc_diag*100,0)` & `r fmt_pct(casc_diag*100,0)`\\
Treated    & `r fmt_pct(casc_diag*casc_t30*100,0)` & `r fmt_pct(casc_diag*casc_t40*100,0)`\\
Controlled & `r fmt_pct(casc_diag*casc_t30*casc_c30*100,1)` & `r fmt_pct(casc_diag*casc_t40*casc_c40*100,1)`\\
\end{tabular}\\[3pt]
{\scriptsize treated $=70\%\times\{30,70\}\%$; controlled $=$ treated $\times\{30,70\}\%$.}
\end{combinedblock}
\end{column}
\end{columns}

\vspace{1pt}
\footnotesize \textbf{Effective modelled coverage} $=$ controlled $+\tfrac{1}{2}\times$ treated-but-uncontrolled $=$ \textbf{`r fmt_pct(casc_eff30*100,2)`} (`r casc_ytar1`) and \textbf{`r fmt_pct(casc_eff40*100,2)`} (`r casc_ytar2`): full effect to controlled patients, half to treated-but-uncontrolled.

\srcmethod{\textbf{Source:} cascade workbook (\texttt{Cascade\_Assumptions}, \texttt{Cascade\_Trajectory}); isolated runner; NCD-RisC coverage inputs. \textbf{Method:} nested targets on the diagnosed/treated/controlled shares; piecewise-linear scale-up, no backsliding.}

## Indonesia's hypertension cascade can progress in two achievable stages

```{r fig-htn-cascade, fig.width=9.4, fig.height=3.7, out.width="0.94\\linewidth"}
htn_long <- melt(htn_sh[, .(year, Diagnosed = diagnosed, Treated = treated, Controlled = controlled)],
                 id.vars = "year", variable.name = "stage", value.name = "share")
htn_long <- htn_long[!is.na(share)]
htn_long[, stage := factor(stage, levels = c("Diagnosed","Treated","Controlled"))]
htn_long[, xlab := factor(year, levels = c(HZ_START, casc_ytar1, casc_ytar2),
                          labels = c(paste0("Baseline\n(", HZ_START, ")"),
                                     paste0(casc_ytar1, "\n(70-30-30)"),
                                     paste0(casc_ytar2, "\n(70-70-70)")))]
casc_blues <- c(Diagnosed = "#9ECAE1", Treated = "#4292C6", Controlled = "#08519C")
ggplot(htn_long, aes(xlab, share, fill = stage)) +
  geom_col(position = position_dodge2(width = 0.8, preserve = "single"), width = 0.8) +
  geom_text(aes(label = paste0(formatC(share * 100, format = "f", digits = 1), "%")),
            position = position_dodge2(width = 0.8, preserve = "single"),
            vjust = -0.4, size = 3.0, fontface = "bold") +
  scale_fill_manual(values = casc_blues) +
  scale_y_continuous(labels = function(x) paste0(x * 100, "%"), limits = c(0, 0.80),
                     expand = expansion(mult = c(0, 0.02))) +
  labs(title = "Hypertension care cascade under the 70-30-30 to 70-70-70 scale-up",
       subtitle = "Diagnosed, treated, and controlled shares of all people with hypertension (intervention scenario)",
       x = NULL, y = "% of people with hypertension") +
  theme_exec(11) +
  theme(legend.position = "top", panel.grid.major.x = element_blank())
```

\srcmethod{\textbf{Source:} cascade \texttt{Cascade\_Trajectory} (sex-averaged) and \texttt{Cascade\_Assumptions}; scale-up begins `r HZ_START+1L`. \textbf{Method:} baseline (`r HZ_START`) shows only measured treated/controlled --- diagnosis is not modelled at baseline, so no diagnosed bar. Type-2 diabetes follows the same targets (baseline treated $\approx$ `r fmt_pct(dm_base$treated*100,1)`, controlled $\approx$ `r fmt_pct(dm_base$controlled*100,1)`).}

## Stronger prevention lowers premature deaths at a rising but bounded cost

```{r fig-casc-2panel, fig.width=9.6, fig.height=3.7, out.width="\\linewidth"}
lab_q  <- c(baseline = "Baseline (current course)", cascade = "Intervention")
qdat <- copy(casc_q)[, series := ifelse(scenario == "baseline", "Baseline (current course)", "Intervention")]
qdat[, panel := "Probability of death from CVD before age 70 (%)"]
setnames(qdat, "q", "value")
cdat <- rbind(
  casc_cost[, .(year, series = "Baseline (current course)", value = cum_base / 1e9)],
  casc_cost[, .(year, series = "Intervention",              value = cum_scen / 1e9)])
cdat[, panel := "Cumulative discounted health-system cost (US$ billions)"]
two <- rbind(qdat[, .(year, series, value, panel)], cdat)
two[, panel := factor(panel, levels = c("Probability of death from CVD before age 70 (%)",
                                         "Cumulative discounted health-system cost (US$ billions)"))]
casc_pal <- c("Baseline (current course)" = col_baseline, "Intervention" = "#D55E00")
ends <- two[, .SD[year == max(year)], by = .(series, panel)]
ggplot(two, aes(year, value, colour = series)) +
  geom_line(linewidth = 1.15) +
  ggrepel::geom_text_repel(data = ends, aes(label = series), size = 2.7, fontface = "bold",
                           hjust = 0, nudge_x = 1.0, direction = "y", segment.colour = NA,
                           max.overlaps = Inf, seed = 3, show.legend = FALSE) +
  facet_wrap(~panel, scales = "free_y") +
  scale_colour_manual(values = casc_pal) +
  scale_x_continuous(limits = c(HZ_START, casc_y1 + 6), breaks = seq(HZ_START, casc_y1, 5)) +
  labs(title = NULL, x = "Year", y = NULL) +
  theme_exec(11) +
  theme(legend.position = "none",
        strip.text = element_text(face = "bold", colour = col_wbnavy, size = 10),
        panel.grid.major.x = element_line(colour = "grey92", linewidth = 0.25))
```

\srcmethod{\textbf{Source:} cascade workbook --- annual probability-of-death sheet (R-reconciled) and \texttt{Annual\_Cost}, cumulatively summed and reconciled to \texttt{Budget\_Impact}. \textbf{Method:} probability over the six CVD causes, ages 30--70; cumulative discounted health-system cost (`r fmt_pct(casc_disc*100,0)`/yr to `r HZ_START`, `r casc_price` US\$) of the CVD primary-care and type-2 diabetes cascades. \costcaveat}

## `r title_cascade_results`

```{r tbl-cascade, results='asis'}
snp <- lapply(casc_tab_years, casc_snapshot); names(snp) <- as.character(casc_tab_years)
mm  <- function(x) formatC(x / 1e6, format = "f", digits = 3, big.mark = ",")   # deaths, millions
bn  <- function(x) formatC(x / 1e9, format = "f", digits = 2)                    # cost, US$ billions
casc_tb <- data.table(
  Metric = c("\\quad Baseline (current course)", "\\quad Intervention",
             "\\quad Baseline (current course)", "\\quad Intervention",
             "Additional cost (US\\$ billion, discounted)", "Deaths prevented (million)"),
  y1 = c(mm(snp[[1]]$d_base), mm(snp[[1]]$d_scen), bn(snp[[1]]$c_base), bn(snp[[1]]$c_scen), bn(snp[[1]]$c_add), mm(snp[[1]]$d_av)),
  y2 = c(mm(snp[[2]]$d_base), mm(snp[[2]]$d_scen), bn(snp[[2]]$c_base), bn(snp[[2]]$c_scen), bn(snp[[2]]$c_add), mm(snp[[2]]$d_av)),
  y3 = c(mm(snp[[3]]$d_base), mm(snp[[3]]$d_scen), bn(snp[[3]]$c_base), bn(snp[[3]]$c_scen), bn(snp[[3]]$c_add), mm(snp[[3]]$d_av)))
kbl(casc_tb, format = "latex", booktabs = TRUE, escape = FALSE,
    col.names = c("Cumulative from 2025 through year", as.character(casc_tab_years)),
    align = "lrrr") |>
  kable_styling(latex_options = c("hold_position"), font_size = 8.5) |>
  column_spec(1, width = "6.6cm") |> column_spec(2:4, width = "2.2cm") |>
  row_spec(0, bold = TRUE) |>
  pack_rows("Projected deaths --- CVD (six causes) + type-2 diabetes (million)", 1, 2, bold = TRUE, latex_gap_space = "0.25em", escape = FALSE) |>
  pack_rows("Projected health-system cost (US\\$ billion, discounted)", 3, 4, bold = TRUE, latex_gap_space = "0.25em", escape = FALSE) |>
  pack_rows("Net effect of the strategy", 5, 6, bold = TRUE, latex_gap_space = "0.25em")
```

\srcmethod{\textbf{Source:} cascade \texttt{Annual\_Mortality} and \texttt{Annual\_Cost}/\texttt{Budget\_Impact}; scale-up begins `r HZ_START+1L`. \textbf{Method:} each column cumulative from `r HZ_START` through that year; deaths prevented $=$ baseline $-$ intervention (six CVD causes $+$ type-2 diabetes); additional cost $=$ intervention $-$ baseline (discounted `r fmt_pct(casc_disc*100,0)` to `r HZ_START`, `r casc_price` US\$). \costcaveat}

# What this means for policy

## Prevention and clinical care together offer the biggest, most affordable gains

\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep4pt
```{r impl-left, results='asis'}
if (has_joint)
  cat(sprintf("  \\item \\textbf{Biggest health gains come from doing both.} The combined package averts far more (\\textasciitilde%s) than either family alone.\n", fmt_short(combo_da)))
cat("  \\item \\textbf{Public-health policies are exceptional value.} Population-wide fiscal/regulatory measures avert deaths at a fraction of clinical cost per death averted.\n")
cat("  \\item \\textbf{Clinical scale-up delivers large, durable gains} in survival and premature mortality.\n")
```
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{itemize}\footnotesize\itemsep4pt
  \item \textbf{Sequence for early wins.} Tobacco taxation and other prevention policies can start cutting risk quickly and cheaply, while clinical capacity scales up.
```{r impl-right, results='asis'}
if (has_joint)
  cat(sprintf("  \\item \\textbf{The premature-mortality target is within reach.} The combined package bends Indonesia's probability of death from CVD before age 70 down by about %s by %d.\n", fmt_pct(combo_qred,0), HZ_END))
```
\end{itemize}
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} results contract (\texttt{09\_deck\_results.rds}); cascade workbook. \textbf{Method:} interpretation of modelled deaths averted, premature-mortality reduction, and cost per death averted; policy translation needs judgement on feasibility, financing, and delivery. \costcaveat}

## What these results can and cannot tell us

\begin{itemize}\footnotesize\itemsep3pt
  \item Results depend on \textbf{modelled epidemiological estimates} (GBD) and calibration; the real burden may differ.
  \item \textbf{Effect sizes, future coverage, and unit costs are uncertain}; interventions are assumed to reach and sustain their targets.
  \item The disease model is \textbf{simplified} (healthy $\to$ living with CVD $\to$ death; no recovery; no comorbidity interactions).
  \item Value for money is expressed as \textbf{cost per death averted}; effects on illness (morbidity) and productivity are not valued in money here.
  \item \textbf{Costing results are preliminary.} Unit costs, service use, and delivery assumptions require further validation against \textbf{BPJS and Ministry of Health} data before policy use.
  \item Projections are \textbf{scenarios, not forecasts or guarantees}; they show what is achievable if interventions are delivered as modelled.
\end{itemize}

\srcmethod{\textbf{Source:} Indonesia CVD model and its input workbooks (GBD 2023; UN WPP 2024; FAIR Choices effects; programme and policy unit costs). \textbf{Method:} qualitative statement of the assumptions and uncertainties that bound the quantitative results above.}

## The case for action: two levers, affordable, and a reachable target

\begin{columns}[T]
\begin{column}{0.33\textwidth}
```{r remember-1, results='asis'}
if (has_joint) {
  cat(sprintf("\\begin{combinedblock}{1. Both, together}\n\\footnotesize Clinical + public health averts \\textbf{\\textasciitilde%s deaths} by %d --- more than either alone.\n\\end{combinedblock}\n", fmt_short(combo_da), HZ_END))
} else {
  cat("\\begin{combinedblock}{1. Two levers}\n\\footnotesize Clinical care and population prevention each cut premature CVD mortality.\n\\end{combinedblock}\n")
}
```
\end{column}
\begin{column}{0.33\textwidth}
```{r remember-2, results='asis'}
cat(sprintf("\\begin{phblock}{2. Prevention is affordable}\n\\footnotesize Population policies avert deaths at about \\textbf{%s} each.\n\\end{phblock}\n", fmt_usd(ph_cpda %+na% NA_real_)))
```
\end{column}
\begin{column}{0.33\textwidth}
```{r remember-3, results='asis'}
if (has_joint) {
  cat(sprintf("\\begin{clinicalblock}{3. A reachable target}\n\\footnotesize The probability of death from CVD before age 70 falls \\textbf{%s} by %d under the combined package.\n\\end{clinicalblock}\n", fmt_pct(combo_qred,0), HZ_END))
} else {
  cat(sprintf("\\begin{clinicalblock}{3. A reachable target}\n\\footnotesize All clinical interventions cut the probability of death from CVD before age 70 by \\textbf{%s} by %d.\n\\end{clinicalblock}\n", fmt_pct(clin_qred,0), HZ_END))
}
```
\end{column}
\end{columns}

\vspace{4pt}
```{r closing-banner, results='asis'}
if (has_joint) {
  cat(sprintf(paste0("\\begin{center}\n\\large Combined action against CVD could avert about \\hlkey{%s deaths} ",
                     "at about \\hlkey{%s per death averted} through %d.\n\\end{center}\n"),
              fmt_short(combo_da), fmt_usd(combo_cpda), HZ_END))
} else {
  cat(sprintf(paste0("\\begin{center}\n\\large Scaling up CVD interventions could avert hundreds of thousands ",
                     "of deaths through %d at a modest cost per death averted.\n\\end{center}\n"), HZ_END))
}
```

\srcmethod{\textbf{Source:} results contract (\texttt{09\_deck\_results.rds}). \textbf{Method:} deaths averted cumulative `r HZ_START`--`r HZ_END`; reduction in the probability of death from CVD before age 70 in `r HZ_END`; cost per death averted uses discounted additional cost. \costcaveat}

# Technical appendix

## Appendix 1 --- Scenario crosswalk

```{r tbl-crosswalk}
## Built dynamically from the current run's combined-workbook scenarios.
fam_order <- c(clinical = 1, public_health = 2, clinical_public_health = 3)
xwd <- copy(met_comb)[, .(scenario, intervention_family, scenario_label, n_interventions)]
xwd[, ford := fam_order[intervention_family]]
xwd[is.na(ford), ford := 9]
setorder(xwd, ford, -n_interventions)
fam_disp <- c(clinical = "Clinical (FAIR Choices)", public_health = "Public health",
              clinical_public_health = "Combined")
xwd[, Family := fam_disp[intervention_family]]
xwd[is.na(Family), Family := intervention_family]
## Only show the family label on its first row (visual grouping).
xwd[, Family := ifelse(duplicated(Family), "", Family)]
xwd[, IDesc := gsub("_", "\\\\_", scenario)]
xw <- xwd[, .(Family, ID = IDesc, Label = scenario_label)]
kbl(xw, format="latex", booktabs=TRUE, escape=FALSE,
    col.names=c("Family","Workbook scenario id","Executive label"), align="lll") |>
  kable_styling(latex_options=c("hold_position"), font_size=5.2) |>
  column_spec(1, bold=TRUE, width="3.0cm") |> column_spec(2, monospace=TRUE, width="4.0cm") |>
  column_spec(3, width="6.6cm") |> row_spec(0, bold=TRUE)
```

\srcmethod{\textbf{Source:} combined workbook \texttt{Scenario\_Catalog} (binding active-scenario contract). \textbf{Method:} identity/labels/family read from the current run; excluded interventions never appear; packages are single joint runs of their children (omitted from main tables to avoid double-counting).}

## Appendix 2 --- Definitions, horizons and assumptions

\footnotesize
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\textbf{Metrics}
\begin{itemize}\scriptsize\itemsep2pt
  \item \textbf{Deaths averted (cumulative)} $=\sum_{t=`r HZ_START`}^{`r HZ_END`}(\text{baseline deaths}_t-\text{scenario deaths}_t)$, across modelled CVD \& diabetes causes.
  \item \textbf{\% reduction in deaths, `r HZ_END`} $=100\times(\text{baseline}-\text{scenario})/\text{baseline}$ in `r HZ_END`.
  \item \textbf{Probability of death from CVD before age 70} $=$ probability of dying of CVD between exact ages 30 and 70 (six CVD causes, both sexes, single-year life table).
  \item \textbf{\% reduction in that probability, `r HZ_END`}: analogous to deaths.
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\textbf{Costing \& value for money}
\begin{itemize}\scriptsize\itemsep2pt
  \item Horizon `r HZ_START`--`r HZ_END`; cost discount rate `r fmt_pct(disc_rate*100,0)`/yr; price year `r price_year`.
  \item Costs in \textbf{market US\$}; cost perspective: `r cost_perspective`.
  \item \textbf{Cost per death averted} $=$ cumulative \textbf{discounted} additional (incremental) cost $\div$ cumulative \textbf{undiscounted} deaths averted.
  \item A ``Dominant'' scenario averts deaths while lowering cost; ``not defined'' means no deaths averted.
\end{itemize}
\end{column}
\end{columns}

\vspace{3pt}
\scriptsize \textbf{Model.} Discrete-time state-transition model; single-year ages 0--95; six CVD causes (IHD, ischemic stroke, intracerebral haemorrhage, hypertensive heart disease, rheumatic heart disease, cardiomyopathy \& myocarditis) plus type-2 diabetes; calibrated to GBD 2023; demography and life expectancy from UN WPP 2024. \textbf{Data provenance:} every ordinary-run number is read from the current-run results contract (\texttt{output/09\_deck\_results.rds}) and the workbooks' literal probability-of-death series; the cascade strategy is read from its isolated workbook; scenario identity follows the \texttt{Scenario\_Catalog} contract (binding \texttt{include\_flag}). No Excel recalculation is used.

## Appendix 3 --- How costs are calculated

\footnotesize Intervention-based costing adapted from the DCP Cost Model of @Watkins2020ResourceRequirements, from a `r cost_perspective` perspective with annual projections. For intervention $i$, sex/age group $s$, year $t$:

\vspace{-2pt}
\begin{columns}[T]
\begin{column}{0.52\textwidth}
{\scriptsize
\begin{align*}
C_{i,s,t} &= PIN_{i,s,t}\times cov_{i,s,t}\times f_i\times uc_i\\[2pt]
\Delta C_{s,t} &= C_{s,t}-C_{0,t}\\[2pt]
PV(\Delta C_s) &= \sum_{t=`r HZ_START+1L`}^{`r HZ_END`}\frac{\Delta C_{s,t}}{(1+r)^{\,t-`r HZ_START`}}\\[2pt]
\text{Cost per death prevented}_s &= \frac{PV(\Delta C_s)}{\sum_{t=`r HZ_START+1L`}^{`r HZ_END`}(D_{0,t}-D_{s,t})}
\end{align*}}
\end{column}
\begin{column}{0.46\textwidth}
\scriptsize
$PIN$ = people in need (population $\times$ epidemiological need); $cov$ = coverage (baseline or scenario); $f_i$ = services per person per year; $uc_i$ = Indonesia-adjusted unit cost; subscript $0$ = baseline (current course), $s$ = scenario; $r$ = `r fmt_pct(disc_rate*100,0)`/yr discount rate; price year `r price_year`; currency market US\$; perspective `r cost_perspective`. Deaths $D$ remain \textbf{undiscounted} in the cost-per-death calculation. Downstream cost offsets are \textbf{excluded}.
\end{column}
\end{columns}

\srcmethod{\textbf{Source:} adapted from Watkins et al. (2020); unit costs, coverage, and people-in-need from the input workbooks and \texttt{Annual\_Cost}; demography from UN WPP 2024. \textbf{Method:} annual costs by intervention $\times$ group $\times$ year, differenced against baseline, discounted to the `r HZ_START` base year. \costcaveat}

## Appendix 4 --- References

\scriptsize

::: {#refs}
:::
