---
title: "Indonesia NCD FAIR Choices: Costing & Economic Analysis"
subtitle: "Cardiovascular intervention package — budget impact and interim cost-effectiveness"
date: "`r format(Sys.Date(), '%d %B %Y')`"
author: "University of Washington · World Bank Indonesia NCD program"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
    code_folding: hide
    self_contained: true
    css: indonesia_ncd_report.css
    mathjax: default
params:
  workbook_path: "output/indonesia_model_cost_value_formulae.xlsx"
  focus_scenario: "all"
  decision_horizon: NA
  show_code: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = params$show_code,
  error   = FALSE,      # knitting must fail loudly on real errors, not swallow them
  message = FALSE,
  warning = FALSE,
  dev     = "png",
  dpi     = 150,
  fig.align  = "center",
  fig.width  = 9,
  fig.height = 5.2
)
```

<style>
/* ---- Report-specific helpers layered on top of indonesia_ncd_report.css ---- */
.badge{display:inline-block;padding:2px 10px;border-radius:999px;font-size:11px;
  font-weight:800;letter-spacing:.04em;color:#fff;vertical-align:middle;}
.badge-pass{background:#1b9e8b;}
.badge-review{background:#c98a00;}
.badge-fail{background:#c0392b;}
.badge-muted{background:#7a8b99;}
.eq-note{margin:-14px 0 26px;color:var(--ink-500);font-size:12.5px;text-align:center;}
.math.display{max-width:820px;margin-left:auto;margin-right:auto;}
.kpi-value small{font-size:14px;font-weight:600;color:var(--ink-500);letter-spacing:0;}
table.compact-kbl{font-size:13px;}
</style>

```{r packages, include=FALSE}
# ---- Dependency guard: detect, never install --------------------------------
.need <- c("here", "readxl", "dplyr", "tidyr", "tibble", "purrr", "stringr",
           "forcats", "scales", "ggplot2", "knitr", "kableExtra", "DT",
           "htmltools", "glue")
.missing <- .need[!vapply(.need, requireNamespace, logical(1), quietly = TRUE)]
if (length(.missing)) {
  stop(
    "The following required packages are not installed:\n  ",
    paste(.missing, collapse = ", "),
    "\nInstall them in R (this report never installs packages itself):\n  ",
    sprintf('install.packages(c(%s))',
            paste0('"', .missing, '"', collapse = ", ")),
    call. = FALSE
  )
}
suppressPackageStartupMessages({
  library(here); library(readxl); library(dplyr); library(tidyr)
  library(tibble); library(purrr); library(stringr); library(forcats)
  library(scales); library(ggplot2); library(knitr); library(kableExtra)
  library(DT); library(htmltools); library(glue)
})
```

```{r resolve-workbook, include=FALSE}
# ---- Locate the single canonical cost-value workbook ------------------------
# Rule: use params$workbook_path when it exists; otherwise glob the canonical
# pattern. Zero matches -> stop. Multiple matches -> only proceed if the exact
# canonical filename is present, else stop and report the ambiguity.
resolve_workbook <- function(rel_path) {
  root <- here::here()
  explicit <- here::here(rel_path)
  if (file.exists(explicit)) return(normalizePath(explicit, winslash = "/"))

  search_dir <- file.path(root, "output")
  hits <- list.files(
    search_dir,
    pattern    = "^indonesia_model_cost_value_formulae.*\\.xlsx$",
    full.names = TRUE
  )
  hits <- hits[!grepl("~\\$", basename(hits))]   # drop Excel lock stubs

  if (length(hits) == 0) {
    stop(glue::glue(
      "No cost-value formulae workbook found.\n",
      "Looked for '{rel_path}' and for pattern ",
      "'indonesia_model_cost_value_formulae*.xlsx' in '{search_dir}'.\n",
      "Run code/cvd-fair-choices/09_cost_value.R to (re)generate it, ",
      "then recalculate and save the workbook so cached formula values are current."
    ), call. = FALSE)
  }
  if (length(hits) == 1) return(normalizePath(hits, winslash = "/"))

  canonical <- hits[basename(hits) == "indonesia_model_cost_value_formulae.xlsx"]
  if (length(canonical) == 1) return(normalizePath(canonical, winslash = "/"))

  stop(glue::glue(
    "Ambiguous workbook selection: {length(hits)} files match ",
    "'indonesia_model_cost_value_formulae*.xlsx' and none is the canonical ",
    "'indonesia_model_cost_value_formulae.xlsx':\n  ",
    "{paste(basename(hits), collapse = '\n  ')}\n",
    "Set params$workbook_path explicitly to the file you intend to use."
  ), call. = FALSE)
}

WB_PATH <- resolve_workbook(params$workbook_path)
WB_REL  <- sub(paste0("^", here::here(), "/?"), "", WB_PATH)
```

```{r ingest, include=FALSE}
# ---- Read every required sheet exactly once ---------------------------------
REQUIRED_SHEETS <- c(
  "README", "Run_Metadata", "Selected_Interventions", "Blocked_Links",
  "Cost_Components", "Annual_Mortality", "Annual_Cost", "Budget_Impact",
  "Cost_Effectiveness", "Economic_Value", "QA_Checks", "Input_Diagnostic",
  "Methods_and_Sources", "Calculation_Assumptions", "Calculation_Map"
)

sheets_present <- readxl::excel_sheets(WB_PATH)
missing_sheets <- setdiff(REQUIRED_SHEETS, sheets_present)
if (length(missing_sheets)) {
  stop(glue::glue(
    "Workbook '{basename(WB_PATH)}' is missing required sheet(s): ",
    "{paste(missing_sheets, collapse = ', ')}.\n",
    "This report reads only the cost-value workbook; regenerate it via Model 09."
  ), call. = FALSE)
}

read_sheet <- function(sheet) {
  suppressWarnings(
    readxl::read_excel(WB_PATH, sheet = sheet, .name_repair = "minimal")
  ) |> as.data.frame(stringsAsFactors = FALSE)
}
WB <- setNames(lapply(REQUIRED_SHEETS, read_sheet), REQUIRED_SHEETS)

# ---- Required-column contracts (validated, never silently renamed) ----------
REQUIRED_COLS <- list(
  Selected_Interventions = c("intervention_id", "intervention_name", "cause_code",
                             "effect_value", "affected_fraction", "baseline_coverage",
                             "target_coverage", "start_year", "target_year",
                             "adjusted_effect_at_target", "cost_scope",
                             "effect_review", "coverage_review"),
  Cost_Components = c("cost_component_key", "intervention_id", "cost_component",
                      "population_in_need_measure", "population_in_need_fraction",
                      "frequency_per_year", "unit_cost_usd", "price_year",
                      "cov_baseline", "cov_target", "cost_review", "cost_ready"),
  Annual_Mortality = c("scenario", "scenario_label", "year", "cause", "cases",
                       "cause_deaths", "base_deaths", "deaths_averted",
                       "base_cases", "cases_averted"),
  Annual_Cost = c("scenario", "year", "intervention_id", "cost_component_key",
                  "coverage_scenario", "annual_cost_baseline", "annual_cost_scenario",
                  "annual_cost_incremental", "disc_cost_incremental",
                  "shared_duplicate_count"),
  Budget_Impact = c("scenario", "year", "baseline_cost", "scenario_cost",
                    "incremental_cost", "disc_incremental_cost",
                    "cumulative_incremental_cost", "cumulative_disc_incremental_cost"),
  Cost_Effectiveness = c("scenario", "scenario_label", "deaths_averted",
                         "cases_averted", "incremental_cost", "disc_incremental_cost",
                         "cost_per_death_averted", "dominance", "reconciliation_status"),
  QA_Checks = c("check", "expected", "actual", "status", "note"),
  Input_Diagnostic = c("scope", "item_key", "field", "problem", "severity"),
  Calculation_Assumptions = c("parameter_id", "value", "unit", "role",
                              "description", "source"),
  Calculation_Map = c("output_sheet", "formula_columns", "depends_on",
                      "calculation", "source_role"),
  Methods_and_Sources = c("method_id", "concept", "formula_or_rule", "source"),
  Run_Metadata = c("item", "value")
)
for (sh in names(REQUIRED_COLS)) {
  miss <- setdiff(REQUIRED_COLS[[sh]], names(WB[[sh]]))
  if (length(miss)) {
    stop(glue::glue(
      "Sheet '{sh}' is missing expected column(s): {paste(miss, collapse = ', ')}.\n",
      "Column names in the workbook are authoritative; this report will not guess.\n",
      "Regenerate the workbook or reconcile the header names."
    ), call. = FALSE)
  }
}
```

```{r cached-value-guard, include=FALSE}
# ---- Confirm cached formula results are present and error-free ---------------
# The report uses Excel's cached values for formula cells. If the workbook was
# saved without recalculating, essential numeric fields would be NA or carry
# spreadsheet error tokens. Fail with an actionable message rather than
# reconstructing results from any other source.
EXCEL_ERRORS <- c("#REF!", "#VALUE!", "#DIV/0!", "#N/A", "#NAME?", "#NULL!",
                  "#NUM!", "#SPILL!", "#CALC!", "#GETTING_DATA")

scan_excel_errors <- function(tbl, sheet) {
  chr <- tbl[vapply(tbl, is.character, logical(1))]
  hits <- unlist(lapply(chr, function(col) col[col %in% EXCEL_ERRORS]), use.names = FALSE)
  if (length(hits)) {
    stop(glue::glue(
      "Sheet '{sheet}' contains spreadsheet error value(s): ",
      "{paste(unique(hits), collapse = ', ')}.\n",
      "Open '{basename(WB_PATH)}' in Excel, press Recalculate (F9 / Calculate Now), ",
      "save, and re-knit. This report does not repair or substitute workbook results."
    ), call. = FALSE)
  }
  invisible(TRUE)
}
invisible(lapply(REQUIRED_SHEETS, function(s) scan_excel_errors(WB[[s]], s)))

assert_populated <- function(tbl, cols, sheet) {
  for (cc in cols) {
    v <- tbl[[cc]]
    if (!is.numeric(v)) v <- suppressWarnings(as.numeric(v))
    if (all(is.na(v))) {
      stop(glue::glue(
        "Essential calculated field '{cc}' on sheet '{sheet}' is empty (all NA).\n",
        "The workbook appears to have been saved without cached formula results.\n",
        "Recalculate and save '{basename(WB_PATH)}', then re-knit."
      ), call. = FALSE)
    }
  }
  invisible(TRUE)
}
assert_populated(WB$Cost_Effectiveness,
                 c("deaths_averted", "disc_incremental_cost", "cost_per_death_averted"),
                 "Cost_Effectiveness")
assert_populated(WB$Budget_Impact,
                 c("incremental_cost", "disc_incremental_cost",
                   "cumulative_disc_incremental_cost"),
                 "Budget_Impact")
assert_populated(WB$Annual_Mortality, c("deaths_averted", "cause_deaths"),
                 "Annual_Mortality")
```

```{r helpers, include=FALSE}
# ---- Assumptions accessor ----------------------------------------------------
ca <- WB$Calculation_Assumptions
assume <- function(id, as_num = FALSE) {
  v <- ca$value[match(id, ca$parameter_id)]
  if (length(v) == 0 || is.na(v)) return(NA)
  if (as_num) suppressWarnings(as.numeric(v)) else as.character(v)
}
meta_val <- function(item) {
  rm <- WB$Run_Metadata
  v <- rm$value[match(item, rm$item)]
  if (length(v) == 0) NA_character_ else as.character(v)
}

START_YEAR   <- assume("analysis_start_year", TRUE)
END_YEAR     <- assume("analysis_end_year",   TRUE)
DISCOUNT     <- assume("cost_discount_rate",  TRUE)
PRICE_YEAR   <- assume("cost_price_year",     TRUE)
CURRENCY     <- assume("currency")
PERSPECTIVE  <- assume("economic_perspective")
SCALE_SHAPE  <- assume("scale_up_shape")
BASE_ID      <- assume("baseline_scenario_id")
TOL          <- assume("formula_tolerance", TRUE)
LOCATION     <- meta_val("Location")

# Decision horizon: default to the workbook end year. Robust to how YAML delivers
# the default — logical NA, the string "NA", empty, or NULL all mean "use end year".
.dh_raw <- params$decision_horizon
.dh_num <- suppressWarnings(as.numeric(.dh_raw))
HORIZON <- if (length(.dh_raw) == 0 || all(is.na(.dh_raw)) ||
               length(.dh_num) == 0 || is.na(.dh_num)) END_YEAR else .dh_num
HORIZON <- max(START_YEAR, min(END_YEAR, HORIZON))   # clamp into the modelled range
HORIZON_DEFAULTED <- identical(as.numeric(HORIZON), as.numeric(END_YEAR))
N_YEARS      <- END_YEAR - START_YEAR + 1L

# ---- Formatting helpers ------------------------------------------------------
fmt_n   <- function(x, d = 0) ifelse(is.na(x), "—",
             formatC(round(as.numeric(x), d), format = "f", big.mark = ",", digits = d))
fmt_pct <- function(x, d = 0) ifelse(is.na(x), "—",
             paste0(formatC(100 * as.numeric(x), format = "f", digits = d), "%"))
usd_big <- function(x, d = 1) {
  x <- as.numeric(x)
  ifelse(is.na(x), "—",
    ifelse(abs(x) >= 1e9, paste0("$", formatC(x/1e9, format="f", digits=d, big.mark=","), "B",
      ""),
    ifelse(abs(x) >= 1e6, paste0("$", formatC(x/1e6, format="f", digits=d, big.mark=","), "M"),
    ifelse(abs(x) >= 1e3, paste0("$", formatC(x/1e3, format="f", digits=d, big.mark=","), "k"),
           paste0("$", formatC(x, format="f", digits=0, big.mark=","))))))
}
usd0 <- function(x) ifelse(is.na(x), "—",
          paste0("$", formatC(round(as.numeric(x)), format = "f", big.mark = ",", digits = 0)))
count_big <- function(x, d = 2) {
  x <- as.numeric(x)
  ifelse(is.na(x), "—",
    ifelse(abs(x) >= 1e6, paste0(formatC(x/1e6, format="f", digits=d, big.mark=","), "M"),
    ifelse(abs(x) >= 1e3, paste0(formatC(x/1e3, format="f", digits=d, big.mark=","), "k"),
           formatC(round(x), format="f", big.mark=",", digits=0))))
}

status_badge <- function(s) {
  s <- toupper(trimws(as.character(s)))
  cls <- dplyr::case_when(
    s == "PASS" ~ "badge-pass",
    s == "REVIEW" ~ "badge-review",
    s == "FAIL" ~ "badge-fail",
    TRUE ~ "badge-muted"
  )
  sprintf('<span class="badge %s">%s</span>', cls, s)
}

# ---- Cause & scenario display maps ------------------------------------------
cause_labels <- c(
  ihd = "Ischemic heart disease", istroke = "Ischemic stroke",
  hstroke = "Intracerebral hemorrhage", hhd = "Hypertensive heart disease",
  rhd = "Rheumatic heart disease", cmd = "Cardiomyopathy & myocarditis",
  dm2 = "Type 2 diabetes"
)
lbl_cause <- function(x) ifelse(x %in% names(cause_labels), cause_labels[x], x)

ce <- WB$Cost_Effectiveness
scen_lab_map <- setNames(ce$scenario_label, ce$scenario)
scen_lab_map[BASE_ID] <- "Baseline (no new intervention)"
lbl_scen  <- function(x) ifelse(x %in% names(scen_lab_map), scen_lab_map[x], x)
wrap_scen <- function(x, w = 18) str_wrap(lbl_scen(x), w)

# ---- Colour-blind-safe palette (Okabe–Ito), stable per scenario -------------
scen_levels <- c("I_CVD_PRIMARY", "I_IHD_SECONDARY", "I_HF_ACUTE", "I_HF_CHRONIC",
                 "I_STROKE_SECONDARY", "I_RHD_SURGERY", "I_RHD_SECONDARY", "all")
scen_cols <- c(
  I_CVD_PRIMARY      = "#0072B2",  # blue
  I_IHD_SECONDARY    = "#D55E00",  # vermillion
  I_HF_ACUTE         = "#009E73",  # bluish green
  I_HF_CHRONIC       = "#56B4E9",  # sky blue
  I_STROKE_SECONDARY = "#CC79A7",  # reddish purple
  I_RHD_SURGERY      = "#E69F00",  # orange
  I_RHD_SECONDARY    = "#8C8C8C",  # grey
  all                = "#000000"   # black — combined bundle
)
cause_cols <- c(
  ihd = "#0072B2", istroke = "#CC79A7", hstroke = "#E69F00",
  hhd = "#009E73", rhd = "#D55E00", cmd = "#56B4E9", dm2 = "#8C8C8C"
)

theme_report <- function(base = 12) {
  theme_minimal(base_size = base) +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.title = element_text(face = "bold", size = base + 2),
      plot.subtitle = element_text(color = "#52697d", size = base - 1),
      plot.caption = element_text(color = "#7a8b99", size = base - 3, hjust = 0),
      legend.position = "bottom",
      legend.title = element_text(face = "bold", size = base - 2),
      legend.text = element_text(size = base - 2),
      axis.title = element_text(color = "#29465b")
    )
}
src_cap <- function(sheet) paste0("Source: ", sheet, " sheet · ", basename(WB_PATH))

# ---- Focus scenario resolution ----------------------------------------------
focus_id <- params$focus_scenario
focus_available <- focus_id %in% ce$scenario
if (!focus_available) {
  # Do not silently substitute: fall back to the highest-impact available row
  # and record that the requested focus was absent.
  focus_id <- ce$scenario[which.max(ce$deaths_averted)]
}
focus_row <- ce[ce$scenario == focus_id, , drop = FALSE]

# ---- Independent QA recomputations (comparison only, never overrides WB) ----
bi <- WB$Budget_Impact
am <- WB$Annual_Mortality
bi_final <- bi |>
  group_by(scenario) |>
  filter(year == max(year)) |>
  summarise(cum_disc = cumulative_disc_incremental_cost,
            cum_undisc = cumulative_incremental_cost, .groups = "drop")
am_tot <- am |>
  filter(scenario != BASE_ID) |>
  group_by(scenario) |>
  summarise(da = sum(deaths_averted), ca = sum(cases_averted), .groups = "drop")

recon <- ce |>
  select(scenario, deaths_averted, disc_incremental_cost, cost_per_death_averted) |>
  left_join(bi_final, by = "scenario") |>
  left_join(am_tot, by = "scenario") |>
  mutate(
    d_cost   = disc_incremental_cost - cum_disc,
    d_deaths = deaths_averted - da,
    cpda_recomp = ifelse(da > 0, cum_disc / da, NA_real_),
    cpda_diff   = cost_per_death_averted - cpda_recomp
  )
recon_ok <- all(abs(recon$d_cost) <= pmax(TOL, 1) , na.rm = TRUE) &&
            all(abs(recon$d_deaths) <= pmax(TOL, 1), na.rm = TRUE)

# ---- Overall QA disposition --------------------------------------------------
qa <- WB$QA_Checks
qa_status_norm <- toupper(trimws(qa$status))
overall_qa <- if (any(qa_status_norm == "FAIL")) "FAIL" else
              if (any(qa_status_norm == "REVIEW")) "REVIEW" else "PASS"
n_pass   <- sum(qa_status_norm == "PASS")
n_review <- sum(qa_status_norm == "REVIEW")
n_fail   <- sum(qa_status_norm == "FAIL")

idg <- WB$Input_Diagnostic
n_diag_fail   <- sum(toupper(trimws(idg$severity)) == "FAIL")
n_diag_review <- sum(toupper(trimws(idg$severity)) == "REVIEW")

# Economic value availability
ev <- WB$Economic_Value
econ_value_available <- !("note" %in% names(ev) && ncol(ev) == 1)
```

```{r derived-tables, include=FALSE}
si <- WB$Selected_Interventions
cc <- WB$Cost_Components
ac <- WB$Annual_Cost

# n distinct interventions / links
n_interventions <- dplyr::n_distinct(si$intervention_id)
n_links         <- nrow(si)
n_blocked       <- nrow(WB$Blocked_Links)
n_components    <- nrow(cc)
n_scen_costed   <- dplyr::n_distinct(bi$scenario)

# Uniqueness safeguards
uniq_bi <- nrow(bi) == nrow(dplyr::distinct(bi, scenario, year))
uniq_am <- nrow(am) == nrow(dplyr::distinct(am, scenario, year, cause))
uniq_ac <- nrow(ac) == nrow(dplyr::distinct(ac, scenario, year, cost_record_id))

# Negative cases-averted flag (preserved, interpreted — never coerced)
neg_ca <- ce |> filter(cases_averted < 0) |> arrange(cases_averted)
has_neg_ca <- nrow(neg_ca) > 0

# Best / worst value among *individual* interventions (exclude the bundle)
ce_ind <- ce |> filter(scenario != "all", deaths_averted > 0) |>
  arrange(cost_per_death_averted)
best_val  <- ce_ind |> slice(1)
worst_val <- ce_ind |> slice(n())
```

<div class="rhd-hero">
<span class="rhd-kicker">Health financing decision brief</span>
<h2>Costing and interim cost-effectiveness of the FAIR Choices cardiovascular package for Indonesia</h2>
<p>Every figure in this brief is read directly from the cost-value workbook
<strong>`r basename(WB_PATH)`</strong> — the single source of truth. No epidemiological,
costing, or economic pipeline is executed while this report is knitted, and no result is
copied by hand. Values reflect the cached Excel formula results as last recalculated and saved.</p>
</div>

`r if (!focus_available) sprintf('<div class="report-callout callout-warning"><div class="callout-title">Focus scenario substituted</div><div class="callout-body">The requested focus scenario <code>%s</code> is not present in <strong>Cost_Effectiveness</strong>. The highest-impact available scenario <code>%s</code> (%s) is shown instead. Set <code>params$focus_scenario</code> to a valid scenario id to override.</div></div>', params$focus_scenario, focus_id, lbl_scen(focus_id)) else ""`

# At a glance

The headline figures below describe the **`r lbl_scen(focus_id)`** scenario
(`r if (focus_id == "all") "the combined package of all selected interventions implemented jointly" else "a single selected intervention"`),
over the analysis horizon **`r START_YEAR`–`r END_YEAR`** in **`r LOCATION`**,
in **`r PRICE_YEAR` `r CURRENCY`** from a **`r gsub("_", " ", PERSPECTIVE)`** perspective.

```{r kpi-cards, results='asis'}
kpi <- function(label, value, detail, tone = "kpi-blue") {
  glue('<div class="kpi-card {tone}"><div class="kpi-label">{label}</div>',
       '<div class="kpi-value">{value}</div><div class="kpi-detail">{detail}</div></div>')
}
qa_tone <- switch(overall_qa, PASS = "kpi-teal", REVIEW = "kpi-gold", FAIL = "kpi-red")
cards <- paste0(
  '<div class="kpi-grid">',
  kpi("Discounted incremental cost",
      usd_big(focus_row$disc_incremental_cost),
      glue("Undiscounted {usd_big(focus_row$incremental_cost)} · {DISCOUNT*100}% discount rate · costs only"),
      "kpi-navy"),
  kpi("Total deaths averted",
      count_big(focus_row$deaths_averted),
      glue("Cumulative {START_YEAR}–{END_YEAR}, vs baseline ‘{BASE_ID}’"),
      "kpi-teal"),
  kpi("Discounted cost per death averted",
      usd0(focus_row$cost_per_death_averted),
      "Interim efficiency metric — <strong>not</strong> cost per DALY",
      "kpi-blue"),
  kpi("Analysis horizon",
      glue("{START_YEAR}–{END_YEAR}"),
      glue("{N_YEARS} model years · linear coverage scale-up"),
      "kpi-blue"),
  kpi("Price year &amp; currency",
      glue("{PRICE_YEAR} {CURRENCY}"),
      glue("{gsub('_',' ',PERSPECTIVE)} perspective · downstream offsets excluded"),
      "kpi-gold"),
  kpi("Overall QA status",
      sprintf('%s', overall_qa),
      glue("{n_pass} PASS · {n_review} REVIEW · {n_fail} FAIL across {nrow(qa)} checks"),
      qa_tone),
  '</div>'
)
cat(cards)
```

::: {.source-note}
**Traceability.** KPI values are drawn from the **Cost_Effectiveness** sheet
(row `scenario = "`r focus_id`"`), the **Calculation_Assumptions** sheet (horizon, discount
rate, price year, perspective), and the **QA_Checks** sheet (status tally). The combined
`all` scenario is the joint effect of implementing every selected intervention together;
because of overlapping at-risk populations and competing risks it is **not** the arithmetic
sum of the individual scenarios and should not be ranked against them as a mutually
exclusive option.
:::

# Decision summary

```{r decision-inline, include=FALSE}
all_row  <- ce[ce$scenario == "all", , drop = FALSE]
cvd_row  <- ce[ce$scenario == "I_CVD_PRIMARY", , drop = FALSE]
# cumulative undiscounted budget over horizon for the bundle
all_bi_final <- bi_final[bi_final$scenario == "all", ]
```

The combined FAIR Choices cardiovascular package (**`all`**) is projected to avert
**`r fmt_n(all_row$deaths_averted)`** cardiovascular deaths in `r LOCATION` over
`r START_YEAR`–`r END_YEAR`, at a **discounted incremental cost of
`r usd_big(all_row$disc_incremental_cost)`** (undiscounted
`r usd_big(all_row$incremental_cost)`; cumulative undiscounted budget impact
`r usd_big(all_bi_final$cum_undisc)`). The corresponding **discounted cost per death averted
is `r usd0(all_row$cost_per_death_averted)`**.

Among the individual interventions, **`r lbl_scen(best_val$scenario)`** is the most efficient
on this interim metric at **`r usd0(best_val$cost_per_death_averted)` per death averted**
(`r fmt_n(best_val$deaths_averted)` deaths averted for
`r usd_big(best_val$disc_incremental_cost)` discounted), while
**`r lbl_scen(worst_val$scenario)`** is the least efficient at
**`r usd0(worst_val$cost_per_death_averted)` per death averted**. The
`r fmt_n(worst_val$deaths_averted)` deaths averted by the latter come at a discounted cost of
`r usd_big(worst_val$disc_incremental_cost)` — a spread of roughly
`r fmt_n(worst_val$cost_per_death_averted / best_val$cost_per_death_averted)`× in cost per death
averted across the portfolio.

::: {.report-callout .callout-warning}
::: {.callout-title}
Cost per death averted is an interim metric
:::
::: {.callout-body}
The efficiency ratio reported throughout is **discounted cumulative incremental cost divided
by cumulative (undiscounted) deaths averted** — *not* cost per DALY, QALY, or life-year.
DALYs, YLL and YLD are explicitly **out of scope** in this workbook stage (see the
**README** and **Run_Metadata** sheets). Ratios here therefore capture mortality benefit only
and cannot be compared against DALY-based cost-effectiveness thresholds. Several
secondary-prevention scenarios also show **negative cases averted**, discussed under
*Results* and *Interpretation*, because keeping patients alive increases the prevalent-case
stock even as deaths fall.
:::
:::

# Purpose, scope & perspective

This report presents the **costing, budget impact, and interim (mortality-based)
cost-effectiveness** of the FAIR Choices cardiovascular intervention package selected for
`r LOCATION`. All quantitative content derives from the cost-value workbook produced by the
CVD FAIR Choices pipeline (`r meta_val("Model / pipeline")`).

```{r scope-grid, results='asis'}
scope_card <- function(h, p) glue('<div class="overview-card"><h3>{h}</h3><p>{p}</p></div>')
cat(paste0(
  '<div class="overview-grid">',
  scope_card("Geography &amp; horizon",
             glue("{LOCATION}, annual projection {START_YEAR}–{END_YEAR} ({N_YEARS} years).")),
  scope_card("Comparator",
             glue("Baseline scenario ‘{BASE_ID}’ — {tolower(lbl_scen(BASE_ID))}. ",
                  "Every costed scenario is differenced against this comparator.")),
  scope_card("Currency &amp; price year",
             glue("{CURRENCY}, {PRICE_YEAR} price year. Discount rate {DISCOUNT*100}% per year, ",
                  "applied to costs only; death counts are undiscounted.")),
  scope_card("Perspective &amp; boundary",
             glue("{stringr::str_to_sentence(gsub('_',' ',PERSPECTIVE))} perspective. ",
                  "Downstream disease-cost offsets excluded; DALYs/economic value deferred.")),
  '</div>'
))
```

::: {.source-note}
Scope fields are taken verbatim from **Calculation_Assumptions** (`analysis_start_year`,
`analysis_end_year`, `cost_discount_rate`, `cost_price_year`, `currency`,
`economic_perspective`, `baseline_scenario_id`, `downstream_cost_offsets`) and
**Run_Metadata**.
:::

# Interventions & costing inputs

`r n_interventions` interventions (`r n_links` intervention–cause links) are selected and
costed across `r n_components` cost components; `r n_blocked` health link(s) are **blocked**
and excluded from results (see *Assumptions & data quality*).

## Selected interventions and cause links

```{r tbl-interventions}
si_tbl <- si |>
  transmute(
    Intervention = intervention_name,
    Cause = lbl_cause(cause_code),
    Transition = paste0(transition_from, " → ", transition_to, " (", model_transition, ")"),
    `Effect` = round(effect_value, 3),
    `Affected frac.` = round(affected_fraction, 3),
    `Cov. base` = baseline_coverage,
    `Cov. target` = target_coverage,
    `Adj. effect @ target` = round(adjusted_effect_at_target, 4),
    `Cost scope` = cost_scope,
    `Effect review` = effect_review,
    `Coverage review` = coverage_review
  )
datatable(
  si_tbl, rownames = FALSE, extensions = "Buttons",
  caption = htmltools::tags$caption(style = "caption-side:top;",
    htmltools::HTML("<span class='dt-caption'>Selected_Interventions — selected FAIR Choices CVD links, effect parameters, coverage and review status</span>")),
  options = list(pageLength = 8, dom = "Bfrtip", scrollX = TRUE,
                 buttons = c("copy", "csv"),
                 columnDefs = list(list(className = "dt-right", targets = 3:7)))
) |>
  formatPercentage(c("Cov. base", "Cov. target"), 0)
```

## Cost components, populations in need and unit costs

```{r tbl-components}
cc_tbl <- cc |>
  transmute(
    Component = cost_component,
    Intervention = intervention_id,
    `PIN measure` = population_in_need_measure,
    `PIN frac.` = round(population_in_need_fraction, 3),
    `Freq./yr` = frequency_per_year,
    `Unit cost (USD)` = round(unit_cost_usd, 2),
    `Price yr` = price_year,
    `Cov. base` = cov_baseline,
    `Cov. target` = cov_target,
    `ID-adj.` = ifelse(indonesia_adjusted_flag == 1, "yes", "no"),
    Review = cost_review,
    Ready = ifelse(cost_ready == 1, "yes", "no")
  )
datatable(
  cc_tbl, rownames = FALSE, extensions = "Buttons",
  caption = htmltools::tags$caption(style = "caption-side:top;",
    htmltools::HTML("<span class='dt-caption'>Cost_Components — component definitions, population-in-need rules, unit costs and readiness flags</span>")),
  options = list(pageLength = 8, dom = "Bfrtip", scrollX = TRUE,
                 buttons = c("copy", "csv"))
) |>
  formatCurrency("Unit cost (USD)", currency = "$", digits = 2) |>
  formatPercentage(c("Cov. base", "Cov. target"), 0)
```

::: {.source-note}
Component rows are read from **Cost_Components**. `PIN measure` maps to eligible population
(`all`), prevalent stock (`prevalence`) or new cases (`incidence`) per the **README** and
**Methods_and_Sources** (M06). `r sum(cc$cost_review == "Review")` of `r n_components`
components carry a `Review` cost flag and `r sum(cc$cost_ready == 0)` are not yet marked
ready — see the data-quality section.
:::

# Traceability & methodology

The workbook is organised as a **dependency chain**: editable controls on
*Calculation_Assumptions* feed R-generated source quantities, which live Excel formulas
transform into the decision tables. The **Calculation_Map** sheet enumerates every link.

```{r pipeline, results='asis'}
cat(paste0(
  '<div class="pipeline-grid">',
  '<div class="pipeline-step"><span>1</span><strong>Assumptions</strong><p>Horizon, discount rate, price year, perspective (Calculation_Assumptions).</p></div>',
  '<div class="pipeline-step"><span>2</span><strong>Selection</strong><p>Interventions, cause links, effects &amp; coverage (Selected_Interventions).</p></div>',
  '<div class="pipeline-step"><span>3</span><strong>Costing inputs</strong><p>PIN rules, unit costs, coverage paths (Cost_Components).</p></div>',
  '<div class="pipeline-step"><span>4</span><strong>Annual layers</strong><p>Mortality &amp; component costs by year (Annual_Mortality, Annual_Cost).</p></div>',
  '<div class="pipeline-step"><span>5</span><strong>Budget impact</strong><p>Annual &amp; cumulative incremental cost (Budget_Impact).</p></div>',
  '<div class="pipeline-step"><span>6</span><strong>Cost-effectiveness</strong><p>Deaths averted &amp; cost per death (Cost_Effectiveness).</p></div>',
  '<div class="pipeline-step"><span>7</span><strong>QA</strong><p>Invariants &amp; Excel-vs-R reconciliation (QA_Checks).</p></div>',
  '</div>'
))
```

## Core equations

The workbook records the following formulas (**Methods_and_Sources**, M01–M10). They are
reproduced here exactly as documented — no alternative method is introduced.

**Coverage scale-up and FAIR-adjusted effect** (M01–M04):

$$
\begin{aligned}
\Delta\mathrm{cov}(t) &= \mathrm{cov}(t) - \mathrm{cov}_{\text{base}} \\[4pt]
e_{\text{adj}}(t) &= \frac{\text{effect}\;\cdot\;\Delta\mathrm{cov}(t)}
                          {1 - \text{effect}\;\cdot\;\mathrm{cov}_{\text{base}}} \\[4pt]
\tau(t) &= e_{\text{adj}}(t)\;\cdot\;f_{\text{affected}} \\[4pt]
p_{\text{scenario}}(t) &= p_{\text{baseline}}(t)\;\bigl(1 - \tau(t)\bigr)
\end{aligned}
$$

<p class="eq-note">Prevention links modify incidence (IR); management links modify case
fatality (CF).</p>

**Population in need and annual component cost** (M06):

$$
\mathrm{PIN}(t) = Q(t)\;\cdot\;\text{pin\_fraction}, \qquad
C(t) = \mathrm{PIN}(t)\;\cdot\;\mathrm{cov}(t)\;\cdot\;\text{freq}\;\cdot\;\text{unit\_cost}
$$

<p class="eq-note">\(Q(t)\) is the eligible population, prevalent stock, or incident cases
per the PIN measure; shared components are counted once (M07).</p>

**Incremental budget impact and discounting** (M08–M09):

$$
\Delta B(t) = C_{\text{scenario}}(t) - C_{\text{baseline}}(t), \qquad
d(t) = \frac{1}{(1 + r)^{\,t - t_0}}
$$

<p class="eq-note">Costs are discounted at \(r = `r DISCOUNT`\) from \(t_0 = `r START_YEAR`\);
death counts are left undiscounted.</p>

**Deaths averted and interim cost-effectiveness** (M10):

$$
\mathrm{DA}(t) = D_{\text{baseline}}(t) - D_{\text{scenario}}(t), \qquad
\text{USD per death averted} = \frac{\displaystyle\sum_t d(t)\,\Delta B(t)}
                                    {\displaystyle\sum_t \mathrm{DA}(t)}
$$

```{r tbl-calcmap}
cm <- WB$Calculation_Map |>
  transmute(`Output sheet` = output_sheet, `Formula columns` = formula_columns,
            `Depends on` = depends_on, Calculation = calculation,
            `Source role` = source_role)
kbl(cm, caption = "Calculation_Map — dependency and traceability chain across sheets",
    align = "lllll", table.attr = 'class="table compact-kbl"') |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 12)
```

# Assumptions & data-quality status

```{r assumptions-grid, results='asis'}
show_ids <- c("analysis_start_year","analysis_end_year","cost_discount_rate",
              "cost_price_year","currency","economic_perspective","scale_up_shape",
              "baseline_scenario_id","downstream_cost_offsets","formula_tolerance")
ca_show <- ca[match(show_ids, ca$parameter_id), ]
ca_tbl <- ca_show |>
  transmute(Parameter = parameter_id, Value = value, Unit = unit, Role = role,
            Description = description)
cat('<div class="table-panel">')
print(kbl(ca_tbl, caption = "Calculation_Assumptions — key formula controls and metadata",
          align = "lllll", table.attr = 'class="table compact-kbl"') |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12))
cat('</div>')
```

## Blocked links and diagnostic issues

`r n_blocked` health link(s) are blocked and **excluded** from all results; the input
diagnostic records `r n_diag_fail` FAIL-level and `r n_diag_review` REVIEW-level item(s).

```{r tbl-blocked}
bl <- WB$Blocked_Links |>
  transmute(Intervention = intervention_id, Cause = cause_id,
            Transition = paste0(transition_from, " → ", transition_to),
            Effect = effect_value, `Affected frac.` = affected_fraction,
            `Baseline cov.` = baseline_coverage, `Target cov.` = target_coverage,
            Problem = problem)
kbl(bl, caption = "Blocked_Links — health-effect links excluded from the analysis",
    align = "llllllll", table.attr = 'class="table compact-kbl"') |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12)
```

```{r tbl-diagnostic}
idg_tbl <- idg |>
  transmute(Scope = scope, Item = item_key, Field = field, Problem = problem,
            Severity = severity)
kbl(idg_tbl, caption = "Input_Diagnostic — FAIL/REVIEW issues affecting interpretation",
    align = "lllll", escape = TRUE, table.attr = 'class="table compact-kbl"') |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12) |>
  column_spec(5, bold = TRUE)
```

::: {.report-callout .callout-danger}
::: {.callout-title}
What the blocked links mean for scope
:::
::: {.callout-body}
The diabetes-treatment intervention (`I_T2D_TREATMENT`, both its IHD and T2D links) is
**blocked for missing baseline coverage** and is therefore absent from every result table
and figure. Its associated cost record (`COST_T2D_TREATMENT`) additionally lacks a coverage
trajectory and is not Indonesia-adjusted. Consequently the costed portfolio is **CVD-focused**
and does not represent the full candidate intervention set. Type 2 diabetes (`dm2`) still
appears as a modelled cause in mortality, but with **zero deaths averted** because no valid
diabetes intervention link survived selection.
:::
:::

# Results

## Scenario comparison

```{r tbl-ce}
ce_tbl <- ce |>
  mutate(is_bundle = scenario == "all") |>
  arrange(desc(is_bundle), cost_per_death_averted) |>
  transmute(
    Scenario = scenario_label,
    `Deaths averted` = deaths_averted,
    `Cases averted` = cases_averted,
    `Incremental cost` = incremental_cost,
    `Disc. incremental cost` = disc_incremental_cost,
    `Disc. cost / death averted` = cost_per_death_averted,
    Basis = dominance,
    Reconciliation = reconciliation_status
  )
datatable(
  ce_tbl, rownames = FALSE, extensions = "Buttons",
  caption = htmltools::tags$caption(style = "caption-side:top;",
    htmltools::HTML("<span class='dt-caption'>Cost_Effectiveness — cumulative deaths/cases averted, incremental cost and interim cost per death averted</span>")),
  options = list(pageLength = 9, dom = "Bfrtip", scrollX = TRUE,
                 buttons = c("copy", "csv"), ordering = FALSE)
) |>
  formatRound(c("Deaths averted", "Cases averted"), 0) |>
  formatCurrency(c("Incremental cost", "Disc. incremental cost"), "$", digits = 0) |>
  formatCurrency("Disc. cost / death averted", "$", digits = 0)
```

::: {.source-note}
The first row is the combined **`all`** bundle (shaded conceptually as a portfolio, not a
mutually exclusive option); the remaining rows are individual interventions ordered from most
to least efficient. Negative *Cases averted* are retained exactly as reported by the workbook.
:::

## Budget impact

```{r fig-budget-annual, fig.height=5.4}
bi_plot <- bi |>
  filter(year <= HORIZON) |>
  mutate(scenario = factor(scenario, levels = scen_levels))
```

::: {.chart-panel}
```{r fig-budget-annual-draw, fig.height=5.4}
ggplot(bi_plot, aes(year, incremental_cost / 1e9, colour = scenario)) +
  geom_hline(yintercept = 0, colour = "#c9d6df", linewidth = 0.4) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.1, alpha = 0.7) +
  scale_colour_manual(values = scen_cols, labels = function(x) lbl_scen(x),
                      name = NULL) +
  scale_x_continuous(breaks = scales::pretty_breaks(6)) +
  scale_y_continuous(labels = function(x) paste0("$", x, "B")) +
  labs(
    title = "Annual incremental budget impact by scenario",
    subtitle = glue("Undiscounted scenario minus baseline cost, {START_YEAR}–{HORIZON}"),
    x = NULL, y = glue("Incremental cost ({CURRENCY} billions, undiscounted)"),
    caption = src_cap("Budget_Impact")
  ) +
  theme_report()
```
:::

::: {.chart-panel}
```{r fig-budget-cumulative, fig.height=5.4}
bi_cum <- bi |>
  filter(year <= HORIZON) |>
  mutate(scenario = factor(scenario, levels = scen_levels))
ggplot(bi_cum, aes(year, cumulative_disc_incremental_cost / 1e9, colour = scenario)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.1, alpha = 0.7) +
  scale_colour_manual(values = scen_cols, labels = function(x) lbl_scen(x), name = NULL) +
  scale_x_continuous(breaks = scales::pretty_breaks(6)) +
  scale_y_continuous(labels = function(x) paste0("$", x, "B")) +
  labs(
    title = "Cumulative discounted incremental cost by scenario",
    subtitle = glue("Running total of discounted incremental cost to {HORIZON} ({DISCOUNT*100}% discount rate)"),
    x = NULL, y = glue("Cumulative discounted cost ({CURRENCY} billions)"),
    caption = src_cap("Budget_Impact")
  ) +
  theme_report()
```
:::

The combined package reaches a cumulative discounted incremental cost of
**`r usd_big(all_bi_final$cum_disc)`** by `r END_YEAR`. `r lbl_scen("I_RHD_SURGERY")`
is the single largest cost driver among the individual interventions
(`r usd_big(bi_final$cum_disc[bi_final$scenario=="I_RHD_SURGERY"])` discounted).

## Mortality and cases

::: {.chart-panel}
```{r fig-deaths-averted, fig.height=5.4}
da_time <- am |>
  filter(scenario != BASE_ID, year <= HORIZON) |>
  group_by(scenario, year) |>
  summarise(deaths_averted = sum(deaths_averted), .groups = "drop") |>
  mutate(scenario = factor(scenario, levels = scen_levels))
ggplot(da_time, aes(year, deaths_averted / 1e3, colour = scenario)) +
  geom_line(linewidth = 0.9) +
  scale_colour_manual(values = scen_cols, labels = function(x) lbl_scen(x), name = NULL) +
  scale_x_continuous(breaks = scales::pretty_breaks(6)) +
  scale_y_continuous(labels = function(x) paste0(x, "k")) +
  labs(
    title = "Annual deaths averted by scenario",
    subtitle = glue("Baseline minus scenario deaths, all causes, {START_YEAR}–{HORIZON}"),
    x = NULL, y = "Deaths averted per year (thousands)",
    caption = src_cap("Annual_Mortality")
  ) +
  theme_report()
```
:::

::: {.chart-panel}
```{r fig-deaths-cause, fig.height=5.4}
da_cause <- am |>
  filter(scenario == focus_id, year <= HORIZON) |>
  group_by(year, cause) |>
  summarise(deaths_averted = sum(deaths_averted), .groups = "drop") |>
  mutate(cause = factor(cause, levels = names(cause_cols)))
# keep only causes with any non-zero contribution for a legible legend
keep_causes <- da_cause |> group_by(cause) |>
  summarise(tot = sum(abs(deaths_averted)), .groups = "drop") |>
  filter(tot > 0) |> pull(cause)
da_cause <- da_cause |> filter(cause %in% keep_causes)
ggplot(da_cause, aes(year, deaths_averted / 1e3, fill = cause)) +
  geom_area(alpha = 0.9, colour = "white", linewidth = 0.2) +
  scale_fill_manual(values = cause_cols, labels = function(x) lbl_cause(as.character(x)),
                    name = "Cause") +
  scale_x_continuous(breaks = scales::pretty_breaks(6)) +
  scale_y_continuous(labels = function(x) paste0(x, "k")) +
  labs(
    title = glue("Deaths averted by cause — {lbl_scen(focus_id)}"),
    subtitle = glue("Cause decomposition of annual deaths averted, {START_YEAR}–{HORIZON}"),
    x = NULL, y = "Deaths averted per year (thousands)",
    caption = src_cap("Annual_Mortality")
  ) +
  theme_report()
```
:::

```{r neg-ca-callout, results='asis'}
if (has_neg_ca) {
  rows <- paste(sprintf("%s (%s)", lbl_scen(neg_ca$scenario), fmt_n(neg_ca$cases_averted)),
                collapse = "; ")
  cat(glue(
    '<div class="report-callout callout-warning"><div class="callout-title">',
    'Negative cases averted are preserved, not zeroed</div><div class="callout-body">',
    '{nrow(neg_ca)} scenario(s) report negative cumulative cases averted: {rows}. ',
    'This is expected for management / secondary-prevention interventions: reducing case ',
    'fatality keeps more people alive in the prevalent-disease (‘sick’) state, so the ',
    'stock of living cases rises even as deaths fall. These values are retained exactly ',
    'as the workbook reports them and are interpreted, never coerced to zero.',
    '</div></div>'))
}
```

## Cost drivers

::: {.chart-panel}
```{r fig-cost-drivers, fig.height=6.2}
drivers <- ac |>
  filter(scenario != "all", year <= HORIZON) |>
  group_by(intervention_id, cost_component_key) |>
  summarise(disc_inc = sum(disc_cost_incremental), .groups = "drop") |>
  filter(disc_inc != 0) |>
  mutate(intervention_id = factor(intervention_id, levels = scen_levels),
         comp = str_replace_all(cost_component_key, "_", " "),
         comp = str_wrap(str_to_title(comp), 26)) |>
  arrange(desc(disc_inc)) |>
  slice_head(n = 18)
ggplot(drivers, aes(x = reorder(comp, disc_inc), y = disc_inc / 1e9,
                    fill = intervention_id)) +
  geom_col(width = 0.72) +
  coord_flip() +
  scale_fill_manual(values = scen_cols, labels = function(x) lbl_scen(x),
                    name = "Intervention") +
  scale_y_continuous(labels = function(x) paste0("$", x, "B")) +
  labs(
    title = "Top cost components by discounted incremental cost",
    subtitle = glue("Cumulative {START_YEAR}–{HORIZON}; individual scenarios only (bundle excluded to avoid double counting)"),
    x = NULL, y = glue("Discounted incremental cost ({CURRENCY} billions)"),
    caption = src_cap("Annual_Cost")
  ) +
  theme_report() +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(colour = "#eef2f5"))
```
:::

::: {.source-note}
Components flagged `shared-count-once` are aggregated exactly as they appear in **Annual_Cost**
(one row per scenario-year); the sheet's own `shared_duplicate_count` control prevents any
double counting across causes. The combined `all` scenario is excluded from this chart so that
shared components are not mixed across the portfolio.
:::

## Coverage trajectories

::: {.chart-panel}
```{r fig-coverage, fig.height=5.6}
cov_traj <- ac |>
  filter(scenario != "all", year <= HORIZON) |>
  distinct(intervention_id, cost_component_key, year, coverage_scenario) |>
  mutate(intervention_id = factor(intervention_id, levels = scen_levels))
ggplot(cov_traj, aes(year, coverage_scenario,
                     group = cost_component_key, colour = intervention_id)) +
  geom_line(linewidth = 0.7, alpha = 0.85) +
  scale_colour_manual(values = scen_cols, labels = function(x) lbl_scen(x), name = NULL) +
  scale_x_continuous(breaks = scales::pretty_breaks(6)) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(
    title = "Coverage trajectories for costed components",
    subtitle = glue("Scenario coverage path per component, {SCALE_SHAPE} scale-up {START_YEAR}–{HORIZON}"),
    x = NULL, y = "Coverage (scenario)",
    caption = src_cap("Annual_Cost")
  ) +
  theme_report()
```
:::

## Cost-effectiveness plane and efficiency ranking

::: {.chart-panel}
```{r fig-ce-plane, fig.height=5.8}
ce_pl <- ce |>
  mutate(scenario = factor(scenario, levels = scen_levels),
         is_bundle = scenario == "all")
ggplot(ce_pl, aes(deaths_averted / 1e6, disc_incremental_cost / 1e9)) +
  geom_point(aes(colour = scenario, shape = is_bundle), size = 3.4) +
  geom_text(aes(label = wrap_scen(as.character(scenario), 16)),
            size = 3, vjust = -0.7, hjust = 0.5, colour = "#29465b",
            lineheight = 0.85, check_overlap = FALSE) +
  scale_colour_manual(values = scen_cols, guide = "none") +
  scale_shape_manual(values = c(`FALSE` = 16, `TRUE` = 17), guide = "none") +
  scale_x_continuous(labels = function(x) paste0(x, "M"),
                     expand = expansion(mult = c(0.08, 0.14))) +
  scale_y_continuous(labels = function(x) paste0("$", x, "B"),
                     expand = expansion(mult = c(0.06, 0.12))) +
  labs(
    title = "Cost-effectiveness plane",
    subtitle = "Discounted incremental cost vs cumulative deaths averted (▲ = combined bundle)",
    x = "Deaths averted (millions, cumulative)",
    y = glue("Discounted incremental cost ({CURRENCY} billions)"),
    caption = src_cap("Cost_Effectiveness")
  ) +
  theme_report()
```
:::

::: {.chart-panel}
```{r fig-cpda-rank, fig.height=5.4}
cpda <- ce |>
  filter(deaths_averted > 0) |>
  mutate(scenario = factor(scenario, levels = scen_levels),
         is_bundle = scenario == "all")
rng <- range(cpda$cost_per_death_averted)
use_log <- (rng[2] / rng[1]) > 50   # decades of range -> log scale for legibility
p <- ggplot(cpda, aes(x = reorder(as.character(scenario), -cost_per_death_averted),
                      y = cost_per_death_averted, fill = scenario)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = usd0(cost_per_death_averted)), hjust = -0.08, size = 3,
            colour = "#29465b") +
  coord_flip() +
  scale_fill_manual(values = scen_cols, guide = "none") +
  scale_x_discrete(labels = function(x) wrap_scen(x, 20)) +
  labs(
    title = "Interim efficiency ranking — discounted cost per death averted",
    subtitle = if (use_log) "Note: horizontal axis is log-scaled (values span several orders of magnitude)"
               else "Discounted cost per death averted, ranked",
    x = NULL, y = glue("Discounted cost per death averted ({CURRENCY}){if (use_log) ', log scale' else ''}"),
    caption = src_cap("Cost_Effectiveness")
  ) +
  theme_report() +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(colour = "#eef2f5"))
if (use_log) {
  p + scale_y_log10(labels = scales::dollar_format(),
                    expand = expansion(mult = c(0, 0.18)))
} else {
  p + scale_y_continuous(labels = scales::dollar_format(),
                         expand = expansion(mult = c(0, 0.18)))
}
```
:::

::: {.source-note}
The efficiency ranking excludes the combined `all` bundle from the mutually-exclusive
comparison logic; it is a portfolio, not a competing option. `r if (exists("use_log") && use_log) "A log scale is used because the workbook's cost-per-death-averted values span several orders of magnitude (from hundreds to millions of USD)." else "A linear scale is used."`
:::

# Technical validation

The workbook recomputes each invariant in Excel and reconciles its headline
cost-effectiveness figures against the R engine values stored on **Calculation_Assumptions**.
Overall disposition: **`r status_badge(overall_qa)`** —
`r n_pass` PASS, `r n_review` REVIEW, `r n_fail` FAIL.

```{r tbl-qa}
qa_tbl <- qa |>
  transmute(Check = check, Expected = expected, Actual = actual,
            Status = status, Note = note)
kbl(qa_tbl, caption = "QA_Checks — invariant recomputation and Excel-vs-R reconciliation",
    align = "lllll", table.attr = 'class="table compact-kbl"') |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12) |>
  column_spec(4, bold = TRUE) |>
  row_spec(which(qa_status_norm == "PASS"),   background = "#f4fbf9") |>
  row_spec(which(qa_status_norm == "REVIEW"), background = "#fffbf1") |>
  row_spec(which(qa_status_norm == "FAIL"),   background = "#fff7f8")
```

## Independent reconciliation (report-side QA)

As an additional check, this report re-derives the cross-sheet totals and compares them with
the workbook's authoritative values. **The workbook value is never overwritten** — any
mismatch is reported, not silently corrected.

```{r tbl-recon}
recon_tbl <- recon |>
  transmute(
    Scenario = lbl_scen(scenario),
    `WB disc. cost` = disc_incremental_cost,
    `BI cumulative` = cum_disc,
    `Δ cost` = d_cost,
    `WB deaths averted` = deaths_averted,
    `AM Σ deaths` = da,
    `Δ deaths` = d_deaths,
    `WB $/death` = cost_per_death_averted,
    `Recomputed $/death` = cpda_recomp
  )
kbl(recon_tbl, caption = "Report-side reconciliation: Cost_Effectiveness vs Budget_Impact and Annual_Mortality",
    align = "lrrrrrrrr", digits = c(0,0,0,4,2,2,6,2,2),
    format.args = list(big.mark = ","), table.attr = 'class="table compact-kbl"') |>
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 11)
```

```{r recon-callout, results='asis'}
tone <- if (recon_ok) "callout-success" else "callout-danger"
title <- if (recon_ok) "Cross-sheet totals reconcile" else "Cross-sheet reconciliation mismatch"
body <- if (recon_ok) glue(
  "Every scenario's discounted incremental cost matches the Budget_Impact cumulative total, ",
  "and every scenario's deaths averted matches the Annual_Mortality sum, within the workbook ",
  "tolerance ({TOL} {CURRENCY}/count). The independently recomputed cost per death averted ",
  "equals the workbook's cost_per_death_averted for all scenarios.") else glue(
  "One or more scenarios differ between Cost_Effectiveness and the annual detail sheets beyond ",
  "the workbook tolerance ({TOL}). The workbook values are retained as authoritative and the ",
  "discrepancy is flagged for investigation — recalculate and save the workbook, then re-knit.")
cat(glue('<div class="report-callout {tone}"><div class="callout-title">{title}</div>',
         '<div class="callout-body">{body}</div></div>'))
```

## Structural integrity checks

```{r integrity, results='asis'}
# Non-finite scan over the numeric fields used in results
nf <- function(tbl, cols) sum(!is.finite(as.matrix(sapply(tbl[cols], as.numeric))))
nf_bi <- nf(bi, c("incremental_cost","disc_incremental_cost","cumulative_disc_incremental_cost"))
nf_am <- nf(am, c("cause_deaths","base_deaths","deaths_averted","cases_averted"))
nf_ac <- nf(ac, c("annual_cost_baseline","annual_cost_scenario","annual_cost_incremental",
                  "disc_cost_incremental","coverage_scenario"))
neg_cost <- sum(ac$annual_cost_baseline < 0, na.rm = TRUE) +
            sum(ac$annual_cost_scenario < 0, na.rm = TRUE)
scen_sets_match <- setequal(unique(bi$scenario), setdiff(unique(am$scenario), BASE_ID)) &&
                   setequal(unique(bi$scenario), unique(ac$scenario))

check_row <- function(label, ok, detail) {
  b <- if (ok) '<span class="badge badge-pass">PASS</span>' else '<span class="badge badge-review">REVIEW</span>'
  glue('<div class="repro-card"><span>{if (ok) "✓" else "!"}</span><h3>{label} {b}</h3><p>{detail}</p></div>')
}
cat(paste0(
  '<div class="repro-grid">',
  check_row("Budget_Impact keys", uniq_bi,
            glue("Scenario × year rows are unique ({nrow(bi)} rows, {n_scen_costed} scenarios × {N_YEARS} years).")),
  check_row("Annual_Mortality keys", uniq_am,
            glue("Scenario × year × cause rows are unique ({nrow(am)} rows) — safe to aggregate to year × cause.")),
  check_row("Annual_Cost keys", uniq_ac,
            glue("Scenario × year × cost record rows are unique ({nrow(ac)} rows) — no component double-counted.")),
  check_row("Finite values", nf_bi + nf_am + nf_ac == 0,
            glue("Non-finite entries in results fields: Budget_Impact {nf_bi}, Annual_Mortality {nf_am}, Annual_Cost {nf_ac}.")),
  check_row("Non-negative costs", neg_cost == 0,
            glue("{neg_cost} negative baseline/scenario annual cost value(s) — costs are expected to be non-negative.")),
  check_row("Consistent scenario set", scen_sets_match,
            "The same scenarios appear across Budget_Impact, Annual_Mortality (excl. baseline) and Annual_Cost."),
  '</div>'
))
```

::: {.source-note}
These are internal-consistency checks the report performs on the workbook tables; they
confirm the sheets can be joined and aggregated safely. Zero/negative deaths averted are
handled explicitly where ratios are computed (the efficiency ranking and recomputed cost per
death averted exclude scenarios with non-positive deaths averted), and negative *cases*
averted are preserved as described under *Results*.
:::

::: {.report-callout .callout-warning}
::: {.callout-title}
Arithmetic PASS ≠ external validity
:::
::: {.callout-body}
`r n_review` QA check(s) are **REVIEW** and the input diagnostic carries `r n_diag_fail`
**FAIL** item(s); the model is therefore **not** described as fully validated. Passing
arithmetic and reconciliation checks confirms internal consistency of the costing and
budget-impact calculations — it does **not** establish epidemiological or external validity
of the underlying effect sizes, coverage assumptions, or unit costs, several of which remain
`Provisional` or `Review` in the source sheets.
:::
:::

# Interpretation & limitations

**Model results, not forecasts.** The figures are projections under a specified set of
interventions, effect sizes, coverage trajectories and unit costs — they describe *what the
model implies if these inputs hold*, not a prediction of realised outcomes. Effect and
coverage parameters are largely flagged **`Provisional`** (`r sum(si$effect_review == "Provisional")`
of `r n_links` links) and most unit costs carry a **`Review`** flag.

```{r limitations-grid, results='asis'}
lim_card <- function(h, p) glue('<div class="limitation-card"><h3>{h}</h3><p>{p}</p></div>')
cat(paste0(
  '<div class="limitations-grid">',
  lim_card("No DALYs / economic value",
           glue("DALYs, YLL, YLD and disability weights are out of scope; the Economic_Value ",
                "sheet is unpopulated. Cost per death averted is the only efficiency metric and ",
                "is not comparable to DALY-based thresholds.")),
  lim_card("Blocked &amp; excluded links",
           glue("{n_blocked} diabetes-treatment link(s) are blocked for missing baseline ",
                "coverage, narrowing the portfolio to CVD interventions.")),
  lim_card("Provisional inputs",
           glue("Effect sizes and coverage are mostly Provisional; unit costs are mostly ",
                "flagged Review. One cost record is not Indonesia-adjusted and uses a ",
                "different price year.")),
  lim_card("Negative cases averted",
           glue("Secondary-prevention scenarios raise the prevalent-case stock (survival ",
                "effect); negative cases averted are retained and must be read as living ",
                "cases, not modelling error.")),
  '</div>'
))
```

::: {.report-callout .callout-info}
::: {.callout-title}
Economic value (VSL/VSLY) unavailable
:::
::: {.callout-body}
`r if (!econ_value_available) paste0("The Economic_Value sheet contains only an omission note: “", WB$Economic_Value$note[1], "” No monetised economic value, benefit–cost ratio, or net benefit is reported, and none is invented here.") else "The Economic_Value sheet is populated; see the workbook for VSL/VSLY-based results."`
The **`downstream_cost_offsets`** control is set to `r assume("downstream_cost_offsets")`
(excluded), so averted downstream disease costs are **not** netted against intervention costs.
No productivity effects, fiscal effects, GDP comparisons, cost-utility thresholds, or
probabilistic uncertainty are included because none are present in this workbook.
:::
:::

# Reproducibility

```{r repro-grid, results='asis'}
repro_card <- function(n, h, p) glue('<div class="repro-card"><span>{n}</span><h3>{h}</h3><p>{p}</p></div>')
sheets_read <- paste(REQUIRED_SHEETS, collapse = ", ")
cat(paste0(
  '<div class="repro-grid">',
  repro_card("1", "Single data source",
             glue("Workbook: <code>{WB_REL}</code>. All 15 sheets read once: {sheets_read}. ",
                  "No CSV, RDS, database, model object or other spreadsheet is read.")),
  repro_card("2", "No upstream execution",
             "The epidemiological, costing and economic pipelines are not sourced or re-run during knitting. Cached Excel formula values are used as-is."),
  repro_card("3", "Parameters",
             glue("focus_scenario = <code>{params$focus_scenario}</code>; ",
                  "decision_horizon = <code>{HORIZON}</code>",
                  "{if (HORIZON_DEFAULTED) ' (default: workbook analysis end year)' else ''}; ",
                  "show_code = <code>{params$show_code}</code>.")),
  repro_card("4", "Rendered",
             glue("Knitted {format(Sys.time(), '%Y-%m-%d %H:%M')} · R {getRversion()} · ",
                  "workbook run date {meta_val('Run date')}.")),
  '</div>'
))
```

```{r session-info}
si_txt <- capture.output(sessionInfo())
cat(paste(si_txt, collapse = "\n"))
```

<div class="report-footer">
<div><strong>Indonesia NCD FAIR Choices — Costing &amp; Economic Analysis.</strong>
Source of truth: <code>`r basename(WB_PATH)`</code>. This report executes no upstream model code.</div>
<div>University of Washington · World Bank Indonesia NCD program</div>
</div>

<script>
/* Robustness guard: on some load orderings htmlwidgets marks a DataTable
   "static-bound" before its render completes, leaving an empty widget div.
   Re-render only the widgets that ended up without a DataTables wrapper. */
(function () {
  function fixWidgets() {
    if (!window.HTMLWidgets || !HTMLWidgets.staticRender) return;
    var widgets = document.querySelectorAll('.html-widget');
    var repaired = false;
    widgets.forEach(function (w) {
      if (w.classList.contains('datatables') && !w.querySelector('.dataTables_wrapper')) {
        w.classList.remove('html-widget-static-bound');
        w.innerHTML = '';
        repaired = true;
      }
    });
    if (repaired) { try { HTMLWidgets.staticRender(); } catch (e) {} }
  }
  if (document.readyState === 'complete') { setTimeout(fixWidgets, 200); }
  else { window.addEventListener('load', function () { setTimeout(fixWidgets, 200); }); }
})();
</script>
