---
title: "Projected Health and Economic Outcomes of Cardiovascular Risk-Factor Control Interventions in Indonesia: A Modelling Study"
date: "`r Sys.Date()`"
author: "University of Washington"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE, echo=FALSE}
rm(list = ls())
knitr::opts_chunk$set(
  echo    = TRUE,
  error   = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.height = 5,
  fig.width  = 9,
  fig.align  = "center"
)
```

```{r libraries, message=FALSE, warning=FALSE, echo=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(kableExtra)
library(DT)
library(scales)
library(openxlsx)
```

```{r paths, message=FALSE, warning=FALSE, echo=FALSE}
wd       <- "C:/Users/wrgar/OneDrive - UW/02Work/WorldBank-Indonesia/uw-wb-indonesia-ncd/"
wd_data  <- paste0(wd, "data/processed/")
wd_outp  <- paste0(wd, "output/")
wd_scen  <- paste0(wd, "scenarios/")
wd_raw <- paste0(wd,"data/raw/")
```

```{r helpers, message=FALSE, warning=FALSE, echo=FALSE}
# ---- Formatters --------------------------------------------------------------
fmt_ci <- function(est, lo, hi, digits = 1) {
  f <- function(x) formatC(round(x / 1e6, digits), format = "f",
                            digits = digits, big.mark = ",")
  paste0(f(est), " (", f(lo), "–", f(hi), ")")
}
fmt_m <- function(x, digits = 1)
  formatC(round(x / 1e6, digits), format = "f", digits = digits, big.mark = ",")
fmt_k <- function(x, digits = 0)
  formatC(round(x / 1e3, digits), format = "f", digits = digits, big.mark = ",")
fmt_M <- function(x, digits = 1)   # alias used by slide-deck section
  formatC(round(x / 1e6, digits), format = "f", digits = digits, big.mark = ",")

# ---- Cause labels ------------------------------------------------------------
cause_map <- c(
  ihd     = "Ischemic heart disease",
  istroke = "Ischemic stroke",
  hstroke = "Intracerebral hemorrhage",
  hhd     = "Hypertensive heart disease"
)

# ---- Intervention display labels --------------------------------------------
# Scenarios from 06_run_scenarios.R (parallel execution block):
#   baseline      -> b.a.u
#   bp_only       -> Antihypertensive Therapy
#   sodium_only   -> Sodium Reduction        (salteff = 0 in current run)
#   tfa_only      -> Trans-Fat Elimination
#   statins_only  -> Statin Therapy
#   bp_sodium     -> HTN + Sodium
#   bp_sodium_tfa -> HTN + Sodium + TFA
#   all_four      -> All Interventions

int_labels <- c(
  "b.a.u"                    = "Business-as-Usual",
  "Antihypertensive Therapy" = "Antihypertensive Therapy",
  "Sodium Reduction"         = "Sodium Reduction",
  "Trans-Fat Elimination"    = "Trans-Fat Elimination",
  "Statin Therapy"           = "Statin Therapy",
  "HTN + Sodium"             = "HTN + Sodium",
  "HTN + Sodium + TFA"       = "HTN + Sodium + TFA",
  "All Interventions"        = "All Interventions"
)

int_order <- c(
  "Antihypertensive Therapy",
  "Sodium Reduction",
  "Trans-Fat Elimination",
  "Statin Therapy",
  "All Interventions"
)

int_order_individual <- c(
  "Antihypertensive Therapy",
  "Sodium Reduction",
  "Trans-Fat Elimination",
  "Statin Therapy"
)

# ---- Colour palettes ---------------------------------------------------------
int_cols <- c(
  "Antihypertensive Therapy" = "#1f78b4",
  "Sodium Reduction"         = "#33a02c",
  "Trans-Fat Elimination"    = "#ff7f00",
  "Statin Therapy"           = "#e31a1c",
  "HTN + Sodium"             = "#a6cee3",
  "HTN + Sodium + TFA"       = "#b2df8a",
  "All Interventions"        = "#6a3d9a",
  "Business-as-Usual"        = "#999999"
)

cause_cols <- c(
  ihd     = "#56B4E9",
  istroke = "#CC79A7",
  hstroke = "#F0E442",
  hhd     = "#E69F00"
)

cause_labels_plot <- c(
  ihd     = "Ischemic heart disease",
  istroke = "Ischemic stroke",
  hstroke = "Intracerebral hemorrhage",
  hhd     = "Hypertensive heart disease"
)

# ---- WHO standard population (ages 20-95+) for ASMR -------------------------
full_who_std <- data.table(
  age    = 0:100,
  weight = c(
    rep(0.013, 10), rep(0.014, 10), rep(0.013, 10), rep(0.012, 10),
    rep(0.011, 10), rep(0.009, 10), rep(0.007, 10), rep(0.005, 10),
    rep(0.003, 10), rep(0.001, 10), 0.001
  )
)
who_std <- full_who_std[age >= 20 & age <= 94]
who_std <- rbind(
  who_std,
  data.table(age = 95, weight = full_who_std[age >= 95, sum(weight)])
)
who_std[, weight := weight / sum(weight)]

calculate_asmr <- function(dt, std_pop = who_std,
                            deaths_col = "deaths_dx",
                            population_col = "Nx",
                            disagg_var = "intervention") {
  dt <- copy(dt)
  dt[, age := as.integer(gsub("\\+", "", age))]
  dt <- merge(dt, std_pop, by = "age", all.x = TRUE)
  dt[, rate          := get(deaths_col) / get(population_col)]
  dt[, weighted_rate := rate * weight]
  group_vars <- c("year", disagg_var)
  dt[, .(ASMR = sum(weighted_rate, na.rm = TRUE) * 1e5), by = group_vars]
}

# ---- Deaths-delayed calculator -----------------------------------------------
deaths_summary <- function(dt,
                           baseline_label = "b.a.u",
                           by_cols = c("age_group", "sex", "year",
                                       "location", "cause", "age"),
                           value_col = "dead") {
  stopifnot(is.data.table(dt))
  n_interventions <- uniqueN(dt[intervention != baseline_label, intervention])
  baseline_cols   <- setdiff(by_cols, "intervention")
  baseline   <- dt[intervention == baseline_label,
                   .(baseline_deaths = sum(get(value_col), na.rm = TRUE)),
                   by = baseline_cols]
  nonbaseline <- dt[intervention != baseline_label,
                    .(deaths_projected = sum(get(value_col), na.rm = TRUE)),
                    by = by_cols]
  if (!("intervention" %in% by_cols))
    nonbaseline[, deaths_projected := deaths_projected / n_interventions]
  merged <- merge(nonbaseline, baseline, by = baseline_cols,
                  all.x = TRUE, allow.cartesian = TRUE)
  merged[, deaths_delayed     := baseline_deaths - deaths_projected]
  merged[, deaths_delayed_ave := deaths_delayed / 25]
  merged[, deaths_delayed_pct := deaths_delayed / baseline_deaths]
  merged[]
}
```

```{r country-metadata, message=FALSE, warning=FALSE, echo=FALSE}
countries_dt <- fread(paste0(wd_data, "Country_groupings_extended.csv"))

# Import WHO regions and update countries dt region column
who_regions <- fread(paste0(wd_raw, "who-regions.csv"))

setnames(who_regions, old = c("Entity","Code", "World regions according to WHO"),
                      new = c("location","iso3", "region_who"))

who_regions[, region_who := gsub("\\s*\\(WHO\\)", "", region_who)]

who_regions[,c("Year","location"):=NULL]

countries_dt <- merge(countries_dt, who_regions, by = "iso3", all.x = TRUE)
# ---- Fill missing WHO regions for specific ISO3s (after merge) --------------
who_fix <- data.table(
  iso3 = c("BMU","GRL","PSE","SDS","TWN","VCT","VIR"),
  region_who_fix = c("Americas",     # Bermuda
                     "Europe",       # Greenland
                     "Eastern Mediterranean", # Palestine / West Bank and Gaza
                     "Africa",       # South Sudan
                     "Western Pacific", # Taiwan
                     "Americas",     # St. Vincent and the Grenadines
                     "Americas")     # Virgin Islands (U.S.)
)

countries_dt <- merge(countries_dt, who_fix, by = "iso3", all.x = TRUE)

# Prefer existing region_who; otherwise use the fix
countries_dt[, region_who := fcoalesce(region_who, region_who_fix)]
countries_dt[, region_who_fix := NULL]

countries_dt[,region := fifelse(!is.na(region_who), region_who, region)]

# remove who region
countries_dt[, region_who := NULL]

# Import world bank income groups 2026
income_groups <- fread(paste0(wd_raw, "World Bank Income Group 2026.csv"))

# if missing income category assign to Low income L
income_groups[location_income=="", location_income := "L"]

# Encode L,H, UMC, HIC to match GBD income groupings
income_groups[, location_income := fcase(
  location_income == "L", "Low income",
  location_income == "LM", "Lower-middle income",
  location_income == "UM", "Upper-middle income",
  location_income == "H", "High income",
  default = location_income
)]
# merge
countries_dt <- merge(countries_dt, income_groups[, .(iso3, location_income)],
                    by = "iso3", all.x = TRUE)

#unique by iso
countries_dt <- unique(countries_dt, by = "location")
```

```{r load-model-output, message=FALSE, warning=FALSE, echo=FALSE}
# ---- Load all country-level model output RDS files --------------------------
# Generated by 06_run_scenarios.R (parallel execution for Indonesia)
out_model_path <- paste0(wd_outp, "out_model/")
files <- list.files(path = out_model_path, pattern = "\\.rds$", full.names = TRUE)

if (length(files) == 0)
  stop(paste0("No .rds files found in: ", out_model_path,
              "\nRun 06_run_scenarios.R first."))

data.out <- rbindlist(
  lapply(files, function(f) { dt <- readRDS(f); setDT(dt); dt }),
  use.names = TRUE, fill = TRUE
)
data.out <- data.out[!is.na(location)]

# Drop effect-ratio tracking columns if present
if (all(c("eff_ir", "eff_cf") %in% names(data.out)))
  data.out[, c("eff_ir", "eff_cf") := NULL]

# ---- Age groups --------------------------------------------------------------
data.out[, age_group := cut(
  age,
  breaks         = c(0, 70, 80, 90, Inf),
  labels         = c("<70", "70-79", "80-89", "90+"),
  right          = FALSE,
  include.lowest = TRUE
)]

# ---- Standardise scenario -> intervention labels ----------------------------
# 06_run_scenarios.R writes "scenario" as the idcol from rbindlist()
if ("scenario" %in% names(data.out) && !"intervention" %in% names(data.out)) {
  setnames(data.out, "scenario", "intervention")
} else if ("scenario" %in% names(data.out)) {
  data.out[, intervention := scenario]
  data.out[, scenario := NULL]
}

# Map scenario codes to display labels
data.out[intervention %in% c("baseline", "BAU", "b.a.u"),
         intervention := "b.a.u"]
data.out[intervention %in% c("bp_only", "antihypertensive",
                              "BP", "Antihypertensive therapy"),
         intervention := "Antihypertensive Therapy"]
data.out[intervention %in% c("sodium_only", "sodium", "Salt"),
         intervention := "Sodium Reduction"]
data.out[intervention %in% c("tfa_only", "tfa", "TFA"),
         intervention := "Trans-Fat Elimination"]
data.out[intervention %in% c("statins_only", "statins", "Statins"),
         intervention := "Statin Therapy"]
data.out[intervention %in% c("bp_sodium", "BP + Salt"),
         intervention := "HTN + Sodium"]
data.out[intervention %in% c("bp_sodium_tfa", "BP + Salt + TFA"),
         intervention := "HTN + Sodium + TFA"]
data.out[intervention %in% c("all_four", "all_interventions",
                              "BP + Salt + TFA + Statins"),
         intervention := "All Interventions"]

# Keep only recognised scenarios
data.out <- data.out[intervention %in% c("b.a.u", int_order)]

# ---- Merge region metadata ---------------------------------------------------
if (nrow(countries_dt) > 0) {
  data.out <- merge(data.out, countries_dt[, .(location, region)],
                    by = "location", all.x = TRUE)
} else {
  data.out[, region := NA_character_]
}
if (!"region" %in% names(data.out)) data.out[, region := NA_character_]

# ---- Cause label map ---------------------------------------------------------
cause_map_full <- c(
  ihd     = "Ischemic heart disease",
  istroke = "Ischemic stroke",
  hstroke = "Intracerebral hemorrhage",
  hhd     = "Hypertensive heart disease"
)
num_causes  <- length(unique(cause_map_full))
study_years <- paste0(min(data.out$year[data.out$year >= 2025]),
                       "–", max(data.out$year))
```

---

# Background

Indonesia faces a rapidly growing burden of noncommunicable diseases (NCDs), driven by population aging, high smoking prevalence, rising obesity and diabetes, and limited access to preventive care. Cardiovascular disease (CVD) is the leading cause of death in Indonesia, accounting for a substantial share of premature mortality and disability.

This report presents modelled projections of CVD mortality and economic impact under a set of risk-factor control interventions for Indonesia over the period `r study_years`. Four causes of CVD death are modelled jointly: ischemic heart disease (IHD), ischemic stroke, intracerebral hemorrhage (hemorrhagic stroke), and hypertensive heart disease (HHD). A discrete-time multi-state transition model (Susceptible $\rightarrow$ Incident $\rightarrow$ Prevalent $\rightarrow$ Dead) is calibrated to GBD 2023 data and projected forward to 2050 using UNWPP 2024 population estimates.

The central research question is:

> *What would be the projected health and economic impact --- in terms of deaths averted, life-years gained, DALYs averted, and monetary value of mortality reduction --- of implementing CVD risk-factor control interventions in Indonesia from 2025 to 2050?*

---

# Intervention Scenarios

The model evaluates seven intervention scenarios against a business-as-usual (BAU) baseline, reflecting four types of CVD risk-factor control policies. These scenarios are defined and parameterised in `code/cvd/06_run_scenarios.R`.

## Individual Interventions

| Scenario | Intervention | Key Parameters |
|----------|-------------|----------------|
| `bp_only` | **Antihypertensive Therapy** | Country-specific HTN control targets (aspirational); linear scale-up from baseline to target, 2026--2040. Effect sizes from Ettehad et al. applied per BP bin. |
| `sodium_only` | **Sodium Reduction** | Infrastructure is in place for future activation of a 30% reduction target. |
| `tfa_only` | **Trans-Fat Elimination** | Elimination of trans-fat intake (target 0% of energy from TFA); policy start year 2028. Affects IHD case fatality via GBD-based age-specific relative risks. |
| `statins_only` | **Statin Therapy** | Scale-up to 60% coverage among eligible adults by 2050; full adherence assumed. Reduces IHD and ischemic stroke incidence and case fatality. |

## Combination Scenarios

| Scenario | Interventions Combined |
|----------|----------------------|
| `bp_sodium` | Antihypertensive Therapy + Sodium Reduction |
| `bp_sodium_tfa` | Antihypertensive Therapy + Sodium Reduction + Trans-Fat Elimination |
| `all_four` | **All Interventions** --- Antihypertensive Therapy + Sodium Reduction + Trans-Fat Elimination + Statin Therapy |

Combination scenarios apply intervention effects multiplicatively: each intervention's effect ratio for incidence and case fatality is combined as a product across active interventions within each scenario.

---

# Policy Scenario Trajectories

## Hypertension Control Scale-up

```{r policy-bp, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
covfxn <- tryCatch(fread(paste0(wd_data, "covfxn2.csv")), error = function(e) NULL)

if (!is.null(covfxn)) {
  dt_plot <- covfxn[, .(Country, Year, scenario_value = Progress)]
  g <- ggplot(dt_plot, aes(x = Year, y = scenario_value * 100, group = Country)) +
    geom_line(color = "grey70", linewidth = 0.6, alpha = 0.7) +
    geom_hline(yintercept = 50, linetype = "dashed", color = "#1f78b4", linewidth = 1) +
    annotate("text", x = 2051, y = 51.5, hjust = 1, size = 3.5,
             label = "50% target (2040)", color = "#1f78b4") +
    scale_x_continuous(breaks = seq(2020, 2050, 5), limits = c(2020, 2051), expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 100), expand = c(0, 0)) +
    labs(
      title    = "Hypertension Control Scale-up — Indonesia",
      subtitle = "Linear scale-up to country-specific aspirational HTN control target by 2040",
      x = "Year", y = "Hypertension Control Rate (%)"
    ) +
    theme_minimal(base_size = 13) +
    theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank())
  print(g)
  ggsave(paste0(wd_outp, "aim1_policy_htn_general.png"), plot = g,
         width = 11, height = 5, dpi = 300)
} else {
  message("covfxn2.csv not found - skipping HTN scale-up plot.")
}
```

## Statin Therapy Scale-up

```{r policy-statins, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
dt_statin_scen <- tryCatch(
  readRDS(paste0(wd_data, "statin_data.rds")),
  error = function(e) NULL
)

if (!is.null(dt_statin_scen)) {
  dt_statin_idn <- dt_statin_scen[location == "Indonesia"]
  if (nrow(dt_statin_idn) > 0) {
    g <- ggplot(dt_statin_idn,
                aes(x = year, y = statins_uptake * 100)) +
      geom_line(color = "#e31a1c", linewidth = 1) +
      geom_hline(yintercept = 50, linetype = "dashed", color = "#e31a1c", linewidth = 0.8) +
      annotate("text", x = 2035, y = 52, hjust = 0.5, size = 3.5,
               label = "50% target (2030)", color = "#e31a1c") +
      scale_x_continuous(breaks = seq(2020, 2050, 5), limits = c(2020, 2051), expand = c(0, 0)) +
      scale_y_continuous(limits = c(0, 100), expand = c(0, 0)) +
      labs(
        title    = "Statin Therapy Scale-up — Indonesia",
        subtitle = "Scale-up to 50% coverage among eligible adults by 2030",
        x = "Year", y = "Risk Population Receiving Statin Therapy (%)"
      ) +
      theme_minimal(base_size = 13) +
      theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank())
    print(g)
    ggsave(paste0(wd_outp, "aim1_policy_statins.png"), plot = g,
           width = 11, height = 5, dpi = 300)
  } else {
    message("No Indonesia data in statin file - skipping plot.")
  }
} else {
  message("Statin data file not found - skipping plot.")
}
```

---

# Disease Trends (Baseline)

## Incidence and Prevalence Over Time

```{r disease-trends, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
dt_trend <- data.out[intervention == "b.a.u" & year >= 2019,
                     .(new_cases = sum(newcases, na.rm = TRUE),
                       sick      = sum(sick,     na.rm = TRUE),
                       pop       = sum(pop,       na.rm = TRUE)),
                     by = year]

dt_melt <- melt(dt_trend, id.vars = "year",
                measure.vars  = c("new_cases", "sick"),
                variable.name = "metric", value.name = "count")
dt_melt[, metric_lbl := fifelse(metric == "new_cases", "Incident cases", "Prevalent cases")]

g <- ggplot(dt_melt, aes(x = year, y = count / 1e6,
                          color = metric_lbl, linetype = year > 2019)) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Incident cases"  = "steelblue",
                                "Prevalent cases" = "tomato"),
                     name = NULL) +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"), guide = "none") +
  scale_y_continuous(labels = comma) +
  labs(
    title    = "Incident and Prevalent CVD Cases in Indonesia — Business-as-Usual",
    subtitle = "All four causes combined; solid = observed, dashed = projected",
    x = "Year", y = "Persons (millions)",
    caption = "Source: Model output; ages 20–95+"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "aim1_disease_trend.png"), plot = g,
       width = 11, height = 5, dpi = 300)
```

## Age-Standardised Mortality Rate by Cause

```{r asmr-cause, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=6}
dt_asmr_in <- data.out[intervention == "b.a.u" & year >= 2019,
                        .(deaths_dx = sum(dead, na.rm = TRUE),
                          well      = sum(well, na.rm = TRUE),
                          sick      = sum(sick, na.rm = TRUE)),
                        by = .(cause, age, year)]
dt_asmr_in[, Nx := well + sick]

asmr_cause <- calculate_asmr(dt_asmr_in, deaths_col = "deaths_dx",
                              population_col = "Nx", disagg_var = "cause")
asmr_base  <- asmr_cause[year == 2019, .(cause, ASMR_base = ASMR)]
asmr_cause <- merge(asmr_cause, asmr_base, by = "cause")
asmr_cause[, index := ASMR / ASMR_base * 100]
asmr_cause[, cause_lbl := fcase(
  cause == "ihd",     "Ischemic heart disease",
  cause == "istroke", "Ischemic stroke",
  cause == "hstroke", "Intracerebral hemorrhage",
  cause == "hhd",     "Hypertensive heart disease",
  default = cause
)]

g <- ggplot(asmr_cause, aes(x = year, y = index,
                             color = cause_lbl, linetype = year > 2019)) +
  geom_line(linewidth = 1) +
  geom_hline(yintercept = 100, linetype = "dotted", color = "grey50") +
  scale_color_manual(values = setNames(cause_cols, cause_labels_plot),
                     name = "Cause of death") +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"), guide = "none") +
  labs(
    title    = "Age-Standardised CVD Mortality Rate Index in Indonesia — Business-as-Usual",
    subtitle = "Index: 2019 = 100; solid = observed, dashed = projected",
    x = "Year", y = "ASMR Index (2019 = 100)",
    caption = "WHO World Standard Population (2000–2025), rescaled to ages 20–95+"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "aim1_asmr_cause.png"), plot = g,
       width = 11, height = 5, dpi = 300)
```

---

# Projected Deaths Averted (2025--2050)

```{r compute-deaths-averted, message=FALSE, warning=FALSE, echo=FALSE}
# ---- Core: deaths delayed by intervention x year ----------------------------
dt_base <- data.out[intervention == "b.a.u" & year >= 2025,
                    .(baseline_deaths = sum(dead, na.rm = TRUE)),
                    by = .(age_group, sex, year, location, cause, age)]

dt_nonbase <- data.out[intervention != "b.a.u" & year >= 2025,
                       .(dead = sum(dead, na.rm = TRUE)),
                       by = .(age_group, sex, year, location, cause, age, intervention)]

dt_merged <- merge(dt_nonbase, dt_base,
                   by = c("age_group", "sex", "year", "location", "cause", "age"),
                   all.x = TRUE)
dt_merged[, deaths_delayed := baseline_deaths - dead]

# ---- Cumulative totals -------------------------------------------------------
dt_cumul <- dt_merged[, .(deaths_delayed = sum(deaths_delayed, na.rm = TRUE)),
                       by = intervention]

# ---- Annual time-series ------------------------------------------------------
dt_annual <- dt_merged[, .(deaths_delayed = sum(deaths_delayed, na.rm = TRUE)),
                        by = .(intervention, year)]
dt_annual[, intervention := factor(intervention, levels = int_order)]
setorder(dt_annual, intervention, year)
dt_annual[, cum_deaths_delayed := cumsum(deaths_delayed), by = intervention]

# ---- Per-scenario totals for inline text ------------------------------------
total_all  <- dt_cumul[intervention == "All Interventions",        deaths_delayed]
total_bp   <- dt_cumul[intervention == "Antihypertensive Therapy", deaths_delayed]
total_salt <- dt_cumul[intervention == "Sodium Reduction",         deaths_delayed]
total_tfa  <- dt_cumul[intervention == "Trans-Fat Elimination",    deaths_delayed]
total_stat <- dt_cumul[intervention == "Statin Therapy",           deaths_delayed]
```

Between 2025 and 2050, implementing all four intervention strategies simultaneously in Indonesia is projected to avert approximately **`r fmt_k(total_all)` thousand deaths**. Antihypertensive therapy alone would avert `r fmt_k(total_bp)` thousand deaths; trans-fat elimination would avert `r fmt_k(total_tfa)` thousand deaths; and statin therapy scale-up would avert `r fmt_k(total_stat)` thousand deaths. The sodium reduction scenario is currently parameterised with zero effect (`salteff = 0`) and therefore contributes no additional mortality reduction in this model run.

## Table 1 --- Cumulative Deaths Averted by Intervention (2025--2050)

```{r table1, message=FALSE, warning=FALSE, echo=FALSE}
tab1 <- copy(dt_cumul)
tab1 <- tab1[intervention != "b.a.u"]
tab1[, intervention := factor(intervention, levels = int_order)]
setorder(tab1, intervention)

tab1[, deaths_k      := round(deaths_delayed / 1e3, 1)]
tab1[, deaths_annual := round(deaths_delayed / 25 / 1e3, 1)]

tab1_fmt <- tab1[, .(
  Intervention       = intervention,
  Deaths_averted_k   = comma(deaths_k,      accuracy = 0.1),
  Avg_per_year       = comma(deaths_annual,  accuracy = 0.1)
)]
setnames(tab1_fmt,
         old = c("Deaths_averted_k", "Avg_per_year"),
         new = c("Deaths averted (thousands, 2025–2050)",
                 "Average per year (thousands)"))

kable(tab1_fmt,
      caption = "Table 1. Cumulative deaths averted in Indonesia (2025–2050) by intervention scenario.",
      align   = "lrr") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) |>
  column_spec(1, width = "8cm") |>
  row_spec(0, bold = TRUE) |>
  footnote(general = paste0(
    "Deaths averted = deaths in business-as-usual scenario minus deaths under each intervention, ",
    "summed over 2025–2050. ",
    "Causes: ischemic heart disease, ischemic stroke, intracerebral hemorrhage, ",
    "hypertensive heart disease. Sodium reduction is parameterised with zero effect in the current run."
  ))

dt_excel_tab1 <- copy(tab1_fmt)
```

## Figure 1 --- Deaths Averted by Intervention (Bar Chart)

```{r figure1, message=FALSE, warning=FALSE, echo=FALSE, fig.width=9, fig.height=5}
dt_fig1 <- tab1[, .(
  intervention   = factor(intervention, levels = rev(int_order)),
  deaths_delayed = deaths_delayed
)]

g <- ggplot(dt_fig1, aes(x = deaths_delayed / 1e3, y = intervention,
                          fill = as.character(intervention))) +
  geom_col(width = 0.65) +
  scale_fill_manual(values = int_cols, guide = "none") +
  scale_x_continuous(expand = c(0, 0), labels = comma) +
  labs(
    title   = "Projected Deaths Averted by Intervention in Indonesia, 2025–2050",
    x       = "Deaths averted (thousands)",
    y       = NULL,
    caption = "Deaths averted vs. business-as-usual scenario; all causes combined."
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"),
        panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank())
print(g)
ggsave(paste0(wd_outp, "aim1_fig1_deaths_intervention.png"), plot = g,
       width = 9, height = 5, dpi = 300)
```

## Table 2 --- Cumulative Deaths Averted by Intervention and Cause

```{r table2, message=FALSE, warning=FALSE, echo=FALSE}
dt_tab2 <- dt_merged[
  intervention %in% int_order,
  .(deaths_delayed = sum(deaths_delayed, na.rm = TRUE)),
  by = .(intervention, cause)
]

dt_tab2[, cause_lbl := fcase(
  cause == "ihd",     "Ischemic heart disease",
  cause == "istroke", "Ischemic stroke",
  cause == "hstroke", "Intracerebral hemorrhage",
  cause == "hhd",     "Hypertensive heart disease",
  default = cause
)]

dt_tab2_wide <- dcast(dt_tab2, intervention ~ cause_lbl,
                      value.var = "deaths_delayed", fill = NA)
dt_tab2_wide[, intervention := factor(intervention, levels = int_order)]
setorder(dt_tab2_wide, intervention)

for (col in names(dt_tab2_wide)[names(dt_tab2_wide) != "intervention"]) {
  dt_tab2_wide[, (col) := fifelse(
    is.na(get(col)) | abs(as.numeric(get(col))) < 1, "–",
    comma(round(get(col) / 1e3, 1), accuracy = 0.1)
  )]
}

kable(dt_tab2_wide,
      caption = "Table 2. Cumulative deaths averted in Indonesia (thousands, 2025–2050) by intervention and cause of death.",
      align   = c("l", rep("r", ncol(dt_tab2_wide) - 1))) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) |>
  add_header_above(setNames(c(1L, ncol(dt_tab2_wide) - 1L),
                            c(" ", "Deaths averted, thousands (2025–2050)"))) |>
  row_spec(0, bold = TRUE) |>
  footnote(general = paste0(
    "– = intervention has negligible or no effect on this cause. ",
    "Statin therapy affects IHD and ischemic stroke only. ",
    "TFA elimination affects IHD case fatality only. ",
    "Antihypertensive interventions affect all four CVD causes."
  ))

dt_excel_tab2 <- copy(dt_tab2_wide)
```

## Figure 2 --- Deaths Averted by Cause (Stacked Bar)

```{r figure2, message=FALSE, warning=FALSE, echo=FALSE, fig.width=9, fig.height=6}
dt_fig2 <- dt_merged[
  intervention %in% int_order & cause %in% names(cause_map_full),
  .(deaths_delayed = sum(deaths_delayed, na.rm = TRUE)),
  by = .(intervention, cause)
]
dt_fig2[, intervention := factor(intervention, levels = int_order)]
dt_fig2[, cause_lbl := cause_labels_plot[cause]]
dt_fig2[, cause_lbl := factor(cause_lbl, levels = unname(cause_labels_plot))]

g <- ggplot(dt_fig2, aes(x = deaths_delayed / 1e3, y = intervention, fill = cause_lbl)) +
  geom_col(width = 0.65) +
  scale_fill_manual(values = setNames(cause_cols, cause_labels_plot),
                    name = "Cause of death") +
  scale_x_continuous(expand = c(0, 0), labels = comma) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  labs(
    title = "Deaths Averted by Cause of Death and Intervention in Indonesia, 2025–2050",
    x     = "Deaths averted (thousands)",
    y     = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position    = "bottom",
        panel.grid.minor   = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.title         = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "aim1_fig2_deaths_cause.png"), plot = g,
       width = 9, height = 6, dpi = 300)
```

## Figure 3 --- Time Series: Annual Deaths Averted by Individual Intervention

```{r figure3-annual, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
g <- ggplot(dt_annual[intervention %in% int_order_individual],
            aes(x = year, y = deaths_delayed / 1e3,
                color = intervention, fill = intervention)) +
  geom_area(position = "stack", alpha = 0.7, color = "grey20", linewidth = 0.15) +
  scale_color_manual(values = int_cols, name = NULL) +
  scale_fill_manual(values  = int_cols, name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Annual Deaths Averted by Individual Intervention in Indonesia, 2026–2050",
    x     = "Year",
    y     = "Deaths averted (thousands)"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "aim1_fig3_annual_averted.png"), plot = g,
       width = 11, height = 5, dpi = 300)
```

## Figure 4 --- Cumulative Deaths Averted Over Time

```{r figure4-cumulative, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
g <- ggplot(dt_annual[intervention %in% int_order_individual],
            aes(x = year, y = cum_deaths_delayed / 1e3, color = intervention)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = int_cols, name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Cumulative Deaths Averted by Individual Intervention in Indonesia, 2026–2050",
    x     = "Year",
    y     = "Cumulative deaths averted (thousands)"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "aim1_fig4_cum_averted.png"), plot = g,
       width = 11, height = 5, dpi = 300)
```

---

# Outputs by Demographics

## By Sex and Age Group

```{r demo-age-sex, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
dt_age_sex <- dt_merged[
  intervention %in% int_order,
  .(deaths_delayed = sum(deaths_delayed, na.rm = TRUE)),
  by = .(intervention, age_group, sex)
]

order_lvls <- dt_age_sex[, .(total = sum(deaths_delayed, na.rm = TRUE)), by = intervention
                         ][order(total), intervention]
dt_age_sex[, intervention := factor(intervention, levels = order_lvls)]

g <- ggplot(dt_age_sex,
            aes(x = deaths_delayed / 1e3, y = intervention, fill = age_group)) +
  geom_col(width = 0.7) +
  facet_wrap(~sex, nrow = 1) +
  scale_x_continuous(expand = c(0, 0), labels = comma) +
  labs(
    title = "Deaths Averted by Age Group, Sex, and Intervention in Indonesia (2025–2050)",
    x     = "Deaths averted (thousands)",
    y     = NULL,
    fill  = "Age group"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position  = "bottom",
        panel.border     = element_rect(color = "black", fill = NA, linewidth = 0.5),
        strip.text       = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        plot.title       = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "aim1_fig5_demo_age.png"), plot = g,
       width = 10, height = 5, dpi = 300)
```

## Table 3 --- Deaths Averted by Age Group and Intervention

```{r table3-age, message=FALSE, warning=FALSE, echo=FALSE}
dt_tab3 <- deaths_summary(
  data.out[year >= 2025],
  baseline_label = "b.a.u",
  by_cols        = c("age_group", "intervention")
)
dt_tab3 <- dt_tab3[intervention %in% int_order]
dt_tab3[, intervention := factor(intervention, levels = int_order)]
setorder(dt_tab3, age_group, intervention)

dt_tab3_fmt <- dt_tab3[, .(
  Age_group          = age_group,
  Intervention       = intervention,
  Projected_deaths   = comma(deaths_projected,   accuracy = 1),
  Baseline_deaths    = comma(baseline_deaths,    accuracy = 1),
  Deaths_averted     = comma(deaths_delayed,     accuracy = 1),
  Deaths_averted_yr  = comma(deaths_delayed_ave, accuracy = 1),
  Pct_averted        = percent(deaths_delayed_pct, accuracy = 0.1)
)]
setnames(dt_tab3_fmt,
         old = c("Age_group", "Projected_deaths", "Baseline_deaths",
                 "Deaths_averted", "Deaths_averted_yr", "Pct_averted"),
         new = c("Age group", "Projected deaths", "Baseline deaths",
                 "Deaths averted", "Deaths averted/year", "% averted"))

kable(dt_tab3_fmt,
      caption = "Table 3. Cumulative deaths averted in Indonesia (2025–2050) by age group and intervention.",
      align   = c("l","l","r","r","r","r","r")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 12) |>
  row_spec(0, bold = TRUE)

dt_excel_tab3 <- copy(dt_tab3_fmt)
```

---

# ASMR Under Intervention Scenarios

```{r asmr-interventions, message=FALSE, warning=FALSE, echo=FALSE, fig.width=12, fig.height=7}
dt_asmr2 <- data.out[year >= 2019,
                      .(dead = sum(dead, na.rm = TRUE),
                        well = sum(well, na.rm = TRUE),
                        sick = sum(sick, na.rm = TRUE)),
                      by = .(intervention, age, year)]
dt_asmr2[, Nx := well + sick]

asmr_int <- calculate_asmr(dt_asmr2, deaths_col = "dead",
                            population_col = "Nx", disagg_var = "intervention")
asmr_int[, intervention := factor(intervention, levels = c("b.a.u", int_order))]
asmr_int[, int_lbl := fcase(intervention == "b.a.u", "Business-as-Usual",
                             default = as.character(intervention))]

g <- ggplot(asmr_int,
            aes(x = year, y = ASMR, color = int_lbl, linetype = year > 2019)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = c("Business-as-Usual" = "#999999", int_cols), name = NULL) +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"), guide = "none") +
  labs(
    title    = "Age-Standardised CVD Mortality Rate by Intervention Scenario — Indonesia",
    subtitle = "All causes combined; solid = calibration period, dashed = projection",
    x        = "Year",
    y        = "ASMR (per 100,000 population)",
    caption  = "WHO World Standard Population (2000–2025), rescaled to ages 20–95+"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "aim1_asmr_interventions.png"), plot = g,
       width = 12, height = 6, dpi = 300)
```

---

# Burden of Disease Results

This section presents burden of disease estimates --- including deaths averted, years of life lost (YLLs) averted, years lived with disability (YLDs) averted, and disability-adjusted life years (DALYs) averted --- for each intervention scenario in Indonesia over the period 2026--2050. All "averted" values are computed as the difference between the business-as-usual scenario and each intervention scenario, cumulated across years.

```{r load-dalys, echo=FALSE, message=FALSE, warning=FALSE}
# ---- Load DALYs output from 07_output_dalys.R --------------------------------
dt_dalys <- readRDS(paste0(wd_outp, "dt_output_dalys.rds"))
setDT(dt_dalys)

# ---- Apply same scenario label mapping as data.out --------------------------
if ("scenario" %in% names(dt_dalys) && !"intervention" %in% names(dt_dalys)) {
  setnames(dt_dalys, "scenario", "intervention")
} else if ("scenario" %in% names(dt_dalys)) {
  dt_dalys[, intervention := scenario]
  dt_dalys[, scenario := NULL]
}

dt_dalys[intervention %in% c("baseline", "BAU", "b.a.u"),
         intervention := "b.a.u"]
dt_dalys[intervention %in% c("bp_only", "antihypertensive", "BP"),
         intervention := "Antihypertensive Therapy"]
dt_dalys[intervention %in% c("sodium_only", "sodium", "Salt"),
         intervention := "Sodium Reduction"]
dt_dalys[intervention %in% c("tfa_only", "tfa", "TFA"),
         intervention := "Trans-Fat Elimination"]
dt_dalys[intervention %in% c("statins_only", "statins", "Statins"),
         intervention := "Statin Therapy"]
dt_dalys[intervention %in% c("bp_sodium", "BP + Salt"),
         intervention := "HTN + Sodium"]
dt_dalys[intervention %in% c("bp_sodium_tfa", "BP + Salt + TFA"),
         intervention := "HTN + Sodium + TFA"]
dt_dalys[intervention %in% c("all_four", "all_interventions",
                              "BP + Salt + TFA + Statins"),
         intervention := "All Interventions"]

# ---- Merge country metadata ---------------------------------------------------
dt_dalys <- merge(dt_dalys,
                  countries_dt[, .(location, region, location_income)],
                  by = "location", all.x = TRUE)

# ---- Compute baseline values per cell ----------------------------------------
daly_base <- dt_dalys[intervention == "b.a.u",
                      .(base_deaths = sum(dead, na.rm = TRUE),
                        base_yll    = sum(yll, na.rm = TRUE),
                        base_yld    = sum(yld, na.rm = TRUE),
                        base_daly   = sum(daly, na.rm = TRUE)),
                      by = .(location, cause, year)]

daly_int <- dt_dalys[intervention != "b.a.u",
                     .(deaths = sum(dead, na.rm = TRUE),
                       yll    = sum(yll, na.rm = TRUE),
                       yld    = sum(yld, na.rm = TRUE),
                       daly   = sum(daly, na.rm = TRUE)),
                     by = .(location, cause, year, intervention)]

dt_bod <- merge(daly_int, daly_base, by = c("location", "cause", "year"), all.x = TRUE)
dt_bod[, `:=`(deaths_averted = base_deaths - deaths,
              yll_averted    = base_yll    - yll,
              yld_averted    = base_yld    - yld,
              daly_averted   = base_daly   - daly)]

# Merge region and income group onto the differenced data
dt_bod <- merge(dt_bod,
                countries_dt[, .(location, region, location_income)],
                by = "location", all.x = TRUE)

# Filer by int_order
dt_bod <- dt_bod[intervention %in% c("b.a.u", int_order)]

```

## Table --- Burden of Disease Averted by Intervention Scenario

Cumulative 2026--2050 deaths, YLLs, YLDs, and DALYs averted by intervention scenario, relative to business-as-usual.

```{r bod-table1, echo=FALSE, message=FALSE, warning=FALSE}
bod_tab1 <- dt_bod[, .(Deaths  = sum(deaths_averted, na.rm = TRUE),
                       YLLs    = sum(yll_averted,    na.rm = TRUE),
                       YLDs    = sum(yld_averted,    na.rm = TRUE),
                       DALYs   = sum(daly_averted,   na.rm = TRUE)),
                   by = intervention]
bod_tab1[, intervention := factor(intervention, levels = int_order)]
setorder(bod_tab1, intervention)

bod_tab1_fmt <- copy(bod_tab1)
bod_tab1_fmt[, Deaths := fmt_k(Deaths)]
bod_tab1_fmt[, YLLs   := fmt_k(YLLs)]
bod_tab1_fmt[, YLDs   := fmt_k(YLDs)]
bod_tab1_fmt[, DALYs  := fmt_k(DALYs)]
setnames(bod_tab1_fmt, "intervention", "Intervention")

kable(bod_tab1_fmt,
      caption = "Cumulative burden of disease averted in Indonesia (thousands, 2026–2050) by intervention scenario.",
      align = c("l", "r", "r", "r", "r")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) |>
  row_spec(0, bold = TRUE) |>
  footnote(general = "Values in thousands. Averted = business-as-usual minus intervention scenario.")
```

## Table --- Burden of Disease Averted by Intervention Scenario and Cause

Cumulative 2026--2050 deaths and DALYs averted by intervention scenario and cause, with percentage contribution of each cause to total reduction within each scenario.

```{r bod-table2, echo=FALSE, message=FALSE, warning=FALSE}
bod_tab2 <- dt_bod[, .(Deaths  = sum(deaths_averted, na.rm = TRUE),
                       YLLs    = sum(yll_averted,    na.rm = TRUE),
                       YLDs    = sum(yld_averted,    na.rm = TRUE),
                       DALYs   = sum(daly_averted,   na.rm = TRUE)),
                   by = .(intervention, cause)]
bod_tab2[, intervention := factor(intervention, levels = int_order)]

# Compute % contribution within each scenario
bod_tab2[, pct_deaths := Deaths / sum(Deaths) * 100, by = intervention]
bod_tab2[, pct_dalys  := DALYs  / sum(DALYs)  * 100, by = intervention]

bod_tab2[, cause_lbl := fcase(
  cause == "ihd",     "Ischemic heart disease",
  cause == "istroke", "Ischemic stroke",
  cause == "hstroke", "Intracerebral hemorrhage",
  cause == "hhd",     "Hypertensive heart disease",
  default = cause
)]
setorder(bod_tab2, intervention, cause)

bod_tab2_fmt <- bod_tab2[, .(
  Intervention        = intervention,
  Cause               = cause_lbl,
  Deaths              = fmt_k(Deaths),
  YLLs                = fmt_k(YLLs),
  YLDs                = fmt_k(YLDs),
  DALYs               = fmt_k(DALYs),
  `% Deaths`          = paste0(round(pct_deaths, 1), "%"),
  `% DALYs`           = paste0(round(pct_dalys,  1), "%")
)]

kable(bod_tab2_fmt,
      caption = "Cumulative burden of disease averted in Indonesia (thousands, 2026–2050) by intervention and cause.",
      align = c("l", "l", "r", "r", "r", "r", "r", "r")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 12) |>
  row_spec(0, bold = TRUE) |>
  footnote(general = paste0(
    "Values in thousands. % columns show each cause’s share of the total reduction ",
    "within that intervention scenario."
  ))
```

## Figure --- DALYs Averted Time Series by Intervention Scenario

Cumulative DALYs averted over time (2026--2050), with one line per intervention scenario.

```{r bod-fig1-dalys-ts, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
bod_annual <- dt_bod[, .(daly_averted = sum(daly_averted, na.rm = TRUE)),
                     by = .(intervention, year)]
bod_annual[, intervention := factor(intervention, levels = int_order)]
setorder(bod_annual, intervention, year)
bod_annual[, cum_daly_averted := cumsum(daly_averted), by = intervention]

g <- ggplot(bod_annual, aes(x = year, y = cum_daly_averted / 1e3,
                             color = intervention)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = int_cols, name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Cumulative DALYs Averted by Intervention Scenario in Indonesia, 2026–2050",
    x     = "Year",
    y     = "Cumulative DALYs averted (thousands)"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "aim1_bod_fig1_dalys_ts.png"), plot = g,
       width = 10, height = 5, dpi = 300)
```

---

# Economic Evaluation

This section presents the economic value of CVD mortality reductions under each intervention scenario in Indonesia. The analysis estimates the monetary benefit of deaths averted using two complementary approaches: the Value of a Statistical Life (VSL) and the Value of a Statistical Life Year (VSLY). Results are produced by `code/cvd/08_economic_value_calculation.R`.

## Methodology

The economic evaluation follows the methodological framework of Robinson & Hammitt (2011) and Robinson et al. (2019):

**Value of a Statistical Life (VSL) transfer.** The VSL for Indonesia is derived from a US reference VSL using an income-based benefit transfer:

$$\text{VSL}_{\text{IDN}} = \text{VSL}_{\text{USA}} \times \left(\frac{\text{GNI}_{\text{IDN}}}{\text{GNI}_{\text{USA}}}\right)^{\varepsilon}$$

where $\varepsilon$ is the income elasticity. The primary estimate uses the differential elasticity recommended by Robinson et al. (2019): $\varepsilon = 0.8$ for countries at or above US income, $\varepsilon = 1.2$ for countries below (Indonesia falls in the latter category). Uniform elasticities of 1.0 and 1.5 are used as sensitivity bounds. A minimum VSL-to-GNI ratio floor of 20 is applied.

**Value of a Statistical Life Year (VSLY).** A constant VSLY is derived as:

$$\text{VSLY} = \frac{\text{VSL}}{\text{LE}_{\text{avg adult}}}$$

where $\text{LE}_{\text{avg adult}}$ is the undiscounted remaining life expectancy at the average age of the adult population (from WPP 2024 life tables). VSLY monetary values are then computed using age-specific life-years gained:

$$\text{VSLY value} = \text{VSLY} \times \sum_{a} \left[\text{deaths averted}(a) \times \text{LE}(a)\right]$$

**Income data and projection.** GNI per capita (PPP, constant international dollars) is sourced from the World Bank. For years beyond the last observed value, GNI is projected forward using SSP2 annual GDP per capita growth rates from IIASA SSP 3.1.

**Discounting.** Annual benefit streams are discounted to a 2026 base year at a primary rate of 3%, with 1% and 5% as sensitivity bounds. The discount factor applied is $d_t = (1 + r)^{-(t - 2026)}$.

**Time horizon and perspective.** Benefits are cumulated over 2026--2050. The analysis takes a societal perspective, valuing all mortality reductions regardless of payer.

```{r load-econ, echo=FALSE, message=FALSE, warning=FALSE}
# ---- Load economic evaluation outputs from 08_economic_value_calculation.R ----
econ_results_file <- paste0(wd_outp, "08_vsl_results.rds")
econ_vsl_file     <- paste0(wd_outp, "08_vsl_summary_table_e1_2_primary.rds")
econ_vsly_file    <- paste0(wd_outp, "08_vsly_summary_table_e1_2_primary.rds")
econ_app_file     <- paste0(wd_outp, "08_vsl_vsly_summary_table_appended.rds")

econ_ok <- all(file.exists(c(econ_results_file, econ_vsl_file, econ_vsly_file, econ_app_file)))

if (!econ_ok) {
  warning("Economic evaluation outputs not found in ", wd_outp,
          ". Run code/cvd/08_economic_value_calculation.R first.")
}
```

```{r econ-results-table, echo=FALSE, message=FALSE, warning=FALSE, eval=econ_ok}
# ---- Load full results and build Indonesia-specific summary ------------------
dt_econ <- as.data.table(readRDS(econ_results_file))

# ---- Scenario label mapping (matches model output scenario names) -----------
econ_scen_lbl <- c(
  bp_only       = "Antihypertensive Therapy",
  sodium_only   = "Sodium Reduction",
  tfa_only      = "Trans-Fat Elimination",
  statins_only  = "Statin Therapy",
  bp_sodium     = "HTN + Sodium",
  bp_sodium_tfa = "HTN + Sodium + TFA",
  all_four      = "All Interventions"
)

dt_econ <- dt_econ[scenario %in% names(econ_scen_lbl)]
dt_econ[, intervention := factor(econ_scen_lbl[scenario], levels = int_order)]

# Filer by int_order
dt_econ <- dt_econ[intervention %in% c("b.a.u", int_order)]

```

## Table --- Economic Value of Deaths Averted (VSL and VSLY, Primary Estimate)

Cumulative discounted value of deaths averted over 2026--2050, using the primary income elasticity (differential $\varepsilon$ = 0.8/1.2) and 3% discount rate.

```{r econ-summary-table, echo=FALSE, message=FALSE, warning=FALSE, eval=econ_ok}
econ_summary <- dt_econ[
  location == "Indonesia",
  .(
    Deaths_averted        = sum(deaths_averted, na.rm = TRUE),
    VSL_value_disc        = sum(economic_value_e1_2_disc_r3, na.rm = TRUE),
    VSLY_value_disc       = sum(vsly_value_e1_2_disc_r3, na.rm = TRUE),
    Life_years_gained     = sum(life_years_gained_undisc, na.rm = TRUE)
  ),
  by = .(intervention)
]
setorder(econ_summary, intervention)

# Format for display
econ_summary_fmt <- econ_summary[, .(
  Intervention       = intervention,
  `Deaths averted`   = comma(round(Deaths_averted), accuracy = 1),
  `Life-years gained` = comma(round(Life_years_gained), accuracy = 1),
  `VSL value (M USD PPP)` = comma(round(VSL_value_disc / 1e6, 1), accuracy = 0.1),
  `VSLY value (M USD PPP)` = comma(round(VSLY_value_disc / 1e6, 1), accuracy = 0.1)
)]

kable(econ_summary_fmt,
      caption = paste0(
        "Economic value of CVD deaths averted in Indonesia (2026–2050). ",
        "Values discounted at 3% to 2026 base year; primary elasticity (e = 0.8/1.2)."
      ),
      align = c("l", "r", "r", "r", "r")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) |>
  row_spec(0, bold = TRUE) |>
  footnote(general = paste0(
    "VSL = Value of a Statistical Life; VSLY = Value of a Statistical Life Year. ",
    "M USD PPP = millions of constant international dollars (purchasing power parity). ",
    "Income elasticity: differential (0.8 HIC / 1.2 LMIC), Robinson et al. (2019)."
  ))
```

## Table --- Economic Value as Share of National Income

```{r econ-share-table, echo=FALSE, message=FALSE, warning=FALSE, eval=econ_ok}
econ_share <- dt_econ[
  location == "Indonesia",
  .(
    VSL_value_disc  = sum(economic_value_e1_2_disc_r3, na.rm = TRUE),
    VSLY_value_disc = sum(vsly_value_e1_2_disc_r3, na.rm = TRUE),
    Total_income    = sum(population * gni_pc_ppp * disc_r3, na.rm = TRUE)
  ),
  by = .(intervention)
]
econ_share[, `:=`(
  VSL_share  = VSL_value_disc / Total_income,
  VSLY_share = VSLY_value_disc / Total_income
)]
setorder(econ_share, intervention)

econ_share_fmt <- econ_share[, .(
  Intervention     = intervention,
  `VSL (% of GNI)` = percent(VSL_share, accuracy = 0.01),
  `VSLY (% of GNI)` = percent(VSLY_share, accuracy = 0.01)
)]

kable(econ_share_fmt,
      caption = paste0(
        "Economic value of deaths averted as share of discounted national income, ",
        "Indonesia 2026–2050. Primary elasticity (e = 0.8/1.2), r = 3%."
      ),
      align = c("l", "r", "r")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) |>
  row_spec(0, bold = TRUE) |>
  footnote(general = paste0(
    "Share of GNI = cumulative discounted VSL or VSLY value divided by cumulative ",
    "discounted national income (population × GNI per capita). Discount rate = 3%."
  ))
```

## Figure --- Economic Value Over Time (VSLY, Primary Estimate)

```{r econ-time-series, echo=FALSE, message=FALSE, warning=FALSE, eval=econ_ok, fig.width=10, fig.height=5}
econ_annual <- dt_econ[
  location == "Indonesia",
  .(vsly_disc = sum(vsly_value_e1_2_disc_r3, na.rm = TRUE)),
  by = .(intervention, year)
]
setorder(econ_annual, intervention, year)
econ_annual[, cum_vsly := cumsum(vsly_disc), by = intervention]

g <- ggplot(econ_annual,
            aes(x = year, y = cum_vsly / 1e6, color = intervention)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = int_cols, name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(
    title    = "Cumulative Economic Value of Deaths Averted in Indonesia (VSLY)",
    subtitle = "Primary elasticity (e = 0.8/1.2), discounted at r = 3%",
    x        = "Year",
    y        = "Cumulative VSLY value (millions USD PPP)"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "aim1_econ_vsly_ts.png"), plot = g,
       width = 10, height = 5, dpi = 300)
```

## Figure --- Sensitivity Analysis: Income Elasticity

```{r econ-sensitivity, echo=FALSE, message=FALSE, warning=FALSE, eval=econ_ok, fig.width=10, fig.height=5}
econ_sens <- dt_econ[
  location == "Indonesia",
  .(
    `e = 1.0 (low)`       = sum(vsly_value_e1_0 * disc_r3, na.rm = TRUE),
    `e = 0.8/1.2 (primary)` = sum(vsly_value_e1_2_disc_r3, na.rm = TRUE),
    `e = 1.5 (high)`      = sum(vsly_value_e1_5 * disc_r3, na.rm = TRUE)
  ),
  by = intervention
]
setorder(econ_sens, intervention)

econ_sens_long <- melt(
  econ_sens,
  id.vars = "intervention",
  variable.name = "Elasticity",
  value.name = "VSLY_value"
)

g <- ggplot(econ_sens_long,
            aes(x = VSLY_value / 1e6, y = intervention, color = Elasticity)) +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c(
    "e = 1.0 (low)"          = "#4393c3",
    "e = 0.8/1.2 (primary)"  = "#2166ac",
    "e = 1.5 (high)"         = "#d6604d"
  )) +
  scale_x_continuous(labels = comma) +
  labs(
    title    = "VSLY Sensitivity to Income Elasticity — Indonesia (2026–2050)",
    subtitle = "Cumulative discounted VSLY value by intervention and elasticity assumption",
    x        = "Cumulative VSLY value (millions USD PPP)",
    y        = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom",
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(wd_outp, "aim1_econ_sensitivity.png"), plot = g,
       width = 10, height = 5, dpi = 300)
```

---

# Excel Export

```{r excel-export, message=FALSE, warning=FALSE, echo=FALSE}
wb <- createWorkbook()

add_sheet <- function(wb, sheet_name, dat, title_str) {
  addWorksheet(wb, sheet_name)
  writeData(wb, sheet_name, title_str, startRow = 1, startCol = 1)
  addStyle(wb, sheet_name,
           style = createStyle(fontSize = 12, textDecoration = "bold"),
           rows = 1, cols = 1)
  writeData(wb, sheet_name, dat, startRow = 3, startCol = 1,
            headerStyle = createStyle(textDecoration = "bold",
                                      fgFill = "#2E75B6", fontColour = "white",
                                      border = "Bottom"))
  setColWidths(wb, sheet_name, cols = seq_len(ncol(dat)), widths = "auto")
}

add_sheet(wb, "T1_Deaths_by_Intervention", as.data.frame(dt_excel_tab1),
          "Table 1. Cumulative deaths averted in Indonesia (2025-2050) by intervention scenario.")
add_sheet(wb, "T2_Deaths_by_Cause", as.data.frame(dt_excel_tab2),
          "Table 2. Deaths averted (thousands) by intervention and cause of death.")
add_sheet(wb, "T3_Deaths_by_Age", as.data.frame(dt_excel_tab3),
          "Table 3. Deaths averted by age group and intervention.")

out_xlsx <- paste0(wd_scen, "aim1_results_tables.xlsx")
saveWorkbook(wb, out_xlsx, overwrite = TRUE)
message("Tables exported to: ", out_xlsx)
```

---

# SLIDE DECK OUTPUTS
# Single source of truth for aim1_executive_slides.Rmd
# Runs after all prior sections; all key objects (dt_merged, dt_annual, dt_cumul,
# asmr_int, total_all, total_bp, total_stat) are
# already in the environment. Writes to output/slides/.

```{r slides-setup, message=FALSE, warning=FALSE, echo=FALSE}
slides_dir <- paste0(wd_outp, "slides/")
if (!dir.exists(slides_dir)) dir.create(slides_dir, recursive = TRUE)
message("Slide artefacts will be written to: ", slides_dir)
```

```{r slides-scalars, message=FALSE, warning=FALSE, echo=FALSE}
# SL-01: Scalar bundle
bau_total_deaths <- data.out[intervention == "b.a.u" & year >= 2025,
                              sum(dead, na.rm = TRUE)]
sl_scalars <- list(
  study_years      = study_years,
  total_all        = total_all,
  total_bp         = total_bp,
  total_salt       = total_salt,
  total_tfa        = total_tfa,
  total_stat       = total_stat,
  bau_total_deaths = bau_total_deaths
)
saveRDS(sl_scalars, paste0(slides_dir, "sl_scalars.rds"))
message("SL-01 saved: sl_scalars.rds")
```

```{r slides-tab-summary, message=FALSE, warning=FALSE, echo=FALSE}
# SL-02: Cumulative summary table with formatted display columns
sl_tab_summary <- copy(dt_cumul)[intervention != "b.a.u"]
sl_tab_summary[, pct    := deaths_delayed / bau_total_deaths * 100]
sl_tab_summary[, cum_K  := paste0(fmt_k(deaths_delayed), " K")]
sl_tab_summary[, avg_yr := paste0(fmt_k(deaths_delayed / 25), " K")]
sl_tab_summary[, pct_s  := paste0(round(pct, 1), "%")]
sl_tab_summary[, intervention := factor(intervention, levels = int_order)]
setorder(sl_tab_summary, intervention)
saveRDS(sl_tab_summary, paste0(slides_dir, "sl_tab_summary.rds"))
message("SL-02 saved: sl_tab_summary.rds")
```

```{r slides-dt-cause, message=FALSE, warning=FALSE, echo=FALSE}
# SL-05: Deaths by cause x intervention
sl_dt_cause <- dt_merged[intervention %in% int_order & cause %in% names(cause_map_full),
                          .(averted = sum(deaths_delayed, na.rm = TRUE)),
                          by = .(cause, intervention)]
sl_dt_cause[, cause_lbl := fcase(
  cause == "ihd",     "IHD",
  cause == "istroke", "Ischemic stroke",
  cause == "hstroke", "Hemorrhagic stroke",
  cause == "hhd",     "Hypertensive HD",
  default = cause
)]
sl_dt_cause[, intervention := factor(intervention, levels = int_order)]
saveRDS(sl_dt_cause, paste0(slides_dir, "sl_dt_cause.rds"))
message("SL-05 saved: sl_dt_cause.rds")
```

```{r slides-dt-age, message=FALSE, warning=FALSE, echo=FALSE}
# SL-06: Deaths by age group x intervention
sl_age_raw <- data.out[year >= 2025,
                        .(dead = sum(dead, na.rm = TRUE)),
                        by = .(year, cause, intervention, age_group)]
sl_bau_ag  <- sl_age_raw[intervention == "b.a.u",
                          .(bau = sum(dead, na.rm = TRUE)),
                          by = .(year, cause, age_group)]
sl_int_ag  <- sl_age_raw[intervention %in% int_order,
                          .(int = sum(dead, na.rm = TRUE)),
                          by = .(year, cause, intervention, age_group)]
sl_dt_age  <- merge(sl_int_ag, sl_bau_ag, by = c("year", "cause", "age_group"))
sl_dt_age  <- sl_dt_age[, .(averted = sum(bau - int, na.rm = TRUE)),
                         by = .(age_group, intervention)]
sl_dt_age[, intervention := factor(intervention, levels = int_order)]
saveRDS(sl_dt_age, paste0(slides_dir, "sl_dt_age.rds"))
message("SL-06 saved: sl_dt_age.rds")
```

```{r slides-dt-asmr, message=FALSE, warning=FALSE, echo=FALSE}
# SL-07: ASMR time series
saveRDS(copy(asmr_int), paste0(slides_dir, "sl_dt_asmr.rds"))
message("SL-07 saved: sl_dt_asmr.rds")
```

```{r slides-dt-annual, message=FALSE, warning=FALSE, echo=FALSE}
# SL-08: Annual deaths-averted time series
saveRDS(copy(dt_annual)[intervention %in% int_order],
        paste0(slides_dir, "sl_dt_annual.rds"))
message("SL-08 saved: sl_dt_annual.rds")
```

```{r slides-fig-baseline-trend, message=FALSE, warning=FALSE, echo=FALSE}
# SL-09: Baseline disease trend PNG (slide-optimised dimensions)
dt_trend_sl <- data.out[intervention == "b.a.u" & year >= 2019,
                         .(new_cases = sum(newcases, na.rm = TRUE),
                           sick      = sum(sick,     na.rm = TRUE)),
                         by = year]
dt_melt_sl  <- melt(dt_trend_sl, id.vars = "year",
                    measure.vars  = c("new_cases", "sick"),
                    variable.name = "metric", value.name = "count")
dt_melt_sl[, metric_lbl := fifelse(metric == "new_cases",
                                    "Incident cases", "Prevalent cases")]

g_trend <- ggplot(dt_melt_sl,
                   aes(x = year, y = count / 1e6,
                       color = metric_lbl, linetype = year > 2019)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = c("Incident cases"  = "steelblue",
                                "Prevalent cases" = "tomato"), name = NULL) +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"),
                        guide = "none") +
  scale_y_continuous(labels = comma) +
  labs(x = "Year", y = "Persons (millions)",
       subtitle = "Solid = calibration; dashed = projection") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.subtitle = element_text(color = "grey40", size = 9))
ggsave(paste0(slides_dir, "sl_fig_baseline_trend.png"), plot = g_trend,
       width = 8, height = 4, dpi = 200)
message("SL-09 saved: sl_fig_baseline_trend.png")
```

```{r slides-fig-asmr, message=FALSE, warning=FALSE, echo=FALSE}
# SL-10: ASMR by intervention
g_asmr_sl <- ggplot(asmr_int,
                     aes(x = year, y = ASMR, color = int_lbl,
                         linetype = year > 2019)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = c("Business-as-Usual" = "#999999", int_cols),
                     name = NULL) +
  scale_linetype_manual(values = c(`FALSE` = "solid", `TRUE` = "dashed"),
                        guide = "none") +
  labs(x = "Year", y = "ASMR (per 100,000)",
       subtitle = "Solid = calibration; dashed = projection") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        panel.grid.minor = element_blank(),
        plot.subtitle = element_text(color = "grey40", size = 8)) +
  guides(color = guide_legend(nrow = 2))
ggsave(paste0(slides_dir, "sl_fig_asmr.png"), plot = g_asmr_sl,
       width = 8, height = 4, dpi = 200)
message("SL-10 saved: sl_fig_asmr.png")
```

```{r slides-fig-deaths-averted, message=FALSE, warning=FALSE, echo=FALSE}
# SL-11: Annual deaths averted area chart (individual interventions only)
g_ann_sl <- ggplot(dt_annual[intervention %in% int_order_individual],
                    aes(x = year, y = deaths_delayed / 1e3, fill = intervention)) +
  geom_area(alpha = 0.85, position = "identity") +
  scale_fill_manual(values = int_cols[int_order_individual], name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(x = "Year", y = "Deaths averted (thousands)") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(nrow = 2))
ggsave(paste0(slides_dir, "sl_fig_deaths_averted.png"), plot = g_ann_sl,
       width = 8, height = 4, dpi = 200)
message("SL-11 saved: sl_fig_deaths_averted.png")
```

```{r slides-fig-by-cause, message=FALSE, warning=FALSE, echo=FALSE}
# SL-12: Deaths averted by cause
g_cause_sl <- ggplot(sl_dt_cause,
                      aes(x = cause_lbl, y = averted / 1e3, fill = intervention)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = int_cols[int_order], name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(x = NULL, y = "Deaths averted (thousands)") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 9)) +
  guides(fill = guide_legend(nrow = 2))
ggsave(paste0(slides_dir, "sl_fig_by_cause.png"), plot = g_cause_sl,
       width = 8, height = 4, dpi = 200)
message("SL-12 saved: sl_fig_by_cause.png")
```

```{r slides-fig-by-age, message=FALSE, warning=FALSE, echo=FALSE}
# SL-13: Deaths averted by age group
g_age_sl <- ggplot(sl_dt_age,
                    aes(x = age_group, y = averted / 1e3, fill = intervention)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = int_cols[int_order], name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(x = "Age group", y = "Deaths averted (thousands)") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 7),
        panel.grid.minor = element_blank()) +
  guides(fill = guide_legend(nrow = 2))
ggsave(paste0(slides_dir, "sl_fig_by_age.png"), plot = g_age_sl,
       width = 8, height = 4, dpi = 200)
message("SL-13 saved: sl_fig_by_age.png")
```

```{r slides-fig-cumulative-time, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=5}
# SL-15: Cumulative deaths averted by individual intervention
g <- ggplot(dt_annual[intervention %in% int_order_individual],
            aes(x = year, y = cum_deaths_delayed / 1e3, color = intervention)) +
  geom_line(linewidth = 1.1) +
  scale_color_manual(values = int_cols, name = NULL) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "",
    x     = "Year",
    y     = "Cumulative deaths averted (thousands)"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
print(g)
ggsave(paste0(slides_dir, "sl_fig_cum_averted.png"), plot = g,
       width = 11, height = 5, dpi = 300)
message("SL-15 saved: sl_fig_cum_averted.png")
```

```{r slides-econ, message=FALSE, warning=FALSE, echo=FALSE}
# SL-ECON: Economic-valuation artefacts for the executive deck.
# Consumes outputs of code/cvd/08_economic_value_calculation.R.

econ_files <- c(
  vsl_prim  = paste0(wd_outp, "08_vsl_summary_table_e1_2_primary.rds"),
  vsly_prim = paste0(wd_outp, "08_vsly_summary_table_e1_2_primary.rds"),
  appended  = paste0(wd_outp, "08_vsl_vsly_summary_table_appended.rds")
)
econ_slide_ok <- all(file.exists(econ_files))

if (econ_slide_ok) {

  econ_scen_lbl_slides <- c(
    bp_only       = "Antihypertensive Therapy",
    sodium_only   = "Sodium Reduction",
    tfa_only      = "Trans-Fat Elimination",
    statins_only  = "Statin Therapy",
    bp_sodium     = "HTN + Sodium",
    bp_sodium_tfa = "HTN + Sodium + TFA",
    all_four      = "All Interventions"
  )

  # ---- Region x scenario tables (primary, e1_2) ----------------------------
  build_econ_tab <- function(path) {
    dt <- as.data.table(readRDS(path))
    dt <- dt[scenario %in% names(econ_scen_lbl_slides)]
    dt[, intervention := factor(econ_scen_lbl_slides[scenario],
                                levels = int_order)]
    setorder(dt, intervention, who_region)
    dt
  }

  sl_econ_vsl_primary  <- build_econ_tab(econ_files[["vsl_prim"]])
  sl_econ_vsly_primary <- build_econ_tab(econ_files[["vsly_prim"]])

  saveRDS(sl_econ_vsl_primary,  paste0(slides_dir, "sl_econ_vsl_primary.rds"))
  saveRDS(sl_econ_vsly_primary, paste0(slides_dir, "sl_econ_vsly_primary.rds"))

  # ---- Headline scalar: Indonesia VSLY share of income (cumulative) --------
  idn_vsly_all <- sl_econ_vsly_primary[
    who_region == "South-East Asia" & intervention == "All Interventions"
  ]
  idn_vsl_all  <- sl_econ_vsl_primary[
    who_region == "South-East Asia" & intervention == "All Interventions"
  ]

  # Fall back to World if South-East Asia not available
  if (nrow(idn_vsly_all) == 0)
    idn_vsly_all <- sl_econ_vsly_primary[
      who_region == "World" & intervention == "All Interventions"
    ]
  if (nrow(idn_vsl_all) == 0)
    idn_vsl_all <- sl_econ_vsl_primary[
      who_region == "World" & intervention == "All Interventions"
    ]

  sl_econ_scalars <- list(
    idn_vsly_share_total = if (nrow(idn_vsly_all)) idn_vsly_all$share_total else NA_real_,
    idn_vsl_share_total  = if (nrow(idn_vsl_all))  idn_vsl_all$share_total  else NA_real_,
    idn_vsly_value_total = if (nrow(idn_vsly_all)) idn_vsly_all$metric_total else NA_real_,
    idn_vsl_value_total  = if (nrow(idn_vsl_all))  idn_vsl_all$metric_total  else NA_real_,
    base_year              = 2026L,
    horizon_end            = 2050L,
    elasticity_primary     = "Differential (e=0.8 HIC / 1.2 LMIC), Robinson et al. 2019",
    discount_rate          = 0.03
  )
  saveRDS(sl_econ_scalars, paste0(slides_dir, "sl_econ_scalars.rds"))

  # ---- Dot plot: VSL/VSLY share of income, by period -----------------------
  dt_app <- as.data.table(readRDS(econ_files[["appended"]]))
  dt_app <- dt_app[scenario %in% names(econ_scen_lbl_slides)]
  dt_app[, intervention := factor(econ_scen_lbl_slides[scenario], levels = int_order)]

  share_cols <- grep("^share_", names(dt_app), value = TRUE)

  sl_econ_dotplot_dt <- melt(
    dt_app,
    id.vars       = c("valuation_type", "elasticity_case", "who_region",
                      "scenario", "intervention"),
    measure.vars  = share_cols,
    variable.name = "period",
    value.name    = "share_income"
  )

  sl_econ_dotplot_dt[, period := fcase(
    period == "share_2026",  "2026",
    period == "share_2030",  "2030",
    period == "share_2040",  "2040",
    period == "share_2050",  "2050",
    period == "share_total", "2026–2050\n(cumulative)",
    default = as.character(period)
  )]
  sl_econ_dotplot_dt[, period := factor(period,
    levels = c("2026", "2030", "2040", "2050", "2026–2050\n(cumulative)")
  )]
  sl_econ_dotplot_dt[, elasticity_label := fcase(
    elasticity_case == "e1_2_primary",     "Primary (e = 0.8/1.2)",
    elasticity_case == "e1_5_sensitivity", "Sensitivity (e = 1.5)",
    default = elasticity_case
  )]

  saveRDS(sl_econ_dotplot_dt, paste0(slides_dir, "sl_econ_dotplot_dt.rds"))

  g_econ_dot <- ggplot(
    sl_econ_dotplot_dt[valuation_type == "VSLY"],
    aes(x = share_income, y = intervention,
        colour = elasticity_label)
  ) +
    geom_point(size = 2.5, alpha = 0.85,
               position = position_dodge(width = 0.55)) +
    facet_grid(who_region ~ period, scales = "free_x") +
    scale_x_continuous(
      labels = percent_format(accuracy = 0.1),
      expand = expansion(mult = c(0.05, 0.15))
    ) +
    scale_colour_manual(
      name   = "Elasticity case",
      values = c("Primary (e = 0.8/1.2)" = "#2166ac",
                 "Sensitivity (e = 1.5)" = "#d6604d")
    ) +
    labs(
      title    = "Economic value of avoidable CVD mortality in Indonesia, proportion of income",
      subtitle = "VSLY only | Discounted at r = 3% (Robinson et al. 2019)",
      x        = "Proportion of discounted GNI",
      y        = NULL
    ) +
    theme_bw(base_size = 9) +
    theme(
      strip.text.y       = element_text(angle = 0, hjust = 0, face = "bold"),
      strip.text.x       = element_text(face = "bold"),
      strip.background   = element_rect(fill = "#f0f0f0", colour = "grey70"),
      panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      legend.position    = "bottom",
      legend.box         = "horizontal",
      plot.title         = element_text(face = "bold"),
      axis.text.y        = element_text(size = 8)
    )

  ggsave(paste0(slides_dir, "sl_econ_dotplot.png"),
         plot = g_econ_dot, width = 11, height = 6.2, dpi = 220)
  message("SL-ECON saved: sl_econ_vsl_primary.rds, sl_econ_vsly_primary.rds, ",
          "sl_econ_scalars.rds, sl_econ_dotplot_dt.rds, sl_econ_dotplot.png")
} else {
  warning("Economic-valuation inputs not found in ", wd_outp,
          " — skipping SL-ECON artefacts. Run code/cvd/08_economic_value_calculation.R first.")
}
```

```{r slides-manifest, message=FALSE, warning=FALSE, echo=FALSE}
# SL-16: Artefact manifest
sl_manifest <- data.table(
  file = c(
    "sl_scalars.rds", "sl_tab_summary.rds",
    "sl_dt_cause.rds", "sl_dt_age.rds",
    "sl_dt_asmr.rds", "sl_dt_annual.rds",
    "sl_fig_baseline_trend.png", "sl_fig_asmr.png",
    "sl_fig_deaths_averted.png", "sl_fig_by_cause.png",
    "sl_fig_by_age.png",
    "sl_fig_cum_averted.png",
    "sl_econ_vsl_primary.rds", "sl_econ_vsly_primary.rds",
    "sl_econ_scalars.rds", "sl_econ_dotplot_dt.rds",
    "sl_econ_dotplot.png"
  ),
  type = c(rep("rds", 6), rep("png", 6), rep("rds", 4), "png"),
  consumed_by = c(
    "inline scalars", "tab-summary kable",
    "fig-by-cause ggplot", "fig-by-age ggplot",
    "fig-asmr ggplot", "fig-deaths-averted ggplot",
    "fig-baseline-trend include_graphics", "fig-asmr include_graphics (fallback)",
    "fig-deaths-averted include_graphics (fallback)",
    "fig-by-cause include_graphics (fallback)",
    "fig-by-age include_graphics (fallback)",
    "fig-cum-averted include_graphics",
    "econ-regional-table kable",
    "econ-regional-table kable (VSLY headline)",
    "inline scalars (Indonesia VSLY share)",
    "econ-dotplot ggplot (alt source)",
    "econ-dotplot include_graphics"
  )
)
saveRDS(sl_manifest, paste0(slides_dir, "sl_manifest.rds"))

kable(sl_manifest,
      caption = "Slide deck artefact manifest. All files written to output/slides/.",
      col.names = c("File", "Type", "Consumed by (slides chunk)")) |>
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = TRUE, font_size = 11) |>
  row_spec(0, bold = TRUE)

message(sprintf("SLIDE DECK OUTPUTS complete. %d artefacts written to %s",
                nrow(sl_manifest), slides_dir))

```

---

# PAPER OUTPUTS
# Artefacts consumed by who_cvd_targets_paper1.Rmd
# Extends slide artefacts with additional scalars and tables needed for the manuscript.

```{r paper-setup, message=FALSE, warning=FALSE, echo=FALSE}
paper_dir <- paste0(wd_outp, "paper/")
if (!dir.exists(paper_dir)) dir.create(paper_dir, recursive = TRUE)
message("Paper artefacts will be written to: ", paper_dir)
```

```{r paper-scalars, message=FALSE, warning=FALSE, echo=FALSE}
# ---- Cause-specific deaths averted (All Interventions) ----------------------
cause_averted_all <- dt_merged[intervention == "All Interventions",
                                .(averted = sum(deaths_delayed, na.rm = TRUE)),
                                by = cause]

# ---- Cause-specific deaths averted by intervention --------------------------
cause_averted_bp <- dt_merged[intervention == "Antihypertensive Therapy",
                               .(averted = sum(deaths_delayed, na.rm = TRUE)),
                               by = cause]
cause_averted_stat <- dt_merged[intervention == "Statin Therapy",
                                 .(averted = sum(deaths_delayed, na.rm = TRUE)),
                                 by = cause]

# ---- BAU cause-specific deaths -----------------------------------------------
bau_by_cause <- data.out[intervention == "b.a.u" & year >= 2025,
                          .(bau_deaths = sum(dead, na.rm = TRUE)),
                          by = cause]

# ---- Percent of BAU averted --------------------------------------------------
pct_all  <- round(total_all  / bau_total_deaths * 100, 1)
pct_bp   <- round(total_bp   / bau_total_deaths * 100, 1)
pct_stat <- round(total_stat / bau_total_deaths * 100, 1)

# ---- Annual deaths averted at key timepoints ---------------------------------
annual_2030_all <- dt_annual[intervention == "All Interventions" & year == 2030,
                              deaths_delayed]
annual_2050_all <- dt_annual[intervention == "All Interventions" & year == 2050,
                              deaths_delayed]
annual_2030_bp  <- dt_annual[intervention == "Antihypertensive Therapy" & year == 2030,
                              deaths_delayed]
annual_2050_bp  <- dt_annual[intervention == "Antihypertensive Therapy" & year == 2050,
                              deaths_delayed]
annual_2030_stat <- dt_annual[intervention == "Statin Therapy" & year == 2030,
                               deaths_delayed]
annual_2050_stat <- dt_annual[intervention == "Statin Therapy" & year == 2050,
                               deaths_delayed]

# ---- Age-group totals (All Interventions) ------------------------------------
age_averted_all <- dt_merged[intervention == "All Interventions",
                              .(averted = sum(deaths_delayed, na.rm = TRUE)),
                              by = age_group]

# ---- Sex totals (All Interventions) ------------------------------------------
sex_averted_all <- dt_merged[intervention == "All Interventions",
                              .(averted = sum(deaths_delayed, na.rm = TRUE)),
                              by = sex]

# ---- BOD summary scalars (All Interventions) ---------------------------------
bod_all <- bod_tab1[intervention == "All Interventions"]
total_ylls_all  <- bod_all$YLLs
total_ylds_all  <- bod_all$YLDs
total_dalys_all <- bod_all$DALYs

bod_bp <- bod_tab1[intervention == "Antihypertensive Therapy"]
total_dalys_bp <- if (nrow(bod_bp) > 0) bod_bp$DALYs else NA_real_

bod_stat <- bod_tab1[intervention == "Statin Therapy"]
total_dalys_stat <- if (nrow(bod_stat) > 0) bod_stat$DALYs else NA_real_

# ---- Save comprehensive paper scalar bundle ----------------------------------
paper_scalars <- list(
  # Core counts
  study_years      = study_years,
  bau_total_deaths = bau_total_deaths,
  total_all        = total_all,
  total_bp         = total_bp,
  total_salt       = total_salt,
  total_tfa        = total_tfa,
  total_stat       = total_stat,

  # Percentages of BAU
  pct_all  = pct_all,
  pct_bp   = pct_bp,
  pct_stat = pct_stat,

  # Annual deaths averted at key years
  annual_2030_all  = annual_2030_all,
  annual_2050_all  = annual_2050_all,
  annual_2030_bp   = annual_2030_bp,
  annual_2050_bp   = annual_2050_bp,
  annual_2030_stat = annual_2030_stat,
  annual_2050_stat = annual_2050_stat,

  # Cause-specific (All Interventions)
  averted_ihd_all     = cause_averted_all[cause == "ihd",     averted],
  averted_istroke_all = cause_averted_all[cause == "istroke", averted],
  averted_hstroke_all = cause_averted_all[cause == "hstroke", averted],
  averted_hhd_all     = cause_averted_all[cause == "hhd",     averted],

  # Cause-specific (Antihypertensive)
  averted_ihd_bp     = cause_averted_bp[cause == "ihd",     averted],
  averted_istroke_bp = cause_averted_bp[cause == "istroke", averted],
  averted_hstroke_bp = cause_averted_bp[cause == "hstroke", averted],
  averted_hhd_bp     = cause_averted_bp[cause == "hhd",     averted],

  # Cause-specific (Statins)
  averted_ihd_stat     = cause_averted_stat[cause == "ihd",     averted],
  averted_istroke_stat = cause_averted_stat[cause == "istroke", averted],

  # Age-group (All Interventions)
  averted_lt70  = age_averted_all[age_group == "<70",   averted],
  averted_70_79 = age_averted_all[age_group == "70-79", averted],
  averted_80_89 = age_averted_all[age_group == "80-89", averted],
  averted_90p   = age_averted_all[age_group == "90+",   averted],

  # Sex (All Interventions)
  averted_male   = sex_averted_all[sex == "Male",   averted],
  averted_female = sex_averted_all[sex == "Female", averted],

  # BOD summary
  total_ylls_all  = total_ylls_all,
  total_ylds_all  = total_ylds_all,
  total_dalys_all = total_dalys_all,
  total_dalys_bp  = total_dalys_bp,
  total_dalys_stat = total_dalys_stat
)

saveRDS(paper_scalars, paste0(paper_dir, "paper_scalars.rds"))
message("Paper scalars saved: paper_scalars.rds")
```

```{r paper-tables, message=FALSE, warning=FALSE, echo=FALSE}
# Save key data tables for the manuscript
saveRDS(copy(dt_cumul),     paste0(paper_dir, "paper_dt_cumul.rds"))
saveRDS(copy(dt_tab2),      paste0(paper_dir, "paper_dt_cause.rds"))
saveRDS(copy(bod_tab1),     paste0(paper_dir, "paper_bod_summary.rds"))
saveRDS(copy(bod_tab2),     paste0(paper_dir, "paper_bod_cause.rds"))
saveRDS(copy(dt_annual),    paste0(paper_dir, "paper_dt_annual.rds"))
saveRDS(copy(dt_age_sex),   paste0(paper_dir, "paper_dt_age_sex.rds"))
message("Paper tables saved to: ", paper_dir)
```

```{r export-slides-bod-scalars, message=FALSE, warning=FALSE, echo=FALSE}
# Exports burden of disease scalars for aim1_executive_slides.RMD

sl_scalars$total_dalys_all  <- total_dalys_all
sl_scalars$total_ylls_all   <- total_ylls_all
sl_scalars$total_ylds_all   <- total_ylds_all
sl_scalars$total_dalys_bp   <- total_dalys_bp
sl_scalars$total_dalys_stat <- total_dalys_stat

saveRDS(sl_scalars, paste0(slides_dir, "sl_scalars.rds"))
message("SL-01 updated: sl_scalars.rds (with BOD scalars)")

# Create and save sl_bod_summary
sl_bod_summary <- copy(bod_tab1)
saveRDS(sl_bod_summary, paste0(slides_dir, "sl_bod_summary.rds"))
message("SL-17 saved: sl_bod_summary.rds")
```

---

*Report generated on `r Sys.time()` using R `r R.version$major`.`r R.version$minor`.*
