---
title: "Cancer Modeling Inputs and Calculations"
author: "Mohamed Albirair"
date: "`r Sys.Date()`"
output:
      html_document:
            toc: true
            toc_float:
                  collapsed: false
            toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r}
## load packages
pacman::p_load(tidyverse)

# Load settings
z <- yaml::read_yaml("../../settings.yml")

# Home folder
# z <- list()
z$home_dir <- "../../"

# R code, inputs, and outputs
if (!exists("root")){
      root <- z$home_dir
}

z$fnx_dir     <- paste0(root, "/fnx")
z$inputs_dir  <- paste0(root, "/input")
z$outputs_dir <- paste0(root, "/outputs")
z$scripts_dir <- paste0(root, "/scripts")

# Source functions
list.files(path       = z$fnx_dir,
           pattern    = "\\.R$",
           full.names = TRUE) %>%
      walk(.f        = safely(.f = ~ source(.x, echo = FALSE)),
           .progress = TRUE)

# URLs
# `not it`
source(z$not_in)

# ggplot theme_caviz
source(z$theme_caviz)
## Basic scenario-defining input values
load("../inputs/all-baseline-data.RData")
load("../inputs/RData/Uganda/bsln/Uganda_bsln_input.RData")
```

```{r}
scrollable <- function(df) {
      # https://stackoverflow.com/questions/45849433/horizontal-scrolling-in-rmarkdown
      kableExtra::kable(df, "html") %>% 
            kableExtra::kable_styling("striped") %>% 
            kableExtra::scroll_box(height = "600px")
}
```

## Baseline Stage Distribution

```{r}
bsln_stg_dist %>% scrollable()
```

## GBD Epi Data

```{r}
country_epi %>% 
      map_depth(.x     = .,
                .depth = 2,
                .f     = function(df) {
                      df[as.character(reference_year)]
                }) %>% 
      imap_dfr(., function(cause_data, cause_name) {
            imap_dfr(cause_data, function(sex_data, sex_name) {
                  imap_dfr(sex_data, function(year_data, year_name) {
                        year_data %>% 
                              mutate(cause = cause_name,
                                     sex   = sex_name,
                                     year  = year_name)
                  })
            })
      }) %>% scrollable()
```

## GBD All-Cause Mortality

```{r}
country_acm %>% 
      map_depth(.x     = .,
                .depth = 2,
                .f     = function(df) {
                      df[as.character(reference_year)]
                }) %>% 
      imap_dfr(., function(cause_data, cause_name) {
            imap_dfr(cause_data, function(sex_data, sex_name) {
                  imap_dfr(sex_data, function(year_data, year_name) {
                        year_data %>% 
                              mutate(cause = cause_name,
                                     sex   = sex_name,
                                     year  = year_name)
                  })
            })
      }) %>% scrollable()
```

## GBD Fertility Data

```{r}
country_fert %>% 
      map_depth(.x     = .,
                .depth = 1,
                .f     = function(df) {
                      df[as.character(reference_year)]
                }) %>% 
      imap_dfr(., function(sex_data, sex_name) {
            imap_dfr(sex_data, function(year_data, year_name) {
                  year_data %>% 
                        mutate(sex   = sex_name,
                               year  = year_name)
            })
      }) %>% scrollable()
```

## No. of Cancer Cases

```{r}
n_cases %>% 
      imap_dfr(., function(cause_data, cause_name) {
            imap_dfr(cause_data, function(sex_data, sex_name) {
                  sex_data %>% 
                        mutate(cause = cause_name,
                               sex   = sex_name)
            })
      }) %>% scrollable()
```

## No. and Rate of Cancer-Specific Mortality

```{r}
n_cmx_deaths %>% 
      imap_dfr(., function(cause_data, cause_name) {
            imap_dfr(cause_data, function(sex_data, sex_name) {
                  sex_data %>% 
                        mutate(cause = cause_name,
                               sex   = sex_name)
            })
      }) %>% scrollable()
```

## No. and Rate of All-Cause Mortality + Total Population Counts

```{r}
n_mx_N_acm %>% 
      imap_dfr(., function(cause_data, cause_name) {
            imap_dfr(cause_data, function(sex_data, sex_name) {
                  sex_data %>% 
                        mutate(cause = cause_name,
                               sex   = sex_name)
            })
      }) %>% scrollable()
```

## Markov Correction for All-Cause and Cancer-Specific Mortality Data
```{r}
markov_correction %>% 
      filter(cause %in% c(cause1[2], cause2[2], cause3[2])) %>% 
      scrollable()
```

## Starting Health State Distribution

```{r}
start_stg_dist %>% 
      imap_dfr(., function(loc_data, loc_name) {
            imap_dfr(loc_data, function(cause_data, cause_name) {
                  imap_dfr(cause_data, function(sex_data, sex_name) {
                        imap_dfr(sex_data, function(year_data, year_name) {
                              imap_dfr(year_data, function(age_data, age_name) {
                                    age_data %>% 
                                          mutate(loc   = loc_name,
                                                 cause = cause_name,
                                                 sex   = sex_name,
                                                 year  = year_name,
                                                 age   = age_name)
                              })    
                        })
                  })
            })
      }) %>% scrollable()
```


## Transition Probability Matrices

### Bladder Cancer (Category 1)

```{r}
bsln_tpms$Uganda$`Bladder cancer`$Female$`2019`$`45`
```

### Breast Cancer (Category 2)

```{r}
bsln_tpms$Uganda$`Breast cancer`$Female$`2019`$`45`
```

### CRC (Category 3)

```{r}
bsln_tpms$Uganda$`Colon and rectum cancer`$Female$`2019`$`45`
```


