---
title: "Cancer Modeling Inputs and Calculations"
author: "Mohamed Albirair"
date: "`r Sys.Date()`"
output:
      html_document:
            toc: true
            toc_float:
                  collapsed: false
            toc_depth: 3
            number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r, eval=TRUE}
## load packages
pacman::p_load(tidyverse)

# Load settings
z <- yaml::read_yaml("../../settings.yml")

# Home folder
# z <- list()
z$home_dir <- "../"

# R code, inputs, and outputs
if (!exists("root")){
      root <- z$home_dir
}

z$fnx_dir     <- paste0(root, "fnx")
z$inputs_dir  <- paste0(root, "input")
z$outputs_dir <- paste0(root, "outputs")
z$scripts_dir <- paste0(root, "scripts")

# Fast sourcing
loc_fnx <- list.files(path       = z$fnx_dir,
                      pattern    = "\\.R$",
                      full.names = TRUE)

for (i in loc_fnx) {
      source(i, echo = FALSE)
}

# URLs
# `not it`
source(z$not_in)

# ggplot theme_caviz
source(z$theme_caviz)
## Basic scenario-defining input values
# load("../inputs/all-baseline-data.RData")
# load("../inputs/RData/Uganda/bsln/Uganda_bsln_input.RData")
```

```{r, eval=TRUE}
scrollable <- function(df) {
      # https://stackoverflow.com/questions/45849433/horizontal-scrolling-in-rmarkdown
      kableExtra::kable(df, "html") %>% 
            kableExtra::kable_styling("striped") %>% 
            kableExtra::scroll_box(height = "600px")
}
```

# Data Prep

```{r, eval=TRUE}
# Create location mapping vector for fast lookup
location_map <- c(
      "Bolivia (Plurinational State of)"      = "Bolivia",
      "Brunei Darussalam"                     = "Brunei",
      "Cabo Verde"                            = "Cape Verde",
      "Côte d'Ivoire"                         = "Cote d'Ivoire",
      "Czechia"                               = "Czech Republic",
      "Federated States of Micronesia"        = "Micronesia",
      "Republic of Moldova"                   = "Moldova",
      "Micronesia (Federated States of)"      = "Micronesia",
      "Iran (Islamic Republic of)"            = "Iran",
      "Lao People's Democratic Republic"      = "Laos",
      "North Macedonia"                       = "Macedonia",
      "Democratic People's Republic of Korea" = "North Korea",
      "Republic of Korea"                     = "South Korea",
      "Syrian Arab Republic"                  = "Syria",
      "Swaziland"                             = "Eswatini",
      "United Republic of Tanzania"           = "Tanzania",
      "Taiwan (Province of China)"            = "Taiwan",
      "The Bahamas"                           = "Bahamas",
      "The Gambia"                            = "Gambia",
      "Venezuela (Bolivarian Republic of)"    = "Venezuela",
      "Viet Nam"                              = "Vietnam",
      "Virgin Islands, U.S."                  = "Virgin Islands",
      "United States"                         = "United States of America",
      "United States Virgin Islands"          = "Virgin Islands"
)

## Cancer Models

# Most basic model (well, sick, dead)
cause1 <- sort(c("Uterine cancer",
                 "Esophageal cancer",
                 "Thyroid cancer",
                 "Stomach cancer",
                 "Liver cancer",
                 "Pancreatic cancer",
                 "Ovarian cancer",
                 "Bladder cancer",
                 "Lip and oral cavity cancer",
                 "Nasopharynx cancer",
                 "Other pharynx cancer"))       # 11 cancer types

# Staged model (well, local, regional, distant, dead)
# Screening-amenable cancers
cause2 <- sort(c("Tracheal, bronchus, and lung cancer",
                 "Prostate cancer",
                 "Breast cancer"))

# Staged model with pre-cancerous lesions
cause3 <- sort(c("Cervical cancer",
                 "Colon and rectum cancer"))

# Female-only cancer types
female_cancers <- c("Cervical cancer", "Ovarian cancer", "Uterine cancer")
```

# Read Data

```{r}
## This is much faster way for loading the data--only keeping the chunks below for referencing

# Build a temporary environment and load the large object data
env <- new.env()
load("../inputs/gbd_dt.RData", envir = env)

# Return the value of the named object
gbd_pop             <- get("gbd_pop", envir = env)
gbd_pop_5yr         <- get("gbd_pop_5yr", envir = env)
gbd_epi_5yr         <- get("gbd_epi_5yr", envir = env)
age_models_dt       <- get("age_models_dt", envir = env)
age_models_pop_dt   <- get("age_models_pop_dt", envir = env)
gbd_epi_1yr_predict <- get("gbd_epi_1yr_predict", envir = env)
```

## GBD Pop

```{r, eval=FALSE}
## Population Counts Data
# https://vizhub.healthdata.org/gbd-results?params=gbd-api-2021-permalink/3931936f370fdaf18adfca2089680981

pop_zip_files <- list.files(path       = "../../Data/GBD 2021/Pop/",
                            pattern    = "\\.zip$",
                            full.names = TRUE)

gbd_pop_raw <- data.table::rbindlist(l         = lapply(X   = pop_zip_files,
                                                        FUN = read_csv_from_zip),
                                     fill      = TRUE,
                                     use.names = TRUE)

## Cancer-Specific Epidemiological Data
# Sep 2025: https://vizhub.healthdata.org/gbd-results?params=gbd-api-2021-permalink/df6164cc40d2ac450659826f45ba1661
# Additional <1 age groups: https://vizhub.healthdata.org/gbd-results?params=gbd-api-2021-permalink/2ed3b5e1505a7cfff44ad2730a065a97

zip_files <- list.files(path       = "../../Data/GBD 2021/16 cancer types/",
                        pattern    = "\\.zip$",
                        full.names = TRUE)

gbd_epi_raw <- data.table::rbindlist(l         = lapply(X   = zip_files,
                                                        FUN = read_csv_from_zip),
                                     fill      = TRUE,
                                     use.names = TRUE)
```

### GBD Pop: 1Y Age Groups

```{r, eval=FALSE}
# Convert to data.table for merging
location_map_dt <- data.table::data.table(
      old_name = names(location_map),
      new_name = unname(location_map)
)

# Prep a separate copy of target dt
gbd_pop <- data.table::copy(gbd_pop_raw)

gbd_pop <- gbd_pop[
      # Rename age to age_cat
      , `:=` (age_cat = age)
][
      # Convert age categories (much faster than case_when)
      , age := data.table::fcase(
            age_cat %in% c("<28 days",
                           "1-5 months",
                           "6-11 months"), 0.01,
            age_cat == "<1 year",          0.00,
            age_cat == "12-23 months",     1.00,
            age_cat == "2-4 years",        3.01,
            age_cat == "<5 years",         4.01,
            default = as.numeric(str_extract(age_cat, "\\d+"))
      )
][
      # Remove unwanted age groups
      age %!in% c(0.01, 3.01, 4.01),
][
      # Sort efficiently
      order(location, year, age)
]

gbd_pop[location_map_dt, location := i.new_name, on = .(location = old_name)]
```

### GBD Pop: 5Y Age Groups

```{r, eval=FALSE}
data.table::setDT(gbd_pop)

gbd_pop_5yr <- gbd_pop[
      , .(pop = sum(val)),
      by = .(
            location, year, sex,
            age_gp = cut(
                  age,
                  breaks = c(-Inf, seq(5, 95, 5), Inf),
                  labels = c("<5 years",
                             paste0(seq(5, 90, 5), "-", seq(9, 94, 5), " years"),
                             "95+ years"),
                  right = FALSE
            )
      )
][
      , age := data.table::fifelse(
            age_gp == "<5 years", 0, as.numeric(str_extract(age_gp, "\\d+"))
      )
]
```

## GBD Epi Data

```{r, eval=FALSE}
# Prep a separate copy of target dt
gbd_epi_5yr <- data.table::copy(gbd_epi_raw)

gbd_epi_5yr <- gbd_epi_5yr[
      # Apply sex-specific filtering first to reduce data size
      (cause %in% female_cancers & sex == "Female") |
            (cause == cause2[2] & sex == "Male") |
            (!(cause %in% female_cancers) & cause != cause2[2])
][
      # Map locations
      , location := data.table::fcoalesce(location_map[location], location)
][
      # Rename age to age_cat
      , age_cat := age
][
      # Convert age categories (much faster than case_when)
      , age := data.table::fcase(
            age_cat %in% c("<28 days",
                           "1-5 months",
                           "6-11 months"), 0.01,
            age_cat == "<1 year",          0.10,
            age_cat == "12-23 months",     1.00,
            age_cat == "2-4 years",        3.01,
            age_cat == "<5 years",         0.00,
            default = as.numeric(str_extract(age, "^\\d+"))
      )
][
      # Remove unwanted age groups#--------------------------------------------# TBU
      age %!in% c(0.01, 0.10, 1.00, 3.01)
][
      , age_mid := data.table::fcase(
            age < 95, age + 2.5,
            default = age
      )
][
      , age_gp := cut(
            age,
            breaks = c(-Inf, seq(5, 95, 5), Inf),
            labels = c("<5 years",
                       paste0(seq(5, 90, 5), "-", seq(9, 94, 5), " years"),
                       "95+ years"),
            right = FALSE
      )
][
      , val := data.table::fifelse(is.na(val), 0, val)
][
      # Sort efficiently
      order(location, year, age, measure, metric)
]
```

### GBD Epi: 1Y Interpolation

#### Step 1: Build `log(Metric) ~ Age` Models

```{r, eval=FALSE}
age_models_dt <- gbd_epi_5yr[
      metric == "Rate" &
            year %in% c(z$sim_start_year, z$reference_year) &
            age_mid < 95,
      .(location, sex, year, cause, measure, age = age_mid,  # removed original age to run the model on age_mid
        log_val = log(data.table::fifelse(val == 0, 1e-16, val)))        # Converted log(val == 0)
][
      , .(model = list(lm(log_val ~ age, data = .SD))),
      by = .(location, sex, year, cause, measure)
]
```

```{r}
# Check:
nrow(age_models_dt) / 
      length(unique(age_models_dt$location)) /  # 204 locations
      nrow(count(age_models_dt, sex, cause)) /  # 28: 12 * 2 + 3 female + 1 male
      length(unique(age_models_dt$year)) /      # 2: 2000 sim start year & 2019 ref year
      length(unique(age_models_dt$measure))     # 3: incid, prev, mort
```

#### Step 2: Incorporate 1Y Age-Group Pop Counts

```{r, eval=FALSE}
age_models_pop_dt <- gbd_pop[
      year %in% c(z$sim_start_year, z$reference_year),
      .(location, sex, year, age, pop = val) # year
][
      age_models_dt,
      on = .(location, sex, year),
      allow.cartesian = TRUE  # Explicitly allow the cartesian product
]
```

```{r}
# Check:
nrow(age_models_pop_dt) / 
      length(unique(age_models_pop_dt$location)) /  # 204 locations
      nrow(count(age_models_pop_dt, sex, cause)) /  # 28: 12 * 2 + 3 female + 1 male
      length(unique(age_models_pop_dt$age)) /       # 96 1-year age groups
      length(unique(age_models_pop_dt$year)) /      # 2: 2000 sim start year & 2019 ref year
      length(unique(age_models_pop_dt$measure))     # 3: incid, prev, mort
```

#### Step 3: Predict 1Y Age-Groups Rates and Counts

```{r, eval=FALSE}
# gbd_epi_1yr_predict <- data.table::copy(age_models_pop_dt)
gbd_epi_1yr_predict <- age_models_pop_dt

gbd_epi_1yr_predict[
      , log_fitted_val := {
            if (!is.null(model[[1]])) {
                  predict(model[[1]], newdata = .SD) 
            } else NA_real_
      },
      by = .(location, sex, cause, year, measure)
][
      , pred_rates := exp(log_fitted_val)              # Calculate predicted rates
][
      , pred_cases := exp(log_fitted_val * pop / 1e5)  # Calculate predicted cases
][
      , model := NULL  # Remove model column
][
      , age_gp := cut(
            age,
            breaks = c(-Inf, seq(5, 95, 5), Inf),
            labels = c("<5 years",
                       paste0(seq(5, 90, 5), "-", seq(9, 94, 5), " years"),
                       "95+ years"),
            right = FALSE
      )
]
```

```{r}
# Check:
nrow(gbd_epi_1yr_predict) / 
      length(unique(gbd_epi_1yr_predict$location)) /  # 204 locations
      nrow(count(gbd_epi_1yr_predict, sex, cause)) /  # 28: 12 * 2 + 3 female + 1 male
      length(unique(gbd_epi_1yr_predict$age)) /       # 96 1-year age groups
      length(unique(gbd_epi_1yr_predict$year)) /      # 2: 2000 sim start year & 2019 ref year
      length(unique(gbd_epi_1yr_predict$measure))     # 3: incid, prev, mort
```

# Outputs

## Predicted `Log(Rates)`

```{r}
gbd_epi_1yr_predict %>%
      filter(sex == "Female",
             cause %in% c(cause2[-2], cause3[2]),
             location == "Uganda",
             measure != "Prevalence",
             year == 2019) %>%

      ggplot() +
      geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
      geom_line(aes(x = age, y = log(pred_rates)), linetype = 1, color = "green") +
      facet_grid(cause ~ measure, scales = "free_y") +  # Free y-axis scales
      labs(title = "Age-Value Regression Lines by Group",
           x = "Age", y = "Log value") +
      theme_caviz
```

## Predicted Rates

```{r}
gbd_epi_1yr_predict %>%
      filter(sex == "Female",
             cause %in% c(cause2[-2], cause3[2]),
             location == "Uganda",
             measure != "Prevalence",
             year == 2019) %>% 
      mutate(cause = case_match(cause,
                                cause2[1] ~ "Breast",
                                cause2[3] ~ "Lung",
                                cause3[2] ~ "CRC",
                                .default = NA)) %>% 

      ggplot() +
      geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
      geom_line(aes(x = age, y = pred_rates), linetype = 1, color = "green") +
      scale_y_continuous(labels = scales::label_number(scale = 1e-5, suffix = "K")) +
      facet_grid(cause ~ measure, scales = "free_y") +  # Free y-axis scales
      labs(title = "Rates (per 100K) among Females in Uganda, 2019",
           x = "Age", y = "Rate (per 100K)") +
      theme_caviz +
      theme(strip.text.y = element_blank()) +
      theme(strip.text.y = element_text(size = 12, angle = 0))
```

## Predicted Counts

```{r}
gbd_epi_1yr_predict %>%
      filter(sex == "Female",
             cause %in% c(cause2[-2], cause3[2]),
             location == "Uganda",
             measure != "Prevalence",
             year == 2019) %>% 
      mutate(cause = case_match(cause,
                                cause2[1] ~ "Breast",
                                cause2[3] ~ "Lung",
                                cause3[2] ~ "CRC",
                                .default = NA)) %>% 
      # mutate(calc_pred_cases = pred_rates * pop / 1e5) %>% 

      ggplot() +
      geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
      geom_line(aes(x = age, y = pred_cases), linetype = 1, color = "green") +
      facet_grid(cause ~ measure, scales = "free_y") +  # Free y-axis scales
      labs(title = "Case counts among Females in Uganda, 2019",
           x = "Age", y = "Case counts") +
      theme_caviz +
      theme(strip.text.y = element_blank()) +
      theme(strip.text.y = element_text(size = 12, angle = 0))
```

## Comparision: GBD vs. Predicted Values

```{r}
gbd_calc_compr <- left_join(

      # Linear model predicted counts (incid and deaths), all ages
      x  = gbd_epi_1yr_predict %>% 
            filter(year == 2019) %>% 
            select(!c(pop, log_fitted_val)) %>% 
            pivot_longer(cols = starts_with("pred_"),
                         names_to = "metric",
                         values_to = "pred_val") %>% 
            mutate(metric = ifelse(metric == "pred_rates", "Rate", "Number")),
                   # source = "Pred"),

      # GBD calculated counts (incid and deaths), all ages
      y = gbd_epi_5yr %>% 
            filter(year == 2019) %>% 
            select(!c(age, upper, lower, age_cat)), # %>% 
            # mutate(source = "GBD"),

      by = join_by(measure, location, year, cause, sex, age_gp, metric)
) %>% 
      pivot_longer(cols = ends_with("val"),
                   names_to = "source",
                   values_to = "val") %>% 
      mutate(source = ifelse(source == "pred_val", "Pred", "GBD"))
      # mutate(counts_diff = pred_cases - val,
      #        counts_ratio = pred_cases / val)
```

### Observed vs. Predicted Rates

#### Incidence

```{r}
subset_example <- gbd_calc_compr %>% 
      filter(sex == "Female",
             location %in% c("Uganda", "United States of America", "India", "China"))

# https://forum.posit.co/t/adding-manual-legend-to-ggplot2/41651/3
scale_col <- c("GBD" = "blue", "Pred" = "green")

subset_example %>% 
      split(.$location) %>% 
      map(.x = .,
          .f = ~ .x %>% 
                ggplot() +
                geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
                geom_point(data = .x %>% 
                                 filter(source == "GBD",
                                        measure == "Incidence",
                                        metric == "Rate"),
                           mapping = aes(x = age_mid, y = log(val / 1e5), color = "GBD")) +
                geom_line(data = .x %>% 
                                filter(source == "Pred",
                                       measure == "Incidence",
                                       metric == "Rate"),
                          mapping = aes(x = age, y = log(val / 1e5), color = "Pred")) +
                scale_color_manual(values = scale_col,
                                   guide = guide_legend()) +
                facet_wrap(. ~ cause, scales = "free_y") +  # Free y-axis scales
                labs(title = "(Log) Predicted Incidence vs. GBD (Log) Incidence Rates",
                     subtitle = paste0(unique(.x$location), ", Females, 2019"),
                     x = "Age", y = " (Log) Incidence rate",
                     color = NULL) +
                theme_minimal() +
                theme(strip.text.y = element_blank()) +
                theme(strip.text.y = element_text(size = 12, angle = 0),
                      legend.position = "bottom"))
```

#### Mortality

```{r}
subset_example %>% 
      split(.$location) %>% 
      map(.x = .,
          .f = ~ .x %>% 
                ggplot() +
                geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
                geom_point(data = .x %>% 
                                 filter(source == "GBD",
                                        measure == "Deaths",
                                        metric == "Rate"),
                           mapping = aes(x = age_mid, y = log(val / 1e5), color = "GBD")) +
                geom_line(data = .x %>% 
                                filter(source == "Pred",
                                       measure == "Deaths",
                                       metric == "Rate"),
                          mapping = aes(x = age, y = log(val / 1e5), color = "Pred")) +
                scale_color_manual(values = scale_col,
                                   guide = guide_legend()) +
                facet_wrap(. ~ cause, scales = "free_y") +  # Free y-axis scales
                labs(title = "(Log) Predicted Mortality vs. GBD (Log) Mortality Rates",
                     subtitle = paste0(unique(.x$location), ", Females, 2019"),
                     x = "Age", y = "(Log) Mortality rate",
                     color = NULL) +
                theme_minimal() +
                theme(strip.text.y = element_blank()) +
                theme(strip.text.y = element_text(size = 12, angle = 0),
                      legend.position = "bottom"))
```

### Observed vs. Predicted Counts

#### Incidence

```{r}
subset_example <- gbd_calc_compr %>% 
      filter(sex == "Female",
             location %in% c("Uganda", "United States of America", "India", "China"))

# https://forum.posit.co/t/adding-manual-legend-to-ggplot2/41651/3
scale_col <- c("GBD" = "blue", "Pred" = "red")

# subset_example %>%
#       filter(measure == "Incidence", 
#              location == "Uganda",
#              cause == cause2[1],
#              source == "Pred",
#              age == 79,
#              metric == "Rate")

subset_example %>% 
      # Use the same incidence and mortality rates for 80+ as for 79
      mutate(val = ifelse(source == "Pred" & metric == "Rate" & age >= 80,
                          val[source == "Pred" & metric == "Rate" & age == 79], val),
             .by = c(measure, location, cause, sex, year)) %>% 

      left_join(x  = .,
                y  = gbd_pop %>% select(location, sex, year, age, pop = val),
                by = join_by(location, sex, year, age)) %>%
      # mutate(val = ifelse(source == "Pred" & metric == "Number" & age >= 80,
      #                     val[source == "Pred" & metric == "Rate" & age == 79] * pop, val),
      #        .by = c(measure, location, cause, sex, year)) %>%
      mutate(val = ifelse(measure == "Number", val * pop, val)) %>%
      
      # filter(location == "Uganda", cause == cause2[1], source == "Pred", age >= 60)
      # filter(location == "Uganda") %>%
      
      split(.$location) %>% 
      map(.x = .,
          .f = ~ .x %>% 
                ggplot() +
                geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
                geom_point(data = .x %>% 
                                 filter(source == "GBD",
                                        measure == "Incidence",
                                        metric == "Number"),
                           mapping = aes(x = age_mid, y = val, color = "GBD")) +
                geom_line(data = .x %>% 
                                filter(source == "Pred",
                                       measure == "Incidence",
                                       metric == "Number"),
                          mapping = aes(x = age, y = val, color = "Pred")) +
                scale_color_manual(values = scale_col,
                                   guide = guide_legend()) +
                facet_wrap(. ~ cause, scales = "free_y") +  # Free y-axis scales
                labs(title = "Predicted Incident vs. GBD Incidence Counts",
                     subtitle = paste0(unique(.x$location), ", Females, 2019"),
                     x = "Age", y = "Case count",
                     color = NULL) +
                theme_minimal() +
                theme(strip.text.y = element_blank()) +
                theme(strip.text.y = element_text(size = 12, angle = 0),
                      legend.position = "bottom"))
```

#### Mortality

```{r}
subset_example %>% 
      # Use the same incidence and mortality rates for 80+ as for 79
      mutate(val = ifelse(source == "Pred" & metric == "Rate" & age >= 80,
                          val[source == "Pred" & metric == "Rate" & age == 79], val),
             .by = c(measure, location, cause, sex, year)) %>% 

      left_join(x  = .,
                y  = gbd_pop %>% select(location, sex, year, age, pop = val),
                by = join_by(location, sex, year, age)) %>%
      # mutate(val = ifelse(source == "Pred" & metric == "Number" & age >= 80,
      #                     val[source == "Pred" & metric == "Rate" & age == 79] * pop, val),
      #        .by = c(measure, location, cause, sex, year)) %>%
      mutate(val = ifelse(measure == "Number", val * pop, val)) %>%
      
      split(.$location) %>% 
      map(.x = .,
          .f = ~ .x %>% 
                ggplot() +
                geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
                geom_point(data = .x %>% 
                                 filter(source == "GBD",
                                        measure == "Deaths",
                                        metric == "Number"),
                           mapping = aes(x = age_mid, y = val, color = "GBD")) +
                geom_line(data = .x %>% 
                                filter(source == "Pred",
                                       measure == "Deaths",
                                       metric == "Number"),
                          mapping = aes(x = age, y = val, color = "Pred")) +
                scale_color_manual(values = scale_col,
                                   guide = guide_legend()) +
                facet_wrap(. ~ cause, scales = "free_y") +  # Free y-axis scales
                labs(title = "Predicted Mortality vs. GBD Mortality Counts",
                     subtitle = paste0(unique(.x$location), ", Females, 2019"),
                     x = "Age", y = "Death count",
                     color = NULL) +
                theme_minimal() +
                theme(strip.text.y = element_blank()) +
                theme(strip.text.y = element_text(size = 12, angle = 0),
                      legend.position = "bottom"))
```

#### 5-Year Counts

##### Incidence

```{r}
subset_example %>% 
      # Use the same incidence and mortality rates for 80+ as for 79
      mutate(val = ifelse(source == "Pred" & metric == "Rate" & age >= 80,
                          val[source == "Pred" & metric == "Rate" & age == 79], val),
             .by = c(measure, location, cause, sex, year)) %>% 

      left_join(x  = .,
                y  = gbd_pop %>% select(location, sex, year, age, pop = val),
                by = join_by(location, sex, year, age)) %>%
      # mutate(val = ifelse(source == "Pred" & metric == "Number" & age >= 80,
      #                     val[source == "Pred" & metric == "Rate" & age == 79] * pop, val),
      #        .by = c(measure, location, cause, sex, year)) %>%
      mutate(val = ifelse(measure == "Number", val * pop, val)) %>% 
      
      # Convert to 5-year age groups
      filter(metric == "Number") %>% 
      group_by(measure, metric, location, cause, sex, year, age_gp, source) %>% 
      reframe(val = ifelse(source == "Pred", sum(val), unique(val)),
              age_mid = unique(age_mid)) %>% 
      
      # filter(location == "Uganda", cause == cause2[1], source == "Pred", age >= 60)
      # filter(location == "Uganda") %>%
      
      split(.$location) %>% 
      map(.x = .,
          .f = ~ .x %>% 
                ggplot() +
                geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
                geom_point(data = .x %>% 
                                 filter(source == "GBD",
                                        measure == "Incidence",
                                        metric == "Number"),
                           mapping = aes(x = age_mid, y = val, color = "GBD")) +
                geom_point(data = .x %>% 
                                filter(source == "Pred",
                                       measure == "Incidence",
                                       metric == "Number"),
                          mapping = aes(x = age_mid, y = val, color = "Pred")) +
                scale_color_manual(values = scale_col,
                                   guide = guide_legend()) +
                facet_wrap(. ~ cause, scales = "free_y") +  # Free y-axis scales
                labs(title = "Predicted Incident vs. GBD Incidence Counts",
                     subtitle = paste0(unique(.x$location), ", Females, 2019"),
                     x = "Age", y = "Case count",
                     color = NULL) +
                theme_minimal() +
                theme(strip.text.y = element_blank()) +
                theme(strip.text.y = element_text(size = 12, angle = 0),
                      legend.position = "bottom"))
```

#### Deaths

```{r}
subset_example %>% 
      # Use the same incidence and mortality rates for 80+ as for 79
      mutate(val = ifelse(source == "Pred" & metric == "Rate" & age >= 80,
                          val[source == "Pred" & metric == "Rate" & age == 79], val),
             .by = c(measure, location, cause, sex, year)) %>% 

      left_join(x  = .,
                y  = gbd_pop %>% select(location, sex, year, age, pop = val),
                by = join_by(location, sex, year, age)) %>%
      # mutate(val = ifelse(source == "Pred" & metric == "Number" & age >= 80,
      #                     val[source == "Pred" & metric == "Rate" & age == 79] * pop, val),
      #        .by = c(measure, location, cause, sex, year)) %>%
      mutate(val = ifelse(measure == "Number", val * pop, val)) %>%
      
      # Convert to 5-year age groups
      filter(metric == "Number") %>% 
      group_by(measure, metric, location, cause, sex, year, age_gp, source) %>% 
      reframe(val = ifelse(source == "Pred", sum(val), unique(val)),
              age_mid = unique(age_mid)) %>% 
      
      
      split(.$location) %>% 
      map(.x = .,
          .f = ~ .x %>% 
                ggplot() +
                geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
                geom_point(data = .x %>% 
                                 filter(source == "GBD",
                                        measure == "Deaths",
                                        metric == "Number"),
                           mapping = aes(x = age_mid, y = val, color = "GBD")) +
                geom_point(data = .x %>% 
                                filter(source == "Pred",
                                       measure == "Deaths",
                                       metric == "Number"),
                          mapping = aes(x = age_mid, y = val, color = "Pred")) +
                scale_color_manual(values = scale_col,
                                   guide = guide_legend()) +
                facet_wrap(. ~ cause, scales = "free_y") +  # Free y-axis scales
                labs(title = "Predicted Mortality vs. GBD Mortality Counts",
                     subtitle = paste0(unique(.x$location), ", Females, 2019"),
                     x = "Age", y = "Death count",
                     color = NULL) +
                theme_minimal() +
                theme(strip.text.y = element_blank()) +
                theme(strip.text.y = element_text(size = 12, angle = 0),
                      legend.position = "bottom"))
```

### ACM

```{r}
# gbd_acm_1yr_predict

gbd_acm_calc_compr <- left_join(

      # Linear model predicted counts (incid and deaths), all ages
      x  = gbd_acm_1yr_predict %>% 
            filter(year == 2019) %>% 
            select(!c(pop, log_fitted_val)) %>% 
            pivot_longer(cols = starts_with("pred_"),
                         names_to = "metric",
                         values_to = "pred_val") %>% 
            mutate(metric = ifelse(metric == "pred_rates", "Rate", "Number")) %>% 
            
            mutate(age_gp = cut(
                  age,
                  breaks = c(-Inf, seq(5, 95, 5), Inf),
                  labels = c("<5 years",
                             paste0(seq(5, 90, 5), "-", seq(9, 94, 5), " years"),
                             "95+ years"),
                  right = FALSE
            )),
                   # source = "Pred"),

      # GBD calculated acm counts, all ages
      y = gbd_acm_5yr %>% #altern_acm_raw %>% # 
            filter(year == 2019) %>%
            # left_join(x  = .,
            #           y  = gbd_acm_raw %>% 
            #                 rename(age_cat = age) %>% 
            #                 filter(age_cat %in% c("<1 year"))) %>% 
            # mutate(age_gp = case_when(age %in% c("<28 days",
            #                                      "1-5 months",
            #                                      "6-11 months")  ~ "<1 year",
            #                           age %in% c("12-23 months",
            #                                      "2-4 years")    ~ "1-5 years",
            #                           TRUE                       ~ age)) %>%
            
            mutate(age_gp = cut(
                  age,
                  breaks = c(-Inf, seq(5, 95, 5), Inf),
                  labels = c("<5 years",
                             paste0(seq(5, 90, 5), "-", seq(9, 94, 5), " years"),
                             "95+ years"),
                  right = FALSE
            )) %>% 
            
            select(!c(age, upper, lower)),
            # select(!c(age, upper, lower)),
            # mutate(source = "GBD"),

      by = join_by(measure, location, year, cause, sex, metric, age_gp)
) %>% 
      pivot_longer(cols = ends_with("val"),
                   names_to = "source",
                   values_to = "val") %>% 
      mutate(source = ifelse(source == "pred_val", "Pred", "GBD"))
```

```{r}
acm_subset_example <- gbd_acm_calc_compr %>% 
      filter(sex == "Female",
             location %in% c("Uganda", "United States of America", "India", "China"))
```

#### 5-Year Death Rates

```{r}
acm_subset_example %>% 
      # Use the same incidence and mortality rates for 80+ as for 79
      # mutate(val = ifelse(source == "Pred" & metric == "Rate" & age >= 80,
      #                     val[source == "Pred" & metric == "Rate" & age == 79], val),
      #        .by = c(measure, location, cause, sex, year)) %>% 

      # left_join(x  = .,
      #           y  = gbd_pop %>% select(location, sex, year, age, pop = val),
      #           by = join_by(location, sex, year, age)) %>%
      # mutate(val = ifelse(source == "Pred" & metric == "Number" & age >= 80,
      #                     val[source == "Pred" & metric == "Rate" & age == 79] * pop, val),
      #        .by = c(measure, location, cause, sex, year)) %>%
      # mutate(val = ifelse(measure == "Number", val * pop, val)) %>% 
      
      filter(metric == "Rate") %>% 
      
      split(.$location) %>% 
      map(.x = .,
          .f = ~ .x %>% 
                ggplot() +
                geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
                geom_point(data = .x %>% 
                                 filter(source == "GBD"),
                           mapping = aes(x = age_mid, y = val, color = "GBD")) +
                geom_line(data = .x %>% 
                                filter(source == "Pred"),
                          mapping = aes(x = age, y = val, color = "Pred")) +
                scale_color_manual(values = scale_col,
                                   guide = guide_legend()) +
                scale_y_continuous(limits = c(0, 35e3),
                                   breaks = seq(0, 3e4, 1e4),
                                   labels = scales::label_comma()) +
                facet_wrap(. ~ cause, scales = "free_y") +  # Free y-axis scales
                labs(title = "Predicted Mortality vs. GBD Mortality Rates",
                     subtitle = paste0(unique(.x$location), ", Females, 2019"),
                     x = "Age", y = "Mortality rate (per 100K)",
                     color = NULL) +
                theme_minimal() +
                theme(strip.text.y = element_blank()) +
                theme(strip.text.y = element_text(size = 12, angle = 0),
                      legend.position = "bottom"))
```

```{r}
altern_acm_raw <- gbd_acm_raw %>% 
      filter(year == 2019) %>% 
      filter(metric == "Rate") %>% 
      filter(location %in% c("Uganda", "China", "India", "United States of America")) %>% 
      mutate(age_fct = factor(age,
                              levels = uniq_raw_age[c(22, 25, 26, 17, 21,
                                                      1:16, 18:20, 23)])) %>% 
      # mutate(age_mid = ifelse(grep("^\\d+-\\d+ years$|^\\d+\\+ years$", age, value = TRUE), as.numeric(parse_number(age)) + 2.5, ))
      
      mutate(age_mid = case_when(age == "<28 days"    ~ 15 / 30 / 12,
                                 age == "1-5 months"  ~ 3 / 12,
                                 age == "6-11 months" ~ 8.5 / 12,
                                 age == "<1 year"     ~ 0.5,
                                 age == "2-4 years"   ~ 3,
                                 age == "<5 years"    ~ 2.5,
                                 TRUE                 ~ 
                                       as.numeric(parse_number(age)) + 2.5))
```

```{r}
uniq_raw_age <- unique(gbd_acm_raw$age)

left_join(
      x  = gbd_acm_raw %>% 
            filter(year == 2019) %>% 
            filter(metric == "Rate") %>% 
            filter(location %in% c("Uganda", "China", "India", "United States of America")) %>% 
            mutate(age_fct = factor(age,
                                    levels = uniq_raw_age[c(22, 25, 26, 17, 21,
                                                            1:16, 18:20, 23)])) %>% 
            # mutate(age_mid = ifelse(grep("^\\d+-\\d+ years$|^\\d+\\+ years$", age, value = TRUE), as.numeric(parse_number(age)) + 2.5, ))
            
            mutate(age_mid = case_when(age == "<28 days"    ~ 15 / 30 / 12,
                                       age == "1-5 months"  ~ 3 / 12,
                                       age == "6-11 months" ~ 8.5 / 12,
                                       age == "<1 year"     ~ 0.5,
                                       age == "2-4 years"   ~ 3,
                                       age == "<5 years"    ~ 2.5,
                                       TRUE                 ~ 
                                             as.numeric(parse_number(age)) + 2.5)),
      
      y = gbd_acm_5yr %>% 
            filter(year == 2019) %>% 
            filter(age_cat %!in% c("<28 days", "1-5 months", "6-11 months",
                                  "12-23 months")) %>% 
            select(!c(age, upper, lower, age_cat))
) %>% 
      
      # filter(between(age_mid, 25, 40))
      
      ggplot(mapping = aes(x = age_mid, y = val, color = sex)) +
      geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
      geom_point(alpha = 0.35) +
      # geom_line(data = .x %>% 
      #                 filter(source == "Pred",
      #                        measure == "Deaths",
      #                        metric == "Rate"),
      #           mapping = aes(x = age, y = val, color = "Pred")) +
      # scale_color_manual(values = scale_col,
      #                    guide = guide_legend()) +
      facet_wrap(. ~ location) +  # Free y-axis scales
      labs(title = "Predicted Mortality vs. GBD Mortality Counts",
           # subtitle = paste0(unique(.x$location), ", Females, 2019"),
           x = "Age", y = "Mortality rate (per 100K)",
           color = NULL) +
      theme_minimal() +
      theme(strip.text.y = element_blank()) +
      theme(strip.text.y = element_text(size = 12, angle = 0),
            legend.position = "bottom")
```

#### 5-Year Death Counts

```{r}
acm_subset_example %>% 
      filter(#source == "GBD",
             metric == "Number",
             age_gp %in% c("<1 year", "1-5 years", "5-9 years"))
```

```{r}
acm_subset_example %>% 
      # Use the same incidence and mortality rates for 80+ as for 79
      # mutate(val = ifelse(source == "Pred" & metric == "Rate" & age >= 80,
      #                     val[source == "Pred" & metric == "Rate" & age == 79], val),
      #        .by = c(measure, location, cause, sex, year)) %>% 

      # left_join(x  = .,
      #           y  = gbd_pop %>% select(location, sex, year, age, pop = val),
      #           by = join_by(location, sex, year, age)) %>%
      
      # mutate(val = ifelse(source == "Pred" & metric == "Number" & age >= 80,
      #                     val[source == "Pred" & metric == "Rate" & age == 79] * pop, val),
      #        .by = c(measure, location, cause, sex, year)) %>%
      filter(metric == "Number") %>% 
      
      # mutate(val = ifelse(test = metric == "Rate" & source == "Pred",
      #                     yes  = val * pop / 1e5,
      #                     no   = val)) %>% 
      # filter(case_when(source == "GBD" ~ metric == "Number"))
      
      # Convert to 5-year age groups
      filter(age_cat %!in% c("<28 days", "1-5 months", "6-11 months",
                            "<1 year", "12-23 months", "2-4 years")) %>% 
      
      # mutate(age_gp = cut(
      #       age,
      #       breaks = c(-Inf, seq(5, 95, 5), Inf),
      #       labels = c("<5 years",
      #                  paste0(seq(5, 90, 5), "-", seq(9, 94, 5), " years"),
      #                  "95+ years"),
      #       right = FALSE
      # )) %>% 
      
      # filter(location == "Uganda", sex == "Female", year == 2019, age < 15)
      
      # Check:
      # split(.$measure, .$metric, .$location, .$cause, .$sex, .$year, .$age_gp) %>% 
      # map(.x = .,
      #     .f = ~ .x %>% 
      #           filter(source == "GBD") %>% 
      #           select(val) %>% 
      #           pull() %>% 
      #           table())
      
      # filter(age < 15)
      # 
      # select(age_mid) %>% 
      # unique()

      group_by(measure, metric, location, cause, sex, year, age_gp, source) %>% 
      reframe(val = ifelse(
            test = source == "Pred",
            yes  = sum(val),
            no   = mean(val))) %>% 
      unique() %>% 
      
      mutate(age_mid = case_when(age_gp == "<5 years"  ~ 2.5,
                                 TRUE                  ~ as.numeric(parse_number(as.character(age_gp))) + 2.5)) %>%
      # filter(location == "Uganda", cause == cause2[1], source == "Pred", age >= 60)
      # filter(location == "Uganda") %>%
      
      split(.$location) %>% 
      map(.x = .,
          .f = ~ .x %>% 
                ggplot() +
                geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
                geom_point(data = .x %>% 
                                 filter(source == "GBD"),
                           mapping = aes(x = age_mid, y = val, color = "GBD")) +
                geom_point(data = .x %>% 
                                filter(source == "Pred"),
                          mapping = aes(x = age_mid, y = val, color = "Pred")) +
                scale_color_manual(values = scale_col,
                                   guide = guide_legend()) +
                scale_y_continuous(# limits = c(0, 13e5),
                                   # breaks = seq(0, 1.25e6, 25e4),
                                   labels = scales::label_comma()) +
                facet_wrap(. ~ cause, scales = "free_y") +  # Free y-axis scales
                labs(title = "Predicted vs. GBD Incidence Counts",
                     subtitle = paste0(unique(.x$location), ", Females, 2019"),
                     x = "Age", y = "Death count",
                     color = NULL) +
                theme_minimal() +
                theme(strip.text.y = element_blank()) +
                theme(strip.text.y = element_text(size = 12, angle = 0),
                      legend.position = "bottom"))
```


#### Subset Cancers

```{r}
uganda_example_subset <- gbd_calc_compr %>% 
      filter(sex == "Female",
             cause %in% c(cause2[-2], cause3[2]),
             location == "Uganda") %>% 
      filter(measure != "Prevalence") %>% 
      mutate(cause = case_match(cause,
                                cause2[1] ~ "Breast",
                                cause2[3] ~ "Lung",
                                cause3[2] ~ "CRC",
                                .default = NA))

      
uganda_example_subset %>% 
      ggplot() +
      geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
      geom_point(data = uganda_example_subset %>% 
                       filter(source == "GBD") %>% 
                       filter(metric == "Rate"),
                 mapping = aes(x = age_mid, y = val), color = "blue", alpha = 0.6) +
      geom_line(data = uganda_example_subset %>% 
                       filter(source == "Pred") %>% 
                       filter(metric == "Rate"),
                 mapping = aes(x = age, y = val), color = "green") +
      # geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
      facet_grid(cause ~ measure, scales = "free_y") +  # Free y-axis scales
      labs(title = "Predicted cases vs. GBD Rates",
           subtitle = "Uganda, Females, 2019",
           x = "Age", y = "Value") +
      theme_caviz +
      theme(strip.text.y = element_blank()) +
      theme(strip.text.y = element_text(size = 12, angle = 0))
```

```{r}
# check:
nrow(uganda_example_subset) /                          # 2304 rows
      length(unique(uganda_example_subset$age)) /      # 96 single-year ages
      length(unique(uganda_example_subset$cause)) /    # 3 causes
      length(unique(uganda_example_subset$measure)) /  # 2 measures: deaths and incid
      length(unique(uganda_example_subset$metric)) /   # 2 metrics: rate and count
      length(unique(uganda_example_subset$source))     # 2 sources: GBD and pred
```


### Observed vs. Predicted Case Counts

```{r}
uganda_example_subset %>% 
      ggplot() +
      geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
      geom_point(data = uganda_example_subset %>% 
                       filter(source == "GBD") %>% 
                       filter(metric == "Number"),
                 mapping = aes(x = age_mid, y = val, color = "GBD"), alpha = 0.6) +
      geom_line(data = uganda_example_subset %>% 
                      filter(source == "Pred") %>% 
                       filter(metric == "Number"),
                 mapping = aes(x = age, y = val, color = "Pred")) +
      scale_color_manual(values = scale_col,
                         guide = guide_legend()) +
      facet_grid(cause ~ measure, scales = "free_y") +  # Free y-axis scales
      labs(title = "Predicted cases vs. GBD case counts",
           subtitle = "Uganda, Females, 2019",
           x = "Age", y = "Value") +
      theme_caviz +
      theme(strip.text.y = element_blank()) +
      theme(strip.text.y = element_text(size = 12, angle = 0),
            legend.position = "bottom")
```

### Observed vs. Predicted Case `Log(Counts)`

```{r}
uganda_example_subset %>% 
      ggplot() +
      geom_hline(yintercept = 0, linetype = 2, color = "darkred") +
      geom_point(data = uganda_example_subset %>% 
                       filter(source == "GBD") %>% 
                       filter(metric == "Number"),
                 mapping = aes(x = age_mid, y = log(val), color = "GBD"), alpha = 0.6) +
      geom_line(data = uganda_example_subset %>% 
                      filter(source == "Pred") %>% 
                       filter(metric == "Number"),
                 mapping = aes(x = age, y = log(val), color = "Pred")) +
      scale_color_manual(values = scale_col,
                         guide = guide_legend()) +
      facet_grid(cause ~ measure, scales = "free_y") +  # Free y-axis scales
      labs(title = "Log predicted cases vs. GBD log case counts",
           subtitle = "Uganda, Females, 2019",
           x = "Age", y = "Value") +
      theme_caviz +
      theme(strip.text.y = element_blank()) +
      theme(strip.text.y = element_text(size = 12, angle = 0),
            legend.position = "bottom")
```

### Histogram of Case Counts

```{r}
gbd_calc_compr %>% 
      group_by(measure, location, cause, sex, source) %>%
      reframe(tot_cases = sum(val, na.rm = TRUE)) %>% 
      
      pivot_wider(id_cols = c(measure, location, cause, sex),
                  names_from = source,
                  values_from = tot_cases) %>% 
      
      ggplot() +
      geom_histogram(aes(x = GBD, fill = "GBD"), alpha = 0.5, position = "identity") +
      geom_histogram(aes(x = Pred, fill = "Pred"), alpha = 0.5, position = "identity") +
      scale_x_continuous(breaks = seq(0, 2e7, 5e6),
                         labels = scales::label_comma(scale = 1e-6, suffix = "M")) +
      scale_fill_manual(values = scale_col,
                        guide = guide_legend()) +
      facet_wrap(. ~ cause) +
      labs(x = "Case counts",
           fill = NULL) +
      theme_minimal() +
      theme(legend.position = "bottom")
```

### Histogram of Difference in Case Counts

```{r}
gbd_calc_compr %>% 
      group_by(measure, location, cause, sex, source) %>%
      reframe(tot_cases = sum(val, na.rm = TRUE)) %>% 
      
      pivot_wider(id_cols = c(measure, location, cause, sex),
                  names_from = source,
                  values_from = tot_cases) %>% 
      mutate(counts_diff = Pred - GBD,
             counts_ratio = Pred / GBD) %>% 
      
      # reframe(range(counts_diff))
      
      ggplot() +
      geom_histogram(aes(x = counts_diff), alpha = 0.5, position = "identity", fill = "red") +
      # geom_histogram(aes(x = counts_ratio), alpha = 0.5, position = "identity", fill = "blue") +
      scale_x_continuous(breaks = seq(-14e6, 12e6, 6e6),
                         labels = scales::label_comma(scale = 1e-6, suffix = "M")) +
      facet_wrap(. ~ cause) +
      labs(x = "Counts difference between GBD and predictions") +
      theme_minimal()
```

### Histogram of Case Counts Ratios

```{r}
gbd_calc_compr %>% 
      group_by(measure, location, cause, sex, source) %>%
      reframe(tot_cases = sum(val, na.rm = TRUE)) %>% 
      
      pivot_wider(id_cols = c(measure, location, cause, sex),
                  names_from = source,
                  values_from = tot_cases) %>% 
      mutate(counts_diff = Pred - GBD,
             counts_ratio = Pred / GBD) %>% 
      
      # reframe(range(counts_diff))
      
      ggplot() +
      # geom_histogram(aes(x = counts_diff), alpha = 0.5, position = "identity", fill = "red") +
      geom_histogram(aes(x = counts_ratio), alpha = 0.5, position = "identity", fill = "blue") +
      # scale_x_continuous(breaks = seq(0, 1.75e3, 500),
      #                    labels = scales::label_comma(scale = 1e-3, suffix = "K")) +
      facet_wrap(. ~ cause) +
      labs(x = "Counts Ratio (predictions to GBD)") +
      theme_minimal()
```

